---
title: "Functional traits"
editor: source
eval: false
echo: true
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 3
---


# Assign functional traits based on Jones et al. 2025. 

```{r, message=FALSE, warning=FALSE}
# install.packages("stringdist")
library(tidyverse)
library(stringdist)
library(data.table)
library(fuzzyjoin)
```

## Import dataframes to align.
```{r}
jones <- read.csv("../tax_annotation_proj/TMDv1.1_Jones.csv") %>% 
  separate(`Grazer.Taxonomic.ID`, into = c("GENUS", "SPECIES"), sep = " ", remove = FALSE) %>% 
  # separate_longer_delim(`Synonym.Basionym`, delim = ", ") %>% 
  # separate(`Synonym.Basionym`, into = c("alt_genus", "alt_species"), sep = " ", remove = FALSE) %>% 
  mutate(SPECIES = str_squish(SPECIES),
         GENUS = str_squish(GENUS)) %>% 
         # alt_species = str_squish(alt_species),
         # alt_genus = str_squish(alt_genus)) %>% 
  mutate(GENUS_SPECIES = paste(GENUS, SPECIES, sep = " "))
         # genus_species = paste(alt_genus, alt_species, sep = " "))# %>% 
  # select(Taxon = `Grazer.Taxonomic.ID`, GENUS_SPECIES, genus_species)
  
# head(jones)

gom <- read.csv("../tax_annotation_proj/tax_to_annotate.csv") %>% 
  # filter(grepl("Dinoflagellata|Ciliophora", Taxon)) %>% 
  separate(Taxon, into = c("Domain", "Supergroup", "Division", "Subdivision", "Class", "Order", "Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
  separate(Species, into = c("GENUS", "SPECIES"), sep="_(?=[^_]+$)") %>% 
  filter(SPECIES != "sp.") %>% 
  unite(GENUS_SPECIES, GENUS, SPECIES, sep = " ")
# head(gom)
# unique(gom$Subdivision)
```

```{r}
# library(fuzzyjoin)
joined <- fuzzyjoin::stringdist_left_join(
  gom, jones,
  by = c("GENUS_SPECIES"),
  distance_col = "DIST",
  ignore_case = TRUE,
  method = "lv"
) %>% 
  mutate(Morphotype = (str_replace_all(Morphotype, "eGNCM", "eSNCM"))) #revision based on Mitra et al. mixotrophy database

head(joined)
# table(joined$GENUS_SPECIES.y)
unique(joined$Morphotype)
```

## Explore joined table.

```{r}
head(joined)
```




Jones et al. imported ciliate and dinoflagellate functional traits. Our Gulf data originally had how many ASVs associated with dinoflagellates or ciliates? and now how many have a functional trait?
```{r}
# Dino ASVs
length(unique((joined %>% 
  filter(Subdivision == "Dinoflagellata"))$FeatureID))

# Cilite ASVs
length(unique((joined %>% 
  filter(Subdivision == "Ciliophora"))$FeatureID))
```
For dinoflagellates & ciliates, get list of what has been assigned.
```{r}
result_joined <- joined %>% 
  filter(Subdivision == "Dinoflagellata" | 
           Subdivision == "Ciliophora") %>% 
  select(FeatureID, Subdivision, Feeding.Mode, Morphotype, GENUS_SPECIES.x, GENUS_SPECIES.y)
```

## Summary of feeding modes and morphotypes
```{r}
table(result_joined %>% select(Subdivision, Feeding.Mode))

table(result_joined %>% select(Subdivision, Morphotype))
```

*CM*: Constitutive mixotroph
*NCM*: non-constitutive mixotroph
*GNCM*: generalist non-constitutive mixotroph
*SNCM*: specialist non-constitutive mixotroph

Functional type definitions (Mitra et al.:
* _CM_: Constitutive mixotroph that has a natural ability to photosynthesize.
* _CM*_: Evidence for constitutive mixotrophy, but not recorded
* _eGNCM_: Acquires photosynthetic machinery from a variety of prey and retains as endosymbionts. 
* _PA_: Photoautotroph
* _pSNCM_: Mixotroph that steals photosynthetic machinery from a specific prey type. plastidic specialist non-constitutive mixotroph.
* _SH_: Strict heterotrophy
* _SH*_: Evidence for strict heterotroph, but not recorded

```{r}
# joined %>% 
#   filter(Morphotype == "eSNCM")
```


```{r}
result_joined %>% 
  group_by(Subdivision) %>% 
  summarise(total = n(),
            No_assignment = sum(is.na(Feeding.Mode))) %>% 
  mutate(Percent_wNone = (No_assignment/total)*100)
```

Will likely focus on dinoflagellates for this work, as 80% of ASVs have been assigned trophic mode. Make trophic mode key for this Gulf study.

```{r}
colnames(joined)
dino_key <- joined %>% 
  filter(Subdivision == "Dinoflagellata") %>% distinct()
# write.csv(dino_key, file = "input-data/dino_fxn_trait.csv")
```

