---
title: "2023 Gulf data analysis"
editor: source
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 4
---

# Contents & introduction

> Physical Connectivity Shapes Signature Microeukaryotic Communities Across the Northern Gulf
> River Plumes & Eddy Fields observations of microeukaryotic biodiversity in the Northern Gulf
> Eddies, Riverine Inputs, and Environmental Niches Shape Signature Microeukaryotic Taxa of the Northern Gulf
> Community connectivity and functional partitioning across plume-to-pelagic gradients


_Core questions:_

(1) How does the single-celled eukaryotic (microeukaryote or protistan) community composition and diversity change from the Mississippi River plume to offshore? 

(2) What environmental factors play a role in shaping the microeukaryotic community dynamics at shallow versus deep depths in the Gulf of Mexico? 

(3) How do putative functional roles among microeukaryotes vary across critical environmental gradients? 

(4) Are there specific taxonomic groups, classes, or species that may be considered indicative of the Northern Gulf ecosystem? Especially along specific temperature, salinity, oxygen, and nutrient trends?

## Set up R environment

Set up R & import starting data

```{r, warning=FALSE, message=FALSE}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

Below is direct from `GoM-2023-amplicons.qmd`
```{r}
load("input-data/gulf-2023-sequence-analysis_02062026.RData", verbose = TRUE)
```


## Ordering & factoring

```{r}
offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5",
                             "S9", "S8", "S7", 
                             "S11", "S12", "S14", "S15")
stn_order_shore_offshore <- c(1, 2, 3, 4, 5, 
                              9, 8, 7, 6, 
                              10, 11, 12, 13, 14, 15, 16, 17)

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(metadata_seqsamples$Depth))))

depth_order_round <- as.character(sort(round(unique(metadata_seqsamples$Depth), digits = 0)))

depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

watercol_features <- c("other", "DCM", "Salinity maximum", "Oxygen minimum", "Salinity maximum, DCM")
watercol_features_label <- c("Depth with no feature", "DCM", "Salinity maximum", "Oxygen minimum", "Salinity maximum, DCM")
watercol_shape <- c(21, 24, 25, 22, 23)
names(watercol_shape) <- watercol_features_label
# watercol_shape
```


## Study stats

```{r}
(unique(asv_long_wtax_clean$Station)) # 12 stations
length(unique(asv_long_wtax_clean$Station)) # 12 stations
length(unique(asv_long_wtax_clean$STN_NISKIN)) # 90 unique samples
```

# K-means clustering by environmental parameters

```{r}
library(anomalize)
library(cluster)
library(factoextra)
set.seed(9450)
```

Normalize across variables. Will not include categorical values. We are grouping STN_NISKIN.

```{r}
env_only_mat <- metadata_seqsamples %>% 
  select(STN_NISKIN, Depth, ENV_VARIABLE, value) %>% distinct() %>%
  pivot_wider(names_from = ENV_VARIABLE, values_from = value) %>% distinct() %>% 
  select(-Prok_count) %>% 
  drop_na() %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix
```

Normalize matrix of numeric metadata.
```{r}
# Normalize
standarize_metadata <- decostand(env_only_mat, MARGIN = 2, method = "standardize")

# Output from here should be stn_niskin
rownames(standarize_metadata) 
```

First group to see how many clusters we need. 
```{r}
kclusts <-
  tibble(k = 2:10) %>%
  mutate(
    kclust = map(k, ~ kmeans(standarize_metadata, .x)),
    glanced = map(kclust, broom::glance))

# Unnest
output_kclusts <- kclusts %>%
  unnest(cols = c(glanced))

# plot
output_kclusts %>%
  ggplot(aes(x = k, y = tot.withinss)) +
    geom_line(alpha = 0.5, linewidth = 1.2, color = "darkgreen") +
    geom_point(size = 2, color = "darkgreen") +
  theme_linedraw()
```
> Above plot shows hypothetical numbers of k clusters (x-axis) with the variance within the cluster (y-axis). We want to isolate the number of clusters where we cover as much variance as possible - BUT if we add more clusters, not much will be helpful. This means "the bend" here. In this case, We are going to go with 6 clusters. 

Use chosen number of clusters (n=6) to compute kmeans.
```{r}
kmeans_env_6 <- kmeans(standarize_metadata, centers = 6)
summary(kmeans_env_6)
# Report some findings using "broom"
broom::tidy(kmeans_env_6)
broom::glance(kmeans_env_6)
```

Combine kmeans clustering with original data.
```{r}
# Combine with data
output_kmeans_env_6 <- broom::augment(kmeans_env_6, standarize_metadata)
head(data.frame(output_kmeans_env_6)) # Prints cluster membership at end of this dataframe.

kmeans_membership_6 <- data.frame(output_kmeans_env_6) %>% 
  select(STN_NISKIN = `.rownames`, CLUSTER = `.cluster`) %>% distinct()

# save(kmeans_membership_6, file = "input-data/kmeans-membership.RData")
```


#### Table with features

```{r}
load("input-data/kmeans-membership.RData", verbose = TRUE)
```

```{r}
head(metadata_stn_info)
head(metadata_seqsamples)
head(kmeans_membership_6)
unique(kmeans_membership_6$STN_NISKIN)
```


```{r}
table_wfeature <- metadata_seqsamples %>% 
  left_join(metadata_stn_info %>% 
  select(Station, Transect, OilRig_Count, MLD, SAL_MAX, O2_MIN, DCM, STN_CATEGORY_3, STN_CATEGORY_4)) %>% 
  left_join(kmeans_membership_6) %>% 
  mutate(SAL_THRESHOLD = case_when(
    Depth > SAL_MAX ~ "Below salinity maximum", #At and below the salinity maximum
    Depth <= SAL_MAX ~ "Above salinity maximum",
  )) %>% 
  mutate(MLD_THRESHOLD = case_when(
    Depth > MLD ~ "Below MLD", #At and below the salinity maximum
    Depth <= MLD ~ "Above MLD",
  )) %>% 
  mutate(DCM_THRESHOLD = case_when(
    Depth > DCM ~ "Below DCM",
    Depth <= DCM ~ "Above DCM", #At and above the DCM
  )) %>% 
  mutate(O2_THRESHOLD = case_when(
    Depth > O2_MIN ~ "Below oxygen minimum",
    Depth <= O2_MIN ~ "Above oxygen minimum", #At and above the O2 minimum
  ))

# head(table_wfeature)

table_wstn_wdepth <- table_wfeature %>%
  select(-ENV_VARIABLE, -value) %>% distinct()

table_supp <- table_wfeature %>%
  select(-Pressure, -datetime_cst, -starts_with("SUN"), -HR_OF_DAY, -stn) %>%
  pivot_wider(names_from = ENV_VARIABLE, values_from = value)
# glimpse(table_supp)
# write.csv(table_supp, file = "output-tables/supp-table-allsampleinfo.csv")
```

# Sample comparison & taxonomic classification

## PCA analysis

Create wide format matrix with sequence counts averaged across replicates (by ASV)
Make wide format and as a matrix.
```{r}
asv_wide_avg_matrix <- asv_long_avg_wtax_clean %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
# dim(asv_wide_avg_matrix)
# Working with 10563 ASVs
# dim(asv_long_avg_wtax_clean)
```

```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_matrix))))
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

#### Plot PCA

> Using MLD

```{r, fig.height=5, fig.width=6}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

PCA_PLOT <- data.frame(pca_clr$x, STN_NISKIN = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin)) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = MLD_THRESHOLD,
             # fill = CLUSTER)) + # Try this for a supplementary figure.
             fill = depth_order)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  # geom_jitter(stroke = 0.6, size = 2.5, width = 0.8, shape = 21) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(21, 23)) +
  scale_fill_manual(values = blues) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
PCA_PLOT
# ggsave("figures/pca-plot.svg", width = 6, height = 5, device = "svg", limitsize = FALSE)
```

### PCA above DCM

```{r}
# head(table1_wfeature)
above_DCM_df <- table_wstn_wdepth %>% 
  filter(DCM_THRESHOLD == "Above DCM")
above_dcm_list <- as.character(unique(above_DCM_df$STN_NISKIN))
```


Make wide format and as a matrix.
```{r}
aboveDMC_wide_matrix <- asv_long_avg_wtax_clean %>% 
  filter(STN_NISKIN %in% above_dcm_list) %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
pca_clr_abovedcm <- prcomp(data.frame(compositions::clr(t(aboveDMC_wide_matrix))))
```

```{r}
variance_clr_abovedcm <- (pca_clr_abovedcm$sdev^2)/sum(pca_clr_abovedcm$sdev^2)

barplot(variance_clr_abovedcm, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```


```{r, fig.height=5, fig.width=6}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

pca_abovedcm <- data.frame(pca_clr_abovedcm$x, STN_NISKIN = rownames(pca_clr_abovedcm$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin)) %>% 
  # mutate(FEATURE = str_replace_na(Features_list, replacement = "other")) %>% 
  # mutate(FEATURE_ORDER = factor(FEATURE, levels = watercol_features, labels = watercol_features_label)) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = ,
             fill = depth_order)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8, shape = 21, color = "black") +
  ylab(paste0('PC2 ',round(variance_clr_abovedcm[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr_abovedcm[1]*100,2),'%')) +
  # scale_shape_manual(values = watercol_shape) +
  scale_fill_manual(values = blues) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
pca_abovedcm
# ggsave("figures/pca-plot.svg", width = 6, height = 5, device = "svg", limitsize = FALSE)
```




# Diversity indices

Use wide dataframe from above.

```{r}
head(asv_wide_avg_matrix)[,1:3]
```

```{r}
shannon_all <- vegan::diversity(asv_wide_avg_matrix, index = "shannon", MARGIN = 2)
```


```{r}
# ?diversity
invsimp_all <- vegan::diversity(asv_wide_avg_matrix, index = "invsimpson", MARGIN = 2)
# head(data.frame(invsimp_all))
```

## Plot diversity


### Shannon

By depth
```{r}
depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")

SHANNON <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = depth_order, 
             x = shannon_all, 
             fill = depth_order,
             shape = MLD_THRESHOLD)) +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 10),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = "Shannon Diversity", y = "") +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))

```

```{r, fig.width=8, fig.height=3}
SHANNON_ <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  mutate(STN_ORDER = factor(as.character(Station), levels = stn_order_shore_offshore)) %>% 
 #
  ggplot(aes(x = STN_ORDER, y = shannon_all, 
             shape = MLD_THRESHOLD, fill = depth_order)) +
  facet_grid(cols = vars(Transect), scale = "free_x") +
  geom_jitter(stroke = 0.6, width = 0.2, size = 2) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 10),
    strip.background = element_blank(),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(y = "Shannon Diversity", x = "Station") +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
SHANNON_
```

### Inverse Simpson
```{r}
SIMPSON <- data.frame(invsimp_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = depth_order, x = invsimp_all, 
             fill = depth_order, shape = MLD_THRESHOLD)) +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 10),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = "Inverse Simpsons", y = "") +
guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
SIMPSON
```

## k-means T-S plot

```{r}
# install.packages("marelac")
# install.packages("plot3D")
# library(marelac);library(plot3D)
# Source - https://stackoverflow.com/a
# Posted by JulianHi
# Retrieved 2026-01-04, License - CC BY-SA 3.0
# install.packages("gsw")
# devtools::install_github("ggTS",username="Davidatlarge")
```


```{r}
# library(tidyverse)
# ts <- table1_wfeature %>% ungroup() %>% select(Temperature, Salinity, CLUSTER, Depth)
# 
# CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9", "#126580")
# CLUSTER_NUM <- c(1, 2, 3, 4, 5, 6)
# names(CLUSTER_NUM) <- CLUSTER_COLOR
# 
# ts_cluster <- table1_wfeature %>%
# ggplot(aes(x = Salinity, y = Temperature, fill = CLUSTER))
# 
# TS <- ggTS(sal = ts$Salinity, pot.temp = ts$Temperature,
# reference.p = 0, col.par = ts$Depth, col.name = "Depth (m)")
# TS
```


```{r}
ts <- table_supp %>% ungroup() %>% select(Temperature, Salinity)
# ts
minT <- min(ts$Temperature)
maxT <- max(ts$Temperature)
minS <- min(ts$Salinity)
maxS <- min(ts$Salinity)

set.seed(1234)
salC <- seq(from=minS,to=maxS,length.out = 156)
tempC <- seq(from=minT,to=maxT,length.out = 156)
```

```{r, fig.width=7, fig.height=6}
# head(table1_wfeature)
CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9", "#126580")
CLUSTER_NUM <- c(1, 2, 3, 4, 5, 6)
names(CLUSTER_NUM) <- CLUSTER_COLOR

ts_cluster <- table_supp %>% 
  ggplot(aes(x = Salinity, y = Temperature, fill = CLUSTER)) +
  # geom_point(stat = "identity", shape = 21, color = "black") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.2, stat = "identity", shape = 21, color = "black") +
  scale_fill_manual(values =CLUSTER_COLOR) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 21, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5))) +
  labs(x = "Salinity (ppt)", y = "Temperature (°C)")
ts_cluster
```

### Combine pca and diversity

Without TS cluster
```{r, fig.width=10, fig.height=9}
(PCA_PLOT + coord_fixed()) /
  ((SHANNON  + theme(legend.position = "none")) |
  (SIMPSON  +  theme(legend.position = "none", axis.text.y = element_blank()))) + 
  plot_layout(axis_titles = "collect", widths = c(1, 2)) + plot_annotation(tag_levels = c("A"))
# ggsave("figures/pca-diversity-plot.svg", width = 10, height = 9, device = "svg", limitsize = FALSE)
```

```{r, fig.width=10, fig.height=9}
# ts + PCA_PLOT + SHANNON_
(PCA_PLOT + coord_fixed()) /
  ((SHANNON  + theme(legend.position = "none")) |
  (SIMPSON  +  theme(legend.position = "none", axis.text.y = element_blank()))) + 
  plot_layout(axis_titles = "collect", widths = c(1, 2)) + plot_annotation(tag_levels = c("A"))
# ggsave("figures/pca-diversity-plot.svg", width = 10, height = 9, device = "svg", limitsize = FALSE)
```


```{r, fig.width=12, fig.height=8}
(ts_cluster ) | (PCA_PLOT + coord_fixed()) / (SHANNON_ + theme(legend.position = "none")) +
  plot_layout(nrow = 2, widths = c(1, 2, 8)) + plot_annotation(tag_levels = c("A"))
```

# Dendrogram by community dist

Take the matrix of all data, divide by column total so dendrogram is based on relative sequence abundance by samples. 

```{r}
library(ggdendro)
# CLR transformations
clr_comm <- (compositions::clr(t(asv_wide_avg_matrix)))

#
clr_eucl_comm <- stats::dist((clr_comm), method = "euclidean")
class(clr_eucl_comm)
# clr_eucl_comm 
# dim(data.frame(clr_eucl_comm))
# Calculate the maximum distance between each sample pair (dist( method = "maximum")). Then take dissimilarity structure and assemble using the average hierarchical clustering as a dendrogram.
# head(clr_comm)
cluster_gom <- hclust(clr_eucl_comm, method = "average")
# head(cluster_gom)
# cluster_gom$labels
# Convert object to a dendrogram object in R
dendro <- as.dendrogram(cluster_gom)
gom_dendro <- dendro_data(dendro, type = "rectangle")
# rownames(segment(gom_dendro))
```

```{r, fig.height=18, fig.width=4}
gom_dendro_plot <- ggplot(segment(gom_dendro)) +
  geom_segment(aes(x = x, y = y, xend = xend, 
    yend = yend)) + 
  coord_flip() + 
  scale_x_continuous(expand = c(0.005, 0.3)) +
  scale_y_reverse(expand = c(0.5, 0.5)) +
  geom_text(size = 3, aes(x = x, y = y, label = str_replace_all(label, "_", " "), angle = 0, hjust = 0), data = label(gom_dendro)) +
  theme_dendro() + labs(y = "Dissimilarity") +
    theme(axis.text.x = element_text(color = "black", size = 8, face = "bold"),
          axis.line.x = element_line(color = "#252525"), 
        axis.ticks.x = element_line(), axis.title.x = element_text(color = "black", 
            size = 8))
# svg('figs/SUPPLEMENTARY-dendrogram-wreps.svg', w = 10, h = 8)
gom_dendro_plot
```

Isolate label names that correspond to dendrogram.

```{r}
tmp <- data.frame(gom_dendro$labels)
class(tmp)
labels_for_dendro <- as.character(tmp$label)
setdiff(labels_for_dendro, table_wfeature$STN_NISKIN)
head(table_wfeature)
```

### Dendrogram annotation

Set up metadata for dendrogram annotation.
```{r}
dendro_annotation <- table_supp %>%
  filter(STN_NISKIN %in% labels_for_dendro) %>% 
  # Factor labeling order
  mutate(label_order = factor(as.character(STN_NISKIN), levels = labels_for_dendro)) %>% 
  distinct()
```

```{r, fig.height=12, fig.width=3}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

DEPTH <- dendro_annotation %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = label_order, x = 1, fill = depth_order)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = blues) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
DEPTH
```

```{r, fig.height=12, fig.width=3}
# (unique(dendro_annotation$Station))
stn_order_seqs <- c(1, 2, 3, 4, 5, 
                              9, 8, 7,
                              11, 12, 14, 15)

stn_order_color <- c("#e8dddc", "#e09f99", "#db756b","#d94638", "#db1e0d", 
                              "#a89492", "#ad6f6a", "#a6453c", 
                              "#4f4948", "#523735", "#4a221e", "#450d08")

names(stn_order_color) <- stn_order_seqs

STN <- dendro_annotation %>% 
  mutate(STN = factor(Station, levels = stn_order_seqs)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = STN)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = stn_order_color) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
STN
```


```{r, fig.height=12, fig.width=3}
CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9", "#126580")
CLUSTER_NUM <- c(1, 2, 3, 4, 5, 6)
names(CLUSTER_NUM) <- CLUSTER_COLOR

KMEANS <- dendro_annotation %>% 
  # Factoring
  mutate(cluster_order = factor(CLUSTER, levels = rev(CLUSTER_NUM))) %>%
  ggplot(aes(y = label_order, x = 1, fill = cluster_order)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = CLUSTER_COLOR) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
KMEANS
```

```{r, fig.height=12, fig.width=3}
DIST <- dendro_annotation %>%
  mutate(dist_outflow = DIST_OUTFLOW/1000) %>% 
  ggplot(aes(y = label_order, x = 1, fill = dist_outflow)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_continuous(option = "magma") +
  scale_fill_viridis_c(option = "magma", direction = -1) +
    # scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "x10^3 km", y = "")
DIST

```

```{r, fig.height=12, fig.width=2}
PROFILES <- dendro_annotation %>%
  ungroup() %>% 
  select(STN_NISKIN, label_order, ends_with("_THRESHOLD")) %>% 
  distinct() %>%
  pivot_longer(cols = ends_with("_THRESHOLD"), names_to = "FEATURE", values_to = "CATEGORY") %>% 
  mutate(VAL = case_when(
    grepl("Above", CATEGORY) ~ "above",
    TRUE ~ "below"
  )) %>%
  mutate(Feature_order = factor(FEATURE, levels = c("MLD_THRESHOLD","DCM_THRESHOLD","SAL_THRESHOLD","O2_THRESHOLD"), labels = c("MLD","DCM","Salinity","Oxycline"))) %>% 
  ggplot(aes(y = label_order, x = Feature_order, fill = VAL)) +
    geom_point(color = "black", shape = 22, size = 4) +
  coord_fixed(ratio = 1) +
  scale_fill_manual(values = c("#d8d8f0", "#181752")) + #above, below
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
    axis.line.x = element_line(color = "black", linewidth = 0.2),
    axis.line.y = element_line(color = "black", linewidth = 0.2),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_text(color = "black", size = 8, angle = 90, hjust = 1, vjust = 0.5)) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
PROFILES

```


```{r, fig.height=12, fig.width=2}
TRANSECT <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = TRANSECT)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  scale_fill_manual(values = c("#e0a8d8", "#9c388d", "#540548")) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
TRANSECT
```

```{r, fig.height=12, fig.width=2}
CATEGORY <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = STN_CATEGORY_4)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  scale_fill_manual(values = c("#e3c69d","#e3ab5d", "#c4770a", "#523711")) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
CATEGORY
```
```{r, fig.height=12, fig.width=2}
OIL <- dendro_annotation %>%
  mutate(OIL_COUNT = as.character(OilRig_Count)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = OIL_COUNT)) +
    geom_point(shape = 22, color = "black", size = 4) +
    scale_fill_manual(values = c("grey90","#e38aa6","#ba4e70","#a13356","#8a1a3d","#54051e"))+
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
OIL
```
```{r, fig.height=12, fig.width=2}
DAY <- dendro_annotation %>%
  ggplot(aes(y = label_order, x = 1, fill = DAY_NIGHT)) +
    geom_point(shape = 22, color = "black", size = 4) +
    scale_fill_manual(values = c("#afa7d1","#413c59"))+
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
DAY
```

```{r, fig.height=12, fig.width=4}
# range(dendro_annotation$DEPTH_MAX)
# MAX <- dendro_annotation %>% 
#   ggplot(aes(y = label_order, x = 1, fill = DEPTH_MAX)) +
#     geom_point(shape = 22, color = "black", size = 4) +
#     scale_fill_viridis_c(option = "mako", direction = -1, breaks = c(0, 50, 300, 500, 1000, 2000, 3000)) +
# scale_y_discrete(position = "right") +
#   theme_void() + theme(
#     legend.position = "right",
#     legend.title = element_blank(),
#     legend.text = element_text(color = "black"),
#     strip.background = element_blank(),
#     axis.ticks = element_blank(),
#     axis.title = element_text(color = "black"),
#     axis.text.y = element_text(color = "black", size = 6),
#     axis.text.x = element_blank()) +
#   # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
#   labs(x = "", y = "")
# MAX
```
```{r, fig.height=12, fig.width=4}
PROK <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = Prok_count)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_fermenter(na.value = "grey50", palette = "Reds", direction = 1) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
PROK
```
# Nutrient profiles for dendrogram

```{r}
head(dendro_annotation)
```

```{r, fig.height=12, fig.width=3}
ta <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = TA)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = "Total Alkalinity", y = "")

dic <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = DIC)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = "DIC", y = "")


sil <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = SIL)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = expression("Si(OH)"["4"] ~ "µmol" ~ L^-1), y = "")

po4 <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = PO4)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = expression("PO"["4"] ~ "µmol" ~ L^-1), y = "")
oxy <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = Oxygen_CTD)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = expression("O"["2"] ~ "ml" ~ L^-1), y = "")

no3 <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = NO3)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = expression("NO"["3"] ~ "µmol" ~ L^-1), y = "")

no2 <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = NO2)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = expression("NO"["2"] ~ "µmol" ~ L^-1), y = "")

nh4 <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = NH4)) +
    geom_point(color = "black") +
  # scale_y_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = expression("NH"["4"] ~ "µmol" ~ L^-1), y = ""); nh4
```

# Taxonomy

Richness by depth, station, and feature.

```{r}
head(table_wstn_wdepth)
```


```{r, fig.height=5, fig.width=7}
asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  group_by(Station, DEPTH_BIN, TRANSECT) %>% 
    summarise(SEQ_SUM = sum(MEAN_REPS_seq),
              ASV_COUNT = n()) %>% 
  mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin))) |> 
  ggplot(aes(x = as.factor(Station), y = depth_bin_order, 
             size = ASV_COUNT, fill = TRANSECT)) +
    geom_jitter(shape = 21, color = "black", stroke = 1.5) +
  scale_x_discrete(position = "top") +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(3,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted")
  ) +
  labs(x = "", y = "Depth (m)")
```

```{r}
tax_color  <- c("#607080","#2f97f7","#2c5c1d","#6da65b", "grey20",
"#de6600", #Gyrista
"#ee429b", #dinos
"#1abc9c", # ciliates
"#23b521","#f5bfb0","#fea02f", "#f7d472",
"#d7b6d7","#460e57", 
"#572f0e","#916643","#c95e06", "#dba272", "#e6d2b8", "#b09b10",
"#ad3d46", #Haptista
"#5e0810", "grey30", "#db9399","#93c47d","#ca7ade")

tax_simplified_list <- as.character(rev(sort(unique(asv_long_avg_wtax_clean $Supergroup_simplified))))
tax_simplified_list
names(tax_color) <- tax_simplified_list
names(tax_simplified_list) <- tax_color
```

Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level.

### For dendrogram

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
TAX <- asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  group_by(STN_NISKIN, Station, Niskin, 
           Depth, DEPTH_BIN, Supergroup_simplified) |> 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro),
         tax_order = factor(Supergroup_simplified, levels = (tax_simplified_list))) %>% 
  # mutate(label_order = str_remove(label_order, "_")) |> 
  ggplot(aes(y = (label_order), x = SUM_SEQ, fill = tax_order)) +
    geom_bar(stat = "identity", position = "fill", 
             color = "grey20", width = 0.7, linewidth = 0.3) +
  scale_fill_manual(values = tax_color) +
  # scale_y_discrete(position = "right") +
    scale_y_discrete(position = "left") +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 5),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Relative Abundance", y = "")
TAX

```

### Plot ASV totals


Total ASVs
```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
ASV_TOTAL <- asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  group_by(STN_NISKIN, Station, Niskin) |> 
  summarise(ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = (label_order), x = ASV_COUNT)) +
    geom_bar(stat = "identity", position = "stack", 
             color = "black", fill = "grey40", width = 0.7, linewidth = 0.3) +
  scale_y_discrete(position = "left") +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 5),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Total ASVs", y = "")
ASV_TOTAL
```


```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
four <- c("TSAR-Dinoflagellata", "TSAR-Radiolaria", "TSAR-Ciliophora", "TSAR-Gyrista")

TAX_TOP_FOUR <- asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  filter(Supergroup_simplified %in% four) %>% 
  group_by(STN_NISKIN, Station, Niskin, Supergroup_simplified) %>% 
  summarise(ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro),
         tax_order = factor(Supergroup_simplified, levels = (tax_simplified_list))) %>% 
  ggplot(aes(y = (label_order), x = ASV_COUNT, fill = tax_order)) +
  geom_bar(stat = "identity", position = "stack", 
             color = "black", width = 0.7, linewidth = 0.3) +
  scale_fill_manual(values = tax_color) +
  facet_grid(cols = vars(tax_order), scales = "free") +
  scale_y_discrete(position = "left") +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 5),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Total ASVs", y = "")
TAX_TOP_FOUR

```





### Plot sequence relative abundance

```{r, fig.width=5, fig.height=14}
asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  group_by(Station, DEPTH_BIN, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin)),
          tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = depth_bin_order, x = SUM_SEQ, fill = tax_order)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.9) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(Station), scales = "free_y", space = "free", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```


### Plot sequence relative abundance with depths

```{r, fig.width=5, fig.height=14}
asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  group_by(Station, Depth, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = as.factor(Depth), x = SUM_SEQ, fill = tax_order)) +
  geom_bar(stat = "identity", position = "fill", color = "white", width = 0.9) +
  scale_fill_manual(values = tax_color) +
  # scale_y_reverse() +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(Station), scales = "free_y", space = "free", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

Plot ASV richness by taxonomic groups

```{r, fig.width=22, fig.height=5}
# tax_color
asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  group_by(DEPTH_BIN, Supergroup_simplified, Class) %>%
  summarise(ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "Eukaryota_X") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin)),
          tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = depth_bin_order, x = Class, fill = tax_order, size = ASV_COUNT)) +
  geom_point(shape = 21, color = "grey10", alpha = 0.7, stroke = 0.5) +
  scale_fill_manual(values = tax_color) +
  scale_size_continuous(range = c(1,10), breaks = c(1, 10, 50, 100, 500, 1000, 2000)) +
  facet_grid(cols = vars(tax_order), scales = "free_x", space = "free_x", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
    strip.text.y = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 8, color = "black", face = "bold"),
    strip.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 8, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold", angle = 90, 
                               hjust = 1, vjust = 0, size = 8)) +
  guides(fill = "none",
         size = guide_legend(nrow = 1, shape = 21, color = "black", fill = "black")) +
  labs(x = "", y = "Depth (m)")

# ggsave("figures/tax-dotplot.svg", width = 22, height = 5, device = "svg", limitsize = FALSE)
```

# Plot Dendrogram compilation

```{r, fig.height=13, fig.width=45}
# (TAX_TOP_FOUR + theme(legend.position = "none"))

gom_dendro_plot + DEPTH + STN + KMEANS + PROFILES + 
  oxy + nh4 + no3 + no2 + po4 + sil + dic + ta + 
  DAY + DIST + 
  CATEGORY + TRANSECT + OIL + 
  PROK + TAX + ASV_TOTAL +
  plot_layout(nrow = 1, 
              widths = c(2.5, 1, 1, 1, 1,
                         0.5, 0.5, 0.5, 0.5, 0.5, 0.5,0.5, 0.5,
                         1, 1, 
                         1, 1, 1,
                         1, 1, 1))
# ggsave("figures/dendrogram-compiled.svg", width = 45, height = 13, device = "svg", limitsize = FALSE)
```

# CLR targets & plots

Confirming what core protistan groups are the most prominent members of the community across all of the samples. 
```{r}
# head(asv_long_avg_wtax)
top_taxa <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(!(grepl("Obazoa", Supergroup_simplified))) %>% 
  group_by(Supergroup_simplified, STN_NISKIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>% 
  ungroup() %>% 
  # Subset the top 2 groups with the highest ASV count
  group_by(STN_NISKIN) %>% 
  slice_max(order_by = ASV_COUNT, n = 2)
# top_taxa
list_top <- unique(top_taxa$Supergroup_simplified)
list_top
```

```{r}
list_top
```
Top taxa by ASV total and location from coastline.
```{r}
# head(asv_long_avg_wtax_clean)
top_taxa_category <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(!(grepl("Obazoa", Supergroup_simplified))) %>% 
  group_by(Supergroup_simplified, STN_NISKIN, STN_CATEGORY_4) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>% 
  ungroup() %>% 
  # Subset the top 2 groups with the highest ASV count
  group_by(STN_NISKIN, STN_CATEGORY_4) %>% 
  slice_max(order_by = ASV_COUNT, n = 2) %>% 
  ungroup()

table(top_taxa_category %>% select(Supergroup_simplified, STN_CATEGORY_4))
```

## Dinoflagellates
```{r}
class_dino_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Dinoflagellata-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix

head(class_dino_mat)

class_dino_clr <- vegan::decostand(class_dino_mat, MARGIN = 1, method = "rclr")
##
colnames(class_dino_clr) <- colnames(class_dino_mat)
rownames(class_dino_clr) <- rownames(class_dino_mat)
# data.frame(class_dino_clr)
```

```{r, fig.height=10, fig.width=3}
plot_dino_clr <- data.frame(class_dino_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
plot_dino_clr
```
#### Syndiniales

```{r}
class_syn_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  filter(Class == "Syndiniales") %>% 
  filter(!is.na(Order)) %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Dinoflagellata-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(Order, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = Order, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix

head(class_syn_mat)

class_syn_clr <- vegan::decostand(class_syn_mat, MARGIN = 1, method = "rclr")
##
colnames(class_syn_clr) <- colnames(class_syn_mat)
rownames(class_syn_clr) <- rownames(class_syn_mat)
```

```{r, fig.height=11, fig.width=3}
plot_syn_clr <- data.frame(class_syn_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
# plot_syn_clr
```


## Radiolaria
```{r}
list_top

  
class_rad_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Radiolaria-Unannotated",
    (is.na(Order) & Class == "X") ~ "Radiolaria-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_rad_mat)
class_rad_clr <- vegan::decostand(class_rad_mat, MARGIN = 1, method = "rclr")
##
colnames(class_rad_clr) <- colnames(class_rad_mat)
rownames(class_rad_clr) <- rownames(class_rad_mat)

# data.frame(class_rad_clr)
```

```{r, fig.height=10, fig.width=3}
plot_rad_clr <- data.frame(class_rad_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```


## Ciliate
```{r}
list_top
tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Ciliophora"); unique(tmp$Class)
  
class_cil_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Ciliophora-Unannotated",
    (is.na(Order) & Class == "X") ~ "Ciliophora-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_cil_mat)
class_cil_clr <- vegan::decostand(class_cil_mat, MARGIN = 1, method = "rclr")
##
colnames(class_cil_clr) <- colnames(class_cil_mat)
rownames(class_cil_clr) <- rownames(class_cil_mat)

# data.frame(class_rad_clr)
```

```{r, fig.height=13, fig.width=5}
plot_cil_clr <- data.frame(class_cil_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```


## Gyrista
```{r}
list_top
tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Gyrista"); unique(tmp$Class)
  
class_gyrista_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Gyrista-Unannotated",
    (is.na(Order) & Class == "X") ~ "Gyrista-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_gyrista_mat)
class_gyrista_clr <- vegan::decostand(class_gyrista_mat, MARGIN = 1, method = "rclr")
##
colnames(class_gyrista_clr) <- colnames(class_gyrista_mat)
rownames(class_gyrista_clr) <- rownames(class_gyrista_mat)
```

```{r, fig.height=13, fig.width=5}
plot_gyrista_clr <- data.frame(class_gyrista_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "") +
  guides(aesthetic = guide_legend(ncol = 1, reverse = FALSE)) 
```

```{r}
list_top
tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Class == "Mediophyceae" | Class == "Coscinodiscophyceae")
unique(tmp$Order)
  
order_gyrista_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Class == "Mediophyceae" | Class == "Coscinodiscophyceae") %>% 
  ungroup() %>% 
  mutate(ORDER = case_when(
    is.na(Order) ~ "Unannotated",
    # (is.na(Order) & Class == "X") ~ "Gyrista-Unannotated",
         TRUE ~ Order)) %>% 
  group_by(Class, ORDER, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  unite(TAXA, Class, ORDER, sep = ";") %>% 
  pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(order_gyrista_mat)
order_gyrista_clr <- vegan::decostand(order_gyrista_mat, MARGIN = 1, method = "rclr")
##
colnames(order_gyrista_clr) <- colnames(order_gyrista_mat)
rownames(order_gyrista_clr) <- rownames(order_gyrista_mat)
```

```{r, fig.height=15, fig.width=5}
plot_gyrista_order_clr <- data.frame(order_gyrista_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  separate(Class, into = c("CLASS", "ORDER"), sep = "\\.") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = ORDER, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), cols = vars(CLASS), space = "free", scale = "free") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          strip.placement = "outside") +
    labs(x = "", y = "") +
  guides(aesthetic = guide_legend(ncol = 1, reverse = FALSE)) 
# plot_gyrista_order_clr
```

## Bigyra
```{r}
list_top
tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Bigyra"); unique(tmp$Class)
  
class_bigyra_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Bigyra-Unannotated",
    (is.na(Order) & Class == "X") ~ "Bigyra-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_bigyra_mat)
class_bigyra_clr <- vegan::decostand(class_bigyra_mat, MARGIN = 1, method = "rclr")
##
colnames(class_bigyra_clr) <- colnames(class_bigyra_mat)
rownames(class_bigyra_clr) <- rownames(class_bigyra_mat)
```

```{r, fig.height=13, fig.width=3}
plot_bigya_clr <- data.frame(class_bigyra_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```




## Haptista
```{r}
# list_top
class_hapto_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Haptista-Unannotated",
    (is.na(Order) & Class == "X") ~ "Haptista-Unannotated",
         TRUE ~ Order)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_hapto_mat)
class_hapto_clr <- vegan::decostand(class_hapto_mat, MARGIN = 1, method = "rclr")
##
colnames(class_hapto_clr) <- colnames(class_hapto_mat)
rownames(class_hapto_clr) <- rownames(class_hapto_mat)

# data.frame(class_hapto_clr)
```

```{r, fig.height=10, fig.width=3}
plot_hapto_clr <- data.frame(class_hapto_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```

Try boxplot version. 
```{r, fig.height=5, fig.width=4.5}
data.frame(class_hapto_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, fill = stn, group = DEPTH_BIN_ORDER)) +
  geom_vline(xintercept = 0, color = "grey40") +
  ggdist::stat_halfeye(side = "top", adjust = 0.5, width = 0.2, 
                       justification = 0, .width = 0, fill = "black", alpha = 0.6,
                       overflow = "compress") +
  # geom_boxplot(width = 0.2, fill = "grey20", color = "grey10",
  #              alpha = 0.6, linewidth = 0.3) +
  ggdist::stat_dots(side = "bottom", layout = "bin",
                    justification = 1.0,
                    overflow = "compress", dotsize = 0.4,
                    fill = "grey40", color = "grey10") +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "", title = "Haptophyte ASVs (clr)")
```


## Archaeplastida
```{r}
class_arch_mat <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "Archaeplastida") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Dinoflagellata-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix

head(class_arch_mat)

class_arch_clr <- vegan::decostand(class_arch_mat, MARGIN = 1, method = "rclr")
##
colnames(class_arch_clr) <- colnames(class_arch_mat)
rownames(class_arch_clr) <- rownames(class_arch_mat)
# data.frame(class_dino_clr)
```

```{r, fig.height=10, fig.width=5}
plot_arch_clr <- data.frame(class_arch_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
# plot_arch_clr
```


## Compile taxa CLR plots
```{r, fig.width=22, fig.height=11}
plot_dino_clr + plot_hapto_clr + plot_rad_clr + plot_cil_clr + plot_gyrista_clr + plot_bigya_clr + plot_layout(nrow = 1, widths = c(1, 2.6, 1.2, 2.6, 3.5, 0.8))
# ggsave("figures/clr-top-taxa.svg", width = 22, height = 11, device = "svg", limitsize = FALSE)
```



# Taxa by depth - curated

#### Area plot w/ bar plot function

```{r}
area_wbar <- function(DF, COLORS){
 area_plot <- (DF %>% 
  filter(DEPTH_BIN == "< 5m") %>% 
    group_by(TAXA, Station, TRANSECT) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  ungroup() %>%
  group_by(Station) %>% 
      mutate(PERC = (SEQ_SUM/sum(SEQ_SUM))) %>%
  select(-SEQ_SUM) %>%
  pivot_wider(names_from = TAXA, values_from = PERC, values_fill = 0) %>%
  mutate(STATION_ORDER = factor(Station, levels = stn_order_shore_offshore)) %>% 
  mutate(TRANSECT_ORDER = factor(TRANSECT, levels = c("transect1", "transect2", "transect3"), labels = c("Transect 1", "Transect 2", "Transect 3"))) %>% 
  pivot_longer(cols = -c(Station, STATION_ORDER, TRANSECT, TRANSECT_ORDER), names_to = "TAXA", values_to = "PERC") %>% 
  ggplot() +
  geom_area(aes(x = as.factor(STATION_ORDER), 
                y = PERC, fill = TAXA, group = TAXA), 
            stat = "identity", position = "stack",
            color = "black") +
    scale_fill_manual(values = COLORS) +
  facet_grid(cols = vars(TRANSECT_ORDER), space = "free", scales = "free_x") +
  theme_classic() +
  scale_x_discrete(expand = c(0.05,0.05)) +
  theme(axis.text.x = element_text(color = "black", size = 10, face = "bold"),
          axis.text.y = element_text(color = "black", size = 10, face = "bold"),
          axis.title = element_text(color = "black", size = 10, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.x = element_text(color = "black", angle = 0, 
                                      size = 10, face = "bold",
                                      hjust = 0, vjust = 1),
          strip.background = element_blank(),
          legend.position = "none",
          legend.title = element_blank()) +
    labs(y = "Sequence Relative Abundance", x = "Station (surface)", title = "") +
     guides(aesthetic = guide_legend(ncol = 1, reverse = FALSE)))
 bar_plot <- (DF %>% 
    group_by(TAXA, DEPTH_BIN, Class) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA, pattern = Class)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
    scale_fill_manual(values = COLORS) +
    scale_x_continuous(expand = c(0,0)) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", size = 10, face = "bold"),
          axis.text.y = element_text(color = "black", size = 10, face = "bold"),
          axis.title = element_text(color = "black", size = 10, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.x = element_text(color = "black", angle = 0, size = 10, face = "bold"),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank(),
          legend.text  = element_text(size = 8, color = "black", face = "bold")) +
    labs(x = "Sequence Relative Abundance", y = "Depth") +
     guides(aesthetic = guide_legend(ncol = 1, reverse = FALSE),
            shape = guide_legend(override.aes = list(size = 0.5)),
            color = guide_legend(override.aes = list(size = 0.5))))
 ##
 area_plot + bar_plot +
   plot_layout(ncol = 1, widths = c(1, 0.6), heights = c(1, 1), guides = 'collect')
}
```

## Dinoflagellata
```{r, fig.width=13, fig.height=4}
DINO <- {mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.8, fill = "grey80", 
                color = "grey10",
                alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Dinoflagellata") +
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    grepl("Dinophyceae_X", TAXA) ~ "Dinophyceae",
    grepl("Dinophyta_X", TAXA) ~ "Dinophyta",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN, Class) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA, pattern = Class)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
     scale_fill_manual(values = c("#db1f7d", "#d16b9e", "#d1a9bd", "#8a1751", "#7a3759", "#78616c", "#540c30",
                                 "#4d917b", "#89c977",
                                 "#bfad0a", "#e8912e","#aba250", "#cc9e6a", "#a34e27")) +
    scale_x_continuous(expand = c(0,0)) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
     guides(aesthetic = guide_legend(ncol = 1, reverse = FALSE)) +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))}

DINO
```
### Dinoflagellate area map

Prep dinoflagellate dataframe
```{r}
dino_tax_df <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    ungroup() %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    grepl("Dinophyceae_X", TAXA) ~ "Dinophyceae",
    grepl("Dinophyta_X", TAXA) ~ "Dinophyta",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>%
    ungroup()

tax_dino_order <- (sort(unique(dino_tax_df$TAXA)))

dino_colors <- c("#db1f7d", "#d16b9e", "#d1a9bd", "#8a1751", "#7a3759", "#78616c", "#540c30",
                                 "#4d917b", "#89c977",
                                 "#bfad0a", "#e8912e","#aba250", "#cc9e6a", "#a34e27")
names(dino_colors) <- tax_dino_order
# dino_colors
```


```{r, fig.width=8, fig.height=9}
DINO_AREA <- area_wbar(dino_tax_df, dino_colors)
DINO_AREA
```


## Radiolaria

```{r, fig.width=13, fig.height=4}
RAD <- {mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Radiolaria") +
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Acantharea", TAXA) ~ "Acantharea",
    grepl("RAD_A", TAXA) ~ "RAD_A",
    grepl("RAD_B", TAXA) ~ "RAD_B",
    grepl("RAD_C", TAXA) ~ "RAD_C",
    grepl("Radiolaria_XX", TAXA) ~ "Radiolaria",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
    scale_fill_manual(values = c("#805edb", "#afe3cd", "#164531", "#e3e340", "#c4c49b", "#5b6940", "#648bb3")) +
    scale_x_continuous(expand = c(0,0)) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))}
# RAD
```

### Radiolaria area map

```{r}
rad_tax_df<-  asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Acantharea", TAXA) ~ "Acantharea",
    grepl("RAD_A", TAXA) ~ "RAD_A",
    grepl("RAD_B", TAXA) ~ "RAD_B",
    grepl("RAD_C", TAXA) ~ "RAD_C",
    grepl("Radiolaria_XX", TAXA) ~ "Radiolaria",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    ungroup()

rad_tax_only <- sort(unique(rad_tax_df$TAXA))
rad_colors <- c("#805edb", "#afe3cd", "#164531", "#e3e340", "#c4c49b", "#5b6940", "#648bb3")
names(rad_colors) <- rad_tax_only
```

```{r, fig.width=8, fig.height=9}
RAD_AREA <- area_wbar(rad_tax_df, rad_colors)
```


## Ciliate

```{r, fig.width=13, fig.height=4}
other_ciliates <- c("Oligohymenophorea;Oligohymenophorea_X", "Litostomatea;Haptoria_6", "Heterotrichea;Heterotrichea_X", "Spirotrichea;Hypotrichia", "Ciliophora_X;Ciliophora_XX")

CILIATE <- {mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Ciliophora") +
    ##
    
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    (TAXA %in% other_ciliates) ~ "Ciliates unannotated",
    grepl("CONTH_", TAXA) ~ "CONThreeP unannotated",
    grepl("CONThreeP_X", TAXA) ~ "CONThreeP unannotated",
    (Class == "Discotrichidae") ~ "CONThreeP;Discotrichidae",
    grepl("Phyllopharyngea;Cyrtophoria", TAXA) ~ "Phyllopharyngea;Cyrtophoria",
    grepl("Spirotrichea;Choreotrichida", TAXA) ~ "Spirotrichea;Choreotrichida",
    grepl("Litostomatea;Litostomatea_X", TAXA) ~ "Litostomatea",
    grepl("Prostomatea_1;Prostomatea_1_X", TAXA) ~ "Prostomatea",
    grepl("Colpodea_X", TAXA) ~ "Colpodea",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
    scale_x_continuous(expand = c(0,0)) +
    scale_fill_manual(values = c("#f5bfb0", "#ed7e5f", "#d13206", 
  "#cf9229", 
  "#cfcf57", "#d1d106", 
  "#accf97", "#7ade3c", "#7ade3c", "#84c2a5", "#2edb8a", "#02703d",
  "#3899a8", "#5402e0", "#916dcf", "#604a70")) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    guides(aesthetic = guide_legend(ncol = 1, reverse = FALSE)) +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))}
# CILIATE
```

### Ciliate area map

Prep dataframe
```{r}
ciliate_tax_df <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    (TAXA %in% other_ciliates) ~ "Ciliates unannotated",
    grepl("CONTH_", TAXA) ~ "CONThreeP unannotated",
    grepl("CONThreeP_X", TAXA) ~ "CONThreeP unannotated",
    (Class == "Discotrichidae") ~ "CONThreeP;Discotrichidae",
    grepl("Phyllopharyngea;Cyrtophoria", TAXA) ~ "Phyllopharyngea;Cyrtophoria",
    grepl("Spirotrichea;Choreotrichida", TAXA) ~ "Spirotrichea;Choreotrichida",
    grepl("Litostomatea;Litostomatea_X", TAXA) ~ "Litostomatea",
    grepl("Prostomatea_1;Prostomatea_1_X", TAXA) ~ "Prostomatea",
    grepl("Colpodea_X", TAXA) ~ "Colpodea",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    ungroup() 


ciliate_tax_only <- rev(sort(unique(ciliate_tax_df$TAXA)))
ciliate_tax_only
ciliate_colors <- c("#f5bfb0", "#ed7e5f", "#d13206", 
  "#cf9229", "#69552a",
  "#cfcf57", "#d1d106", 
  "#accf97", "#7ade3c", "#7ade3c", "#84c2a5", "#2edb8a", "#02703d",
  "#3899a8", "#445cad", "#5402e0", "#916dcf", "#604a70", "#7b857e")

names(ciliate_colors) <- ciliate_tax_only

```

```{r, fig.width=8, fig.height=9}
CILIATE_AREA <- area_wbar(ciliate_tax_df, ciliate_colors)
```

## Gyrista

```{r, fig.width=17, fig.height=4}
gyrista_less <- c("Pelagophyceae unannotated", "Pelagophyceae;NA","Coscinodiscophyceae;Coscinodiscales", "Bacillariophyceae;Licmophorales", "Eustigmatophyceae;Eustigmatophyceae_X", "Pirsoniales;Pirsoniales_X","Peronosporomycetes;Peronosporomycetes_X", "Bacillariophyceae;Cymbellales") 

GYRISTA <- {mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Gyrista") +
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>%
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  mutate(TAXA = case_when(
    is.na(Class) ~ "Gyrista other",
    (TAXA %in% gyrista_less) ~ "Gyrista other",
    grepl(";NA", TAXA) ~ paste(Class, "unannotated", sep = " "),
    (TAXA == "Gyrista_X;Gyrista_XX") ~ "Gyrista other",
    grepl("Dictyochophyceae", TAXA) ~ "Dictyochophyceae",
    grepl("Chrysophyceae_Clade-EC2", TAXA) ~ "Chrysophyceae_Clade-EC2",
    grepl("Bacillariophyceae;Bacillariophyceae_X", TAXA) ~ "Bacillariophyceae unannotated",
    grepl("Pelagophyceae", TAXA) ~ "Pelagophyceae;Pelagomonadales",
    grepl("MOCH-", TAXA) ~ "MOCH clades 2-5",
    (TAXA == "Mediophyceae;Mediophyceae_X" | TAXA == "Mediophyceae;NA") ~ "Mediophyceae unannotated",
    TRUE ~ TAXA
  )) %>%
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
    scale_fill_manual(values = c("#93c47d", "#7db7c4", "#a27dc4", 
  "#d65c1a", "#c98d6d", "#662a0a", "#805944", "#a3690b", 
  "#d93529", "#736b6b","#d4cf44",
  "#8ce663", "#38a605", 
  "#0ba7d6","#1836cc", "#6abbd4", "#5669c4","#97b8c2",
  "#5cd1b8", 
  "#d596e3", "#b82fd6", "#a10b83","#a85b99")) +
    scale_x_continuous(expand = c(0,0)) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    guides(aesthetic = guide_legend(ncol = 1)) +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))}
# GYRISTA
```

### Gyrista area map

Prep dataframe

```{r}
gyrista_less <- c("Pelagophyceae unannotated", "Pelagophyceae;NA","Coscinodiscophyceae;Coscinodiscales", "Bacillariophyceae;Licmophorales", "Eustigmatophyceae;Eustigmatophyceae_X", "Pirsoniales;Pirsoniales_X","Peronosporomycetes;Peronosporomycetes_X", "Bacillariophyceae;Cymbellales") 

gyrista_df <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>%
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  mutate(TAXA = case_when(
    is.na(Class) ~ "Gyrista other",
    (TAXA %in% gyrista_less) ~ "Gyrista other",
    grepl(";NA", TAXA) ~ paste(Class, "unannotated", sep = " "),
    (TAXA == "Gyrista_X;Gyrista_XX") ~ "Gyrista other",
    grepl("Dictyochophyceae", TAXA) ~ "Dictyochophyceae",
    grepl("Chrysophyceae_Clade-EC2", TAXA) ~ "Chrysophyceae_Clade-EC2",
    grepl("Bacillariophyceae;Bacillariophyceae_X", TAXA) ~ "Bacillariophyceae unannotated",
    grepl("Pelagophyceae", TAXA) ~ "Pelagophyceae;Pelagomonadales",
    grepl("MOCH-", TAXA) ~ "MOCH clades 2-5",
    (TAXA == "Mediophyceae;Mediophyceae_X" | TAXA == "Mediophyceae;NA") ~ "Mediophyceae unannotated",
    TRUE ~ TAXA
  ))

gyrista_tax_order <- rev(sort(unique(gyrista_df$TAXA)))

gyrista_colors <- c("#93c47d", "#de5472","#7db7c4", "#a27dc4", 
  "#d65c1a", "#c98d6d", "#662a0a", "#805944", "#bd5f17","#a3690b", 
  "#d93529", "#736b6b","#d4cf44",
  "#225706","#8ce663", "#38a605", 
  "#0ba7d6","#1836cc", "#6abbd4", "#5669c4","#97b8c2","#0d3142",
  "#5cd1b8", 
  "#d596e3", "#b82fd6", "#a10b83","#a85b99")

names(gyrista_colors) <- gyrista_tax_order
```


```{r, fig.width=8, fig.height=9}
GYRISTA_AREA <- area_wbar(gyrista_df, gyrista_colors) + guides(fill = guide_legend(ncol = 1))
# GYRISTA_AREA
# GYRISTA_AREA + theme(legend.position = "none")
```

## Bigyra
```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Bigyra") +
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Order, Family, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))
```

### Bigyra tree map

```{r}
bigyra_tax_df <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Order, Family, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra")
    
```



## Haptista

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "Haptista") +
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "Haptista") %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Pterocystida", TAXA) ~ "Pterocystida",
    grepl("Haptophyta_Clade_", TAXA) ~ "Haptophyta Clades-3/4",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))
# ggsave("figures/haptista_clr_richness_relabun.svg", width = 13, height = 4, device = "svg", limitsize = FALSE)
```

### Haptophyte tree map

```{r}
haptophy_tax_df <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Pterocystida", TAXA) ~ "Pterocystida",
    grepl("Haptophyta_Clade_", TAXA) ~ "Haptophyta Clades-3/4",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
    ungroup()

hapto_tax_order <- sort(unique(haptophy_tax_df$TAXA))
hapto_colors <- c("#851287", "#601ed4",
                  "#a3d9be", "#1da862", "#468063", "#13b0a8","#2a6e6a", "#5ba1bd", "#c478b0")
names(hapto_colors) <- hapto_tax_order
```

```{r, fig.width=8, fig.height=9}
area_wbar(haptophy_tax_df, hapto_colors)
# ggsave("figures/haptophyte_area.svg", width = 8, height = 9, device = "svg", limitsize = FALSE)
```


## Archaeplastida

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
  filter(Supergroup_simplified == "Archaeplastida") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "Archaeplastida") +
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Supergroup_simplified == "Archaeplastida") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>%  
    mutate(TAXA = Class) %>% 
  # unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  # mutate(TAXA = case_when(
  #   grepl("Pterocystida", TAXA) ~ "Pterocystida",
  #   grepl("Haptophyta_Clade_", TAXA) ~ "Haptophyta Clades-3/4",
  #   TRUE ~ TAXA
  # )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "Archaeplastida") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))
# ggsave("figures/arche_clr_richness_relabun.svg", width = 13, height = 4, device = "svg", limitsize = FALSE)
```

## Save taxa designations
```{r}
# save(dino_tax_df,rad_tax_df,ciliate_tax_df,gyrista_df,haptophy_tax_df,bigyra_tax_df, file = "input-data/tax-level-assessment.RData")
```


### Compile area plots
```{r, fig.width=18, fig.height=15}
(DINO_AREA - GYRISTA_AREA) / (CILIATE_AREA - RAD_AREA) + plot_annotation(tag_levels = "A")
# ggsave("figures/areaplots-alltaxa.svg", width = 18, height = 15, device = "svg", limitsize = FALSE)
```

# RDA analysis

RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it.
Null H = variances across groups are equal
p-value of < 0.05 means there is a significant difference in variance among groups.
```{r}
# library(car)
# Levene's test with one independent variable
# leveneTest(weight ~ group, data = PlantGrowth)
# Levene's test with multiple independent variables
# leveneTest(len ~ supp*dose, data = ToothGrowth)
```

```{r}
set.seed(1234)
head(table_wstn_wdepth)
# head(table_wfeature)
head(table_supp)
```

## All

Isolate categorical vs. numeric metadata. For numeric data, create a normalized scale. 

```{r}
# Make categorical data table from metadata (Table 1)
categorical <- table_wstn_wdepth %>% 
  select(STN_NISKIN, STN_ORDER, TRANSECT, OilRig_Count, DAY_NIGHT, HR_OF_DAY, ends_with("_THRESHOLD")) %>% distinct() %>% 
  arrange(STN_NISKIN)

# Generate a matrix of numerical data from metadata
metadata_for_rda_0 <- table_supp %>% 
  select(STN_NISKIN, Depth, DIST_OUTFLOW, NH4, NO2, NO3, Oxygen_CTD, PO4, SIL, Salinity, Temperature) %>% distinct() %>% 
  # pivot_longer(cols = -c(STN_NISKIN), names_to = "VARIABLES", values_to = "VALUE") %>% 
  # group_by(STN_NISKIN, VARIABLES) %>% 
  #   summarise(MEAN_VAL = mean(VALUE)) %>% 
  # pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>%
  as.matrix
```

Normalize matrix of numeric metadata then re-combine with categorical metadata.
```{r}
# Normalize
standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")

# Combine normalized metadata with categorical data
metadata_normalized <- categorical %>% 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN")
# head(metadata_normalized)

# Output from here should be the same rownames for ASV table below.
rownames(metadata_normalized) 
```
Generate a matrix from ASV counts
```{r}
# head(asv_long_avg_wtax_clean)
asv_wide_avg_reps_matrix <- asv_long_avg_wtax_clean %>% 
  ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix()

rownames(asv_wide_avg_reps_matrix)
```

```{r}
clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_reps_matrix))

# Confirm that all 

asv_list <- colnames(asv_wide_avg_reps_matrix)
# asv_list
meta_list <- rownames(metadata_normalized)
clr_list <- rownames(clr_asv_matrix)
setdiff(meta_list, clr_list)
```

Run RDA analysis

```{r}
# | eval: false
# rda_all_gom <- vegan::rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
# rda_all_gom
# ?rda
```

```{r}
# | eval: false
# adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
# # RsquareAdj(rda_all_gom)
# adj_rda_out
# anova_out <- anova.cca(rda_all_gom, step = 100, by = "term")
# ?anova.cca()
```

Large f-ratio means there is a lot of between group variance compared to within-group variance (independent variables have a greater influence on depdendent variables)
Smaller f-ratios indicate the groups are similar to one another. So differences in groups cannot be differentatied from "random"
Variance indicates the overall spread of data


## RDA function
First create a function for running RDA analysis.

```{r}
# asv_list <- colnames(asv_wide_avg_reps_matrix)
# asv_list
# meta_list <- rownames(metadata_normalized)
# clr_list <- rownames(clr_asv_matrix)
# save(table_supp, table_wstn_wdepth, metadata_normalized, asv_long_avg_wtax_clean, file = "RDA_Robjects_02072026.RData")
```

```{r}
# How to use:
## Modify and subset table1 or asv_long_avg_wtax completely before use.
## Check to make sure the same STN_NISKIN IDs are present in both.
run_rda <- function(mod_metadata, mod_asv_long, 
                    DATASET, features = c("include", "exclude")){
  # Select categorical data
  # mod_metadata <- surf_table
  # mod_asv_long <- surf_asv
  if (features == "include") {
    categorical <- mod_metadata %>% 
             select(STN_NISKIN, STN_ORDER, TRANSECT, 
                    OilRig_Count, DAY_NIGHT, ends_with("_THRESHOLD")) %>% 
      distinct()
    } else {
      categorical <- mod_metadata %>% 
             select(STN_NISKIN, STN_ORDER, TRANSECT, 
                    OilRig_Count, DAY_NIGHT) %>% distinct()}
  #
  # Isolate numerical data as matrix
  metadata_for_rda_0 <- mod_metadata %>% 
    select(STN_NISKIN, Depth, DIST_OUTFLOW, NH4, NO2, NO3, Oxygen_CTD, PO4, SIL, Salinity, Temperature) %>% distinct() %>% 
    pivot_longer(cols = -c(STN_NISKIN), names_to = "VARIABLES", values_to = "VALUE") %>% 
    group_by(STN_NISKIN, VARIABLES) %>% 
      summarise(MEAN_VAL = mean(VALUE)) %>% 
    pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
    arrange(STN_NISKIN) %>% 
    column_to_rownames(var = "STN_NISKIN") %>% as.matrix
  # Normalize
  standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")
  # Combine normalized numerical data with categorical
  metadata_normalized <- categorical %>% 
    right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
    arrange(STN_NISKIN) %>% 
    column_to_rownames(var = "STN_NISKIN")
  #
  ## Prepare ASV data
  asv_wide_avg_reps_matrix <- mod_asv_long %>% 
    ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
  # Transform
  asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix
  #
  # CLR normalization
  clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))
  #
  # setdiff(rownames(metadata_normalized), (rownames(clr_asv_matrix)))
  # RDA step
  rda_output <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
  # rda_output
  #
  # R-squared value
  adj_rda_out <- RsquareAdj(rda_output)$adj.r.squared #adj_rda_out
  # Run ANOVA
  anova_out <- anova.cca(rda_output, step = 100, by = "term")
  #
  # Get stats
  ## Percentage constrained
  total_amount <- (rda_output$tot.chi)
  constrained <- (rda_output$CCA$tot.chi)
  perc_constratained <- constrained/total_amount
  # Formula
  formula <- as.character(rda_output$call)[2]
  # Create data frame output for RDAs run
  rda_results_df <- data.frame(scores(anova_out)) %>% 
    rownames_to_column(var = "VARIABLES") %>% 
    select(VARIABLES, P_VALUE = `Pr..F.`) %>% 
    add_column(TEST = DATASET,
               PERC_CONSTRAINED = perc_constratained,
               ADJ_R_SQUARED = adj_rda_out,
               FORMULA = formula)
  #
  return(rda_results_df)
}
# class(as.character(rda_output$call)[2])
```


```{r}
# All
all_samples <- table_supp
  # Filter low abundance ASVs
  # filter(DCM != "Above DCM") %>% ungroup()
all_samples_list <- as.character(unique(table_supp$STN_NISKIN))

asvs_all_samples <- asv_long_avg_wtax_clean %>%
  filter(STN_NISKIN %in% all_samples_list) %>% ungroup()

# all_rda <- run_rda(all_samples, asvs_all_samples,
#                    "All samples", features = "include")
# all_rda
```

## RDA compile from HPRC

`fxn-runRDA_1000_HPRC.R`
`RDA_Robjects_02072026.RData`

```{r}
# install.packages("gt")
library(gt)
```


```{r}
load("output-tables/df_rda_output_0207202.RData", verbose = TRUE)
```
```{r}
head(df_rda_results)
```


```{r}
rda_summary_table <- gt(df_rda_results %>% 
     select(-FORMULA, -Residual, `Input data` = TEST,
            `Percent constrained` = PERC_CONSTRAINED,
            `Adjusted R-squared` = ADJ_R_SQUARED,
            `MLD` = MLD_THRESHOLD, `DCM` = DCM_THRESHOLD, 
            `Salinity maximum` = SAL_THRESHOLD, 
            `Oxygen minimum depth` = O2_THRESHOLD, 
            Station = STN_ORDER, `Depth (m)` = Depth,
            Salinity, Temperature, NH4, NO2, NO3, Oxygen, PO4, SIL)) %>% 
            # `Day vs. night` = DAY_NIGHT,
            # `Number of oil rigs` = OilRig_Count)) %>% 
  data_color(columns = -c(`Input data`, `Percent constrained`, `Adjusted R-squared`),
             palette = "Reds", apply_to = c("fill"),
             na_color = "grey30",
             colors = scales::col_bin(
      palette = c("transparent", "#730303", "#d62d2d", "#bd6666", "transparent" ,"transparent"),
      bins = c(0, 0.001, 0.01, 0.05, 0.1, 1),
      na.color = "grey20"
    ))
rda_summary_table
```

```{r}
gtsave(rda_summary_table, filename = "output-tables/RDA-gt-table.html")
gtsave(rda_summary_table, filename = "output-tables/RDA-gt-table.docx")
```

# Indicator Species

## indicspecies
https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html

```{r}
library(indicspecies)
```

multipatt is the name of the command from the indicspecies package. The mulitpatt command results in lists of species that are associated with a particular group of samples. If your group has more than 2 categories, multipatt will also identify species that are statistically more abundant in combinations of categories.

The parameters for multipatt are as follows:

the community abundance matrix:"abund"
the vector that contains your sample grouping information: "time"
the function that multipatt is using to identify indicator species: func = "r.g"
the number of permutations used in the statistical test: control = how(nperm=9999)

https://cran.r-universe.dev/indicspecies/doc/manual.html


```{r}
list_top
range(table_supp$Oxygen_CTD)
```

Isolate the top taxa noted before performing Indicator Species analysis.
```{r}
# | eval: false
# | echo: true
indic_input <- asv_long_avg_wtax_clean %>%
  left_join(table_supp) %>% 
  # filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified %in% list_top) %>% 
  filter(!grepl("_X", Supergroup_simplified)) %>%
   # Add additional bins
  mutate(SAL_BIN = case_when(
    Salinity < 29 ~ "< 29",
    Salinity >= 29 & Salinity < 31 ~ "29-31",
    Salinity >= 31 & Salinity < 34 ~ "31-34",
    Salinity >= 34 ~ "> 34")) %>% 
  mutate(O2_BIN = case_when(
    Oxygen_CTD <= 3 ~ "< 3",
    Oxygen_CTD > 3  ~ "> 3")) %>% 
  # Add additional bins
  mutate(TEMP_BIN = case_when(
    Temperature < 5 ~ "< 5",
    Temperature >=5 & Temperature < 15 ~ "5-15",
    Temperature >= 15 & Temperature < 20 ~ "15-20",
    Temperature >= 20 & Temperature < 25 ~ "20-25",
    Temperature >= 25 ~ "> 25")) %>% 
  mutate(NO3_BIN = case_when(
    NO3 < 10 ~ "< 10",
    NO3 >= 10 & NO3 < 20 ~ "10-20",
    NO3 >= 20 ~ "> 20")) %>% 
  mutate(OIL_BIN = case_when(
      OilRig_Count > 0 ~ "Oil Rig",
      TRUE ~ "No oil rig"
    )) %>% select(STN_NISKIN, Station, DEPTH_BIN, ends_with("_THRESHOLD"), 
         SAL_BIN, TEMP_BIN, NO3_BIN, OIL_BIN, OilRig_Count, O2_BIN, 
         MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, names_prefix = "ASV_", values_from = MEAN_REPS_seq, values_fill = 0)
```

"Indval" two species are not strongly correlated and their frequency in the data set is larger than the frequency of sites belonging to the site group, the joint occurrence of the two species has higher positive predictive value for the site group than the two species taken independently

‘d’ of species as the product of the relative frequency and relative average abundance in clusters.

## Fxn summary of indicator species

```{r}
# INV$sign[1]
summary_fxn_indic_spp <- function(clust, abun, param){
  df_inv <- multipatt(abun, clust, 
                      func = "IndVal", 
                      duleg = TRUE, 
                      control = how(nperm = 99))
  #
  prelim <- (as.data.frame(df_inv$sign))
  max_index <- max(prelim$index)
  summary_inv_df <- (as.data.frame(df_inv$sign)) %>% 
      rownames_to_column(var = "FeatureID") %>% 
    pivot_longer(cols = starts_with("s."), names_to = "Indicative_of", values_to = "P_A") %>% 
    filter(P_A == 1) %>% 
    filter(p.value <= 0.05) %>% 
    mutate(indic_index = (index/max_index),
           FeatureID = str_remove_all(FeatureID, "ASV_")) %>% 
    select(Indicative_of, FeatureID, indic_index, p.value, stat) %>% 
    add_column(PARAM_INDV = param)
}
# ?multipatt()
```

`summary_fxn_indic_spp(clust, abun, param)`

```{r}
abun <- indic_input %>% 
  select(STN_NISKIN, starts_with("ASV_")) %>% 
  column_to_rownames(var = "STN_NISKIN")

# Subset indicator species to all above the DCM?
# head(indic_input)
# unique(indic_input$DEPTH_BIN)
# abun_abovedcm <- indic_input %>% 
#   filter(DCM_THRESHOLD == "Above DCM") %>% 
#   select(STN_NISKIN, starts_with("ASV_")) %>% 
#   column_to_rownames(var = "STN_NISKIN") 
# 
# indic_input_abovedcm <- indic_input %>% 
#   filter(DCM_THRESHOLD == "Above DCM") %>% 
#   mutate(NO3_aboveDCM = case_when(
#     NO3 == 0 ~ "0",
#     (NO3 > 0 & NO3 < 1) ~ "0-1",
#     (NO3 >= 1) ~ ">=1",
#   )) %>% select(-starts_with("ASV_"))
# View(indic_input_abovedcm)

##
# abun_surface <- indic_input %>% 
#   filter(DEPTH_BIN == "< 5m") %>% 
#   select(STN_NISKIN, starts_with("ASV_")) %>% 
#   column_to_rownames(var = "STN_NISKIN")
# 
# indic_input_surface <- indic_input %>% 
#   filter(DEPTH_BIN == "< 5m") %>% 
#   mutate(NO3_aboveDCM = case_when(
#     NO3 == 0 ~ "0",
#     (NO3 > 0 & NO3 < 1) ~ "0-1",
#     (NO3 >=1) ~ ">=1",
#   )) %>% select(-starts_with("ASV_")) %>% 
#   mutate(TEMP_surface = round(Temperature, digits = 0))
```

All depths and stations
```{r}
# depth_indic_bin <- summary_fxn_indic_spp((indic_input$DEPTH_BIN), abun, "Depth")
# dcm_bin <- summary_fxn_indic_spp((indic_input$DCM_THRESHOLD), abun, "DCM")
# mld_bin <- summary_fxn_indic_spp((indic_input$MLD_THRESHOLD), abun, "MLD")
# o2min_bin <- summary_fxn_indic_spp((indic_input$O2_THRESHOLD), abun, "O2_MIN")
# o2_bin <- summary_fxn_indic_spp((indic_input$O2_BIN), abun, "O2")
# no3_bin <- summary_fxn_indic_spp((indic_input$NO3_BIN), abun, "NO3")
# temp_bin <- summary_fxn_indic_spp((indic_input$TEMP_BIN), abun, "TEMP")
# sal_bin <- summary_fxn_indic_spp((indic_input$SAL_BIN), abun, "SAL")
# stn_bin <- summary_fxn_indic_spp((indic_input$Station), abun, "Station")
```

Surface
```{r}
# View(indic_input_surface)
# surface_no3_bin <- summary_fxn_indic_spp((indic_input_surface$NO3_aboveDCM), abun_surface, "NO3")
# surface_temp_bin <- summary_fxn_indic_spp((indic_input_surface$TEMP_surface), abun_surface, "TEMP")
# surface_stn_bin <- summary_fxn_indic_spp((indic_input_surface$Station), abun_surface, "Station")
# surface_sal_bin <- summary_fxn_indic_spp((indic_input_surface$SAL_BIN), abun_surface, "SAL")
```

Above DCM
```{r}
# View(indic_input_abovedcm)
# abovedcm_depth_indic_bin <- summary_fxn_indic_spp((indic_input_abovedcm$DEPTH_BIN), abun_abovedcm, "Depth")
# abovedcm_no3_bin <- summary_fxn_indic_spp((indic_input_abovedcm$NO3_aboveDCM), abun_abovedcm, "NO3")
# abovedcm_temp_bin <- summary_fxn_indic_spp((indic_input_abovedcm$TEMP_BIN), abun_abovedcm, "TEMP")
# abovedcm_sal_bin <- summary_fxn_indic_spp((indic_input_abovedcm$SAL_BIN), abun_abovedcm, "SAL")
# abovedcm_stn_bin <- summary_fxn_indic_spp((indic_input_abovedcm$Station), abun_abovedcm, "Station")
```


To show information, or report, from indicator species analyses, we want to see which taxa were selected for these analyses.


```{r}
tax_annot <- asv_long_avg_wtax_clean %>% 
  select(FeatureID, Supergroup_simplified, SUPERGROUP_CLASS, Taxon) %>% distinct()
```


### All indicator spp results
```{r}
# results_indic_summary <- rbind(depth_indic_bin, dcm_bin, mld_bin, o2min_bin, o2_bin, no3_bin, temp_bin, sal_bin,stn_bin) %>% 
#   left_join(tax_annot) %>% 
#   group_by(Supergroup_simplified, SUPERGROUP_CLASS, Indicative_of, PARAM_INDV) %>% 
#     summarise(COUNT = n()) %>% 
#   mutate(INDIC = str_replace_all(Indicative_of, "s\\.", paste("Indicator_", PARAM_INDV, "_", sep = ""))) %>% ungroup() %>% 
#   select(-Indicative_of) %>% 
#   pivot_wider(names_from = INDIC, values_from = COUNT)
# 
# # View(results_indic_summary)
# results_indic_full <- rbind(depth_indic_bin, dcm_bin, mld_bin, o2min_bin, o2_bin, no3_bin, temp_bin, sal_bin,stn_bin) %>%
#   left_join(tax_annot) %>%
#   group_by(Supergroup_simplified, SUPERGROUP_CLASS, FeatureID, Taxon, Indicative_of, PARAM_INDV) %>%
#     summarise(COUNT = n()) %>%
#   mutate(INDIC = str_replace_all(Indicative_of, "s\\.", paste("Indicator_", PARAM_INDV, "_", sep = ""))) %>%
#   select(-Indicative_of) %>%
#   pivot_wider(names_from = INDIC, values_from = COUNT)

# write.csv(results_indic_summary, file = "output-tables/summary-indic-spp.csv")
# write.csv(results_indic_full, file = "output-tables/fulllist-indic-spp.csv")
```

#### questions regarding indicator spp.

Diatoms
```{r}
# glimpse(results_indic_full)
# View(results_indic_full %>%
#        filter((Indicative_of == "s.2") |(Indicative_of == "s.1")))
       # filter(grepl("Hemiaulales", Taxon)))
      # filter(grepl("Rhizosoleniaceae", Taxon)))
# filter(grepl("Pseudo-nitzschia", Taxon)))

```


Specific stations
```{r}
# glimpse(results_indic_aboveDCM_full)
# glimpse(results_indic_full)
# View(results_indic_full %>%
# # Oil spill
# filter(!is.na(Indicator_Station_9)))
       # eddy center
       # filter(!is.na(Indicator_Station_7)))
       # eddy edge effect
        # filter(!is.na(Indicator_Station_3)))
        # filter(!is.na(Indicator_Station_1)))

# View(results_indic_aboveDCM_full %>%
       # filter(!is.na(Indicator_Station_8)))
```

```{r}
# View(asv_long_avg_wtax %>% 
#        filter(grepl("Acanthochiasma", Taxon)))
```



```{r}
# formatted_indic_table <- results_indic_summary %>% 
#   gt(rowname_col = "TAXA",
#      groupname_col = "Supergroup_simplified") %>% 
#   tab_spanner(label = "DCM",
#               columns = c("Indicator of Above DCM", "Indicator of Below DCM")) %>% 
#   tab_spanner(label = "Oxygen minimum",
#               columns = c("Indicator of Above oxygen minimum", "Indicator of Below oxygen minimum")) %>%
#   tab_spanner(label = "Temperature °C",
#               columns = c("Indicator of < 8", "Indicator of 8-15", 
#                           "Indicator of 15-20", "Indicator of 20-25", "Indicator of > 25")) %>% 
#   tab_spanner(label = "Depth",
#               columns = c("Indicator of < 5m","Indicator of 5-25m","Indicator of 25-60m",
#                           "Indicator of 60-270m","Indicator of 270-400m","Indicator of 400-800m",
#                           "Indicator of 800-1200m","Indicator of 1200-2000m","Indicator of 2000-3000m")) %>% 
#   tab_spanner(label = "Nitrate µmol/L", 
#               columns = c("Indicator of < 10", "Indicator of 10-20", "Indicator of > 20")) %>% 
#   sub_missing(
#     columns = everything(),
#     rows = everything(),
#     missing_text = "---")
# 
# formatted_indic_table
# gtsave(formatted_indic_table, filename = "output-tables/sppindicator-gt-table.html")
# gtsave(formatted_indic_table, filename = "output-tables/sppindicator-gt-table.docx")
```

```{r}
# results_indic_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_DCM"), starts_with("Indicator_O2")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
  
```

```{r}
# colnames(results_indic_full)
# results_indic_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_TEMP"), starts_with("Indicator_NO3"), `Indicator_O2_> 3`, `Indicator_O2_< 3`) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```


```{r}
# results_indic_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_NO3"), starts_with("Indicator_Station")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```

```{r}
# results_indic_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_NO3"), starts_with("Indicator_Depth")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```

```{r}
# results_indic_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_NO3"), starts_with("Indicator_MLD")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```


### Above DCM indicator spp results
```{r}
# results_indic_aboveDCM_summary <- rbind(abovedcm_depth_indic_bin, abovedcm_sal_bin,abovedcm_no3_bin, abovedcm_temp_bin, abovedcm_stn_bin) %>% 
#   left_join(tax_key_annot) %>%  
#   group_by(Supergroup_simplified, Indicative_of, PARAM_INDV) %>% 
#     summarise(COUNT = n()) %>% 
#   mutate(INDIC = str_replace_all(Indicative_of, "s\\.", paste("Indicator_", PARAM_INDV, "_", sep = ""))) %>% 
#   select(-Indicative_of) %>% 
#   pivot_wider(names_from = INDIC, values_from = COUNT)
# # View(results_indic_aboveDCM_summary)
# 
# results_indic_aboveDCM_full <- rbind(abovedcm_depth_indic_bin, abovedcm_sal_bin,abovedcm_no3_bin, abovedcm_temp_bin, abovedcm_stn_bin) %>% 
#   left_join(tax_key_annot) %>%   
#   group_by(Supergroup_simplified, FeatureID, Taxon, Indicative_of, PARAM_INDV) %>% 
#     summarise(COUNT = n()) %>% 
#   mutate(INDIC = str_replace_all(Indicative_of, "s\\.", paste("Indicator_", PARAM_INDV, "_", sep = ""))) %>%  
#   select(-Indicative_of) %>% ungroup() %>% 
#   pivot_wider(names_from = INDIC, values_from = COUNT)
# View(results_indic_aboveDCM_full)
```

```{r}
# head(results_indic_aboveDCM_full)
```


```{r}
# results_indic_aboveDCM_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_NO3"), starts_with("Indicator_TEMP")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```


```{r}
# results_indic_aboveDCM_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_SAL"), starts_with("Indicator_Station")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```

```{r}
# results_indic_aboveDCM_full %>% ungroup() %>% 
#   filter(!is.na(Supergroup_simplified)) %>% 
#   select(Supergroup_simplified, FeatureID, starts_with("Indicator_SAL"), starts_with("Indicator_NO3")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Supergroup_simplified, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Supergroup_simplified, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Supergroup_simplified)) +
#   scale_x_upset(n_intersections = 25)
```

# Dinoflagellate & ciliate functional traits

```{r}
tmp <- asv_long_avg_wtax_clean %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") #%>% 
  # filter(Class == "Syndiniales")
length(unique(asv_long_avg_wtax_clean$FeatureID))
length(unique(tmp$FeatureID))

# 10563 # all ASVs
# 3611 # Syndiniales Class
# 5521 # Dinoflagellates
# 5521 / 10563 # Dino percentage of ASVs
# 3611 / 5521 # Syndiniales percentage of dino ASVs
head(asv_long_avg_wtax_clean)
```

```{r}
non_syn_dinos_cil <- asv_long_avg_wtax_clean %>%
  filter((Supergroup_simplified == "TSAR-Dinoflagellata") | (Supergroup_simplified == "TSAR-Ciliophora")) %>% 
  filter(Class != "Syndiniales") %>% 
  select(FeatureID, Supergroup_simplified, Order, Family, Genus, Species) %>% distinct() %>% 
  drop_na()
table(non_syn_dinos_cil$Order)

# Export for manual curation
# write.csv(non_syn_dinos_cil, "non_syn_dinos_cil.csv")
```

Manual curation done, compared to Jones et al. table.

```{r}
manual_curated_fxn <- read.csv("non_syn_dinos_cil.csv")
jones_db <- read.csv("../tax_annotation_proj/TMDv1.1_Jones.csv")

trait_asv <- asv_long_avg_wtax_clean %>% 
  left_join(manual_curated_fxn) %>% 
  mutate(Phylum = case_when(
    Class == "Syndiniales" ~ "Dinoflagellata",
    TRUE ~ Phylum
  ),
  `Grazer.Taxonomic.ID` = case_when(
    Class == "Syndiniales" ~ "Syndiniales (Class)",
    TRUE ~ `Grazer.Taxonomic.ID`
  )) %>% 
  left_join(jones_db) %>% 
  filter((Supergroup_simplified == "TSAR-Dinoflagellata") | (Supergroup_simplified == "TSAR-Ciliophora")) %>% 
  mutate(MORPHOTYPE = case_when(
    Morphotype == "PA" ~ "Parasite",
    Morphotype == "CM" ~ "Constitutive mixotroph",
    Morphotype == "CM*" ~ "Constitutive mixotroph",
    Morphotype == "CM/PA" ~ "Constitutive mixotroph/parasite",
    Morphotype == "GH/PA" ~ "Generalist heterotroph/parasite",
    Morphotype == "GH" ~ "Generalist Heterotroph",
    Morphotype == "SH" ~  "Selective heterotroph",
    Morphotype == "CV" ~  "Generalist, carnivore heterotroph",
    Morphotype == "pSNCM" ~ "Plastid bearing Specialist Non-Constitutive Mixotroph",
    grepl("GNCM", Morphotype) ~ "Generalist non-constitutive mixotroph",
    `Feeding.Mode` == "Autotroph" ~ "Autotroph",
    TRUE ~ Morphotype))
glimpse(trait_asv)
```

```{r}
# write.csv(trait_asv, file = "output-tables/fxntrait-ciliate-dino-summary.csv")
```

### Export sankey

> Use copilot to help me write out?

```{r}
head(trait_asv)
# tax_color
dino_colors
# ciliate_colors
first <- trait_asv %>% group_by(DIVISION) %>% summarise(COUNT = n()) %>% 
  mutate(COLOR = case_when(
    DIVISION == "Dinoflagellata" ~ "#ee429b",
    DIVISION == "Ciliophora" ~ "#1abc9c",
  )) %>% 
  mutate(SANKEY = paste("All", " [", COUNT, "] ", DIVISION, COLOR, sep = "")) %>%
  ungroup() %>%
  select(SANKEY)

```

```{r}
# dino_colors
syn_split <- trait_asv %>% 
  filter(DIVISION == "Dinoflagellata") %>% 
  mutate(div = case_when(
    Class == "Syndiniales" ~ "Syndiniales",
    TRUE ~ "Non-Syndiniales"
  )) %>% 
  group_by(DIVISION, div) %>% 
  summarise(COUNT = n()) %>%
  mutate(COLOR = case_when(
    div == "Syndiniales" ~ "#bfad0a",
    TRUE ~ "#ee429b")) %>% 
  mutate(SANKEY = paste(DIVISION, " [", COUNT, "] ", div, " ", COLOR, sep = "")) %>% ungroup() %>%
  select(SANKEY)
```
```{r}
# head(trait_asv)
dino_split <- trait_asv %>% 
  filter((DIVISION == "Dinoflagellata")) %>% 
  mutate(ORDER = case_when(
    is.na(Order) ~ "Other",
    # Order == "Dinophyceae" ~ "Other",
    # Order == "Dinophyta"~ "Other",
    Order == "Dinophyta_XX" ~ "Dinophyta",
    Order == "Dinophyceae_X"~ "Dinophyceae",
    TRUE ~ Order
  )) %>% 
  unite(TAXA, Class, ORDER, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    grepl("Dinophyceae_X", TAXA) ~ "Dinophyceae",
    grepl("Dinophyta_X", TAXA) ~ "Dinophyta",
    TRUE ~ TAXA
  )) %>% 
  mutate(div = case_when(
    Class != "Syndiniales" ~ "Non-Syndiniales",
    Class == "Syndiniales" ~ "Syndiniales"
  )) %>% 
  mutate(TAXA = case_when(
    TAXA == "Dinophyceae;Dinophyceae" ~ "Dinophyceae",
    TAXA == "Dinophyceae;Other" ~ "Dinophyceae",
    TAXA == "Syndiniales;Other" ~ "Syndiniales",
    TAXA == "Dinophyta" ~ "Dinophyceae",
    TRUE ~ TAXA
  )) %>% 
  group_by(div, TAXA) %>% 
  summarise(COUNT = n()) %>% 
  drop_na() %>% 
  mutate(COLOR = factor(TAXA, levels = names(dino_colors), labels = dino_colors)) %>% 
  mutate(TAXA = str_remove_all(TAXA, "Syndiniales;")) %>% 
  mutate(TAXA = str_remove_all(TAXA, "Dinophyceae;")) %>% 
  mutate(TAXA = str_remove_all(TAXA, "Noctilucophyceae;")) %>% 
  mutate(SANKEY = paste(div, " [", COUNT, "] ", TAXA, " ", COLOR, sep = "")) %>% 
  ungroup() %>%
  select(SANKEY)

```
```{r}
# head(trait_asv)
morpho <- as.character(unique(trait_asv$MORPHOTYPE))[-3]
morpho
morpho_color <- c("#bf4e40", "#36305a", "#d19328", "#dbbd88", "#a35527",
                  "#5c5cbd", "#1d6e4f", "#d45984", "#94d459")
names(morpho_color) <- morpho

dino_morpho <- trait_asv %>% 
  filter((DIVISION == "Dinoflagellata")) %>% 
  mutate(ORDER = case_when(
    is.na(Order) ~ "Dinophyceae",
    # Order == "Dinophyceae" ~ "Other",
    # Order == "Dinophyta"~ "Other",
    Order == "Dinophyta_XX" ~ "Dinophyceae",
    Order == "Dinophyceae_X"~ "Dinophyceae",
    TRUE ~ Order
  )) %>%
  group_by(ORDER, MORPHOTYPE) %>% 
  summarise(COUNT = n()) %>% 
  mutate(COLOR = factor(MORPHOTYPE, levels = names(morpho_color), labels = morpho_color)) %>% 
  mutate(SANKEY = paste(ORDER, " [", COUNT, "] ", MORPHOTYPE, " ", COLOR, sep = "")) %>% ungroup() %>%
  select(SANKEY)
```


```{r}
ciliate_colors
other_ciliates <- c("Oligohymenophorea;Oligohymenophorea_X", "Litostomatea;Haptoria_6", "Heterotrichea;Heterotrichea_X", "Spirotrichea;Hypotrichia", "Ciliophora_X;Ciliophora_XX")
ciliate_order <- trait_asv %>% 
  filter((DIVISION != "Dinoflagellata")) %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    (TAXA %in% other_ciliates) ~ "Ciliates unannotated",
    grepl("CONTH_", TAXA) ~ "CONThreeP unannotated",
    grepl("CONThreeP_X", TAXA) ~ "CONThreeP unannotated",
    (Class == "Discotrichidae") ~ "CONThreeP;Discotrichidae",
    grepl("Phyllopharyngea;Cyrtophoria", TAXA) ~ "Phyllopharyngea;Cyrtophoria",
    grepl("Spirotrichea;Choreotrichida", TAXA) ~ "Spirotrichea;Choreotrichida",
    grepl("Litostomatea;Litostomatea_X", TAXA) ~ "Litostomatea",
    grepl("Prostomatea_1;Prostomatea_1_X", TAXA) ~ "Prostomatea",
    grepl("Colpodea_X", TAXA) ~ "Colpodea",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  group_by(DIVISION, TAXA) %>% 
  summarise(COUNT = n()) %>% 
  mutate(COLOR = factor(TAXA, levels = names(ciliate_colors), labels = ciliate_colors)) %>% 
  mutate(SANKEY = paste(DIVISION, " [", COUNT, "] ", TAXA, " ", COLOR, sep = "")) %>% ungroup() %>%
  select(SANKEY)
```
```{r}
ciliate_morpho <- trait_asv %>% 
  filter((DIVISION != "Dinoflagellata")) %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    (TAXA %in% other_ciliates) ~ "Ciliates unannotated",
    grepl("CONTH_", TAXA) ~ "CONThreeP unannotated",
    grepl("CONThreeP_X", TAXA) ~ "CONThreeP unannotated",
    (Class == "Discotrichidae") ~ "CONThreeP;Discotrichidae",
    grepl("Phyllopharyngea;Cyrtophoria", TAXA) ~ "Phyllopharyngea;Cyrtophoria",
    grepl("Spirotrichea;Choreotrichida", TAXA) ~ "Spirotrichea;Choreotrichida",
    grepl("Litostomatea;Litostomatea_X", TAXA) ~ "Litostomatea",
    grepl("Prostomatea_1;Prostomatea_1_X", TAXA) ~ "Prostomatea",
    grepl("Colpodea_X", TAXA) ~ "Colpodea",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  group_by(TAXA, MORPHOTYPE) %>% 
  summarise(COUNT = n()) %>% 
  mutate(COLOR = factor(MORPHOTYPE, levels = names(morpho_color), labels = morpho_color)) %>% 
  mutate(SANKEY = paste(TAXA, " [", COUNT, "] ", MORPHOTYPE, " ", COLOR, sep = "")) %>% ungroup() %>% 
  select(SANKEY)
```

```{r}
# rbind(first,syn_split, dino_split, dino_morpho, ciliate_order, ciliate_morpho)
```

## Get stats on functional trait profiles

```{r}
# head(trait_asv)
```


### Vertical profiles for dinoflagellates.
```{r}
# head(dino_trait_asv)
# depth_bin <- c("< 5m","5-25m","25-60m",
#                           "60-270m","270-400m","400-800m",
#                           "800-1200m","1200-2000m","2000-3000m")
# # depth_bin
# # Dinophyceae;Gymnodiniales
# # Dinophyceae;Gonyaulacales
# head(dino_trait_asv)

```

```{r}
# dino_trait_asv %>% 
#   filter(!is.na(feeding_mode)) %>% 
#   filter(!grepl("Syndiniales", Taxon)) %>% # no parasites
#   mutate(Order_feedingmode = paste(Order, feeding_mode, sep = " ")) %>% 
#   separate(STN_NISKIN, into = c("Stn", "Nisk"), remove = FALSE, sep = "_") %>% 
#   mutate(STN_ORDER = factor(Stn, levels = offshore_on_shore_order)) %>% 
#   mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
#   group_by(Order_feedingmode, Order, STN_ORDER, DEPTH_BIN_ORDER) %>% 
#     summarise(SUM = sum(MEAN_REPS_seq),
#               ASV = n()) %>% 
#   ggplot(aes(x = Order_feedingmode, y = DEPTH_BIN_ORDER, fill = ASV)) +
#   geom_tile(color = "grey30", linewidth = 0.4) +
#     facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
#     scale_x_discrete(position = "top") +
#     scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
#     theme_classic() +
#     theme(axis.ticks = element_blank(),
#           axis.text.x = element_text(color = "black", angle = 45, size = 7,
#                                      hjust = 0, vjust = 0.5),
#           axis.text.y = element_text(color = "black"),
#           strip.text.y = element_text(color = "black", angle = 0),
#           strip.background = element_blank()) +
#     labs(x = "", y = "")
```
```{r}
# dino_trait_asv %>% 
#   filter(!is.na(feeding_mode)) %>% 
#   filter(!grepl("Syndiniales", Taxon)) %>% # no parasites
#   mutate(Order_feedingmode = paste(Order, feeding_mode, MORPHOTYPE,sep = " ")) %>% 
#   separate(STN_NISKIN, into = c("Stn", "Nisk"), remove = FALSE, sep = "_") %>% 
#   mutate(STN_ORDER = factor(Stn, levels = offshore_on_shore_order)) %>% 
#   mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
#   group_by(Order_feedingmode, Order, DEPTH_BIN_ORDER) %>% 
#     summarise(SUM = sum(MEAN_REPS_seq),
#               ASV = n()) %>% 
#   ggplot(aes(x = Order_feedingmode, y = DEPTH_BIN_ORDER, fill = ASV)) +
#   geom_tile(color = "grey30", linewidth = 0.4) +
#     # facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
#     scale_x_discrete(position = "top") +
#     scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
#     theme_classic() +
#     theme(axis.ticks = element_blank(),
#           axis.text.x = element_text(color = "black", angle = 45, size = 7,
#                                      hjust = 0, vjust = 0.5),
#           axis.text.y = element_text(color = "black"),
#           strip.text.y = element_text(color = "black", angle = 0),
#           strip.background = element_blank()) +
#     labs(x = "", y = "")
```

```{r}
# dino_trait_asv %>% 
#   filter(grepl("Gymnodiniales", Taxon)) %>% 
#   # filter(grepl("Gymnodiniales", Taxon) | grepl("Gonyaulacales", Taxon)) %>% 
#   filter(!is.na(MORPHOTYPE)) %>% 
#   group_by(STN_CATEGORY_4, MORPHOTYPE, FeatureID, DEPTH_BIN, Taxon, Genus, Species) %>% 
#     summarise(ASV_COUNT = n()) %>% 
#   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
#   ggplot(aes(x = FeatureID, y = depth_bin_order, fill = MORPHOTYPE)) +
#     geom_tile() +
#   # geom_path(aes(group = FeatureID)) +
#   # geom_point(stat = "identity", shape = 21, color = "black") +
#   # scale_y_reverse() +
#   facet_grid(rows = vars(STN_CATEGORY_4), cols = vars(Genus), space = "free", scale = "free_x") +
#   theme_bw() +
#   theme(axis.text.x = element_blank())
```

## Incorporate Indicator spp results.
From indicator results, how many have a known feeding mode
```{r}
# glimpse(results_indic_aboveDCM_full)
# # length(unique(results_indic_aboveDCM_full$FeatureID))
# # 558 observations
# head(dino_trait_asv)
# # glimpse(dino_trait_asv)
# # length(unique(dino_trait_asv$FeatureID))
# # 5521
# indic_plus_trait <- dino_trait_asv %>% 
#   left_join(results_indic_aboveDCM_full) %>% 
#   filter(!is.na(feeding_mode)) %>% 
#   filter(!is.na(Indicative_of))
# length(unique(indic_plus_trait$FeatureID))
# # View(indic_plus_trait)
# unique(indic_plus_trait$Order)
# unique(indic_plus_trait$feeding_mode)
# colnames(indic_plus_trait)
```

14 total ASVs that also had functional traits, what were these ASVs indicator for?
```{r}
# indic_plus_trait %>% 
#   mutate(Order_feedingmode = paste(Order, feeding_mode, sep = " ")) %>% 
#   ungroup() %>% 
#   select(Order_feedingmode, FeatureID, 
#          # starts_with("Indicator_SAL"),
#          starts_with("Indicator_Depth")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "indicator_list", values_to = "asv_count") %>% 
#   filter(!is.na(asv_count)) %>% distinct() %>% 
#   group_by(Order_feedingmode, FeatureID, indicator_list) %>% 
#     summarise(SUM = sum(asv_count)) %>% 
#   ungroup() %>% 
#   group_by(Order_feedingmode, FeatureID) %>% 
#   summarise(INDICATOR_CATEGORY = list(indicator_list)) %>% 
#   ggplot(aes(x = INDICATOR_CATEGORY)) +
#   geom_bar(aes(fill = Order_feedingmode)) +
#   scale_x_upset(n_intersections = 25)
```
```{r}
# indic_plus_trait %>% 
#   # mutate(Order_feedingmode = paste(Order, feeding_mode, sep = " ")) %>% 
#   select(Order, feeding_mode, MORPHOTYPE, starts_with("Indicator_")) %>% 
#   pivot_longer(cols = starts_with("Indicator_"), names_to = "Indicator", values_to = "asv_count") %>% 
#   mutate(Indicator = str_replace_all(Indicator, "Indicator_", "")) %>% 
#   mutate(Indicator = str_replace_all(Indicator, "_", " ")) %>% 
#   filter(!is.na(asv_count)) %>% 
#   group_by(Order, feeding_mode, MORPHOTYPE, Indicator) %>% 
#     summarise(ASV_total = sum(asv_count)) %>% ungroup() %>% 
#   group_by(Order, feeding_mode, MORPHOTYPE) %>% 
#     summarise(Category = paste(Indicator, collapse = ", "))
```


# Session information

```{r}
sessionInfo()
```
