---
title: "Analysis of 18S rRNA metabarcoding"
editor: visual
---

# Contents & introduction

# Set up R environment

Set up R & import starting data

```{r}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

# Import prior data

From sequencing

```{r}
# | eval: false
# save(samples_too_low, env_info, asv_wide_cleaned, asv_long_cleaned, file =  "input-data/asv_wtax_qc_GoM23_062025.RData")
```

Import sequence R objects

```{r}
load(file = "input-data/asv_wtax_qc_GoM23_062025.RData", verbose = TRUE)
env_info <- read.csv("input-data/metadata-gom-23.csv")
tmp <- read.csv("cellcounts-sept22.csv")
prok <- tmp %>% select(Station, Niskin, PROK_COUNT = `prok.count`)
# head(env_info)
```

```{r}
# head(prok)
env_info <- env_info %>% 
  left_join(prok)
```

## Station information, ordering, location

Distance from shore

```{r}
library(geosphere)
# Long, Lat order
outflow_lat <- 28.95
outflow_long <- 89.39
outflow <- c(outflow_long, outflow_lat)
outflow
```

Set up ordering of stations and locations.

```{r}
stn_info <- read.csv("input-data/manual_stn_classification.csv")

offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5","S9", "S8", "S7", "S11", "S12", "S14", "S15")

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(env_info$DEPTH))))

depth_order_round <- as.character(sort(round(unique(env_info$DEPTH), digits = 0)))

depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")

depth_bin_CHARC <- c("..5m","5.25m","25.60m",
                          "60.270m","270.400m","400.800m",
                          "800.1200m","1200.2000m","2000.3000m")

env_info_mod <- env_info %>% 
  left_join(stn_info) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order),
         TRANSECT = factor(stn, levels = offshore_on_shore_order, labels = transect_labels),
         DEPTH_ORDER = factor(DEPTH, levels = rev(depth_order), labels = rev(depth_order_round))) %>% 
  mutate(DIST_OUTFLOW = distHaversine(cbind(Longitude, Latitude), cbind(outflow_long, outflow_lat))) %>% 
  mutate(DEPTH_BIN = case_when(
    DEPTH <= 5 ~ "< 5m",
    DEPTH > 5 & DEPTH <=25 ~ "5-25m",
    DEPTH > 25 & DEPTH <=60 ~ "25-60m",
    DEPTH > 60 & DEPTH <=270 ~ "60-270m",
    DEPTH > 270 & DEPTH <=400 ~ "270-400m",
    DEPTH > 400 & DEPTH <=800 ~ "400-800m",
    DEPTH > 800 & DEPTH <=1200 ~ "800-1200m",
    DEPTH > 1200 & DEPTH <=2000 ~ "1200-2000m",
    DEPTH > 2000 & DEPTH <=3000 ~ "2000-3000m",
  ))
head(env_info_mod)
# # unique(env_info_mod$DIST_OUTFLOW)
# table(env_info_mod$DEPTH)
# unique(env_info_mod$STN_ORDER)
```

How many stations?
```{r}
head(env_info_mod)
length(unique(env_info_mod$SAMPLES)) # 189 total samples
length(unique(env_info_mod$stn)) # 13 stations
length(unique(env_info_mod$DEPTH)) # 92 different depth.s
# table((env_info_mod %>% select(stn, DEPTH)))
# tmp <- env_info_mod %>% filter(rep != "orig") %>% distinct()
# table(tmp$stn)
```

Create table
```{r}
head(env_info)

table1 <- env_info %>% 
  filter(stn != "S0") %>% arrange(Station, DEPTH) %>% 
  select(stn, nisk, DEPTH, Date, Time_UTC, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Latitude, Longitude) %>% distinct()
# head(table1)
# write.csv(table1, file = "output-tables/table-samplelist.csv")
```


# Curate taxonomic assignment

```{r}
# head(asv_long_cleaned)
# unique(asv_long_cleaned$Domain) # take EUKS
# unique(asv_long_cleaned$Supergroup)
# unique(asv_long_cleaned$Division)
# View(asv_long_cleaned %>% 
#        select(Domain, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct())
```

```{r}
asv_long_clean_wtax <- asv_long_cleaned %>% 
  filter(Domain == "Eukaryota") %>% 
  filter(Supergroup != "nucl") %>% 
  mutate(DIVISION = case_when(
    Subdivision == "X" ~ "Unannotated",
    is.na(Subdivision) ~ "Unannotated",
    TRUE ~ Subdivision
  )) %>% 
  # Goal is to combine Subdivision-Class
  mutate(SUPERGROUP_CLASS = case_when(
    ## High level NAs at the supergroup and division level
    (is.na(Division) & !is.na(Supergroup)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Division) & is.na(Supergroup)) ~ "Eukaryote-Unannotated",
    (is.na(Subdivision)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Class) & !is.na(Subdivision)) ~ paste(Subdivision, "Unannotated", sep = "-"),
    #
    (Subdivision == "X") ~ paste(Division, "Unannotated", sep = "-"),
    (Class == "X") ~ paste(Subdivision, "Unannotated", sep = "-"),
    TRUE ~ paste(Subdivision, Class, sep = "-"))) %>% 
  mutate(Supergroup_simplified = case_when(
    Supergroup == "TSAR" ~ paste("TSAR", Subdivision, sep = "-"),
    Supergroup == "Obazoa" ~ paste("Obazoa", DIVISION, sep = "-"),
    TRUE ~ Supergroup
  )) %>% 
  left_join(env_info_mod) 

# 
# sort(unique(tmp$DIVISION))
# sort(unique(tmp$Subdivision))
# unique(tmp$SUPERGROUP_CLASS)
# head(asv_long_clean_wtax)
tax_key <- asv_long_clean_wtax %>% 
  select(FeatureID, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct()
```


## Normalization and averaging across replicates

```{r}
asv_long_avg_wtax <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, PROX_TOSHORE, DIST_OUTFLOW, STN_ORDER, STN_CATEGORY_3, STN_CATEGORY_4, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, DEPTH_BIN, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup_simplified, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_ID, stn, nisk, sep = " ", remove = FALSE)

head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$STN_CATEGORY_4)
```

> Create a figure with data points for each station and depth. Size of data point = \# of ASVs

```{r, fig.height=10, fig.width=8}
asv_long_avg_wtax %>% 
  group_by(DEPTH_ORDER, STN_ORDER, Station, DEPTH) %>% 
    summarise(SEQ_SUM = sum(MEAN_REPS_seq),
              ASV_COUNT = n()) %>% 
  ggplot(aes(x = STN_ORDER, y = DEPTH, size = ASV_COUNT)) +
    geom_jitter(shape = 21, color = "black", stroke = 2) +
  scale_y_reverse() +
  scale_x_discrete(position = "top") +
     scale_size(range=c(1,12), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    panel.grid.major.x = element_line()
  ) +
  labs(x = "", y = "Depth (m)")
```

```{r}
tax_color  <- c("#db1f7d", "#2c5c1d", "#806f8c",
                "#f5bfb0", "#93c47d","#7c52bf", "#007a7a", 
                "#34495e", "#9b59b6", "#fea02f","#ee429b",
                "#ffe8a4","#ff6f61","#fcf517","#8a0622",
                "#ff7530", "#2986cc", "#e71414","#de6600", 
                "#1abc9c", "#a0b59f", "#d7b6d7", "#23b521", "#5b545c", "black", "white", "grey10", "grey60")


tax_simplified_list <- as.character((unique(asv_long_avg_wtax$Supergroup_simplified)))
names(tax_color) <- tax_simplified_list
```

Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level.

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
asv_long_avg_wtax %>%
  group_by(PROX_TOSHORE, TRANSECT, STN_ORDER, DEPTH_ORDER, SAMPLE_ID, DEPTH, TEMP, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  ggplot(aes(y = (DEPTH_ORDER), x = SUM_SEQ, fill = Supergroup_simplified)) +
    geom_bar(stat = "identity", position = "fill", color = "black") +
  facet_grid(rows = vars(PROX_TOSHORE, STN_ORDER), space = "free", scales = "free") +
  scale_fill_manual(values = tax_color) +
  theme_classic() +theme(
    strip.background = element_blank(),
    axis.text = element_text(color = "black")
  )
```

```{r, fig.width=4, fig.height=7}
asv_long_avg_wtax %>%
  group_by(STN_ORDER, DEPTH_ORDER, TEMP, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  ggplot(aes(y = (DEPTH_ORDER), x = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.8) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(STN_ORDER), scales = "free_y", space = "free", axis.labels = "all_x") +
  # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

## Only surface parameters

```{r, fig.width=8, fig.height=3}
asv_long_avg_wtax %>%
  filter(DEPTH < 5) %>% 
  group_by(stn, STN_ORDER, Supergroup_simplified, TRANSECT) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  ggplot(aes(x = STN_ORDER, y = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.6) +
  scale_fill_manual(values = tax_color) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(TRANSECT), scales = "free", space = "free") +
  # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
  theme_classic() + theme(
    legend.position = "left",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Station", y = "Relative Sequence Abundance")
```

# PCA analysis

Create wide format matrix with sequence counts averaged across replicates (by ASV)

```{r}
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_reps_matrix))))
# ?prcomp
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

```{r, fig.height=5, fig.width=6}
blues <- c("#d0d1e6","#a6bddb","#74a9cf","#3690c0", "#0570b0", "#045a8d", "#023858")
# scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) 

data.frame(pca_clr$x, SAMPLE = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, SAMPLE) %>% 
  separate(SAMPLE, c("STN_ORDER", "DEPTH_ORDER"), sep = "_", remove = FALSE) %>% left_join(env_info_mod) %>% 
  mutate(Depth = as.numeric(DEPTH_ORDER)) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, shape = TRANSECT, fill = Depth)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_point(size = 2.5, stroke = 0.4) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(21, 23, 24, 25)) +
  scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=10),
        legend.title = element_blank(),
        axis.title = element_text(color="black", size=8),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(shape = guide_legend(override.aes = list(fill = "black")))
```

> PCA results to consider. Communities are more similar to one another at depth, surface samples are influenced by other parameters. Should do RDA with all data and then shallower than 500 meters.

```{r}
library(ggdendro)
rel_abun <- decostand(asv_wide_avg_reps_matrix, MARGIN = 2, method = "total")
# Cluster dendrogram (average hierarchical clustering)
cluster_gom <- hclust(dist(t(rel_abun)), method = "average")
dendro <- as.dendrogram(cluster_gom)
gom_dendro <- dendro_data(dendro, type = "rectangle")
```

```{r, fig.height=18, fig.width=4}
gom_dendro_plot <- ggplot(segment(gom_dendro)) +
  geom_segment(aes(x = x, y = y, xend = xend, 
    yend = yend)) + 
  coord_flip() + 
  scale_y_reverse(expand = c(0.2, 0.5), breaks = c(0, 0.2, 0.4, 0.6, 0.8)) +
  geom_text(aes(x = x, y = y, label = label, angle = 0, hjust = 0), data = label(gom_dendro)) +
  theme_dendro() + labs(y = "Dissimilarity") +
    theme(axis.text.x = element_text(color = "black", size = 14), axis.line.x = element_line(color = "#252525"), 
        axis.ticks.x = element_line(), axis.title.x = element_text(color = "black", 
            size = 14))
# svg('figs/SUPPLEMENTARY-dendrogram-wreps.svg', w = 10, h = 8)
gom_dendro_plot
```

Isolate label names that correspond to dendrogram. 
```{r}
tmp <- data.frame(gom_dendro$labels)
class(tmp)
labels_for_dendro <- as.character(tmp$label)
# labels_for_dendro
```
Replot taxonomy relative abundance

```{r}
head(asv_long_avg_wtax)
```


```{r}
head(asv_long_avg_wtax)
```


```{r, fig.height=12, fig.width=4}
blues <- c("#d0d1e6","#a6bddb","#74a9cf","#3690c0", "#0570b0", "#045a8d", "#023858")
# 

DEPTH <- asv_long_avg_wtax %>%
  unite(LABELS, STN_ORDER, DEPTH_ORDER, sep = "_", remove = FALSE) %>% 
  select(LABELS, STN_ORDER, DEPTH_ORDER, DIST_OUTFLOW, TRANSECT, DEPTH, TEMP, SALINITY, OXYGEN) %>% distinct() %>% 
  mutate(label_order = factor(LABELS, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = DEPTH)) +
    geom_point(shape = 21, color = "black", size = 4) +
  scale_y_discrete(position = "right") +
    scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
```

```{r, fig.height=12, fig.width=4}
DIST <- asv_long_avg_wtax %>%
  unite(LABELS, STN_ORDER, DEPTH_ORDER, sep = "_", remove = FALSE) %>% 
  select(LABELS, STN_ORDER, DEPTH_ORDER, DIST_OUTFLOW, TRANSECT, DEPTH, TEMP, SALINITY, OXYGEN) %>% distinct() %>% 
  mutate(dist_outflow = DIST_OUTFLOW/1000) %>% 
  mutate(label_order = factor(LABELS, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = dist_outflow)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_continuous(option = "magma") +
  scale_fill_viridis_c(option = "magma", direction = -1) +
    # scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
DIST
```

```{r, fig.height=12, fig.width=4}
TRANSECT <- asv_long_avg_wtax %>%
  unite(LABELS, STN_ORDER, DEPTH_ORDER, sep = "_", remove = FALSE) %>% 
  select(LABELS, STN_ORDER, DEPTH_ORDER, DIST_OUTFLOW, TRANSECT, DEPTH, TEMP, SALINITY, OXYGEN) %>% distinct() %>% 
  # mutate(dist_outflow = DIST_OUTFLOW/1000) %>% 
  mutate(label_order = factor(LABELS, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = TRANSECT)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
TRANSECT
```


```{r, fig.width=17, fig.height=15}
gom_dendro_plot + (asv_long_avg_wtax %>%
  unite(LABELS, STN_ORDER, DEPTH_ORDER, sep = "_", remove = FALSE) %>% 
  group_by(STN_ORDER, LABELS, DEPTH_ORDER, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  mutate(label_order = factor(LABELS, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.8) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0), position = "top") +
  scale_y_discrete(position = "right") +
  # facet_grid(rows = vars(STN_ORDER), scales = "free_y", space = "free", axis.labels = "all_x") +
  # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_text(color = "black",)) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")) + 
  DEPTH + TRANSECT + DIST +
  plot_layout(nrow = 1)
```


# CLR taxonomy
```{r}
surface_matrix<- asv_long_avg_wtax %>%
  filter(DEPTH < 5) %>% 
  group_by(stn, STN_ORDER, Supergroup_simplified, TRANSECT) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  ungroup() %>% 
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Provora") %>% 
  # filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  filter(!grepl("Obazoa-", Supergroup_simplified)) %>% 
  select(STN_ORDER, Supergroup_simplified, SUM_SEQ) %>% 
  pivot_wider(names_from = Supergroup_simplified, values_from = SUM_SEQ, values_fill = 0) %>%
  column_to_rownames(var = "STN_ORDER") %>% 
  as.matrix
# View(surface_matrix)
```
```{r}
clr_surf <- vegan::decostand(surface_matrix, MARGIN = 1, method = "rclr")
colnames(clr_surf) <- colnames(surface_matrix)
rownames(clr_surf) <- rownames(surface_matrix)
# colnames(surface_matrix)
# colnames(clr_surf)
```

CLR plot
```{r, fig.height=7, fig.width=6}
data.frame(clr_surf) %>% 
  rownames_to_column(var = "STN") %>% 
  pivot_longer(cols = -c(STN), names_to = "TAXA", values_to = "CLR") %>% 
  mutate(stn_order = factor(STN, levels = offshore_on_shore_order)) %>% 
  mutate(transect = factor(STN, levels = offshore_on_shore_order, labels = transect_labels)) %>% 
  ggplot(aes(x = stn_order, y = TAXA, fill = CLR)) +
  scale_y_discrete(position = "right") +
  scale_x_discrete(position = "top") +
  geom_tile(color = "black") +
  scale_fill_fermenter(type = "div", palette = 5) +
  coord_fixed() +
  # facet_grid(cols = vars(transect), scale = "free_x", switch = "both") +
  theme_classic() +
  labs(x = "Station", y = "") +
  theme(axis.text.x = element_text(color = "black", size = 10, face = "bold"),
        axis.text.y = element_text(color = "black", size = 10, face = "bold"),
        strip.background = element_blank())

ggsave("figures/clr-surface.svg", width = 6, height = 7, device = "svg", limitsize = FALSE)
```




# RDA analysis

RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it.

### All samples

```{r}
# | eval: false
head(env_info_mod)

categorical <- env_info_mod %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, TRANSECT) %>% distinct()

metadata_for_rda_0 <- env_info_mod %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL) %>% distinct() %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(SAMPLE_NAME, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")
range(standarize_metadata)
# head(categorical)
metadata_normalized <- categorical %>% 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "SAMPLE_NAME")) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME")
head(metadata_normalized)
```

Isolate all samples from ASV data

```{r}
# | eval: false
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
# | eval: false
# head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>% 
  as.matrix
  
clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))

rownames(clr_asv_matrix)
rownames(metadata_normalized)
```

```{r}
# | eval: false
rda_all_gom <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
rda_all_gom
```

23% of the environmental variables explained changes in community composition for all the samples.

```{r}
# | eval: false
adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
# RsquareAdj(rda_all_gom)
adj_rda_out
anova.cca(rda_all_gom, step = 100, by = "term")
```


## Coastal and surface samples

RDA for shallow and transect 1

```{r}
head(env_info_mod)
# station_jgi

surf_categorical <- env_info_mod %>% 
  # filter(TRANSECT == "transect1") %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, PROX_TOSHORE) %>% distinct()

metadata_for_rda_0_surface <- env_info_mod %>% 
  # filter(TRANSECT == "transect1") %>%  
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, DIST_OUTFLOW) %>% distinct() %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(SAMPLE_NAME, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

# rownames(metadata_for_rda_0_surface)
standarize_metadata_surf <- decostand(metadata_for_rda_0_surface, MARGIN = 2, method = "standardize")
# range(standarize_metadata)

metadata_normalized_surf <- surf_categorical %>% 
  right_join(as.data.frame(standarize_metadata_surf) %>% rownames_to_column(var = "SAMPLE_NAME")) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME")

rownames(metadata_normalized_surf)

subset_sample_surf <- as.character(unique(rownames(metadata_normalized_surf)))
subset_sample_surf
```

```{r}
# head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0_surf <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  filter(SAMPLE_NAME %in% subset_sample_surf) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>% 
  as.matrix
  
clr_asv_matrix_surf <- data.frame(compositions::clr(asv_wide_avg_rda_0_surf))

rownames(clr_asv_matrix_surf)
rownames(metadata_normalized_surf)
```

```{r}
rda_surf_gom <- rda(clr_asv_matrix_surf ~ ., data = metadata_normalized_surf, na.action = na.exclude)
rda_surf_gom
```

```{r}
adj_rda_surf <- RsquareAdj(rda_surf_gom)$adj.r.squared
RsquareAdj(rda_surf_gom)
anova.cca(rda_surf_gom, step = 100, by = "term")
```
Model: rda(formula = clr_asv_matrix_surf ~ PROX_TOSHORE + DIST_OUTFLOW + NH4 + NO2 + OXYGEN + PO4 + SALINITY + SIL + TEMP, data = metadata_normalized_surf, na.action = na.exclude)
             Df Variance      F Pr(>F)    
PROX_TOSHORE  1    32.29 1.3575  0.046 *  
DIST_OUTFLOW  1    26.40 1.1099  0.212    
NH4           1    33.81 1.4212  0.022 *  
NO2           1    80.91 3.4015  0.001 ***
OXYGEN        1    58.15 2.4444  0.001 ***
PO4           1   133.70 5.6204  0.001 ***
SALINITY      1    52.74 2.2172  0.002 ** 
SIL           1    39.10 1.6437  0.005 ** 
TEMP          1    36.47 1.5330  0.020 *  
Residual     80  1903.02                  

# Targeted CLR for all samples

```{r}
head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$Supergroup_simplified)
# unique(asv_long_avg_wtax$SUPERGROUP_CLASS)
```


```{r}
asv_wide_supergroup_matrix <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 10) %>%
  # filter(TRANSECT == "transect1") %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH, sep = "_") %>%
  group_by(SAMPLE_NAME, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  ungroup() %>%
  select(SAMPLE_NAME, Supergroup_simplified, SUM_SEQ) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "Supergroup_simplified") %>% 
  as.matrix
```
```{r}
clr_supergroup_all <- vegan::decostand(asv_wide_supergroup_matrix, MARGIN = 2, method = "rclr")
colnames(clr_supergroup_all) <- colnames(asv_wide_supergroup_matrix)
rownames(clr_supergroup_all) <- rownames(asv_wide_supergroup_matrix)
depth_bin
depth_bin_CHARC

```


CLR plot
```{r}
clr_df <- data.frame(clr_supergroup_all) %>% 
  rownames_to_column(var = "Supergroup_simplified") %>% 
  pivot_longer(cols = -c(Supergroup_simplified), names_to = "STN_NI", values_to = "CLR") %>% 
  separate(STN_NI, into = c("STN", "depth"), sep = "_", remove = FALSE) %>% 
  mutate(DEPTH = as.numeric(depth)) %>% 
  left_join((env_info_mod %>% select(DEPTH, DEPTH_BIN) %>% distinct())) %>% 
  # mutate(DEPTH_CATEGORY = factor(DEPTH, levels = depth_bin_CHARC, labels = depth_bin)) %>% 
  mutate(STN_ORDER = factor(STN, levels = offshore_on_shore_order),
         TRANSECT = factor(STN, levels = offshore_on_shore_order, labels = transect_labels))
head(clr_df)
```

```{r, fig.height=20, fig.width=14}
# clr_df %>% 
#   # mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
#   ggplot(aes(x = depth, y = Supergroup_simplified, fill = CLR)) +
#   # facet_grid(rows = vars(DEPTH_BIN, STN), space = "free", scale = "free") +
#   geom_tile(color = "black") +
#   # facet_grid(rows = vars( ), cols = vars(Class), space = "free", scale = "free") +
#   theme(axis.text.x = element_blank()) +
#   scale_fill_fermenter(type = "div", palette = 5) +
#   coord_flip() +
#   theme(axis.ticks = element_blank(),
#         strip.background = element_blank()) +
#   labs(x = "Depth (m)", y = "")

```

```{r, fig.height=5, fig.width=13}
depth_order <- as.character(unique(sort(as.numeric(clr_radio_df$DEPTH_ORDER))))

unique(clr_radio_df$Class)
# View(clr_radio_df)
clr_radio_df %>% 
  # left_join(env_info_mod) %>% 
  # filter(TRANSECT == "transect1") %>% 
  filter(Class != "RAD") %>% 
  filter(Class != "Radiolaria") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "white") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```

```{r, fig.height=2, fig.width=13}
depth_order <- as.character(unique(sort(as.numeric(clr_cil_df$DEPTH_ORDER))))

unique(clr_cil_df$Class)
# View(clr_radio_df)
clr_cil_df %>% 
  # left_join(env_info_mod) %>% 
  filter(STN == "S5") %>%
  filter(Class == "Phyllopharyngea" |
           Class == "Litostomatea") %>% 
  # filter(Class != "RAD") %>% 
  # filter(Class != "Radiolaria") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "black") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```



# Targeted CLR for individual taxonomic groups

```{r}
head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$Supergroup_simplified)
```


Make matrix:
TSAR-Gyrista
TSAR-Radiolaria
```{r}
asv_wide_gyrista_matrix <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 10) %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  filter(TRANSECT == "transect1") %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  group_by(FeatureID, SAMPLE_NAME, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, SUM_SEQ) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix


asv_wide_radio_matrix <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 500 ) %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
  filter(TRANSECT == "transect1") %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  group_by(FeatureID, SAMPLE_NAME, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, SUM_SEQ) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix
# head(asv_wide_gyrista_matrix)
# unique(asv_long_avg_wtax$Supergroup_simplified)
asv_wide_cil_matrix <- asv_long_avg_wtax %>%
  # filter(MEAN_REPS_seq >= 500 ) %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
  filter(TRANSECT == "transect1") %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  group_by(FeatureID, SAMPLE_NAME, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, SUM_SEQ) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix
```
```{r}
clr_gyrista <- vegan::decostand(asv_wide_gyrista_matrix, MARGIN = 2, method = "rclr")
colnames(clr_gyrista) <- colnames(asv_wide_gyrista_matrix)
rownames(clr_gyrista) <- rownames(asv_wide_gyrista_matrix)

clr_radio <- vegan::decostand(asv_wide_radio_matrix, MARGIN = 2, method = "rclr")
colnames(clr_radio) <- colnames(asv_wide_radio_matrix)
rownames(clr_radio) <- rownames(asv_wide_radio_matrix)


clr_cil <- vegan::decostand(asv_wide_cil_matrix, MARGIN = 2, method = "rclr")
colnames(clr_cil) <- colnames(asv_wide_cil_matrix)
rownames(clr_cil) <- rownames(asv_wide_cil_matrix)
# colnames(asv_long_avg_wtax)
```

```{r}
head(env_info_mod)
```

CLR plot
```{r}
clr_radio_df <- data.frame(clr_radio) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  pivot_longer(cols = -c(FeatureID), names_to = "STN_NI", values_to = "CLR") %>% 
  separate(STN_NI, into = c("STN", "DEPTH_ORDER"), remove = FALSE) %>% 
  left_join(tax_key) %>% distinct()
# unique(clr_radio_df$Family)
   
dim(clr_radio);dim(clr_radio_df)

clr_gyrista_df <- data.frame(clr_gyrista) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  pivot_longer(cols = -c(FeatureID), names_to = "STN_NI", values_to = "CLR") %>% 
  separate(STN_NI, into = c("STN", "DEPTH_ORDER"), remove = FALSE) %>% 
  left_join(tax_key) %>% distinct() %>% 
  filter(!is.na(Class))

clr_cil_df <- data.frame(clr_cil) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  pivot_longer(cols = -c(FeatureID), names_to = "STN_NI", values_to = "CLR") %>% 
  separate(STN_NI, into = c("STN", "DEPTH_ORDER"), remove = FALSE) %>% 
  left_join(tax_key) %>% distinct() %>% 
  filter(!is.na(Class))  


dim(clr_gyrista);dim(clr_gyrista_df)
head(clr_gyrista_df)
unique(clr_gyrista_df$Class)
```

```{r, fig.height=5, fig.width=12}
depth_order <- as.character(unique(sort(as.numeric(clr_gyrista_df$DEPTH_ORDER))))
# View(clr_gyrista_df)
unique(clr_gyrista_df$Class)
rm <- c("Hyphochytriomyceta","MOCH","Pinguiophyceae","Pelagophyceae")
clr_gyrista_df %>% 
  filter(!(Class %in% rm)) %>% 
  # left_join(env_info_mod) %>% 
  # filter(TRANSECT == "transect1") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "white") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```

```{r, fig.height=5, fig.width=13}
depth_order <- as.character(unique(sort(as.numeric(clr_radio_df$DEPTH_ORDER))))

unique(clr_radio_df$Class)
# View(clr_radio_df)
clr_radio_df %>% 
  # left_join(env_info_mod) %>% 
  # filter(TRANSECT == "transect1") %>% 
  filter(Class != "RAD") %>% 
  filter(Class != "Radiolaria") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "white") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```

```{r, fig.height=2, fig.width=13}
depth_order <- as.character(unique(sort(as.numeric(clr_cil_df$DEPTH_ORDER))))

unique(clr_cil_df$Class)
# View(clr_radio_df)
clr_cil_df %>% 
  # left_join(env_info_mod) %>% 
  filter(STN == "S5") %>%
  filter(Class == "Phyllopharyngea" |
           Class == "Litostomatea") %>% 
  # filter(Class != "RAD") %>% 
  # filter(Class != "Radiolaria") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "black") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```


# Cell counts
```{r}
head(asv_long_avg_wtax)
prok <- asv_long_avg_wtax %>% 
  select(PROK_COUNT, SAMPLE_ID, stn, nisk, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL) %>% 
  distinct() %>% 
  filter(!(is.na(PROK_COUNT)))
# head(prok)
```
```{r, fig.height=5, fig.width=4}
prok %>% 
  ggplot(aes(x = PROK_COUNT, y = DEPTH, fill = stn, shape = stn, group = stn)) +
  geom_path(linetype = 2, color = "grey70") +
  geom_point(stat = "identity", color = "black", size = 3) +
  scale_x_log10(limits=c(10^3, 10^6), breaks=10^(0:7)) +
  scale_shape_manual(values = c(21, 24, 23)) +
  scale_y_reverse() +
  theme_classic() +
  scale_fill_manual(values = c("#4432a8", "#b15bc7", "#2d807d")) +
  labs(x = "prokaryotic cells per ml (log)", y = "Depth (m)") +
  theme(panel.grid.major = element_line())

```


# Map

```{r}
# install.packages(c("googleway", "ggrepel", 
# "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))

library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
```

## Map of whole Gulf region

```{r}
ggulf <- ggplot(data = world) +
  geom_sf(aes(fill = region_wb)) +
  annotate(geom = "text", x = -90, y = 26, label = "", 
     fontface = "italic", color = "grey22", size = 6) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  theme(legend.position = "none", axis.title.x = element_blank(), 
     axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
     panel.border = element_rect(fill = NA))
ggulf
```

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show:
  # coord_sf(xlim = c(-90, -86.5), ylim = c(27.5, 30), expand = FALSE) + # Region of study
  coord_sf(xlim = c(-110, -70), ylim = c(7, 47), expand = FALSE) + # Larger region
  # Add data points for stations
  # geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10", alpha = 0.9) + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA))
```

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show
  # coord_sf(xlim = c(-90, -86.5), ylim = c(27.5, 30), expand = FALSE) + # Region of study
  coord_sf(xlim = c(-100, -84), ylim = c(21, 33), expand = FALSE) + # Larger region
  # Add data points for stations
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10") + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA))
```

# ODV plots in R

Extract points you want and dimensions. Interpolate, subtract bathymetry if needed and move along from there.

```{r}
# Manually extracted hexidecimal ODV colour palette
ODV_colours <- c("#feb483", "#d31f2a", "#ffc000", "#27ab19", "#0db5e6", "#7139fe", "#d16cfa")
```

## Surface

### Temperature & Long

```{r}
head(asv_long_clean_wtax)
```

```{r}
temp_odv <- asv_long_clean_wtax %>% 
  select(Transect, TRANSECT, Station, DEPTH, stn, Latitude, Longitude, 
         TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT) %>% 
  distinct()
```

Non-interpolated scatterplot

```{r}
# temp_odv %>%
#   ggplot(aes(x = Longitude, y = DEPTH)) +
#   # geom_point(aes(color = TEMP), size = 2) +
#   # geom_point(aes(color = PROK_COUNT), size = 2) +
#   scale_colour_gradientn(colours = rev(ODV_colours)) +
#   labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
#   scale_y_reverse() + scale_x_reverse() +
#   theme_classic()
```

Interpolate step

```{r}
library(MBA)
# ?mba.surf
head(temp_odv)
# Interpolate
temp_mba <- mba.surf(temp_odv[c("Longitude", "DEPTH", "TEMP")], no.X = 300, no.Y = 300, extend = T)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(temp_mba$xyz.est$z) <- list(temp_mba$xyz.est$x, temp_mba$xyz.est$y)
# head(temp_mba)

temp_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c('LONGITUDE', 'DEPTH'), value.name = 'TEMPERATURE') %>% 
  filter(DEPTH > 0) %>%
  mutate(TEMPERATURE = round(TEMPERATURE, 1))

```

```{r}
ggplot(data = temp_interpol, aes(x = LONGITUDE, y = DEPTH)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  scale_fill_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_y_reverse() + scale_x_reverse() +
  theme_classic() +
  coord_cartesian(expand = F)
```

### Surface params, lat, long

```{r}
head(asv_long_clean_wtax)
```

```{r}
surface_odv <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, SURFACE_TEMP, Transect, TRANSECT) %>%
  summarise(ASV_COUNT = n())
```

Non-interpolated scatterplot

```{r}
surface_odv %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
  geom_point(aes(color = SURFACE_TEMP), size = 2) +
  scale_colour_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic()
```

Interpolate step

```{r}
library(MBA)
# ?mba.surf
head(surface_odv)

# Interpolate
temp_mba <- mba.surf(surface_odv[c("Longitude", "Latitude", "SURFACE_TEMP")], 
                     no.X = 300, no.Y = 300, extend = F)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(temp_mba$xyz.est$z) <- list(temp_mba$xyz.est$x, temp_mba$xyz.est$y)
# head(temp_mba)
# 
temp_surface_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c("Longitude", "Latitude"), value.name = 'TEMPERATURE') %>% 
  mutate(TEMPERATURE = round(TEMPERATURE, 1))
head(temp_surface_interpol)
```

```{r}
surface_temperature <- temp_surface_interpol %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  # Add data points
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  coord_cartesian(expand = F) +
  theme(axis.text = element_text(size = 10, colour="black"))
  
```

##### Temperature with map

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = (world %>% filter(subunit == "United States"))) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
  # Set lat long region to show:
  coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
  # Overlay with temperature 
  geom_raster(data = temp_surface_interpol, aes(fill = TEMPERATURE, x = Longitude, y = Latitude)) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
  # Add data points for stations
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)

```

##### Temperature Surface

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = (world %>% filter(subunit == "United States"))) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
  # Set lat long region to show:
  coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
  # Overlay with temperature 
  geom_raster(data = temp_surface_interpol, aes(fill = TEMPERATURE, x = Longitude, y = Latitude)) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
  # Add data points for stations
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)

```

#### Fxn for ther surface params, lat, long

Set up a data frame with all parameters as columns.

```{r}
surface_odv <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SAL = mean(SALINITY),
            SURFACE_OXY = mean(OXYGEN),
            SURFACE_NH4 = mean(NH4),
            SURFACE_NO2 = mean(NO2),
            SURFACE_NO3 = mean(NO3),
            SURFACE_PO4 = mean(PO4),
            SURFACE_PROK = mean(PROK_COUNT),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_NO3, SURFACE_PO4, SURFACE_PROK, Transect, TRANSECT) %>%
  summarise(ASV_COUNT = n())
```

```{r}
# Put PARAM in quotes
interpolate_lat_long_env <- function(df, PARAM, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)], 
                     no.X = 300, no.Y = 300, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Latitude"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Latitude, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1))
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature 
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
    # Add data points for stations
    geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
    geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 1, color= "black") +
    scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1,
        legend.background = element_blank(),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 1, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.83,.75))
}
# interpolate_lat_long_env(surface_odv, "SURFACE_TEMP", "Temperature (°C)")
```

Parameters to use in function: SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4

```{r}
surf_temp <- interpolate_lat_long_env(surface_odv, "SURFACE_TEMP", "Temperature (°C)")
surf_sal <- interpolate_lat_long_env(surface_odv, "SURFACE_SAL", "Salinity")
surf_oxy <- interpolate_lat_long_env(surface_odv, "SURFACE_OXY", "Oxygen")
```

now: SURFACE_NH4, SURFACE_NO2, SURFACE_PO4

```{r}
# expression("S"["in"] ~ "[W" ~ m^-2~"]")
surf_nh4 <- interpolate_lat_long_env(surface_odv, "SURFACE_NH4", expression("NH"["4"]))
surf_no2 <- interpolate_lat_long_env(surface_odv, "SURFACE_NO2", expression("NO"["2"]))
surf_no3 <- interpolate_lat_long_env(surface_odv, "SURFACE_NO2", expression("NO"["3"]))
surf_po4 <- interpolate_lat_long_env(surface_odv, "SURFACE_PO4", expression("PO"["4"]))
```



```{r, fig.height=13, fig.width=18, warning=FALSE}
# surf_temp + surf_sal + surf_oxy + surf_nh4 + surf_no3 + surf_no2 + surf_po4 + surf_all_asv + plot_annotation(tag_levels = "a")
# ?plot_annotation
```
*TO DO* Figure out how to put the prok counts from the surface here too!! 

## Repeat with ASV

Surface lat/long

```{r}
surface_odv <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SAL = mean(SALINITY),
            SURFACE_OXY = mean(OXYGEN),
            SURFACE_NH4 = mean(NH4),
            SURFACE_NO2 = mean(NO2),
            SURFACE_PO4 = mean(PO4),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4, Transect, TRANSECT) %>%
  summarise(ASV_COUNT = n())
```

Interpolate function for ASVs:

```{r}
# Put PARAM in quotes
interpolate_lat_long_asv <- function(df, PARAM, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)], 
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Latitude"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Latitude, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1))
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature 
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = NA) +
    # Add data points for stations
    geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
    geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 1, color= "black") +
    scale_fill_fermenter(type = "seq", palette = "Reds", direction = 1, na.value = NA) +
    # scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1,
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.background = element_blank(),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 1, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.83,.75))
}
```

### Total ASVs

```{r, warning=FALSE}
surf_all_asv<- interpolate_lat_long_asv(surface_odv, "ASV_COUNT", "Total ASVs")
surf_all_asv
```

```{r}
surface_odv_tax_simple <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, Transect, TRANSECT, Supergroup_simplified) %>%
  summarise(ASV_COUNT = n())

unique(surface_odv_tax_simple$Supergroup_simplified)
```

#### Taxa specific ASV profiles

```{r, fig.height=10, fig.width=18}
interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Gyrista")), 
                         "ASV_COUNT", "Gyrista") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Dinoflagellata")), 
                         "ASV_COUNT", "Dinoflagellate") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Ciliophora")), 
                         "ASV_COUNT", "Ciliates") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Radiolaria")), 
                         "ASV_COUNT", "Radiolaria") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "Haptista")), 
                         "ASV_COUNT", "Haptista") + plot_layout()
```

```{r}
dinos <- interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Ciliophora")), 
                         "ASV_COUNT", "Total ASVs")
```

```{r}
surface_odv_tax_class <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, Transect, TRANSECT, SUPERGROUP_CLASS) %>%
  summarise(ASV_COUNT = n())

sort(unique(surface_odv_tax_class$SUPERGROUP_CLASS))
```

```{r}
# "Dinoflagellata-Syndiniales"
# "Dinoflagellata-Dinophyceae"
# "Dinoflagellata-Dinophyta" 
# "Dinoflagellata-Noctilucophyceae"
interpolate_lat_long_asv((surface_odv_tax_class %>% 
                            filter(SUPERGROUP_CLASS == "Dinoflagellata-Dinophyceae")), 
                         "ASV_COUNT", "Total Dinophyceae ASVs") +
  interpolate_lat_long_asv((surface_odv_tax_class %>% 
                            filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), 
                         "ASV_COUNT", "Total Syndiniales ASVs") +
    interpolate_lat_long_asv((surface_odv_tax_class %>% 
                            filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), 
                         "ASV_COUNT", "Total Syndiniales ASVs")
```


## Depth

Depth params, lat, long

```{r}
head(asv_long_clean_wtax)
```

```{r}
depth_odv <- asv_long_clean_wtax %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT, DEPTH) %>% 
  summarise(TEMP = mean(TEMP),
            SAL = mean(SALINITY),
            OXY = mean(OXYGEN),
            NH4 = mean(NH4),
            NO2 = mean(NO2),
            NO3 = mean(NO3),
            PO4 = mean(PO4),
            MEAN_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, TEMP, SAL, 
           OXY, NH4, NO2, NO3, PO4, Transect, TRANSECT, DEPTH) %>%
  summarise(ASV_COUNT = n())
head(depth_odv)
```

Non-interpolated scatterplot

```{r}
depth_odv %>% 
  filter(TRANSECT == "transect1") %>% 
  ggplot(aes(x = Longitude, y = DEPTH)) +
  geom_point(aes(color = TEMP), size = 2) +
  scale_colour_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_x_reverse() + scale_y_reverse() +
  theme_classic()
```

Interpolate step

```{r}
# library(MBA)
# ?mba.surf
head(depth_odv)

one <- depth_odv %>% 
  filter(TRANSECT == "transect1")
# head(one)
# Interpolate
trn1_mba <- mba.surf(one[c("Longitude", "DEPTH", "TEMP")], 
                     no.X = 300, no.Y = 300, extend = F)

# range(one$Longitude)
max_long <- max(one$Longitude)
min_long <- min(one$Longitude)
max_depth <- max(one$DEPTH)
# max_depth <- max(one$DEPTH)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(trn1_mba$xyz.est$z) <- list(trn1_mba$xyz.est$x, trn1_mba$xyz.est$y)
# head(temp_mba)
# 
temp_trn1_interpol <- reshape2::melt(trn1_mba$xyz.est$z, varnames = c("Longitude", "Depth"), value.name = 'TEMPERATURE') %>% 
  mutate(TEMPERATURE = round(TEMPERATURE, 1)) %>% 
  filter(Longitude > 88 & Longitude < 90)

# 88.136 89.460
# head(temp_trn1_interpol)
# range(temp_trn1_interpol$Depth)
```

```{r}
temp_trn1_interpol %>% 
  ggplot(aes(x = Longitude, y = Depth)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  # Add data points
  geom_point(data = one, aes(x = Longitude, y = DEPTH)) +
  geom_text(data = one, aes(x = Longitude, y = -30, label = one$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 0, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Depth (m)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() + scale_y_reverse() +
  theme_classic() +
  # coord_cartesian(expand = F) +
  theme(axis.text = element_text(size = 10, colour="black"))
  
```

### Fxn for depth parameters


```{r}
# df_all <- depth_odv
# transect <- "transect1"
# PARAM <- "TEMP"
# TITLE <- "Temperature"
# Put PARAM in quotes

interpolate_depth_long_env <- function(df_all, transect, PARAM, TITLE){
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "DEPTH", PARAM)], 
                     no.X = 300, no.Y = 300, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = "black") +
    # Add data points
    geom_point(data = df, aes(x = Longitude, y = DEPTH)) +
    geom_text(data = df, 
              aes(x = Longitude, y = -30, label = df$stn), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse() + scale_y_reverse() +
    theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_rect(fill = NA),
          aspect.ratio = 1,
          legend.background = element_blank(),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 0, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.2,0.26))
}

# interpolate_depth_long_env(depth_odv, "transect1","TEMP", "Temperature (°C)")
```

Parameters to use in function: SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4

df_all, transect, PARAM, TITLE

```{r}
# colnames(depth_odv)
# unique(depth_odv$TRANSECT)
```

Transect 1

```{r, fig.height=13, fig.width=18, warning=FALSE}
(interpolate_depth_long_env(depth_odv, "transect1","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, "transect1","SAL", "Salinity")) +
  (interpolate_depth_long_env(depth_odv, "transect1","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(depth_odv, "transect1","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(depth_odv, "transect1","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(depth_odv, "transect1","NO3", expression("NO"["3"]))) +
  (interpolate_depth_long_env(depth_odv, "transect1","PO4", expression("PO"["4"]))) +
    plot_annotation(tag_levels = "a", title = "Transect 1")
```

Transect 2

```{r, fig.height=13, fig.width=18, warning=FALSE}
(interpolate_depth_long_env(depth_odv, "transect2","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, "transect2","SAL", "Salinity")) +
  (interpolate_depth_long_env(depth_odv, "transect2","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(depth_odv, "transect2","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(depth_odv, "transect2","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(depth_odv, "transect2","NO3", expression("NO"["3"]))) +
  (interpolate_depth_long_env(depth_odv, "transect2","PO4", expression("PO"["4"]))) +
    plot_annotation(tag_levels = "a", title = "Transect 2")
```
```{r, fig.height=13, fig.width=18, warning=FALSE}
(interpolate_depth_long_env(depth_odv, "transect3","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, "transect3","SAL", "Salinity")) +
  (interpolate_depth_long_env(depth_odv, "transect3","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(depth_odv, "transect3","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(depth_odv, "transect3","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(depth_odv, "transect3","NO3", expression("NO"["3"]))) +
  (interpolate_depth_long_env(depth_odv, "transect3","PO4", expression("PO"["4"]))) +
    plot_annotation(tag_levels = "a", title = "Transect 3")
```
### Repeat for upper water column

```{r}
# head(depth_odv)
# range(depth_odv$DEPTH)
shallow <- depth_odv %>% 
  filter(DEPTH <= 300)
```

Transect 1

```{r, fig.height=12, fig.width=17}
(interpolate_depth_long_env(shallow, "transect1","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(shallow, "transect1","SAL", "Salinity")) +
  (interpolate_depth_long_env(shallow, "transect1","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(shallow, "transect1","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(shallow, "transect1","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(shallow, "transect1","PO4", expression("PO"["4"]))) +
    plot_layout(guides = "collect")

```

Transect 2

```{r, fig.height=12, fig.width=17}
(interpolate_depth_long_env(shallow, "transect2","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(shallow, "transect2","SAL", "Salinity")) +
  (interpolate_depth_long_env(shallow, "transect2","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(shallow, "transect2","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(shallow, "transect2","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(shallow, "transect2","PO4", expression("PO"["4"]))) +
    plot_layout(guides = "collect")

```

## Repeat with ASVs

```{r}
# Put PARAM in quotes
interpolate_depth_long_asv <- function(df_all, transect, PARAM, TITLE){
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "DEPTH", PARAM)], 
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = NA) +
    # Add data points
    geom_point(data = df, aes(x = Longitude, y = DEPTH)) +
    geom_text(data = df, 
              aes(x = Longitude, y = -30, label = df$stn), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_fermenter(type = "seq", palette = "Reds", direction = 1, na.value = NA) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse() + scale_y_reverse() +
    theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_rect(fill = NA),
          aspect.ratio = 1,
          legend.background = element_blank(),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 0, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.2,0.26))
}
```

Transect 1 with all ASVs

```{r}
interpolate_depth_long_asv(depth_odv, "transect1","ASV_COUNT", "Total ASVs")
```

Transects

```{r, fig.width=16, fig.height=7, warning=FALSE}
interpolate_depth_long_asv(depth_odv, "transect1","ASV_COUNT", "Total ASVs") +
interpolate_depth_long_asv(depth_odv, "transect2","ASV_COUNT", "Total ASVs") + 
  interpolate_depth_long_asv(depth_odv, "transect3","ASV_COUNT", "Total ASVs") +
  plot_layout(nrow = 1)

```

```{r}
depth_odv_tax_simple <- asv_long_clean_wtax %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT, DEPTH) %>% 
  summarise(MEAN_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, Transect, TRANSECT, DEPTH, Supergroup_simplified) %>%
  summarise(ASV_COUNT = n())

unique(depth_odv_tax_simple$Supergroup_simplified)
```

interpolate_depth_long_asv

by taxa, use `depth_odv_tax_simple`

```{r, fig.height=10, fig.width=18}
interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Gyrista")), "transect1", 
                         "ASV_COUNT", "Gyrista") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Dinoflagellata")), "transect1",
                         "ASV_COUNT", "Dinoflagellate") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Ciliophora")), "transect1",
                         "ASV_COUNT", "Ciliates") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Radiolaria")), "transect1",
                         "ASV_COUNT", "Radiolaria") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "Haptista")), "transect1",
                         "ASV_COUNT", "Haptista") + plot_layout()
```

# Identify ASVs with highest number of sequences throughout data

```{r}
dominant_taxa <- asv_long_avg_wtax %>% 
  group_by(Taxon, Supergroup_simplified, DEPTH_ORDER) %>% 
  summarise(SUM_SEQS = sum(MEAN_REPS_seq)) %>% 
  ungroup() %>% 
  group_by(Supergroup_simplified) %>% 
  slice_max(order_by = SUM_SEQS, n = 15)

```


# Linear regression
```{r}
head(asv_long_avg_wtax)
# PARAMS: DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL

```
```{r}
library(broom)
regression_gom_tmp <- asv_long_avg_wtax %>%
  select(Supergroup_simplified, MEAN_REPS_seq, Station, Niskin, DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL) %>% 
  group_by(Station, Niskin, Supergroup_simplified, DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL) %>% 
summarise(seq_sum = sum(MEAN_REPS_seq),
          asv_rich = n()) %>% 
  # filter(ASV_rich > 1 &)
  pivot_longer(cols = c(DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL), names_to = "PARAMS", values_to = "VALUES") %>% 
  ungroup() 
dim(regression_gom_tmp)
head(regression_gom_tmp)
# unique(regression_gom_tmp$Class)
```
```{r}
part1<- regression_gom_tmp %>% 
# For taxonomic groups, group linear regression by ASV Richness
 group_by(Supergroup_simplified, PARAMS) %>% 
  nest(-Supergroup_simplified, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
    tidied = map(lm_fit, tidy)) %>% unnest(tidied) %>%
  select(Supergroup_simplified, PARAMS, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
    select(everything(), SLOPE = VALUES) %>% data.frame
```
```{r}
part2 <- regression_gom_tmp %>% 
# For taxonomic groups, group linear regression by ASV Richness
 group_by(Supergroup_simplified, PARAMS) %>% 
  nest(-Supergroup_simplified, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
    glanced = map(lm_fit, glance)) %>% unnest(glanced) %>%
  select(Supergroup_simplified, PARAMS, r.squared, adj.r.squared) %>%
  right_join(part1) %>% 
  right_join(regression_gom_tmp)

```
```{r}
head(part2)
unique(part2$Supergroup_simplified)
```

```{r, fig.width=10, fig.height=5}
part2 %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "Ciliate")
```
```{r, fig.width=10, fig.height=5}
part2 %>% 
  filter(Supergroup_simplified == "Haptista") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "Haptista")
```
```{r}
#TSAR-Dinoflagellata
part2 %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "TSAR-Dinoflagellata")
```

## Linear regression function
`function(df, TITLE)`
```{r}
library(broom)
lm_gom <- function(df, TITLE) {
  df_tmp <- df %>%
  select(TAXA, MEAN_REPS_seq, Station, Niskin, DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL) %>% 
  group_by(TAXA, Station, Niskin, DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL) %>% 
  summarise(seq_sum = sum(MEAN_REPS_seq),
            asv_rich = n()) %>% 
  pivot_longer(cols = c(DIST_OUTFLOW, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL), names_to = "PARAMS", values_to = "VALUES") %>% 
  ungroup() 
#
df_tmp_tidy <- df_tmp %>% 
  # For taxonomic groups, group linear regression by ASV Richness
  group_by(TAXA, PARAMS) %>% 
  nest(-TAXA, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
         tidied = map(lm_fit, tidy)) %>% unnest(tidied) %>%
  select(TAXA, PARAMS, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
  select(everything(), SLOPE = VALUES) %>% data.frame

df_lm <- df_tmp %>% 
  # For taxonomic groups, group linear regression by ASV Richness
  group_by(TAXA, PARAMS) %>% 
  nest(-TAXA, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
         glanced = map(lm_fit, glance)) %>% unnest(glanced) %>%
  select(TAXA, PARAMS, r.squared, adj.r.squared) %>%
  right_join(df_tmp_tidy) %>% 
  right_join(df_tmp)

df_lm %>% 
  ggplot(aes(x = VALUES, y = asv_rich)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
                                3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
  theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
        strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black",                                                     size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = TITLE)
}
```

```{r}
head(asv_long_avg_wtax)
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  add_column(TAXA = "All ASVs")), "All ASVs")

lm_gom((asv_long_avg_wtax %>% 
          filter(DEPTH < 10) %>% 
  add_column(TAXA = "Surface ASVs")), "Surface ASVs")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  filter(Class == "Syndiniales") %>% 
  add_column(TAXA = "Syndiniales")), "Syndiniales")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 50) %>% 
  filter(Class == "Dinophyceae") %>% 
  add_column(TAXA = "Dinophyceae")), "Dinophyceae")
```

```{r}
sort(unique(asv_long_avg_wtax$Class))
```


```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Subdivision == "Gyrista") %>% 
  add_column(TAXA = "Gyrista")), "Gyrista")
#Bacillariophyceae
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Bacillariophyceae") %>% 
  add_column(TAXA = "Bacillariophyceae")), "Bacillariophyceae")


lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Syndiniales") %>% 
  add_column(TAXA = "Syndiniales")), "Syndiniales")

lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Dinophyceae") %>% 
  add_column(TAXA = "Dinophyceae")), "Dinophyceae")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  filter(Subdivision == "Ciliophora") %>% 
  add_column(TAXA = "Ciliates")), "Ciliates")
```

```{r}
unique(asv_long_avg_wtax$Subdivision)
```


```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH > 400) %>%
  filter(Subdivision == "Cercozoa") %>% 
  add_column(TAXA = "Cercozoa")), "Cercozoa")

lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH > 400) %>%
  filter(Subdivision == "Radiolaria") %>% 
  add_column(TAXA = "Radiolaria")), "Radiolaria")
```

# Shannon Diversity

Surface stations only

```{r}
asv_wide_surf_matrix <- asv_long_clean_wtax %>%
  filter(DEPTH < 5) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  # First average ASV sequence count across replicates
  group_by(FeatureID, SAMPLE_NAME) %>%
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>%
  as.matrix()

asv_wide_all_matrix <- asv_long_clean_wtax %>%
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  # First average ASV sequence count across replicates
  group_by(FeatureID, SAMPLE_NAME) %>%
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>%
  as.matrix()
```

Calculate Shannon diversity

```{r, fig.width=5, fig.height=3}
# ?diversity()
surface_shannon <- diversity(asv_wide_surf_matrix, index = "shannon", MARGIN = 2)
data.frame(surface_shannon) %>% 
  rownames_to_column(var = "stn_niskin") %>% 
  separate(stn_niskin, into = c("STN", "NISKIN"), remove = FALSE) %>% 
  mutate(STN_ORDER = factor(STN, levels = offshore_on_shore_order),
         TRANSECT = factor(STN, levels = offshore_on_shore_order, labels = transect_labels)) %>% 
  ggplot(aes(x = STN_ORDER, y = surface_shannon)) +
  geom_point(shape = 21, fill = "grey30", color = "black", size = 3) +
  facet_grid(cols = vars(TRANSECT), scales = "free", space = "free") +
  theme_classic() + 
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    panel.background = element_rect(color = "grey80"),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
    labs(x = "Station", y = "Shannon Diversity")
  
```


# Indicator Species

```{r}
# install.packages("labdsv")
library(labdsv)
```


https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html

```{r}
library(indicspecies)
```

Format input table

```{r}
head(asv_long_avg_wtax)
```

```{r}
# save(asv_long_avg_wtax, file = "asv_gom_input_indicator.RData")
head(asv_long_avg_wtax)
# hist(asv_long_avg_wtax$MEAN_REPS_seq)
```

Below is taking much too long on the HPRC and not working.

```{r}
head(asv_long_avg_wtax)
```


```{r}
indic_input <- asv_long_avg_wtax %>%
  # filter(DIVISION == "Dinoflagellata") %>%
  # filter(DEPTH < 5) %>%
  filter(MEAN_REPS_seq > 5) %>%
  # filter(TRANSECT == "transect1") %>%
  # filter(stn == "S1" | stn == "S2") %>% #TESTING
  select(SAMPLE_ID, DEPTH, TEMP, SALINITY, MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fill = 0)
  # select(-TEMP, -SALINITY)
head(indic_input)
```

## IndVal

two species are not strongly correlated and their frequency in the data set is larger than the frequency of sites belonging to the site group, the joint occurrence of the two species has higher positive predictive value for the site group than the two species taken independently

‘d’ of species as the product of the relative frequency and relative average abundance in clusters.

```{r}
# head(indic_input[,1])
# ?cut()
# ?indval()
# params <- indic_input %>% 
#   mutate(depth = as.numeric(DEPTH),
#   temp = as.numeric(TEMP),
#   sal = as.numeric(SALINITY)) %>%
#   select(SAMPLE_ID, depth, temp) %>% 
#   column_to_rownames(var = "SAMPLE_ID") %>% 
#   as.matrix

clust <- cut(indic_input$TEMP, 5, labels = FALSE)

abun <- indic_input %>% 
  select(-DEPTH, -TEMP, -SALINITY) %>% 
  column_to_rownames(var = "SAMPLE_ID")

dino_surf <- indval(x = abun, clustering = clust, comm = abun)
# View(dino_surf$indval)
summary(dino_surf)

# head(dino_surf$indval)
tmp_out <- data.frame(dino_surf$indval) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  left_join(tax_key) %>% 
  arrange(X1)

# range(tmp_out$X1)
# View(tmp_out)

```



Try paring down the data more.

```{r}
indic_input <- asv_long_avg_wtax %>%
  # filter(DEPTH < 800) %>% # Surface only
  filter(MEAN_REPS_seq >= 100) %>% # Subset of tASVs
  filter(TRANSECT == "transect1") %>% # Only the 1 transect
  # filter(stn == "S1" | stn == "S2") %>% #TESTING
  select(SAMPLE_ID, DEPTH, TEMP, SALINITY, MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fill = 0)
head(indic_input)
dim(indic_input)
# ?multipatt
```

```{r}
asv_abun <- indic_input %>%
  select(-SAMPLE_ID, -DEPTH, -TEMP, -SALINITY)

# inc_sampleid <- multipatt(asv_abun, indic_input$DEPTH, func = "r.g", control = how(nperm = 10))
# # inc_sampleid <- multipatt(asv_abun, indic_input$DEPTH_ORDER, func = "IndVal.g", control = how(nperm = 10))
# 
# summary(inc_sampleid)
```


Re-try with normalization step
```{r}
# | eval: false
head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>%
  rownames_to_column(var = "SAMPLE_NAME") %>%
  arrange(SAMPLE_NAME) %>%
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))
# 
rownames(clr_asv_matrix)
rownames(metadata_normalized)
head(clr_asv_matrix)
```

```{r}

# clr_asv_matrix
# data = metadata_normalized, na.action = na.exclude)
```

```{r}
asv_abun <- indic_input %>%
  select(-SAMPLE_ID, -DEPTH, -TEMP, -SALINITY)
# head(asv_abun)

# head(clr_asv_matrix)
# head(metadata_normalized)
# tmp_100_asv <- clr_asv_matrix %>% select(1:50)
# tmp_100_meta <- metadata_normalized

inc_sampleid <- multipatt(tmp_100_asv, tmp_100_meta$DEPTH, func = "r.g", control = how(nperm = 10))
# # inc_sampleid <- multipatt(asv_abun, indic_input$DEPTH_ORDER, func = "IndVal.g", control = how(nperm = 10))
# # inc_sampleid <- multipatt(asv_abun, indic_input$TEMP, func = "r.g", control = how(nperm = 100))
# 
# summary(inc_sampleid)
```

```{r}
# library(stats)
# output_stats <- as.data.frame(inc_sampleid$sign, keep.rownames = TRUE)
# # Add adj p-values
# output_stats[ ,p.value.bh = p.adjust(p.value, method="BH")]
# # range(output_stats$p.value)
# ??p.adjust()
```

`r.g` - indicator species test called a "point biserial correlation coefficient". It measures the correlation between two binary vectors.

```{r}
#extract table of stats
# indisp.sign<-as.data.table(indisp$sign, keep.rownames=TRUE)
# #add adjusted p-value
# indisp.sign[ ,p.value.bh:=p.adjust(p.value, method="BH")]
# #now can select only the indicators with adjusted significant p-values
# indisp.sign[p.value.bh<=0.05, ]
```

# Linear Discriminent Analysis Effect Size (LDA-LEfSe)

-   https://www.bioconductor.org/packages/release/bioc/vignettes/lefser/inst/doc/lefser.html

-   chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://www.bioconductor.org/packages/release/bioc/manuals/lefser/man/lefser.pdf

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("lefser")
library(lefser)

```

# CRT

```{r}
# install.packages("tree")
library(rpart)
library(tree)
```

```{r}
data("iris")
head(iris)
iris_df <- iris
colnames(iris_df) <- c("sepall", "sepalw", "petall", "petalw", "class")
```


CRT, on the other side, focuses on creating predictive models for classifying and prioritizing data as “key indicator variables”. CRT employs a decision tree approach, partitioning the dataset based on informative environmental variables to create a hierarchical structure of decision rules (Alkhasawneh et al., 2014). We used CRT to identify specific taxonomic groups as “key indicator variables” that show ability to indicate different DLs.

```{r}
xtabs( ~ class, data = iris_df)
```

```{r}
# execute tree classification
iris.tree <- tree::tree(class ~ ., data = iris_df)
# ?tree()
head(iris_df)
```

Above step fits a classification or regression tree with data. 

```{r}
summary(iris.tree)
# Only 3 of 4 variables were used for the tree.
# misclassification error rate? 
```

```{r}
plot(iris.tree); text(iris.tree, all = T)
```
For this classification tree, petall has a value < 2.45, therefore suggesting it has a more important classification. 
Sepall <5.15 indicates that is has a less important criterion

Next we "prune" the tree to reduce complexity. For this, we use a function to identify the number of significant notes. The cross validation will find the number of misclassifications.

```{r}
tree::cv.tree(iris.tree)
```

We prune the small branches. Therefore, we want to remove the "Branches" from the tree with the fewest data points. Therefore, we look at the output values of k and see when the values start to really increase. In this example, it is 3. A tree size of 3 is therefore suitable. THe _K_ is considered the "cost-complexity parameter" (which is the "Best" number of nodes/branches to include. )

Pruning the tree:
```{r}
iris.tree.prune <- prune.misclass(iris.tree, best = 3)
summary(iris.tree.prune)
```
```{r}
plot(iris.tree.prune);text(iris.tree.prune, all = T)
```
The tree is much less complex. 

### CRT all samples

```{r}
# head(asv_long_avg_wtax)
```


```{r}
# | eval: false
# input_CRT <- asv_long_avg_wtax %>% 
#   select(FeatureID, DEPTH, TEMP, SALINITY, OXYGEN, NH4, MEAN_REPS_seq) %>% distinct()
# 
# toy_input_CRT <- input_CRT %>% 
#   sample_n(1000)
# 
# test_tree <- tree(FeatureID ~ , toy_input_CRT)
# summary(test_tree)

# test_tree$frame
# plot(test_tree); text(test_tree, all = T)
```

```{r}
# install.packages("devtools")
# require("devtools")
# devtools::install_github(repo = 'zellerlab/siamcat')
```


# Session Information

```{r}
sessionInfo()
```
