---
title: "2023 Gulf data analysis"
editor: source
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 4
---

# Context & introduction

*Core scientific questions:*

(1) How does the single-celled eukaryotic (microeukaryote or protistan) community composition and diversity change from the Mississippi River plume to offshore?

(2) What environmental factors play a role in shaping the microeukaryotic community dynamics at shallow versus deep depths in the Gulf of Mexico?

(3) How do putative functional roles among microeukaryotes vary across critical environmental gradients?

(4) Are there specific taxonomic groups, classes, or species that may be considered indicative of the Northern Gulf ecosystem? Especially along specific temperature, salinity, oxygen, and nutrient trends?


# Set up R environment

Set up R & import starting data. Installation prompts.

```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("decontam")
# install.packages("compositions")
# install.packages("ggupset")
# install.packages("gt")
# install.packages("ggrepel")
# install.packages("ploty")
# install.packages("viridis")
# install.packages("vegan")
# install.packages("anomalize")
# install.packages("cluster")
# install.packages("factoextra")
# install.packages("gt")

# if(!requireNamespace("BiocManager")){
#   install.packages("BiocManager")
# }
# BiocManager::install("phyloseq")
# install.packages("indicspecies")
```


```{r, warning=FALSE, message=FALSE}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt); library(ggrepel)
library(viridis); library(vegan); 
library(treemapify);library(gt)
#
library(anomalize)
library(cluster)
library(factoextra)
library(phyloseq)
library(indicspecies)
```

## Import required data

Below is direct from `GoM-2023-amplicons.qmd`

```{r}
load("input-data/gulf-2023-sequence-analysis_02062026.RData", verbose = TRUE)
```

## Set ordering & factoring

```{r}
offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5",
                             "S9", "S8", "S7", 
                             "S11", "S12", "S14", "S15")
stn_order_shore_offshore <- c(1, 2, 3, 4, 5, 
                              9, 8, 7, 6, 
                              10, 11, 12, 13, 14, 15, 16, 17)

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(metadata_seqsamples$Depth))))

depth_order_round <- as.character(sort(round(unique(metadata_seqsamples$Depth), digits = 0)))

depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

watercol_features <- c("other", "DCM", "Salinity maximum", "Oxygen minimum", "Salinity maximum, DCM")
watercol_features_label <- c("Depth with no feature", "DCM", "Salinity maximum", "Oxygen minimum", "Salinity maximum, DCM")
watercol_shape <- c(21, 24, 25, 22, 23)
names(watercol_shape) <- watercol_features_label
# watercol_shape

water_mass_order <- c("AAIW", "CSWr", "GCW", "NADW", "NASUW", "TACW", "TACWn", "Uncharacterized Upper", "Uncharacterized Deep") 
water_mass_color <- c("darkblue", "orange", "forestgreen", "lightgreen", "lightblue", "darkred", "pink", "grey80", "grey20")
names(water_mass_color) <- water_mass_order
```

### Study stats

```{r}
(unique(asv_long_wtax_clean$Station)) # 12 stations
length(unique(asv_long_wtax_clean$Station)) # 12 stations
length(unique(asv_long_wtax_clean$STN_NISKIN)) # 90 unique samples
```

# K-means clustering by sequence data

Normalize raw sequence data around samples. The goal is to compare this to water masses and group sample types.

```{r}
# install.packages("anomalize")
# install.packages("cluster")
# install.packages("factoextra")
# library(anomalize)
# library(cluster)
# library(factoextra)
set.seed(3500)
```

Normalize across variables. Will not include categorical values. We are grouping STN_NISKIN.

```{r}
head(asv_long_avg_wtax_clean)
# unique(asv_long_avg_wtax_clean$Supergroup_simplified)
```

```{r}
seq_only_mat <- asv_long_avg_wtax_clean %>% 
  select(STN_NISKIN, FeatureID, MEAN_REPS_seq) %>% distinct() %>%
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fill = 0) %>% distinct() %>% 
  drop_na() %>%
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix
```

Normalize matrix of numeric metadata.

```{r}
# Using clr approach, margin is set to 2 to perform this across each ASV ID, so the relationships between samples can be determined. 
standarize_seq <- decostand(seq_only_mat, MARGIN = 2, method = "rclr")
# standarize_seq <- decostand(seq_only_mat, MARGIN = 2, method = "range")
# ?decostand
# Output from here should be stn_niskin
rownames(standarize_seq) <- rownames(seq_only_mat)
colnames(standarize_seq) <- colnames(seq_only_mat)
```

First group to see how many clusters we need.

```{r}
kclusts <-
  tibble(k = 2:16) %>%
  mutate(
    kclust = map(k, ~ kmeans(standarize_seq, .x)),
    glanced = map(kclust, broom::glance))

# Unnest
output_kclusts <- kclusts %>%
  unnest(cols = c(glanced))
# ?kmeans
# plot
output_kclusts %>%
  ggplot(aes(x = k, y = tot.withinss)) +
    geom_line(alpha = 0.5, linewidth = 1.2, color = "darkgreen") +
    geom_point(size = 2, color = "darkgreen") +
  theme_linedraw()
```

> Above plot shows hypothetical numbers of k clusters (x-axis) with the variance within the cluster (y-axis). We want to isolate the number of clusters where we cover as much variance as possible - BUT if we add more clusters, not much will be helpful. This means "the bend" here. In this case, We are going to go with 9 clusters.

Use chosen number of clusters (n=9) to compute kmeans.

```{r}
kmeans_seq_5 <- kmeans(standarize_seq, centers = 9)
summary(kmeans_seq_5)

# Report some findings using "broom"
broom::tidy(kmeans_seq_5)
broom::glance(kmeans_seq_5)
```

Combine kmeans clustering with original data.

```{r}
# Combine with data
output_kmeans_seq5 <- broom::augment(kmeans_seq_5, standarize_seq)
head(data.frame(output_kmeans_seq5)) # Prints cluster membership at end of this dataframe.

kmeans_membership_5 <- data.frame(output_kmeans_seq5) %>% 
  select(STN_NISKIN = `.rownames`, CLUSTER = `.cluster`) %>% distinct()

# Samples in each cluster
summary(kmeans_membership_5$CLUSTER)
# save(kmeans_membership_5, file = "input-data/kmeans-membership-seqs-9.RData")
```

#### Table with features

```{r}
load("input-data/kmeans-membership-seqs-9.RData", verbose = TRUE)
# Import table from Portela et al.
water_mass <- read.csv("input-data/water-mass-data.csv")
```

```{r}
head(metadata_seqsamples)
ts_only <- metadata_seqsamples %>% 
  select(STN_NISKIN, ENV_VARIABLE, value) %>% 
  filter(ENV_VARIABLE == "Temperature" | ENV_VARIABLE == "Salinity") %>% 
  distinct() %>% 
  pivot_wider(names_from = ENV_VARIABLE, values_from = value)
# head(ts_only)  
head(water_mass)
```

```{r}
# add id for tracking
# p2 <- penguins |> mutate(id = row_number(), .before = 1)

tmp <- ts_only %>%
  mutate(ID = row_number(), .before = 1) %>%
  left_join(water_mass, 
            join_by(Temperature >= TEMP_MIN,
                    Temperature < TEMP_MAX,
                    Salinity >= SAL_MIN,
                    Salinity < SAL_MAX)) %>% 
  reframe(Watermass = paste(Watermass, collapse = ", "), .by = ID)

ts_mass <- ts_only %>%
  mutate(ID = row_number(), .before = 1) %>% 
  left_join(tmp)

# unique(ts_mass$Watermass)
```

```{r}
table_wfeature <- metadata_seqsamples %>% 
  left_join(metadata_stn_info %>% 
  select(Station, OilRig_Count, MLD, SAL_MAX, O2_MIN, DCM, STN_CATEGORY_3, STN_CATEGORY_4)) %>% 
  left_join(kmeans_membership_5, by = join_by(STN_NISKIN)) %>% 
  # Water mass
  left_join((ts_mass %>% select(STN_NISKIN, Watermass)), by = join_by(STN_NISKIN)) %>% 
  # Refine water mass
  mutate(WATER_MASS = case_when(
    (Watermass == "TACW, TACWn") ~ "TACWn",
    # (Watermass == "NA") ~ "Uncharacterized",
    (Watermass == "NA" & Depth > 150) ~ "Uncharacterized Deep",
    (Watermass == "NA" & Depth <= 150) ~ "Uncharacterized Upper",
    TRUE ~ Watermass)) %>% 
  mutate(SAL_THRESHOLD = case_when(
    Depth > SAL_MAX ~ "Below salinity maximum", #At and below the salinity maximum
    Depth <= SAL_MAX ~ "Above salinity maximum",
  )) %>% 
  mutate(MLD_THRESHOLD = case_when(
    Depth > MLD ~ "Below MLD", #At and below the salinity maximum
    Depth <= MLD ~ "Above MLD",
  )) %>% 
  mutate(DCM_THRESHOLD = case_when(
    Depth > DCM ~ "Below DCM",
    Depth <= DCM ~ "Above DCM", #At and above the DCM
  )) %>% 
  mutate(O2_THRESHOLD = case_when(
    Depth > O2_MIN ~ "Below oxygen minimum",
    Depth <= O2_MIN ~ "Above oxygen minimum", #At and above the O2 minimum
  ))

table_wstn_wdepth <- table_wfeature %>%
  select(-ENV_VARIABLE, -value, -Watermass) %>% distinct()
table(table_wstn_wdepth$WATER_MASS)
```

```{r}
head(table_wstn_wdepth)
```


### Create Table S1

```{r}
# Import sequence and ASV stats for a supplementary table
tmp_seq_stats <- read.csv(file = "output-tables/seq-asv-stats.csv") %>% 
  unite(STN_NISKIN, STN, NISKIN, sep = "_") %>% 
  select(STN_NISKIN, Sequence_files, Total_ASVs, Total_sequences)
# head(tmp_seq_stats)
```

```{r}
table_supp <- table_wfeature %>%
  select(-Pressure, -datetime_cst, -starts_with("SUN"), -HR_OF_DAY, -stn) %>%
  pivot_wider(names_from = ENV_VARIABLE, values_from = value) %>% 
  left_join(tmp_seq_stats)

glimpse(table_supp)

```

Repeat and format for supplementary table

```{r}
table_supp1 <- table_wfeature %>%
  select(-Pressure, -datetime_cst, -starts_with("SUN"), -HR_OF_DAY, -stn) %>%
  pivot_wider(names_from = ENV_VARIABLE, values_from = value) %>% 
  left_join(tmp_seq_stats) %>% 
  select(Station, Niskin, STN_NISKIN, Depth, DEPTH_BIN, Latitude, Longitude, 
         Date, Time_UTC, DAY_NIGHT, TRANSECT, DIST_OUTFLOW, STN_CATEGORY_4, CLUSTER,
         OilRig_Count, MLD, SAL_MAX, O2_MIN, DCM, Temperature, Salinity, 
         DIC, NH4, NO2, NO3, PO4, SIL, Oxygen_CTD, TA, pH, WATER_MASS, Prok_count, Total_ASVs, Total_sequences)

write.csv(table_supp1, file = "output-tables/supp-table-allsampleinfo.csv")
```

```{r}
head(table_supp1)
unique(table_supp1$Sequence_files)
```

```{r}
# save(table_supp1, table_wstn_wdepth, table_supp, file = "output-tables/table_supp.RData")
```


# Sample comparison & taxonomic classification

## PCA analysis

Create wide format matrix with sequence counts averaged across replicates (by ASV) Make wide format and as a matrix.

```{r}
asv_wide_avg_matrix <- asv_long_avg_wtax_clean %>% 
  filter(STN_NISKIN != "S0_N0") %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
# dim(asv_wide_avg_matrix)
# Working with 10563 ASVs
# dim(asv_long_avg_wtax_clean)
```

```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_matrix))))
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

#### Plot PCA

##### PCA with cluster

```{r, fig.height=5, fig.width=6}
# CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9")
# CLUSTER_NUM <- c(1, 2, 3, 4, 5)
CLUSTER_COLOR <- c("#bd2a2a","#c85e82", "#efc85a", "#df9147","#47c475","#8b94b5","#6e9b4a", "#606aa9", "#524740")
CLUSTER_NUM <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
names(CLUSTER_NUM) <- CLUSTER_COLOR

PCA_PLOT_kmeans <- data.frame(pca_clr$x, STN_NISKIN = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin)) %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  mutate(SURFACE = case_when(
    DEPTH_BIN == "< 5m" ~ "Surface",
    TRUE ~ "sub-surface"
  )) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = SURFACE,
             fill = cluster_order)) + # Try this for a supplementary figure.
             # fill = depth_order)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  # geom_jitter(stroke = 0.6, size = 2.5, width = 0.8, shape = 21) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(21, 23)) +
  scale_fill_manual(values = CLUSTER_COLOR) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
PCA_PLOT_kmeans
```

##### PCA with depth

```{r, fig.height=5, fig.width=6}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

PCA_PLOT_depth <- data.frame(pca_clr$x, STN_NISKIN = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin)) %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  mutate(SURFACE = case_when(
    DEPTH_BIN == "< 5m" ~ "Surface",
    TRUE ~ "sub-surface"
  )) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = SURFACE,
             fill = depth_order)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  # geom_jitter(stroke = 0.6, size = 2.5, width = 0.8, shape = 21) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(21, 23)) +
  scale_fill_manual(values = blues) +
  coord_fixed() +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        # plot.margin = margin(2, 1, 2, 1, "cm")) +
  plot.margin = margin(b = 0, unit = "pt")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
PCA_PLOT_depth
```

### PCA above DCM

```{r}
# head(table1_wfeature)
above_DCM_df <- table_wstn_wdepth %>% 
  filter(DCM_THRESHOLD == "Above DCM")
above_dcm_list <- as.character(unique(above_DCM_df$STN_NISKIN))
```

Make wide format and as a matrix.

```{r}
aboveDMC_wide_matrix <- asv_long_avg_wtax_clean %>% 
  filter(STN_NISKIN %in% above_dcm_list) %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
pca_clr_abovedcm <- prcomp(data.frame(compositions::clr(t(aboveDMC_wide_matrix))))
```

```{r}
variance_clr_abovedcm <- (pca_clr_abovedcm$sdev^2)/sum(pca_clr_abovedcm$sdev^2)

barplot(variance_clr_abovedcm, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

```{r, fig.height=5, fig.width=6}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

pca_abovedcm <- data.frame(pca_clr_abovedcm$x, STN_NISKIN = rownames(pca_clr_abovedcm$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin)) %>% 
  # mutate(FEATURE = str_replace_na(Features_list, replacement = "other")) %>% 
  # mutate(FEATURE_ORDER = factor(FEATURE, levels = watercol_features, labels = watercol_features_label)) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = ,
             fill = depth_order)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8, shape = 21, color = "black") +
  ylab(paste0('PC2 ',round(variance_clr_abovedcm[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr_abovedcm[1]*100,2),'%')) +
  # scale_shape_manual(values = watercol_shape) +
  scale_fill_manual(values = blues) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
pca_abovedcm
# ggsave("figures/pca-plot.svg", width = 6, height = 5, device = "svg", limitsize = FALSE)
```

# Diversity indices

Use wide dataframe from above.

```{r}
shannon_all <- vegan::diversity(asv_wide_avg_matrix, index = "shannon", MARGIN = 2)
```

```{r}
# ?diversity
invsimp_all <- vegan::diversity(asv_wide_avg_matrix, index = "invsimpson", MARGIN = 2)
# head(data.frame(invsimp_all))
```

## Plot diversity

### Shannon

By depth

```{r}
depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")

SHANNON <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = depth_order, 
             x = shannon_all, 
             fill = depth_order,
             shape = MLD_THRESHOLD)) +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 10),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = "Shannon Diversity", y = "") +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
```
```{r, fig.width=6, fig.height=4}
SHANNON_cluster <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = (depth_bin))) %>% 
  mutate(STN_ORDER = factor(as.character(Station), levels = stn_order_shore_offshore)) %>% 
  mutate(transect_order = factor(TRANSECT, levels = c("transect1","transect2","transect3"), labels = c("Transect 1", "Transect 2", "Transect 3"))) %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  mutate(SURFACE = case_when(
    DEPTH_BIN == "< 5m" ~ "Surface",
    TRUE ~ "sub-surface"
  )) %>% 
 #
  ggplot(aes(x = STN_ORDER, y = shannon_all, 
             shape = SURFACE,
             fill = cluster_order)) +
  facet_grid(cols = vars(transect_order), scale = "free_x") +
  geom_jitter(stroke = 0.6, width = 0.2, size = 3) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = CLUSTER_COLOR) +
  theme_bw() + theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        # plot.margin = margin(2, 1, 2, 1, "cm"),
        plot.margin = margin(b = 0, unit = "pt"),
        strip.text = element_text(color="black", size=12),
        strip.background = element_blank()) +
    labs(y = "Shannon Diversity", x = "Station") +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5, reverse = FALSE)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))


SHANNON_cluster
```
```{r, fig.width=6, fig.height=4}
SHANNON_depth <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = (depth_bin))) %>% 
  mutate(STN_ORDER = factor(as.character(Station), levels = stn_order_shore_offshore)) %>% 
  mutate(transect_order = factor(TRANSECT, levels = c("transect1","transect2","transect3"), labels = c("Transect 1", "Transect 2", "Transect 3"))) %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  mutate(SURFACE = case_when(
    DEPTH_BIN == "< 5m" ~ "Surface",
    TRUE ~ "sub-surface"
  )) %>% 
 #
  ggplot(aes(x = STN_ORDER, y = shannon_all, 
             shape = SURFACE,
             fill = depth_order)) +
  facet_grid(cols = vars(transect_order), scale = "free_x") +
  geom_jitter(stroke = 0.6, width = 0.2, size = 3) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_bw() + theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        # plot.margin = margin(2, 1, 2, 1, "cm"),
        plot.margin = margin(b = 0, unit = "pt"),
        strip.text = element_text(color="black", size=12),
        strip.background = element_blank()) +
    labs(y = "Shannon Diversity", x = "Station") +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5, reverse = FALSE)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))


SHANNON_depth
```
```{r, fig.height=4, fig.width=6}
SHANNON_depth_profile <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = (depth_bin))) %>% 
  mutate(STN_ORDER = factor(as.character(Station), levels = stn_order_shore_offshore)) %>% 
  mutate(transect_order = factor(TRANSECT, levels = c("transect1","transect2","transect3"), labels = c("Transect 1", "Transect 2", "Transect 3"))) %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  mutate(SURFACE = case_when(
    DEPTH_BIN == "< 5m" ~ "Surface",
    TRUE ~ "sub-surface"
  )) %>% 
 #
  ggplot(aes(y = Depth, x = shannon_all, 
             shape = SURFACE,
             fill = depth_order)) +
  facet_grid(cols = vars(transect_order), scale = "free_y") +
  geom_hline(yintercept = table_wstn_wdepth$DCM, colour = "red", alpha = 0.2) +
  geom_point(stroke = 0.6, size = 3) +
  scale_y_reverse() +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_bw() + theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        # plot.margin = margin(2, 1, 2, 1, "cm"),
        plot.margin = margin(b = 0, unit = "pt"),
        strip.text = element_text(color="black", size=12),
        strip.background = element_blank()) +
    labs(x = "Shannon Diversity", y = "Depth (m)") +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5, reverse = FALSE)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
```


### Inverse Simpson

```{r}
SIMPSON <- data.frame(invsimp_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(table_wstn_wdepth) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = depth_order, x = invsimp_all, 
             fill = depth_order, shape = MLD_THRESHOLD)) +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  scale_shape_manual(values = c(21, 23)) +
    scale_fill_manual(values = blues) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    axis.title = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 10),
    panel.grid.major = element_line(color = "grey80", linetype = "dashed")) +
    labs(x = "Inverse Simpsons", y = "") +
guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
SIMPSON
```

## T-S plot

```{r}
# Import table from Portela et al.
water_mass <- read.csv("input-data/water-mass-data.csv")
```

```{r}
water_mass %>% 
  ggplot() +
  geom_rect(data = water_mass, aes(xmin = SAL_MIN, xmax = SAL_MAX, ymin = TEMP_MIN, ymax = TEMP_MAX,
                                   linetype = Watermass, color = Watermass),
            alpha = 0.1) +
  theme_minimal() +
  scale_color_manual(values = c("#7116c7", "black", "black", "#f571be", "#db3e25", "#339ef5", "#339ef5")) +
  scale_linetype_manual(values = c("solid", "longdash", "dotted", "solid", "dotted", "longdash", "dotted"))
```

Portela, E., M. Tenreiro, E. Pallàs-Sanz, T. Meunier, A. Ruiz-Angulo, R. Sosa-Gutiérrez, and S. Cusí. 2018. Hydrography of the central and western gulf of Mexico. Journal of Geophysical Research. Oceans, 123(8):5134–5149, https://doi.org/10.1029/2018jc013813.

AAIW - Antarctic Intermediate Water CSWr - Caribbean Surface Water Remnant GCW - Gulf Common Water NADW - North Atlantic Deep Water NASUW - North Atlantic Subtropical Underwater TACW - Tropical Atlantic Central Water TACWn - Tropical Atlantic Central Water (nucleus)

## Plot K-means T-S plotß

```{r, fig.width=6, fig.height=6}
# head(table1_wfeature)
# CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9")
# CLUSTER_NUM <- c(1, 2, 3, 4, 5)
CLUSTER_COLOR <- c("#bd2a2a","#c85e82", "#efc85a", "#df9147","#47c475","#8b94b5","#6e9b4a", "#606aa9", "#524740")
CLUSTER_NUM <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
names(CLUSTER_NUM) <- CLUSTER_COLOR

ts_cluster_df <- table_supp %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  mutate(SURFACE = case_when(
    DEPTH_BIN == "< 5m" ~ "Surface",
    TRUE ~ "sub-surface"
  )) %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin))

water_mass_mod <- water_mass %>% 
  mutate(SAL_MIN_2 = case_when(
    SAL_MIN < 25 ~ 25,
    TRUE ~ SAL_MIN
  ))

ts_cluster <- ggplot() +
  # water masses
  geom_rect(data = water_mass_mod, aes(xmin = SAL_MIN_2, xmax = SAL_MAX, ymin = TEMP_MIN, ymax = TEMP_MAX, color = Watermass, linetype = Watermass),
            fill = "grey80", linewidth = 0.5, alpha = 0.2) +
    scale_color_manual(values = c("#7116c7", "black", "black", "#f571be", "#db3e25", "#339ef5", "#339ef5")) +
  scale_linetype_manual(values = c("solid", "longdash", "dotted", "solid", "dotted", "longdash", "dotted")) +
  geom_text_repel(data = water_mass_mod, size = 3, fontface = "bold", color = "black", 
                  aes(label = Watermass, x = SAL_MIN_2, y = TEMP_MIN),
                  box.padding = 1.5, nudge_x = .25, nudge_y = .5,
                  segment.linetype = 3, segment.alpha = 0.8,
                  segment.curvature = -1e-20, force = 1,
                  point.padding = 0.5) +
  # samples
  geom_jitter(data = ts_cluster_df, aes(x = Salinity, y = Temperature, fill = cluster_order, shape = SURFACE), stroke = 0.6, size = 2.5, width = 0.2, stat = "identity", color = "black") +
  scale_fill_manual(values = CLUSTER_COLOR) +
  scale_shape_manual(values = c(21, 23)) +
  scale_x_continuous(limits = c(25, 41)) +
  # scale_y_log10() +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        # plot.margin = margin(2, 1, 2, 1, "cm")) +
  plot.margin = margin(b = 0, unit = "pt")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 21, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5))) +
  labs(x = "Salinity (ppt)", y = "Temperature (°C)")
ts_cluster
```

```{r}
ts_depth <- ggplot() +
  # water masses
  geom_rect(data = water_mass_mod, aes(xmin = SAL_MIN_2, xmax = SAL_MAX, ymin = TEMP_MIN, ymax = TEMP_MAX, color = Watermass, linetype = Watermass),
            fill = "grey80", linewidth = 0.5, alpha = 0.2) +
    scale_color_manual(values = c("#7116c7", "black", "black", "#f571be", "#db3e25", "#339ef5", "#339ef5")) +
  scale_linetype_manual(values = c("solid", "longdash", "dotted", "solid", "dotted", "longdash", "dotted")) +
  geom_text_repel(data = water_mass_mod, size = 3, fontface = "bold", color = "black", 
                  aes(label = Watermass, x = SAL_MIN_2, y = TEMP_MIN),
                  box.padding = 1.5, nudge_x = .25, nudge_y = .5,
                  segment.linetype = 3, segment.alpha = 0.8,
                  segment.curvature = -1e-20, force = 1,
                  point.padding = 0.5) +
  # samples
  geom_jitter(data = ts_cluster_df, aes(x = Salinity, y = Temperature, fill = depth_order, shape = SURFACE), stroke = 0.6, size = 2.5, width = 0.2, stat = "identity", color = "black") +
  scale_fill_manual(values = blues) + 
  scale_shape_manual(values = c(21, 23)) +
  scale_x_continuous(limits = c(25, 41)) +
  # scale_y_log10() +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        # plot.margin = margin(2, 1, 2, 1, "cm")) +
  plot.margin = margin(b = 0, unit = "pt")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 21, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5))) +
  labs(x = "Salinity (ppt)", y = "Temperature (°C)")
ts_depth
```


### Combine pca and diversity

#### By depth

```{r, fig.width=8, fig.height=7}
(SHANNON_depth_profile + theme(legend.position = "none")) / (PCA_PLOT_depth) + 
  plot_layout(nrow = 2, widths = c(1, 0.5)) + 
  plot_annotation(tag_levels = c("A"))
# ggsave("figures/diversity-pca-by-DEPTH.svg", width = 8, height = 7, device = "svg", limitsize = FALSE)
```


```{r, fig.width=8, fig.height=7}
(SHANNON_depth) / ((PCA_PLOT_depth) + theme(legend.position = "none") + (ts_depth + theme(legend.position = "right"))) + plot_layout(nrow = 2, heights = c(0.8, 1, 1)) + plot_annotation(tag_levels = c("A"))
# ggsave("figures/diversity-pca-ts-plot-DEPTH.svg", width = 8, height = 7, device = "svg", limitsize = FALSE)
```
#### By cluster
```{r, fig.width=8, fig.height=7}
(SHANNON_cluster) / ((PCA_PLOT_kmeans) + theme(legend.position = "none") + (ts_cluster + theme(legend.position = "right"))) + plot_layout(nrow = 2, heights = c(0.8, 1, 1)) + plot_annotation(tag_levels = c("A"))
# ggsave("figures/diversity-pca-ts-plot-CLUSTER.svg", width = 8, height = 7, device = "svg", limitsize = FALSE)
```

# Dendrogram by k-mean cluster

Import kmeans clustering. Order all samples by k-mean cluster.

Get order of kmeans clusters 1 to 5, with station niskin order.

```{r}
table_kmeans_order <- table_supp %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  select(cluster_order, CLUSTER, STN_NISKIN, Depth) %>% distinct() %>% 
  arrange(cluster_order, Depth)
stn_bykmeans <- rev(as.character(table_kmeans_order$STN_NISKIN))
stn_bykmeans
```

Get color assignments

```{r}
# CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9")
# CLUSTER_NUM <- c(1, 2, 3, 4, 5)
CLUSTER_COLOR <- c("#bd2a2a","#c85e82", "#efc85a", "#df9147","#47c475","#8b94b5","#6e9b4a", "#606aa9", "#524740")
CLUSTER_NUM <- c(1, 2, 3, 4, 5, 6, 7, 8, 9)
names(CLUSTER_NUM) <- CLUSTER_COLOR

blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

stn_order_seqs <- c(1, 2, 3, 4, 5, 
                    9, 8, 7,
                    11, 12, 14, 15)

stn_order_color <- c("#e8dddc", "#e09f99", "#db756b","#d94638", "#db1e0d", 
                     "#a89492", "#ad6f6a", "#a6453c", 
                     "#4f4948", "#523735", "#4a221e", "#450d08")

names(stn_order_color) <- stn_order_seqs
```

Dendrogram assign lists

```{r}
table_supp_dendro <- table_supp %>% 
  mutate(cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>%
  mutate(stn_order = factor(STN_NISKIN, levels = (stn_bykmeans))) %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  mutate(water_mass_ORDER = factor(WATER_MASS, levels = water_mass_order)) %>% 
  mutate(dist_outflow = DIST_OUTFLOW/1000) %>%
  mutate(STN = factor(Station, levels = stn_order_seqs)) %>% 
  mutate(OIL_COUNT = as.character(OilRig_Count)) %>% 
  mutate(MLD_THRES = case_when(
    grepl("Above", MLD_THRESHOLD) ~ "Above",
    TRUE ~ "Below"),
    SAL_THRES = case_when(
    grepl("Above", SAL_THRESHOLD) ~ "Above",
    TRUE ~ "Below"),
    DCM_THRES = case_when(
    grepl("Above", DCM_THRESHOLD) ~ "Above",
    TRUE ~ "Below"),
    O2_THRES = case_when(
    grepl("Above", O2_THRESHOLD) ~ "Above",
    TRUE ~ "Below"))
```

Set variables.

```{r}
variable_fill <- list(
  "K-means" = "cluster_order",
  "Depth" = "depth_order",
  "Water mass" = "water_mass_ORDER",
  "Station" = "STN",
  "MLD" = "MLD_THRES",
  "DCM" = "DCM_THRES",
  "O2 min" = "O2_THRES",
  "Sal max" = "SAL_THRES",
  "Distance" = "dist_outflow",
  "Transect" = "TRANSECT",
  "Oil Rigs" = "OIL_COUNT",
  "Day Night" = "DAY_NIGHT",
  "Cells/ml" = "Prok_count",
  # Nutrients
  "Oxygen" = "Oxygen_CTD",
  "NO2" = "NO2",
  "NO3" = "NO3",
  "NH4" = "NH4",
  "PO4" = "PO4",
  "SIL" = "SIL",
  "DIC" = "DIC",
  "TA" = "TA"
)
```

Set scales.

```{r}
scale_fill <- list(
  scale_fill_manual(values = CLUSTER_COLOR, guide = guide_legend(order = 1)), #kmeans
  scale_fill_manual(values = blues, guide = guide_legend(reverse = TRUE, order = 2)), # depth
  scale_fill_manual(values = water_mass_color, guide = guide_legend(reverse = TRUE, order = 3)), #water mass
  scale_fill_manual(values = stn_order_color, guide = guide_legend(reverse = FALSE, order = 4)), # station
  scale_fill_manual(values = c("#d8d8f0", "#181752"), guide = guide_legend(order = 5)), #MLD
  scale_fill_manual(values = c("#d8d8f0", "#181752")), #DCM
  scale_fill_manual(values = c("#d8d8f0", "#181752")), #O2
  scale_fill_manual(values = c("#d8d8f0", "#181752")), #Sal
  scale_fill_viridis_c(option = "magma", direction = -1, guide = guide_legend(reverse = FALSE, order = 6)), #dist
  scale_fill_manual(values = c("#efaaf2", "#f003fc", "#470d4a"), guide = guide_legend(reverse = FALSE, order = 7)), #transect
  scale_fill_manual(values = c("grey90","#e38aa6","#ba4e70","#a13356","#8a1a3d","#54051e"), guide = guide_legend(reverse = FALSE, order = 8)), #Oil rig count
  scale_fill_manual(values = c("#96bbe0","#0a3157"), guide = guide_legend(reverse = TRUE, order = 9)), #DAy night
  scale_fill_fermenter(na.value = "grey50", palette = "Reds", direction = 1, guide = guide_legend(order = 10)), # proks
  ## Nutrients
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 11)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 12)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 13)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 14)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 15)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 16)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 17)),
  scale_fill_viridis(option = "plasma", guide = guide_legend(order = 18))
)
```

Align variables and scales

```{r}
names(scale_fill) <- names(variable_fill)
var_levels <- names(variable_fill)
var_levels
```

Set plot object.

```{r}
heatmap_combined_vars <- ggplot(table_supp_dendro, aes(y = stn_order)) +
  purrr::imap(
    variable_fill, \(x, y) {
      list(
        geom_tile(
          aes(
            x = y,
            fill = .data[[x]]
          ),
          color = "black",
          width = .95
        ),
        scale_fill[[y]],
        ggnewscale::new_scale_fill()
      )
    }
  ) +
  scale_x_discrete(
    limits = var_levels
  ) +
  # coord_fixed(ratio = 1) +
  facet_grid(rows = vars(cluster_order), scales = "free", space = "free") +
  theme_void() +
  theme(
    axis.text.y = element_blank(),
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    ),
    axis.title = element_blank(),
    strip.text = element_blank(),
    legend.position = "left",
    panel.border = element_rect(fill = NA, linewidth = 0.8),
    panel.background = element_rect(colour = NA, fill = "transparent"),
    plot.background = element_rect(colour = NA, fill = "transparent"),
    plot.margin = margin(0, 0, 0, 0, "cm"),
    panel.spacing = unit(0, "cm"),
    legend.title = element_blank()
  )
```

```{r, fig.height=25, fig.width=8}
heatmap_combined_vars
```

### k-mean tax

```{r}
tax_color  <- c("#607080","#2f97f7","#2c5c1d","#6da65b", "grey20",
"#de6600", #Gyrista
"#ee429b", #dinos
"#1abc9c", # ciliates
"#23b521","#f5bfb0","#fea02f", "#f7d472",
"#d7b6d7","#460e57", 
"#572f0e","#916643","#c95e06", "#dba272", "#e6d2b8", "#b09b10",
"#ad3d46", #Haptista
"#5e0810", "grey30", "#db9399","#93c47d","#ca7ade")

tax_simplified_list <- as.character(rev(sort(unique(asv_long_avg_wtax_clean $Supergroup_simplified))))
tax_simplified_list
names(tax_color) <- tax_simplified_list
names(tax_simplified_list) <- tax_color
```

```{r, fig.height=12, fig.width=6}
# Sum by taxonomic groups
bykmeans_tax <- asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  group_by(STN_NISKIN, Station, Niskin, 
           Depth, DEPTH_BIN, Supergroup_simplified, CLUSTER) |> 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(stn_order = factor(STN_NISKIN, levels = (stn_bykmeans)),
         tax_order = factor(Supergroup_simplified, levels = (tax_simplified_list)),
         cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  ggplot(aes(y = (stn_order), x = SUM_SEQ, fill = tax_order)) +
    geom_bar(stat = "identity", position = "fill", 
             color = "grey20", width = 0.7, linewidth = 0.3) +
  scale_fill_manual(values = tax_color) +
    scale_y_discrete(position = "left") +
  facet_grid(rows = vars(cluster_order), scales = "free", space = "free") +
  scale_x_continuous(expand = c(0,0)) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 5),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Relative Abundance", y = "")
bykmeans_tax
```

### k-mean asv total

Total ASVs

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
ASV_TOTAL <- asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  group_by(STN_NISKIN, Station, Niskin, CLUSTER) |> 
  summarise(ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(stn_order = factor(STN_NISKIN, levels = (stn_bykmeans)),
         cluster_order = factor(CLUSTER, levels = (CLUSTER_NUM))) %>% 
  ggplot(aes(y = stn_order, x = ASV_COUNT)) +
    geom_bar(stat = "identity", position = "stack", 
             color = "black", fill = "grey40", width = 0.7, linewidth = 0.3) +
  scale_y_discrete(position = "left") +
  scale_x_continuous(expand = c(0,0)) +
    facet_grid(rows = vars(cluster_order), scales = "free", space = "free") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    strip.text = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 5),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Total ASVs", y = "")
ASV_TOTAL
```

## Plot Dendrogram compilation

```{r, fig.height=18, fig.width=9}
heatmap_combined_vars + 
  (bykmeans_tax + theme(axis.text.y = element_text(color = "black", size = 8, hjust = 1, vjust = 0.5))) + 
  (ASV_TOTAL+ theme(axis.text.y = element_blank())) + 
  plot_layout(nrow = 1, widths = c(1, 0.3, 0.3), guides = "collect")
# ggsave("figures/heatmap-kmeans-all.svg", width = 9, height = 18, device = "svg", limitsize = FALSE)
# ggsave("figures/heatmap-kmeans-legend.svg", width = 9, height = 30, device = "svg", limitsize = FALSE)
```

# Taxonomy

Richness by depth, station, and feature.

```{r}
head(table_wstn_wdepth)
```

```{r}
tax_color  <- c("#607080","#2f97f7","#2c5c1d","#6da65b", "grey20",
"#de6600", #Gyrista
"#ee429b", #dinos
"#1abc9c", # ciliates
"#23b521","#f5bfb0","#fea02f", "#f7d472",
"#d7b6d7","#460e57", 
"#572f0e","#916643","#c95e06", "#dba272", "#e6d2b8", "#b09b10",
"#ad3d46", #Haptista
"#5e0810", "grey30", "#db9399","#93c47d","#ca7ade")

tax_simplified_list <- as.character(rev(sort(unique(asv_long_avg_wtax_clean $Supergroup_simplified))))
tax_simplified_list
names(tax_color) <- tax_simplified_list
names(tax_simplified_list) <- tax_color
# view(tax_simplified_list)
```

Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level.

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
four <- c("TSAR-Dinoflagellata", "TSAR-Radiolaria", "TSAR-Ciliophora", "TSAR-Gyrista")
```


### Plot sequence relative abundance

```{r, fig.width=5, fig.height=14}
asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  group_by(Station, DEPTH_BIN, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin)),
          tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = depth_bin_order, x = SUM_SEQ, fill = tax_order)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.9) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(Station), scales = "free_y", space = "free", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

### Plot sequence relative abundance with depths

```{r, fig.width=5, fig.height=14}
asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  group_by(Station, Depth, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = as.factor(Depth), x = SUM_SEQ, fill = tax_order)) +
  geom_bar(stat = "identity", position = "fill", color = "white", width = 0.9) +
  scale_fill_manual(values = tax_color) +
  # scale_y_reverse() +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(Station), scales = "free_y", space = "free", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

Plot ASV richness by taxonomic groups

```{r, fig.width=22, fig.height=5}
# tax_color
asv_long_avg_wtax_clean %>%
  left_join(table_wstn_wdepth) %>%
  group_by(DEPTH_BIN, Supergroup_simplified, Class) %>%
  summarise(ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "Eukaryota_X") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin)),
          tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = depth_bin_order, x = Class, fill = tax_order, size = ASV_COUNT)) +
  geom_point(shape = 21, color = "grey10", alpha = 0.7, stroke = 0.5) +
  scale_fill_manual(values = tax_color) +
  scale_size_continuous(range = c(1,10), breaks = c(1, 10, 50, 100, 500, 1000, 2000)) +
  facet_grid(cols = vars(tax_order), scales = "free_x", space = "free_x", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
    strip.text.y = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 8, color = "black", face = "bold"),
    strip.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 8, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold", angle = 90, 
                               hjust = 1, vjust = 0, size = 8)) +
  guides(fill = "none",
         size = guide_legend(nrow = 1, shape = 21, color = "black", fill = "black")) +
  labs(x = "", y = "Depth (m)")

# ggsave("figures/tax-dotplot.svg", width = 22, height = 5, device = "svg", limitsize = FALSE)
```
# K-means plot with taxonomy

```{r}
# head(table_supp1)
kmeans_depth <- table_supp1 %>% ungroup() %>% 
  select(CLUSTER, STN_NISKIN, STN_CATEGORY_4, ends_with("_THRESHOLD"), DAY_NIGHT, DEPTH_BIN, WATER_MASS) %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  distinct() %>% 
  ggplot(aes(y = (as.factor((CLUSTER))))) +
    geom_bar(stat = "count", aes(fill = depth_order), color = "white", width = 0.6) +
    scale_fill_manual(values = blues, guide = guide_legend(reverse = TRUE, order = 2)) +
  scale_x_continuous(expand = c(0,0)) +
    theme_classic() +
    theme(axis.text = element_text(color="black", size=12),
          axis.text.y = element_blank(),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(b = 0, unit = "pt")) +
  labs(x = "Number of samples", y = "")
```

```{r}
kmeans_category <- table_supp1 %>% ungroup() %>% 
  select(CLUSTER, STN_NISKIN, STN_CATEGORY_4, ends_with("_THRESHOLD"), DAY_NIGHT, DEPTH_BIN, WATER_MASS) %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  distinct() %>% 
  ggplot(aes(y = (as.factor((CLUSTER))))) +
    geom_bar(stat = "count", aes(fill = STN_CATEGORY_4), color = "white", width = 0.6) +
    scale_fill_manual(values = c("grey5", "grey20", "grey50", "grey80")) +
  scale_x_continuous(expand = c(0,0)) +
    theme_classic() +
    theme(axis.text = element_text(color="black", size=12),
          # axis.text.y = element_blank(),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(b = 0, unit = "pt")) +
  labs(x = "Number of samples", y = "k-means cluster")
```

```{r}
kmeans_water <- table_supp1 %>% ungroup() %>% 
  select(CLUSTER, STN_NISKIN, STN_CATEGORY_4, ends_with("_THRESHOLD"), DAY_NIGHT, DEPTH_BIN, WATER_MASS) %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  mutate(WATER_MASS_ORDER = factor(WATER_MASS, levels = water_mass_order)) %>% 
  distinct() %>% 
  ggplot(aes(y = (as.factor((CLUSTER))))) +
    geom_bar(stat = "count", aes(fill = WATER_MASS_ORDER), color = "white", width = 0.6) +
    scale_fill_manual(values = water_mass_color) +
  scale_x_continuous(expand = c(0,0)) +
    theme_classic() +
    theme(axis.text.x = element_text(color="black", size=12),
          axis.text.y = element_blank(),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(b = 0, unit = "pt")) +
  labs(x = "Number of samples", y = "")
```

```{r}
kmeans_tax <- asv_long_avg_wtax_clean %>%
  left_join(table_supp1 %>% 
              select(STN_NISKIN, CLUSTER) %>% 
              distinct()) %>%
  group_by(CLUSTER, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = as.factor(CLUSTER), x = SUM_SEQ, fill = tax_order)) +
  geom_bar(stat = "identity", position = "fill", color = "white", width = 0.6) +
  scale_fill_manual(values = tax_color) +
  # scale_y_reverse() +
  scale_x_continuous(expand = c(0,0)) +
    theme_classic() +
    theme(axis.text.x = element_text(color="black", size=12),
          axis.text.y = element_blank(),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey80", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(b = 0, unit = "pt")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "")
```



```{r, fig.width=14, fig.height=4}
kmeans_category + kmeans_depth + kmeans_water + kmeans_tax + plot_layout(nrow = 1, guides = "collect")
ggsave("figures/kmeans-barplot-composition.svg", width = 14, height = 4, device = "svg", limitsize = FALSE)
```

# Taxa by depth - curated

## Curate specific taxa datasets

```{r}
stn_long <- table_supp %>% select(DIST_OUTFLOW, Station) %>% distinct() %>% 
  mutate(Station = as.character(Station))
```

### Dinos

```{r}
dinos_subset <- c("Dinoflagellata-Dinophyceae","Dinoflagellata-Syndiniales")

seq_total <- asv_long_avg_wtax_clean %>% 
  # filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  filter(SUPERGROUP_CLASS %in% dinos_subset) %>% 
  group_by(FeatureID) %>% 
  summarise(SAMPLE_COUNT = n_distinct(STN_NISKIN),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  filter(SAMPLE_COUNT >= 9)
#
keep_asvs <- as.character(unique(seq_total$FeatureID))
#
dinos_curated <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  # filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  filter(FeatureID %in% keep_asvs) %>% 
  # left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    grepl("Dinophyceae_X", TAXA) ~ "Dinophyceae",
    grepl("Dinophyta_X", TAXA) ~ "Dinophyta",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  select(FeatureID, TAXA, Station, STN_NISKIN, Depth, DEPTH_BIN, TRANSECT, MEAN_REPS_seq) %>% distinct()
# head(dinos_curated)
#
tmp <- dinos_curated %>% 
  group_by(TAXA) %>% 
  summarise(SUM = sum(MEAN_REPS_seq)) %>% 
  slice_max(n = 8, order_by = SUM)

top_5 <- as.character(unique(tmp$TAXA))

dinos_curated_5 <- dinos_curated %>% 
  filter(TAXA %in% top_5)
```

```{r}
mat_tax_df <- dinos_curated %>% ungroup() %>% 
  # Save taxa id curated and ASV ID
  unite(sample, STN_NISKIN, DEPTH_BIN, sep = ";") %>%
  group_by(TAXA, sample) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "sample") %>% 
    as.matrix 
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tax_df, MARGIN = 2, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tax_df)
  rownames(clr_tmp) <- rownames(mat_tax_df)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "sample") %>% 
    pivot_longer(cols = -c(sample), names_to = "TAXA", values_to = "CLR") %>% 
    separate(sample, into = c("STN_NISKIN", "DEPTH_BIN"), sep = ";") %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = TAXA, fill = CLR)) +
    geom_tile(color = "black") + 
    scale_fill_fermenter(type = "div",
      palette = "RdBu",
      limits = c(-4, 4)) +
    coord_fixed(ratio = 1) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 90, size = 8,
                                     hjust = 1, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          legend.position = "right") +
    labs(x = "", y = "", title = "TSAR-Dinoflagellata")
```

### Rad

```{r}
# tax_simplified_list
rad_subset <- c("Radiolaria-Polycystinea","Radiolaria-Acantharea","Radiolaria-RAD-A","Radiolaria-RAD-B","Radiolaria-RAD-C")
```

```{r}
seq_total <- asv_long_avg_wtax_clean %>% 
  # filter(Supergroup_simplified == "TSAR-Radiolaria") %>%  
  filter(SUPERGROUP_CLASS %in% rad_subset) %>% 
  group_by(FeatureID) %>% 
  summarise(SAMPLE_COUNT = n_distinct(STN_NISKIN),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  filter(SAMPLE_COUNT >= 9)
#
keep_asvs <- as.character(unique(seq_total$FeatureID))
#
rad_curated <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(FeatureID %in% keep_asvs) %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Acantharea", TAXA) ~ "Acantharea",
    grepl("RAD_A", TAXA) ~ "RAD_A",
    grepl("RAD_B", TAXA) ~ "RAD_B",
    grepl("RAD_C", TAXA) ~ "RAD_C",
    grepl("Radiolaria_XX", TAXA) ~ "Radiolaria",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>%
  select(FeatureID, TAXA, Station, STN_NISKIN, Depth, DEPTH_BIN, TRANSECT, MEAN_REPS_seq) %>% distinct()
#
tmp <- rad_curated %>% 
  group_by(TAXA) %>% 
  summarise(SUM = sum(MEAN_REPS_seq)) %>% 
  slice_max(n = 5, order_by = SUM)

top_5 <- as.character(unique(tmp$TAXA))

rad_curated_5 <- rad_curated %>% 
  filter(TAXA %in% top_5)
```

```{r, fig.height=15, fig.width=5}
mat_tax_df_surface <- rad_curated_5 %>% ungroup() %>% 
  filter(DEPTH_BIN == "< 5m") %>% 
  unite(sample, Station, TRANSECT, sep = ";") %>%
    group_by(TAXA, sample) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "sample") %>% 
    as.matrix 
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tax_df_surface, MARGIN = 2, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tax_df_surface)
  rownames(clr_tmp) <- rownames(mat_tax_df_surface)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "sample") %>% 
    pivot_longer(cols = -c(sample), names_to = "TAXA", values_to = "CLR") %>% 
    separate(sample, into = c("Station", "Transect"), sep = ";") %>% 
    left_join(stn_long) %>% 
    mutate(Station_Order = factor(Station, levels = (stn_order_shore_offshore))) %>% 
    ggplot(aes(x = (DIST_OUTFLOW/1000), y = CLR, shape = Transect)) +
    geom_hline(yintercept = 0) +
    geom_line(aes(group = Transect)) +
    geom_jitter(color = "black", fill = "#ee429b", size = 2) +
    scale_shape_manual(values = c(21, 22, 23)) +
    facet_grid(rows = vars(TAXA), scales = "free_x") +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0.5, vjust = 0.5),
          axis.text.y = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 1, vjust = 0.5),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          strip.text.y = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          strip.background = element_blank(),
          panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
          legend.position = "left",
          aspect.ratio = 1/2) +
    labs(x = "km", y = "Center log-ratio", title = "")
```

### Ciliate

```{r}
other_ciliates <- c("Oligohymenophorea;Oligohymenophorea_X", "Litostomatea;Haptoria_6", "Heterotrichea;Heterotrichea_X", "Spirotrichea;Hypotrichia", "Ciliophora_X;Ciliophora_XX")
ciliate_subset <- c("Ciliophora-Spirotrichea", "Ciliophora-Phyllopharyngea", "Ciliophora-Oligohymenophorea", "Ciliophora-CONThreeP")

# Curate asvs to keep
seq_total <- asv_long_avg_wtax_clean %>% 
  # filter(Supergroup_simplified == "TSAR-Ciliophora") %>%  
  filter(SUPERGROUP_CLASS %in% ciliate_subset) %>% 
  group_by(FeatureID) %>% 
  summarise(SAMPLE_COUNT = n_distinct(STN_NISKIN),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  filter(SAMPLE_COUNT >= 9)
#
keep_asvs <- as.character(unique(seq_total$FeatureID))
#
ciliate_curated <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  # filter(Supergroup_simplified == "TSAR-Ciliophora") %>%  
  filter(FeatureID %in% keep_asvs) %>% 
  # left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    (TAXA %in% other_ciliates) ~ "Ciliates unannotated",
    grepl("CONTH_", TAXA) ~ "CONThreeP unannotated",
    grepl("CONThreeP_X", TAXA) ~ "CONThreeP unannotated",
    (Class == "Discotrichidae") ~ "CONThreeP;Discotrichidae",
    grepl("Phyllopharyngea;Cyrtophoria", TAXA) ~ "Phyllopharyngea;Cyrtophoria",
    grepl("Spirotrichea;Choreotrichida", TAXA) ~ "Spirotrichea;Choreotrichida",
    grepl("Litostomatea;Litostomatea_X", TAXA) ~ "Litostomatea",
    grepl("Prostomatea_1;Prostomatea_1_X", TAXA) ~ "Prostomatea",
    grepl("Colpodea_X", TAXA) ~ "Colpodea",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  select(FeatureID, TAXA, Station, STN_NISKIN, Depth, DEPTH_BIN, TRANSECT, MEAN_REPS_seq) %>% distinct()

head(ciliate_curated)

tmp <- ciliate_curated %>% 
  group_by(TAXA) %>% 
  summarise(SUM = sum(MEAN_REPS_seq)) %>% 
  slice_max(n = 5, order_by = SUM)

top_5 <- as.character(unique(tmp$TAXA))

ciliate_curated_5 <- ciliate_curated %>% 
  filter(TAXA %in% top_5)
```

### Gyrista

```{r}
gyrista_less <- c("Pelagophyceae unannotated", "Pelagophyceae;NA","Coscinodiscophyceae;Coscinodiscales", "Bacillariophyceae;Licmophorales", "Eustigmatophyceae;Eustigmatophyceae_X", "Pirsoniales;Pirsoniales_X","Peronosporomycetes;Peronosporomycetes_X", "Bacillariophyceae;Cymbellales") 

# Curate asvs to keep
seq_total <- asv_long_avg_wtax_clean %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  group_by(FeatureID) %>% 
  summarise(SAMPLE_COUNT = n_distinct(STN_NISKIN),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  filter(SAMPLE_COUNT >= 9)
#
keep_asvs <- as.character(unique(seq_total$FeatureID))
#
gyrista_curated <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  filter(FeatureID %in% keep_asvs) %>% 
  # left_join(table_wstn_wdepth) %>%  
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>%
  mutate(TAXA = case_when(
    is.na(Class) ~ "Gyrista;Unannotated",
    (TAXA %in% gyrista_less) ~ "Gyrista;Unannotated",
    grepl(";NA", TAXA) ~ paste(Class, "Unannotated", sep = ";"),
    (TAXA == "Gyrista_X;Gyrista_XX") ~ "Gyrista;Unannotated",
    grepl("Dictyochophyceae", TAXA) ~ "Dictyochophyceae",
    grepl("Chrysophyceae_Clade-EC2", TAXA) ~ "Chrysophyceae;Clade-EC2",
    grepl("Bacillariophyceae;Bacillariophyceae_X", TAXA) ~ "Bacillariophyceae;Unannotated",
    grepl("Pelagophyceae", TAXA) ~ "Pelagophyceae;Pelagomonadales",
    grepl("MOCH-", TAXA) ~ "Gyrista;MOCH clades 2-5",
    (TAXA == "Mediophyceae;Mediophyceae_X" | TAXA == "Mediophyceae;NA") ~ "Mediophyceae;Unannotated",
    TRUE ~ TAXA
  )) %>%
  select(FeatureID, TAXA, Station, STN_NISKIN, Depth, DEPTH_BIN, TRANSECT, MEAN_REPS_seq) %>% distinct()

head(gyrista_curated)

tmp <- gyrista_curated %>% 
  group_by(TAXA) %>% 
  summarise(SUM = sum(MEAN_REPS_seq)) %>% 
  slice_max(n = 5, order_by = SUM)

top_5 <- as.character(unique(tmp$TAXA))

gyrista_curated_5 <- gyrista_curated %>% 
  filter(TAXA %in% top_5)
```

### Bigyra

```{r}
# Curate asvs to keep
seq_total <- asv_long_avg_wtax_clean %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
  group_by(FeatureID) %>% 
  summarise(SAMPLE_COUNT = n_distinct(STN_NISKIN),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  filter(SAMPLE_COUNT >= 9)
#
keep_asvs <- as.character(unique(seq_total$FeatureID))
#
bigyra_curated <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
  filter(FeatureID %in% keep_asvs) %>% 
  unite(TAXA, Order, Family, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  select(FeatureID, TAXA, Station, STN_NISKIN, DEPTH_BIN, TRANSECT, MEAN_REPS_seq) %>% distinct()

tmp <- bigyra_curated %>% 
  group_by(TAXA) %>% 
  summarise(SUM = sum(MEAN_REPS_seq)) %>% 
  slice_max(n = 5, order_by = SUM)

top_5 <- as.character(unique(tmp$TAXA))

bigyra_curated_5 <- bigyra_curated %>% 
  filter(TAXA %in% top_5)
```

### Haptista

```{r}
# Curate asvs to keep
seq_total <- asv_long_avg_wtax_clean %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
  group_by(FeatureID) %>% 
  summarise(SAMPLE_COUNT = n_distinct(STN_NISKIN),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  filter(SAMPLE_COUNT >= 9)
#
keep_asvs <- as.character(unique(seq_total$FeatureID))
#
haptista_curated <- asv_long_avg_wtax_clean %>% 
  left_join(table_wstn_wdepth) %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
  filter(FeatureID %in% keep_asvs) %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Pterocystida", TAXA) ~ "Pterocystida",
    grepl("Haptophyta_Clade_", TAXA) ~ "Haptophyta Clades-3/4",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  select(FeatureID, TAXA, Station, STN_NISKIN, DEPTH_BIN, TRANSECT, MEAN_REPS_seq) %>% distinct()

tmp <- haptista_curated %>% 
  group_by(TAXA) %>% 
  summarise(SUM = sum(MEAN_REPS_seq)) %>% 
  slice_max(n = 5, order_by = SUM)

top_5 <- as.character(unique(tmp$TAXA))

haptista_curated_5 <- haptista_curated %>% 
  filter(TAXA %in% top_5)
```


## Taxa by cluster

Create tree plots for key taxa

```{r}
dino_colors <- c("#db1f7d", "#d1a9bd", "#8a1751", "#540c30",
                 "#4d917b", "#89c977",
                 "#e8912e","#d9bfa3","#aba250", "#cc9e6a","#436e16", "#a34e27")
# names(dino_colors) <- tax_dino_order

tree_dino <- dinos_curated %>% ungroup() %>% 
  left_join((table_wstn_wdepth %>% select(STN_NISKIN, CLUSTER) %>% distinct())) %>% 
  # filter(Depth >= 200) %>% 
  group_by(TAXA, CLUSTER) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            COUNT = n()) %>% 
  separate(TAXA, into = c("GROUP", "SUBGROUP"), sep = ";", remove = FALSE) %>% 
  mutate(SUBGROUP = str_replace_na(SUBGROUP, "Other")) %>% 
  ggplot(aes(area = SUM, fill = SUBGROUP, 
                                        # label = SUBGROUP,
                                        subgroup = GROUP, 
                                        subgroup2 = TAXA)) + 
  geom_treemap() +
  scale_fill_manual(values = dino_colors) +
  facet_grid(cols = vars(CLUSTER)) +
  geom_treemap_subgroup_border(colour = "white", size = 3) +
  geom_treemap_subgroup2_border(colour = "white", size = 1) +
  # geom_treemap_subgroup2_text(colour = "white", alpha = 0.8, fontface = "bold") +
  theme_minimal() +
  theme(strip.text = element_text(color = "black", face = "bold",
                                  hjust = 0, vjust = 0),
        legend.title = element_blank())


```


```{r}
rad_colors <- c("#805edb", "#afe3cd", "#164531", "#e3e340", "#c4c49b", "#5b6940", "#648bb3")
# names(rad_colors) <- tax_rad_order

tree_rad <- rad_curated %>% ungroup() %>% 
  left_join((table_wstn_wdepth %>% select(STN_NISKIN, CLUSTER) %>% distinct())) %>% 
  # filter(Depth >= 200) %>% 
  group_by(TAXA, CLUSTER) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            COUNT = n()) %>% 
  separate(TAXA, into = c("GROUP", "SUBGROUP"), sep = ";", remove = FALSE) %>% 
  mutate(SUBGROUP = str_replace_na(SUBGROUP, "Other")) %>% 
  ggplot(aes(area = SUM, fill = SUBGROUP, 
                                        # label = SUBGROUP,
                                        subgroup = GROUP, 
                                        subgroup2 = TAXA)) + 
  geom_treemap() +
  scale_fill_manual(values = rad_colors) +
  facet_grid(cols = vars(CLUSTER)) +
  geom_treemap_subgroup_border(colour = "white", size = 3) +
  geom_treemap_subgroup2_border(colour = "white", size = 1) +
  # geom_treemap_subgroup2_text(colour = "white", alpha = 0.8, fontface = "bold") +
  theme_minimal() +
  theme(strip.text = element_text(color = "black", face = "bold",
                                  hjust = 0, vjust = 0),
        legend.title = element_blank())


```

```{r}
gyrista_colors <- c("#93c47d", "#de5472","#7db7c4", "#a27dc4", 
                    "#d65c1a", "#c98d6d", "#662a0a", "#805944", "#bd5f17","#a3690b", 
                    "#d93529", "#736b6b","#d4cf44",
                    "#225706","#8ce663", "#38a605", 
                    "#0ba7d6","#1836cc", "#6abbd4", "#5669c4","#97b8c2","#0d3142",
                    "#5cd1b8", 
                    "#d596e3", "#b82fd6", "#a10b83","#a85b99")

# names(gyrista_colors) <- tax_gyrista_order


tree_gyrista <- gyrista_curated %>% ungroup() %>% 
  left_join((table_wstn_wdepth %>% select(STN_NISKIN, CLUSTER) %>% distinct())) %>% 
  # filter(Depth >= 200) %>% 
  group_by(TAXA, CLUSTER) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            COUNT = n()) %>% 
  separate(TAXA, into = c("GROUP", "SUBGROUP"), sep = ";", remove = FALSE) %>% 
  mutate(SUBGROUP = str_replace_na(SUBGROUP, "Other")) %>% 
  ggplot(aes(area = SUM, fill = SUBGROUP, 
                                        # label = SUBGROUP,
                                        subgroup = GROUP, 
                                        subgroup2 = TAXA)) + 
  geom_treemap() +
  scale_fill_manual(values = gyrista_colors) +
  facet_grid(cols = vars(CLUSTER)) +
  geom_treemap_subgroup_border(colour = "white", size = 3) +
  geom_treemap_subgroup2_border(colour = "white", size = 1) +
  # geom_treemap_subgroup2_text(colour = "white", alpha = 0.8, fontface = "bold") +
  theme_minimal() +
  theme(strip.text = element_text(color = "black", face = "bold",
                                  hjust = 0, vjust = 0),
        legend.title = element_blank())


```

```{r}
ciliate_colors <- c("#f5bfb0", "#ed7e5f", "#d13206", 
                    "#cf9229", "#69552a",
                    "#cfcf57", "#84c2a5","#604a70","#d1d106", 
                    "#accf97", "#7ade3c",  "#2edb8a", "#02703d",
                    "#3899a8", "#445cad", "#5402e0", "#916dcf", "#604a70", "#7b857e")

# names(ciliate_colors) <- tax_ciliate_order

tree_ciliate <- ciliate_curated %>% ungroup() %>% 
  left_join((table_wstn_wdepth %>% select(STN_NISKIN, CLUSTER) %>% distinct())) %>% 
  # filter(Depth >= 200) %>% 
  group_by(TAXA, CLUSTER) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            COUNT = n()) %>% 
  separate(TAXA, into = c("GROUP", "SUBGROUP"), sep = ";", remove = FALSE) %>% 
  mutate(SUBGROUP = str_replace_na(SUBGROUP, "Other")) %>% 
  ggplot(aes(area = SUM, fill = SUBGROUP, 
                                        # label = SUBGROUP,
                                        subgroup = GROUP, 
                                        subgroup2 = TAXA)) + 
  geom_treemap() +
  scale_fill_manual(values = ciliate_colors) +
  facet_grid(cols = vars(CLUSTER)) +
  geom_treemap_subgroup_border(colour = "white", size = 3) +
  geom_treemap_subgroup2_border(colour = "white", size = 1) +
  # geom_treemap_subgroup2_text(colour = "white", alpha = 0.8, fontface = "bold") +
  theme_minimal() +
  theme(strip.text = element_text(color = "black", face = "bold",
                                  hjust = 0, vjust = 0),
        legend.title = element_blank())


```

### Compile tree plot

```{r, fig.width=17, fig.height=9}
tree_dino + tree_rad + tree_gyrista + tree_ciliate + plot_layout(ncol = 1)
```


## Function CLR relationship

```{r}
# # requires:
#   # depth_bin
#   # stn_order_shore_offshore
#   # stn_long
toptaxa_depth_clr <- function(DF, TITLE){
  # Get matrix for depth profile
  mat_tax_df <- DF %>% ungroup() %>% 
    unite(sample, STN_NISKIN, DEPTH_BIN, sep = ";") %>%
    group_by(TAXA, sample) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
      pivot_wider(names_from = TAXA, values_from = SUM_SEQ, 
                  values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "sample") %>% 
    as.matrix 
  # colnames(mat_tmp)
  clr_depth <- vegan::decostand(mat_tax_df, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_depth) <- colnames(mat_tax_df)
  rownames(clr_depth) <- rownames(mat_tax_df)
  #
    (data.frame(clr_depth) %>% 
    rownames_to_column(var = "sample") %>% 
    pivot_longer(cols = -c(sample), names_to = "TAXA", values_to = "CLR") %>% 
    separate(sample, into = c("STN_NISKIN", "DEPTH_BIN"), sep = ";") %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = TAXA, fill = CLR)) +
    geom_tile(color = "black") + 
        scale_fill_steps2(limits = c(-2.5,2.5), 
                          show.limits = TRUE,
                          low = "#273fa1",
                          mid = "#ffffff",
                          high = "#a12727") +
    #     scale_fill_gradientn(
    #       colors = c("blue", "dodgerblue", "white", "orange", "red"),
    # rescaler = ~ scales::rescale_mid(.x, mid = 0)
    #     ) +
    # scale_fill_gradient2(
      # low = LO,
      # low = "#3843e0",
      # mid = MID,
      # mid = "#ede4e4",
      # high = HI,
      # high = "#d12b1f",
      # midpoint = 0,
      # na.value = "grey50",
      # aesthetics = "fill",
      # limits = c(-4, 4)
    # ) +
    coord_fixed(ratio = 1) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 90, size = 8,
                                     hjust = 1, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          legend.position = "right") +
    labs(x = "", y = "", title = TITLE))
}
```

```{r}
# requires:
  # depth_bin
  # stn_order_shore_offshore
  # stn_long

toptaxa_surf_clr <- function(DF, TITLE){
  mat_tax_df_surface <- DF %>% ungroup() %>% 
    filter(DEPTH_BIN == "< 5m") %>% 
    unite(sample, Station, TRANSECT, sep = ";") %>%
    group_by(TAXA, sample) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, 
                values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "sample") %>% 
    as.matrix 
  # mat_tax_df_surface
  clr_surf <- vegan::decostand(mat_tax_df_surface, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_surf) <- colnames(mat_tax_df_surface)
  rownames(clr_surf) <- rownames(mat_tax_df_surface)
  ##
  (data.frame(clr_surf) %>% 
    rownames_to_column(var = "sample") %>% 
    pivot_longer(cols = -c(sample), names_to = "TAXA", values_to = "CLR") %>% 
    separate(sample, into = c("Station", "Transect"), sep = ";") %>% 
    left_join(stn_long) %>% 
    mutate(Station_Order = factor(Station, levels = rev(stn_order_shore_offshore))) %>% 
    # ggplot(aes(y = Station_Order, x = TAXA, fill = CLR)) +
    # geom_tile(color = "black") + 
    #     scale_fill_steps2(limits = c(-3,3), show.limits = TRUE) +
    ggplot(aes(x = (DIST_OUTFLOW/1000), y = CLR, fill = TAXA)) +
    geom_hline(yintercept = 0) +
    # geom_line(aes(group = Transect, linetype = Transect)) +
    geom_jitter(color = "white", size = 2, shape = 21) +
    # scale_shape_manual(values = c(21, 22, 23)) +
    # facet_grid(rows = vars(TAXA), scales = "free_x") +
    # coord_fixed(ratio = 1) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 90, size = 8,
                                     hjust = 1, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          legend.position = "right") +
    labs(x = "", y = "", title = TITLE)) 
}
```

### Dinos

```{r, fig.height=8, fig.width=10}
toptaxa_surf_clr(dinos_curated_5, "Dinoflagellate") + 
  toptaxa_depth_clr(dinos_curated_5,  "Dinoflagellates") +
  plot_layout(nrow = 1, heights = c(1, 0.9))
```

### Rad

```{r, fig.height=7, fig.width=10}
toptaxa_surf_clr(rad_curated_5, "Rhizaria") +
  toptaxa_depth_clr(rad_curated_5, "Rhizaria") +
  plot_layout(nrow = 1, heights = c(1, 0.6))
```

### Ciliate

```{r, fig.height=7, fig.width=10}
toptaxa_surf_clr(ciliate_curated_5, "Ciliates") +
  toptaxa_depth_clr(ciliate_curated_5, "Ciliates") +
  plot_layout(nrow = 1, heights = c(1, 0.6))
```

### Gyrista

```{r, fig.height=7, fig.width=10}
toptaxa_surf_clr(gyrista_curated_5, "Gyrista") +
  toptaxa_depth_clr(gyrista_curated_5,  "Gyrista") +
  plot_layout(nrow = 1, heights = c(1, 0.6))
```

### Bigyra

```{r, fig.height=7, fig.width=10}
toptaxa_surf_clr(bigyra_curated_5, "Bigyra") +
  toptaxa_depth_clr(bigyra_curated_5, "Bigyra") +
  plot_layout(nrow = 1, heights = c(1, 0.6))
```

### Haptista

```{r, fig.height=7, fig.width=10}
toptaxa_surf_clr(haptista_curated_5, "Haptista") +
  toptaxa_depth_clr(haptista_curated_5, "Haptista") +
  plot_layout(nrow = 1, heights = c(1, 0.6))
```

## Compile main text - tax specific

```{r, fig.height=25, fig.width=25}
toptaxa_surf_clr(dinos_curated_5, "Dinoflagellate") +
toptaxa_depth_clr(dinos_curated_5,  "Dinoflagellates") +
  toptaxa_surf_clr(ciliate_curated_5, "Ciliates") +
  toptaxa_depth_clr(ciliate_curated_5, "Ciliates") +
    toptaxa_surf_clr(rad_curated_5, "Rhizaria") +
    toptaxa_depth_clr(rad_curated_5, "Rhizaria") +
      toptaxa_surf_clr(gyrista_curated_5, "Gyrista") +
      toptaxa_depth_clr(gyrista_curated_5,  "Gyrista") +
  plot_layout(nrow = 2, ncol = 4, guides = "collect")
  
```

```{r, fig.height=8, fig.width=12}
toptaxa_depth_clr(dinos_curated_5,  "Dinoflagellates") +
  toptaxa_depth_clr(rad_curated_5, "Rhizaria") +
  toptaxa_depth_clr(gyrista_curated_5,  "Gyrista") +
  toptaxa_depth_clr(ciliate_curated_5, "Ciliates") +
  plot_layout(nrow = 1, ncol = 4, guides = "collect")
# ggsave("figures/depth-clr-main-four.svg", width = 12, height = 8, device = "svg", limitsize = FALSE)
```

# RDA analysis

RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it. Null H = variances across groups are equal p-value of \< 0.05 means there is a significant difference in variance among groups.

```{r}
# library(car)
# Levene's test with one independent variable
# leveneTest(weight ~ group, data = PlantGrowth)
# Levene's test with multiple independent variables
# leveneTest(len ~ supp*dose, data = ToothGrowth)
```

```{r}
set.seed(1234)
head(table_wstn_wdepth)
# head(table_wfeature)
head(table_supp)
```

## All

Isolate categorical vs. numeric metadata. For numeric data, create a normalized scale.

```{r}
head(table_wstn_wdepth)
head(table_supp)
```

```{r}
# Make categorical data table from metadata (Table 1)
categorical <- table_wstn_wdepth %>% 
  select(STN_NISKIN, CLUSTER, WATER_MASS, STN_ORDER, TRANSECT, OilRig_Count, DAY_NIGHT, HR_OF_DAY, ends_with("_THRESHOLD")) %>% distinct() %>% 
  arrange(STN_NISKIN)

# Generate a matrix of numerical data from metadata
metadata_for_rda_0 <- table_supp %>% 
  # Ratios for consideration
  mutate(SIL_NO3 = (SIL/NO3)) %>% 
  mutate(T_S = (Temperature/Salinity)) %>% 
  select(STN_NISKIN, Depth, DIST_OUTFLOW, DIC, NH4, NO2, NO3, Oxygen_CTD, PO4, SIL, Salinity, TA, Temperature, pH, SIL_NO3, T_S) %>% distinct() %>% 
  mutate_if(is.numeric, list(~na_if(., Inf))) %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>%
  as.matrix
```

Normalize matrix of numeric metadata then re-combine with categorical metadata.

```{r}
# Normalize
standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")

# Combine normalized metadata with categorical data
metadata_normalized <- categorical %>% 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN")
head(metadata_normalized)

# Output from here should be the same rownames for ASV table below.
rownames(metadata_normalized) 
```

Generate a matrix from ASV counts

```{r}
# head(asv_long_avg_wtax_clean)
asv_wide_avg_reps_matrix <- asv_long_avg_wtax_clean %>% 
  ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix()

rownames(asv_wide_avg_reps_matrix)
```

```{r}
clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_reps_matrix))

# Confirm that all 

asv_list <- colnames(asv_wide_avg_reps_matrix)
# asv_list
meta_list <- rownames(metadata_normalized)
clr_list <- rownames(clr_asv_matrix)
setdiff(meta_list, clr_list)
```

Run RDA analysis
```{r}
# surf_table <- table_supp %>% 
#   filter(grepl("Above",DCM_THRESHOLD)) %>% 
#   ungroup()
# surf_subset <- as.character(unique(surf_table$STN_NISKIN))
# surf_subset
```

```{r}
# | eval: false
# rda_all_gom <- vegan::rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
# rda_all_gom
# ?rda
```

```{r}
# | eval: false
# adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
# # RsquareAdj(rda_all_gom)
# adj_rda_out
# anova_out <- anova.cca(rda_all_gom, step = 100, by = "term")
# ?anova.cca()
```

Large f-ratio means there is a lot of between group variance compared to within-group variance (independent variables have a greater influence on depdendent variables) Smaller f-ratios indicate the groups are similar to one another. So differences in groups cannot be differentiated from "random" Variance indicates the overall spread of data

## RDA function

First create a function for running RDA analysis.

```{r}
asv_list <- colnames(asv_wide_avg_reps_matrix)
# asv_list
meta_list <- rownames(metadata_normalized)
clr_list <- rownames(clr_asv_matrix)
save(table_supp, table_wstn_wdepth, metadata_normalized, asv_long_avg_wtax_clean, file = "RDA_Robjects_07012026.RData")
```

```{r}
head(metadata_normalized)
```



```{r}
# all_samples <- table_supp %>% ungroup()
# all_samples_list <- as.character(unique(table_supp$STN_NISKIN))
# asvs_all_samples <- asv_long_avg_wtax_clean %>%
#   filter(STN_NISKIN %in% all_samples_list) %>% ungroup()
# 
# all_rda <- run_rda(all_samples, asvs_all_samples,
#                    "All samples", features = "include")

# surf_table <- table_supp %>% 
#   filter(grepl("Above",DCM_THRESHOLD)) %>% 
#   ungroup()
# surf_subset <- as.character(unique(surf_table$STN_NISKIN))
# surf_asv <- asv_long_avg_wtax_clean %>%
#   filter(STN_NISKIN %in% surf_subset) %>% ungroup()
# above_dcm_rda <- run_rda(surf_table, surf_asv, "Above DCM", features = "exclude")
```


## RDA compile from HPRC

`fxn-runRDA_1000_HPRC.R` `RDA_Robjects_06132026.RData`

```{r}
# # install.packages("gt")
# library(gt)
```

```{r}
load("output-tables/df_rda_output_06162026.RData", verbose = TRUE)
```

```{r}
head(df_rda_results)
```

```{r}
rda_summary_table <- gt(df_rda_results %>% 
     select(-FORMULA, -Residual, `Input data` = TEST,
            `Percent constrained` = PERC_CONSTRAINED,
            `Adjusted R-squared` = ADJ_R_SQUARED,
            `MLD` = MLD_THRESHOLD, `DCM` = DCM_THRESHOLD, 
            `Salinity maximum` = SAL_THRESHOLD, 
            `Oxygen minimum depth` = O2_THRESHOLD, 
            `Depth (m)` = Depth, 
            `Distance to shore` = DIST_OUTFLOW,
            `Water mass` = WATER_MASS, `Oil Rigs` = OilRig_Count, `Time of day` = DAY_NIGHT, `Dissolved oxygen` = Oxygen_CTD, `DIC:TA` = DIC_TA, Temperature, `Temp:salinity` = T_S, `Silicate:nitrate` = SIL_NO3, NH4, NO2, NO3, PO4, pH)) %>% 
            # `Day vs. night` = DAY_NIGHT,
            # `Number of oil rigs` = OilRig_Count)) %>% 
  data_color(columns = -c(`Input data`, `Percent constrained`, `Adjusted R-squared`),
             palette = "Reds", apply_to = c("fill"),
             na_color = "grey30",
             fn = scales::col_bin(
      palette = c("transparent", "#730303", "#d62d2d", "#bd6666", "transparent" ,"transparent"),
      bins = c(0, 0.001, 0.01, 0.05, 0.1, 1),
      na.color = "grey20"
    ))
rda_summary_table
```

```{r}
gtsave(rda_summary_table, filename = "output-tables/RDA-gt-table.html")
gtsave(rda_summary_table, filename = "output-tables/RDA-gt-table.docx")
```

# Indicator Species

## indicspecies

https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html

```{r}
# install.packages("indicspecies")
# library(indicspecies)
```

multipatt is the name of the command from the indicspecies package. The mulitpatt command results in lists of species that are associated with a particular group of samples. If your group has more than 2 categories, multipatt will also identify species that are statistically more abundant in combinations of categories.

The parameters for multipatt are as follows:

the community abundance matrix:"abund" the vector that contains your sample grouping information: "time" the function that multipatt is using to identify indicator species: func = "r.g" the number of permutations used in the statistical test: control = how(nperm=9999)

https://cran.r-universe.dev/indicspecies/doc/manual.html

```{r}
# list_top
range(table_supp$Oxygen_CTD)
```

Isolate the top taxa noted before performing Indicator Species analysis.

```{r}
# | eval: false
# | echo: true
indic_input <- asv_long_avg_wtax_clean %>%
  left_join(table_supp) %>% 
  # filter(Supergroup_simplified != "TSAR-NA") %>% 
  # filter(Supergroup_simplified %in% list_top) %>% 
  filter(!grepl("_X", Supergroup_simplified)) %>%
   # Add additional bins
  mutate(SAL_BIN = case_when(
    Salinity < 29 ~ "< 29",
    Salinity >= 29 & Salinity < 31 ~ "29-31",
    Salinity >= 31 & Salinity < 34 ~ "31-34",
    Salinity >= 34 ~ "> 34")) %>% 
  mutate(O2_BIN = case_when(
    Oxygen_CTD <= 3 ~ "< 3",
    Oxygen_CTD > 3  ~ "> 3")) %>% 
  # Add additional bins
  mutate(TEMP_BIN = case_when(
    Temperature < 5 ~ "< 5",
    Temperature >=5 & Temperature < 15 ~ "5-15",
    Temperature >= 15 & Temperature < 20 ~ "15-20",
    Temperature >= 20 & Temperature < 25 ~ "20-25",
    Temperature >= 25 ~ "> 25")) %>% 
  mutate(NO3_BIN = case_when(
    NO3 < 10 ~ "< 10",
    NO3 >= 10 & NO3 < 20 ~ "10-20",
    NO3 >= 20 ~ "> 20")) %>% 
  mutate(OIL_BIN = case_when(
      OilRig_Count > 0 ~ "Oil Rig",
      TRUE ~ "No oil rig"
    )) %>% select(STN_NISKIN, Station, CLUSTER, DEPTH_BIN, ends_with("_THRESHOLD"), 
         SAL_BIN, TEMP_BIN, NO3_BIN, OIL_BIN, WATER_MASS, OilRig_Count, O2_BIN, 
         MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, names_prefix = "ASV_", values_from = MEAN_REPS_seq, values_fill = 0)
```

"Indval" two species are not strongly correlated and their frequency in the data set is larger than the frequency of sites belonging to the site group, the joint occurrence of the two species has higher positive predictive value for the site group than the two species taken independently

‘d’ of species as the product of the relative frequency and relative average abundance in clusters.

## Fxn summary of indicator species

```{r}
# INV$sign[1]
summary_fxn_indic_spp <- function(clust, abun, param){
  df_inv <- multipatt(abun, clust, 
                      func = "IndVal", 
                      duleg = TRUE, 
                      control = how(nperm = 99))
  #
  prelim <- (as.data.frame(df_inv$sign))
  max_index <- max(prelim$index)
  summary_inv_df <- (as.data.frame(df_inv$sign)) %>% 
      rownames_to_column(var = "FeatureID") %>% 
    pivot_longer(cols = starts_with("s."), names_to = "Indicative_of", values_to = "P_A") %>% 
    filter(P_A == 1) %>% 
    filter(p.value <= 0.05) %>% 
    mutate(indic_index = (index/max_index),
           FeatureID = str_remove_all(FeatureID, "ASV_")) %>% 
    select(Indicative_of, FeatureID, indic_index, p.value, stat) %>% 
    add_column(PARAM_INDV = param)
}
# ?multipatt()
```

`summary_fxn_indic_spp(clust, abun, param)`

```{r}
abun <- indic_input %>% 
  select(STN_NISKIN, starts_with("ASV_")) %>% 
  column_to_rownames(var = "STN_NISKIN")
```

All depths and stations

```{r}
cluster_indic_bin <- summary_fxn_indic_spp((indic_input$CLUSTER), abun, "Cluster")
# cluster_indic_bin

watermass_indic_bin <- summary_fxn_indic_spp((indic_input$WATER_MASS), abun, "Water mass")
# watermass_indic_bin
```


```{r}
tax_annot <- asv_long_avg_wtax_clean %>% 
  select(FeatureID, Supergroup_simplified, SUPERGROUP_CLASS, Taxon) %>% distinct()
```

## Cluster-specific indicator results

```{r}
# head(cluster_indic_bin)
indicator_barplot_cluster <- cluster_indic_bin %>%
  left_join(tax_annot) %>%
  group_by(Supergroup_simplified, Indicative_of, PARAM_INDV) %>%
    summarise(COUNT = n()) %>%
  mutate(indic_cluster = str_replace_all(Indicative_of,"s\\.", "")) %>%
  mutate(cluster_order = factor(indic_cluster, levels = rev(CLUSTER_NUM))) %>%
  mutate(tax_order = factor(Supergroup_simplified, levels = (tax_simplified_list))) %>%
  ggplot(aes(y = (cluster_order), x = COUNT, fill = tax_order)) +
    geom_bar(stat = "identity", position = "stack",
             color = "grey20", width = 0.8, linewidth = 0.3) +
  scale_fill_manual(values = tax_color) +
  # scale_y_discrete(position = "right") +
    scale_y_discrete(position = "left") +
  scale_x_continuous(expand = c(0,0), limits = c(0,300)) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 8),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Number of indicator ASVs", y = "k-means cluster")
```

## Water mass-specific indicator results

```{r}
# head(watermass_indic_bin)
indicator_bar_watermass <- watermass_indic_bin %>%
  left_join(tax_annot) %>%
  group_by(Supergroup_simplified, Indicative_of, PARAM_INDV) %>%
    summarise(COUNT = n()) %>%
  mutate(water_mass = str_replace_all(Indicative_of,"s\\.", "")) %>%
  mutate(water_order = factor(water_mass, levels = rev(water_mass_order))) %>%
  mutate(tax_order = factor(Supergroup_simplified, levels = (tax_simplified_list))) %>%
  ggplot(aes(y = (water_order), x = COUNT, fill = tax_order)) +
    geom_bar(stat = "identity", position = "stack",
             color = "grey20", width = 0.6, linewidth = 0.3) +
  scale_fill_manual(values = tax_color) +
  scale_y_discrete(position = "left") +
  scale_x_continuous(expand = c(0,0), limits = c(0,300)) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 1, size = 8),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Number of indicator ASVs", y = "Water mass")
```
Supplemental figure for water masses and kmean clustering indicator species.
```{r, fig.height=4, fig.width=10}
 (indicator_barplot_cluster + theme(legend.position = "none")) + indicator_bar_watermass + plot_layout(ncol = 2)
ggsave("figures/indicator-2panel.svg", width = 10, height = 4, device = "svg", limitsize = FALSE)
```


# Dinoflagellate & ciliate functional traits

```{r}
tmp <- asv_long_avg_wtax_clean %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") #%>% 
  # filter(Class == "Syndiniales")
length(unique(asv_long_avg_wtax_clean$FeatureID))
length(unique(tmp$FeatureID))

# 10563 # all ASVs
# 3611 # Syndiniales Class
# 5521 # Dinoflagellates
# 5521 / 10563 # Dino percentage of ASVs
# 3611 / 5521 # Syndiniales percentage of dino ASVs
head(asv_long_avg_wtax_clean)
```

```{r}
non_syn_dinos_cil <- asv_long_avg_wtax_clean %>%
  filter((Supergroup_simplified == "TSAR-Dinoflagellata") | (Supergroup_simplified == "TSAR-Ciliophora")) %>% 
  filter(Class != "Syndiniales") %>% 
  select(FeatureID, Supergroup_simplified, Order, Family, Genus, Species) %>% distinct() %>% 
  drop_na()
table(non_syn_dinos_cil$Order)

# Export for manual curation
# write.csv(non_syn_dinos_cil, "non_syn_dinos_cil.csv")
```

Manual curation done, compared to Jones et al. table.

```{r}
manual_curated_fxn <- read.csv("non_syn_dinos_cil.csv")
jones_db <- read.csv("../tax_annotation_proj/TMDv1.1_Jones.csv")

trait_asv <- asv_long_avg_wtax_clean %>% 
  left_join(manual_curated_fxn) %>% 
  mutate(Phylum = case_when(
    Class == "Syndiniales" ~ "Dinoflagellata",
    TRUE ~ Phylum
  ),
  `Grazer.Taxonomic.ID` = case_when(
    Class == "Syndiniales" ~ "Syndiniales (Class)",
    TRUE ~ `Grazer.Taxonomic.ID`
  )) %>% 
  left_join(jones_db) %>% 
  filter((Supergroup_simplified == "TSAR-Dinoflagellata") | (Supergroup_simplified == "TSAR-Ciliophora")) %>% 
  mutate(MORPHOTYPE = case_when(
    Morphotype == "PA" ~ "Parasite",
    Morphotype == "CM" ~ "Constitutive mixotroph",
    Morphotype == "CM*" ~ "Constitutive mixotroph",
    Morphotype == "CM/PA" ~ "Constitutive mixotroph/parasite",
    Morphotype == "GH/PA" ~ "Generalist heterotroph/parasite",
    Morphotype == "GH" ~ "Generalist Heterotroph",
    Morphotype == "SH" ~  "Selective heterotroph",
    Morphotype == "CV" ~  "Generalist, carnivore heterotroph",
    Morphotype == "pSNCM" ~ "Plastid bearing Specialist Non-Constitutive Mixotroph",
    grepl("GNCM", Morphotype) ~ "Generalist non-constitutive mixotroph",
    `Feeding.Mode` == "Autotroph" ~ "Autotroph",
    TRUE ~ Morphotype))
glimpse(trait_asv)
```

```{r}
# write.csv(trait_asv, file = "output-tables/fxntrait-ciliate-dino-summary.csv")
```

## Get stats on functional trait profiles

```{r}
head(trait_asv)

tmp <- trait_asv %>% select(Feeding.Mode, FeatureID, DIVISION) %>% distinct() %>% 
  mutate(assigned = case_when(
    is.na(Feeding.Mode) ~ "no",
    TRUE ~ "yes"
  )) %>% select(-Feeding.Mode, -FeatureID)

table(tmp)
237/(562 + 237)

3716/(1805+3716)
```

```{r}
tmp <- trait_asv %>% 
  # filter(Class == "Syndiniales") %>% 
  filter(Subdivision == "Dinoflagellata") %>% 
  distinct()
length(unique(tmp$FeatureID))

```

### Total ASVs, trophic mode, location

```{r}
head(trait_asv)
unique(trait_asv$Class)
```

```{r, fig.width=9, fig.height=4}
# unique(trait_asv$SUPERGROUP_CLASS)
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

trait_asv %>% 
  left_join(table_wstn_wdepth) %>%
  mutate(TAX_CLASS = case_when(
    Class == "Syndiniales" ~ "Syndiniales",
    TRUE ~ Subdivision
  )) %>% 
  group_by(TAX_CLASS, MORPHOTYPE, DEPTH_BIN) %>% 
  summarise(ASV_TOTAL = n()) %>% 
  drop_na() %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = depth_order, x = ASV_TOTAL, fill = MORPHOTYPE)) + 
  geom_bar(stat = "identity", position = "dodge", color = "grey20") +
  # scale_fill_manual(values = blues) +
  facet_grid(cols = vars(TAX_CLASS), scales = "free") +
  theme_classic() +
  labs(x = "Total ASVs", y = "Depth")
  
```


```{r, fig.width=5, fig.height=4}
trait_asv %>% 
  left_join(table_wstn_wdepth) %>%
  mutate(TAX_CLASS = case_when(
    Class == "Syndiniales" ~ "Syndiniales",
    TRUE ~ Subdivision
  )) %>% 
  group_by(TAX_CLASS, MORPHOTYPE, DEPTH_BIN) %>% 
  summarise(ASV_TOTAL = n()) %>% 
  drop_na() %>% 
  filter(TAX_CLASS == "Ciliophora") %>% 
  mutate(depth_order = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = depth_order, x = ASV_TOTAL, fill = MORPHOTYPE)) + 
  geom_bar(stat = "identity", position = "dodge", color = "grey20") +
  scale_fill_manual(values = c("#F6CD86","#93CBF9", "#6F9A49", "#C97E28", "#808D7E")) +
  scale_x_continuous(position = "top", expand = c(0,0)) +
  # facet_grid(cols = vars(TAX_CLASS), scales = "free") +
  theme_classic() +
  theme(panel.grid.major = element_line(color = "grey80")) +
  labs(x = "Total ASVs", y = "Depth")
  
```

## Explore specific traits within list
```{r}
head(trait_asv)
unique(trait_asv$Order)
# unique(trait_asv$)

```
```{r}
prey_types_df <- trait_asv %>% 
  select(SUPERGROUP_CLASS, Order, Feeding.Mode, MORPHOTYPE, STN_NISKIN, Grazer.Taxonomic.ID, Synonym.Basionym, Unknown, Bacteria, Detritus, Diatoms, Haptophytes, Chlorophytes, Cryptophytes, Raphidophytes, Pelagophytes, Dinoflagellates, Ciliates, Copepods, `Copepod.Eggs...Nauplii`, `Metazoan.Host`, Other) %>% 
  distinct() %>% 
  pivot_longer(cols = c(Unknown, Bacteria, Detritus, Diatoms, Haptophytes, Chlorophytes, Cryptophytes, Raphidophytes, Pelagophytes, Dinoflagellates, Ciliates, Copepods, `Copepod.Eggs...Nauplii`, `Metazoan.Host`, Other), names_to = "PUTATIVE PREY", values_to = "option", values_drop_na = TRUE) %>% 
  mutate(PREY_EVIDENCE = case_when(
    option == "+" ~ 1,
    TRUE ~ NA
  )) %>% 
  filter(!is.na(PREY_EVIDENCE)) %>% select(-option) %>% distinct() %>% 
  left_join(table_wstn_wdepth)
```
Bubble plot putative prey types
```{r}
prey_types_df %>% 
  group_by(SUPERGROUP_CLASS, `PUTATIVE PREY`) %>% 
  summarise(SUM = sum(PREY_EVIDENCE)) %>% 
  ggplot(aes(y = SUPERGROUP_CLASS, x = `PUTATIVE PREY`, fill = `PUTATIVE PREY`)) +
  geom_point(aes(size = SUM)) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 90, size = 8,
                                     hjust = 1, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          legend.position = "right") +
    labs(x = "Prey", y = "Predator")
```

```{r}
prey_types_df %>% 
  group_by(SUPERGROUP_CLASS, `PUTATIVE PREY`, DEPTH_BIN) %>% 
  summarise(SUM = sum(PREY_EVIDENCE)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = `PUTATIVE PREY`, fill = `PUTATIVE PREY`)) +
  geom_point(aes(size = SUM)) +
  facet_grid(cols = vars(SUPERGROUP_CLASS)) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 90, size = 8,
                                     hjust = 1, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          legend.position = "right") +
    labs(x = "", y = "")
```
```{r}
head(trait_asv)
```

## Save Table S2
```{r}
colnames(trait_asv)
table_s2 <- trait_asv |> 
  select(FeatureID, Taxon, Supergroup_simplified, SUPERGROUP_CLASS, Feeding.Mode, Morphotype, MORPHOTYPE) |> 
  distinct() |> 
  group_by(Classification = Supergroup_simplified, Supergroup_class = SUPERGROUP_CLASS, Feeding_mode = Feeding.Mode, Morphotype, MORPHOTYPE) |> 
  summarise(ASV_count = n(),
            Collapsed_lineage = paste(unique(Taxon), collapse = ", "), .groups = "drop")

write.csv(table_s2, file = "output-tables/ciliate_dino_summary_TableS2.csv")
```



# Network analysis

```{r}
head(table_supp)
head(asv_long_avg_wtax_clean)
```

ASVs have to appear in at least 20% of the samples.

```{r}
length(unique(table_supp$STN_NISKIN))
# length((table_supp$STN_NISKIN))
90*0.20
# Samples must be in at least 18 samples

tmp <- asv_long_avg_wtax_clean %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% distinct() %>% 
  group_by(FeatureID) %>% 
  summarise(NUM_SAMPLES = n()) %>% 
  filter(NUM_SAMPLES >= 18)

# hist(tmp$NUM_SAMPLES)

length(tmp$FeatureID)
# length(unique(tmp$FeatureID))

keep_ASVs_analysis <- as.character(unique(tmp$FeatureID))
```

## Import data as phyloseq object

```{r}
# if(!requireNamespace("BiocManager")){
#   install.packages("BiocManager")
# }
# BiocManager::install("phyloseq")
# library(phyloseq)
```

```{r}
phylo_asv_in <- asv_long_avg_wtax_clean %>% 
  filter(FeatureID %in% keep_ASVs_analysis) %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% distinct() %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fill = 0) %>% 
  select((order(colnames(.)))) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix

phylo_tax_in <- asv_long_avg_wtax_clean %>% 
  filter(FeatureID %in% keep_ASVs_analysis) %>% 
  select(FeatureID, Taxon) %>% distinct() %>% 
  select((order(colnames(.)))) %>%
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix
# dim(phylo_tax_in)
# dim(phylo_asv_in)
# head(phylo_tax_in)
# head(phylo_asv_in)

# Feature IDs need to be in exact same order

# Set as phyloseq object
asv_table <- otu_table(phylo_asv_in, taxa_are_rows = TRUE)
tax_table <- tax_table(phylo_tax_in)
phylo_asv_tax <- phyloseq(asv_table, tax_table)
```

## SPIEC-EASI code

```{r}
# save(phylo_asv_tax, file = "gom-asvs-for-network.R")
```

SLURM script needs

```{r}
# module load GCC/13.3.0  OpenMPI/5.0.3 R_tamu/4.4.2
# R --rtamuenvs=spiec_easi_apr26
```

Rscript:

```{r}
# install.packages(c('huge', 'pulsar', 'VGAM'))
# install.packages("../../SpiecEasi-1.1.1.tar.gz")
# if (!require("BiocManager", quietly = TRUE)) install.packages("BiocManager")
# BiocManager::install("SpiecEasi")
# library(SpiecEasi)
# Cross Domain approach
# se_GR <-spiec.easi(list(bac_phy, euk_phy), method = 'mb', nlambda = 40,
# lambda.min.ratio = 1e-2, pulsar.params = list(thresh = 0.05))
# 
# # Check output getStability(se_GR) sum(getRefit(se_GR))/2
# 
# # Extract weighted matrix se.beta <- as.matrix(symBeta(getOptBeta(se_GR)))
# df_beta <- as.data.frame(se.beta)
# 
# # Extract adajency matrix adj.mat <- getRefit(se_GR) df_adj <-
# # as.data.frame(as.matrix(adj.mat))
# 
# # Assign names from original dataframes colnames(df_beta) <-
# colnames(se_GR$est$data) colnames(df_adj) <- colnames(se_GR$est$data)
# row.names(df_adj) <- colnames(se_GR$est$data) row.names(df_beta) <-
# colnames(se_GR$est$data)
# 
# # Save output
# save(df_adj, df_beta, se_GR, file = 'gr-spieceasi-output-20-08-2020.RData')
```

Load spieceasi output:
```{r}
load("input-data/spieceasi-output-04-30-2026.RData", verbose = TRUE)
```
```{r}
head(df_adj)
head(df_beta)
hist(df_adj$`006f5664d25385baa7bc38462247b4eb`)
```
```{r}
# head(tax_key)
tax_A <- tax_key
colnames(tax_A) <- paste("A_", colnames(tax_A), sep = "")
# head(tax_A)
tax_B <- tax_key
colnames(tax_B) <- paste("B_", colnames(tax_B), sep = "")
# head(tax_B)
```


```{r}
head(df_adj)
df_adj %>% 
  rownames_to_column(var = "A_ASV") %>% 
  pivot_longer(cols = -c(A_ASV), names_to = "B_ASV", values_to = "ADJ") %>%
  left_join(tax_A, by = c("A_ASV" = "A_FeatureID")) %>% 
  filter(ADJ == 1)
```

# Session information

```{r}
sessionInfo()
```
