---
title: "Analysis of 18S rRNA metabarcoding"
editor: visual
---

# Contents & introduction


# Set up R environment

Set up R & import starting data
```{r}
# | eval: false
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```


# Import prior data

From sequencing
```{r}
# | eval: false
# save(samples_too_low, env_info, asv_wide_cleaned, asv_long_cleaned, file =  "input-data/asv_wtax_qc_GoM23_062025.RData")
```

Import sequence R objects
```{r}
load(file = "input-data/asv_wtax_qc_GoM23_062025.RData", verbose = TRUE)
# env_info <- read.csv("input-data/metadata-gom-23.csv")
# head(env_info)
```

## Station information, ordering, location

Distance from shore
```{r}
library(geosphere)
# Long, Lat order
outflow_lat <- 28.95
outflow_long <- 89.39
outflow <- c(outflow_long, outflow_lat)
outflow
```


Set up ordering of stations and locations. 
```{r}
stn_info <- read.csv("input-data/manual_stn_classification.csv")

offshore_on_shore_order <- c(1, 2, 3, 4, 5, 9, 8, 7, 6, 10, 11, 12, 13, 14, 15)

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(env_info$DEPTH))))
depth_order_round <- as.character(sort(round(unique(env_info$DEPTH), digits = 0)))

env_info_mod <- env_info %>% 
  left_join(stn_info) %>% 
  mutate(STN_ORDER = factor(Station, levels = offshore_on_shore_order),
         TRANSECT = factor(Station, levels = offshore_on_shore_order, transect_labels),
         DEPTH_ORDER = factor(DEPTH, levels = rev(depth_order), labels = rev(depth_order_round))) %>% 
  mutate(DIST_OUTFLOW = distHaversine(cbind(Longitude, Latitude), cbind(outflow_long, outflow_lat)))
# head(env_info_mod)
# # unique(env_info_mod$DIST_OUTFLOW)
table(env_info_mod$DEPTH)
```

# Curate taxonomic assignment

```{r}
# head(asv_long_cleaned)
# unique(asv_long_cleaned$Domain) # take EUKS
# unique(asv_long_cleaned$Supergroup)
# unique(asv_long_cleaned$Division)
# View(asv_long_cleaned %>% 
#        select(Domain, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct())
```

```{r}
asv_long_clean_wtax <- asv_long_cleaned %>% 
  filter(Domain == "Eukaryota") %>% 
  filter(Supergroup != "nucl") %>% 
  mutate(DIVISION = case_when(
    Subdivision == "X" ~ "Unannotated",
    is.na(Subdivision) ~ "Unannotated",
    TRUE ~ Subdivision
  )) %>% 
  # Goal is to combine Subdivision-Class
  mutate(SUPERGROUP_CLASS = case_when(
    ## High level NAs at the supergroup and division level
    (is.na(Division) & !is.na(Supergroup)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Division) & is.na(Supergroup)) ~ "Eukaryote-Unannotated",
    (is.na(Subdivision)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Class) & !is.na(Subdivision)) ~ paste(Subdivision, "Unannotated", sep = "-"),
    #
    (Subdivision == "X") ~ paste(Division, "Unannotated", sep = "-"),
    (Class == "X") ~ paste(Subdivision, "Unannotated", sep = "-"),
    TRUE ~ paste(Subdivision, Class, sep = "-"))) %>% 
  mutate(Supergroup_simplified = case_when(
    Supergroup == "TSAR" ~ paste("TSAR", Subdivision, sep = "-"),
    Supergroup == "Obazoa" ~ paste("Obazoa", DIVISION, sep = "-"),
    TRUE ~ Supergroup
  )) %>% 
  left_join(env_info_mod) 

# 
# sort(unique(tmp$DIVISION))
# sort(unique(tmp$Subdivision))
# unique(tmp$SUPERGROUP_CLASS)
# head(asv_long_clean_wtax)
```
## Normalization and averaging across replicates

```{r}
colnames(asv_long_clean_wtax)
```


```{r}
asv_long_avg_wtax <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, PROX_TOSHORE, STN_ORDER, STN_CATEGORY_3, STN_CATEGORY_4, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup_simplified, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_ID, stn, nisk, sep = " ", remove = FALSE)

head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$STN_CATEGORY_4)
```
> Create a figure with data points for each station and depth. Size of data point = # of ASVs

```{r, fig.height=10, fig.width=8}
asv_long_avg_wtax %>% 
  group_by(DEPTH_ORDER, STN_ORDER, Station, DEPTH) %>% 
    summarise(SEQ_SUM = sum(MEAN_REPS_seq),
              ASV_COUNT = n()) %>% 
  ggplot(aes(x = STN_ORDER, y = DEPTH, size = ASV_COUNT)) +
    geom_jitter(shape = 21, color = "black", stroke = 2) +
  scale_y_reverse() +
  scale_x_discrete(position = "top") +
     scale_size(range=c(1,12), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    panel.grid.major.x = element_line()
  ) +
  labs(x = "", y = "Depth (m)")
```



Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level. 
```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
asv_long_avg_wtax %>%
  group_by(PROX_TOSHORE, TRANSECT, STN_ORDER, DEPTH_ORDER, SAMPLE_ID, DEPTH, TEMP, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  ggplot(aes(y = (DEPTH_ORDER), x = SUM_SEQ, fill = Supergroup_simplified)) +
    geom_bar(stat = "identity", position = "fill", color = "black") +
  facet_grid(rows = vars(PROX_TOSHORE, STN_ORDER), space = "free", scales = "free") +
  theme_classic() +theme(
    strip.background = element_blank(),
    axis.text = element_text(color = "black")
  )
```

### CAREER Proposal 

Station 1-5 only. Heat map from surface to deep. 

```{r}
head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$Subdivision)
```

```{r}
tax_color  <- c("#db1f7d", "#2c5c1d", "#806f8c",
                "#f5bfb0", "#93c47d","#7c52bf", "#007a7a", 
                "#34495e", "#9b59b6", "#fea02f","#ee429b",
                "#ffe8a4","#ff6f61","#fcf517","#8a0622",
                "#ff7530", "#2986cc", "#e71414","#de6600", 
                "#1abc9c", "#a0b59f", "#d7b6d7", "#23b521", "#5b545c", "black")
```


```{r, fig.width=4, fig.height=7}
asv_long_avg_wtax %>%
  # filter(STN_ORDER %in% c(1:5)) %>% 
  group_by(STN_ORDER, DEPTH_ORDER, TEMP, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  ggplot(aes(y = (DEPTH_ORDER), x = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.8) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(STN_ORDER), scales = "free_y", space = "free", axis.labels = "all_x") +
  # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

```{r, fig.width=6, fig.height=18}
# asv_long_avg_wtax %>%
#   # filter(STN_ORDER %in% c(1:5)) %>% 
#   group_by(STN_ORDER, DEPTH, DEPTH_ORDER, TEMP, Supergroup, Supergroup_simplified) %>%
#   summarise(SUM_SEQ = sum(MEAN_REPS_seq),
#             ASV_COUNT = n()) %>%
#   filter(Supergroup_simplified != "Eukaryota") %>% 
#   filter(Supergroup_simplified != "TSAR-NA") %>% 
#   filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
#   ggplot(aes(x = as.numeric(DEPTH), y = SUM_SEQ, fill = Supergroup_simplified)) +
#   geom_area(stat = "identity", position = "fill", color = "black") +
#   coord_flip() +
#   # geom_bar(stat = "identity", position = "fill", color = "black", width = 0.8) +
#   scale_fill_manual(values = tax_color) +
#   scale_x_reverse() +
#   # scale_x_continuous(expand = c(0,0)) +
#   facet_grid(rows = vars(STN_ORDER), scales = "free_y", space = "free", axis.labels = "all_x") +
#   # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
#   theme_classic() + theme(
#     legend.position = "right",
#     legend.title = element_blank(),
#     legend.text = element_text(color = "black", face = "bold"),
#     strip.background = element_blank(),
#     strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
#     axis.title = element_text(color = "black", face = "bold"),
#     axis.text.y = element_text(color = "black", face = "bold"),
#     axis.text.x = element_text(color = "black", face = "bold")) +
#   guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
#   labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

### JGI CSP

Isolate close to MI River delta samples
```{r}
station_jgi <- c(1, 2, 3, 4, 9, 8, 7)
head(asv_long_clean_wtax)
unique(asv_long_clean_wtax$Supergroup_simplified)
# class(asv_long_clean_wtax$DEPTH)
unique(asv_long_clean_wtax$DIST_OUTFLOW)
unique(asv_long_clean_wtax$STN_CATEGORY_4)

jgi_tmp <- asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) 
# unique(jgi_tmp$DEPTH)
```

### Plot salinity, temperature, ASV total.

```{r, fig.height=3, fig.width=3}
asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>% 
  group_by(Station, DEPTH, FeatureID, Supergroup_simplified, TEMP, SALINITY, OXYGEN, STN_CATEGORY_4, stn) %>% 
  # Average across replicates
  summarise(MEAN_SEQ_COUNT = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # mutate(depth_m = as.numeric(DEPTH)) %>% 
  group_by(Station, DEPTH, TEMP, SALINITY, STN_CATEGORY_4, stn) %>% 
  summarise(ASV_COUNT = n()) %>% 
  ggplot(aes(x = SALINITY, y = TEMP, color = DEPTH, shape = stn)) +
  geom_jitter(color = "black", width = 0.4, size = 2) +
  scale_fill_fermenter(direction = 1) +
  # scale_shape_manual(values = c(0, 1, 2, 15, 16, 17)) +
  scale_x_continuous(position = "top") +
  theme_classic() +
  theme(legend.title = element_blank()) +
  labs(x = "Salinity (ppt)", y = "Temperature °C")
```

```{r}
# unique(asv_long_clean_wtax$stn)
```


```{r, fig.height=4, fig.width=5}
asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  # filter(Station == 1) %>% 
  select(Station, DEPTH, TEMP, SALINITY, OXYGEN, STN_CATEGORY_4, stn) %>% distinct() %>% 
  filter(DEPTH < 500) %>% 
  ggplot(aes(x = SALINITY, y = DEPTH, fill = stn)) +
  # geom_path() +
  geom_jitter(color = "black", width = 0.4, size = 2, shape = 21) +
  # scale_fill_distiller() +
  # scale_fill_fermenter(direction = 1) +
  scale_y_reverse() +
  # scale_shape_manual(values = c(21, 22, 24, 23)) +
  theme_classic() +
  labs(x = "Salinity (ppt)", y = "Depth (m)")
```


Something with diversity.
```{r, fig.width=9, fig.height=5}
asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>% 
  group_by(Station, FeatureID, Supergroup_simplified, DEPTH_ORDER, TRANSECT, stn) %>% 
  # Average across replicates
  summarise(MEAN_SEQ_COUNT = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  group_by(Station, Supergroup_simplified, DEPTH_ORDER, TRANSECT, stn) %>% 
  summarise(ASV_COUNT = n(),
            SEQ_SUM = sum(MEAN_SEQ_COUNT)) %>% 
  filter(ASV_COUNT > 50) %>% 
  filter(SEQ_SUM > 100) %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Provora") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  mutate(STN_ORDER = factor(Station, levels = station_jgi)) %>% 
  ggplot(aes(x = SEQ_SUM, y = DEPTH_ORDER, fill = Supergroup_simplified)) +
  # geom_tile(color = "white") +
  # coord_fixed(ratio = 1) +
  geom_bar(stat = "identity", position = "fill", color = "white", width = 0.7) +
  facet_wrap(vars(STN_ORDER, TRANSECT), ncol = 4, scales = "free") +
  # geom_point(color = "black", width = 0.4, shape = 21) +
  # facet_grid(cols = vars(STN_ORDER, TRANSECT)) +
  scale_fill_brewer(palette = "Dark2") +
  # scale_shape_manual(values = c(21, 22, 24, 23)) +
  theme_classic() +
  theme(strip.background = element_blank()) +
  labs(x = "Sequence relative abundance", y = "Depth (m)")
```


#### Hypotheses to test

* Archaeplastida and opisthokonta found at oil contaminated sites
* ciliates, apusomonadidae, cercoza were sensitive
* polluted sites have less diversity, as they are highly selective to a species that would be tolerant.
* Diversity changes will be related to relative pollution?
* Salt is a primary drive

# Ordinations / dendrograms

### All samples.
Create wide format matrix with sequence counts averaged across replicates (by ASV)
```{r}
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```


```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_reps_matrix))))
# ?prcomp
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
    cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

```{r}
# head(env_info_mod)
```


```{r, fig.height=5, fig.width=5}
blues <- c("#d0d1e6","#a6bddb","#74a9cf","#3690c0", "#0570b0", "#045a8d", "#023858")

data.frame(pca_clr$x, SAMPLE = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, SAMPLE) %>% 
      separate(SAMPLE, c("STN_ORDER", "DEPTH_ORDER"), sep = "_", remove = FALSE) %>% left_join(env_info_mod) %>% 
  mutate(Depth = as.numeric(DEPTH_ORDER)) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, shape = STN_CATEGORY_4, fill = Depth)) +
    geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
    geom_point(size = 2.5, stroke = 0.4) +
    ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
    xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
    scale_shape_manual(values = c(21, 23, 24, 25)) +
    scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
    theme_bw() +
    theme(axis.text = element_text(color="black", size=10),
          legend.title = element_blank(),
          axis.title = element_text(color="black", size=8),
          legend.text = element_text(color = "black", size = 8, face = "bold"),
          plot.margin = margin(2, 1, 2, 1, "cm")) +
    guides(shape = guide_legend(override.aes = list(fill = "black")))
```


### Samples for JGI

Create wide format matrix with sequence counts averaged across replicates (by ASV)
```{r}
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```


```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_reps_matrix))))
# ?prcomp
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
    cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

```{r}
head(env_info_mod)
unique(env_info_mod$DIST_OUTFLOW)
```
Pre-process
```{r}
pca_jgi <- data.frame(pca_clr$x, SAMPLE = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, SAMPLE) %>% 
      separate(SAMPLE, c("STN_ORDER", "DEPTH_ORDER"), sep = "_", remove = FALSE) %>%
  left_join(env_info_mod) %>% 
  mutate(Depth = as.numeric(DEPTH_ORDER)) %>%
  mutate(Salinity = as.numeric(SALINITY))
# view(pca_jgi)
```


```{r, fig.height=4.5, fig.width=4.5}
# blues <- c("#d0d1e6","#a6bddb","#74a9cf","#3690c0", "#0570b0", "#045a8d", "#023858")
blues <- c("#d0d1e6","#74a9cf","#3690c0", "#0570b0", "#045a8d", "#023858")
pca_jgi %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, color = Depth, fill = Depth, shape = as.factor(Station))) +
    geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
    geom_point(size = 2.5, stroke = 0.8) +
    ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
    xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
    scale_shape_manual(values = c(23, 21, 25, 24,
                                  2, 6, 1)) +
    scale_color_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
    scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
    theme_bw() +
    theme(axis.text = element_text(color="black", size=10),
          legend.title = element_blank(),
          axis.title = element_text(color="black", size=8),
          legend.text = element_text(color = "black", size = 8, face = "bold"),
          plot.margin = margin(2, 1, 2, 1, "cm")) +
    guides(shape = guide_legend(override.aes = list(fill = "black")))
```



# RDA analysis
RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it.

### All samples
```{r}
head(env_info_mod)

categorical <- env_info_mod %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, STN_CATEGORY_3, PROX_TOSHORE) %>% distinct()

metadata_for_rda_0 <- env_info_mod %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, PO4, SIL) %>% distinct() %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(SAMPLE_NAME, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")
range(standarize_metadata)
# head(categorical)
metadata_normalized <- categorical %>% 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "SAMPLE_NAME")) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME")
head(metadata_normalized)
```

Isolate all samples from ASV data
```{r}
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
# head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>% 
  as.matrix
  
clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))

rownames(clr_asv_matrix)
rownames(metadata_normalized)
```

```{r}
rda_all_gom <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
rda_all_gom
```
23% of the environmental variables explained changes in community composition for all the samples. 

```{r}
# plot(rda_all_gom, scaling = 2)
```

```{r}
adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
adj_rda_out
anova.cca(rda_all_gom, step = 100, by = "term")
```
> OUTPUT:
anova.cca(rda_all_gom, step = 100, by = "term")
Permutation test for rda under reduced model
Terms added sequentially (first to last)
Permutation: free
Number of permutations: 999

Show in New Window
Loading objects:
  samples_too_low
  env_info
  asv_wide_cleaned
  asv_long_cleaned
Show in New Window
A tibble:6 × 30
TRANSECT
<fctr>
PROX_TOSHORE
<int>
STN_ORDER
<fctr>
STN_CATEGORY_3
<chr>
transect1	1	1	Coastal	
transect1	1	1	Coastal	
transect1	1	1	Coastal	
transect1	1	1	Coastal	
transect1	1	1	Coastal	
transect1	1	1	Coastal	
6 rows | 1-4 of 30 columns
tbl_df
6 x 30
 [1] "Dinoflagellata"    "Gyrista"           "Ciliophora"        "Bigyra"           
 [5] "Metazoa"           "Chlorophyta"       "Choanoflagellata"  "Radiolaria"       
 [9] "Centroplasthelida" "Telonemia"         NA                  "Cercozoa"         
[13] "Haptophyta"        "Cryptophyta"       "Kathablepharida"   "Fungi"            
[17] "Opisthokonta"      "Picozoa"           "Prasinodermophyta" "Rhodophyta"       
[21] "Ancyromonadida"    "Stramenopiles"     "Discoba"           "Apusomonada"      
[25] "Ichthyosporea"     "Apicomplexa"       "X"                 "Chrompodellids"   
[29] "Perkinsea"         "Collodictyonidae"  "Alveolata"         "Streptophyta"     
[33] "Discosea"          "Euglenozoa"        "Nebulidia"         "Tubulinea"        
[37] "Nibbleridia"       "Evosea"           
R Console
 [1] "Dinoflagellata"    "Gyrista"           "Ciliophora"        "Bigyra"           
 [5] "Metazoa"           "Chlorophyta"       "Choanoflagellata"  "Radiolaria"       
 [9] "Centroplasthelida" "Telonemia"         NA                  "Cercozoa"         
[13] "Haptophyta"        "Cryptophyta"       "Kathablepharida"   "Fungi"            
[17] "Opisthokonta"      "Picozoa"           "Prasinodermophyta" "Rhodophyta"       
[21] "Ancyromonadida"    "Stramenopiles"     "Discoba"           "Apusomonada"      
[25] "Ichthyosporea"     "Apicomplexa"       "X"                 "Chrompodellids"   
[29] "Perkinsea"         "Collodictyonidae"  "Alveolata"         "Streptophyta"     
[33] "Discosea"          "Euglenozoa"        "Nebulidia"         "Tubulinea"        
[37] "Nibbleridia"       "Evosea"           
Show in New Window
`summarise()` has grouped output by 'STN_ORDER', 'DEPTH_ORDER', 'TEMP', 'Supergroup'.
You can override using the `.groups` argument.
R Console


Show in New Window
 [1]   2.482  18.603 353.454   2.900 269.068 149.888  24.170   2.579 401.190 150.128
[11]  64.991  29.003  25.322   2.231  70.191  59.472   9.882   2.746 399.476  98.397
[21]  15.176   2.641 498.597 109.038  76.433  45.819  65.347   2.195
Show in New Window
`summarise()` has grouped output by 'Station', 'DEPTH', 'FeatureID',
'Supergroup_simplified', 'TEMP', 'SALINITY', 'OXYGEN', 'STN_CATEGORY_4'. You can
override using the `.groups` argument.
`summarise()` has grouped output by 'Station', 'DEPTH', 'TEMP', 'SALINITY',
'STN_CATEGORY_4'. You can override using the `.groups` argument.
R Console

Warning	
G2;H2;Warningh: [38;5;251mThe shape palette can deal with a maximum of 6 discrete values because more than 6 becomes difficult to discriminate [36mℹ[38;5;251m you have requested 7 values. Consider specifying shapes manually if you need that many of them.[39mg
G2;H2;Warningh: [38;5;251mRemoved 2 rows containing missing values or values outside the scale range (`geom_point()`).[39mg

Show in New Window
 [1] "S15" "S1"  "S11" "S12" "S14" "S2"  "S3"  "S4"  "S5"  "S7"  "S8"  "S9" 
Show in New Window

Show in New Window
`summarise()` has grouped output by 'Station', 'FeatureID', 'Supergroup_simplified',
'DEPTH_ORDER', 'TRANSECT'. You can override using the `.groups` argument.
`summarise()` has grouped output by 'Station', 'Supergroup_simplified', 'DEPTH_ORDER',
'TRANSECT'. You can override using the `.groups` argument.
R Console


Show in New Window
`summarise()` has grouped output by 'TRANSECT', 'STN_ORDER', 'DEPTH_ORDER',
'FeatureID', 'stn', 'nisk', 'Station', 'Niskin', 'DEPTH', 'TEMP', 'SALINITY',
'OXYGEN', 'NH4', 'NO2', 'NO3', 'PO4', 'SIL', 'Taxon', 'DIVISION', 'SUPERGROUP_CLASS',
'Supergroup', 'Division', 'Subdivision'. You can override using the `.groups`
argument.
Show in New Window
`summarise()` has grouped output by 'TRANSECT', 'STN_ORDER', 'DEPTH_ORDER',
'FeatureID', 'stn', 'nisk', 'Station', 'Niskin', 'DEPTH', 'TEMP', 'SALINITY',
'OXYGEN', 'NH4', 'NO2', 'NO3', 'PO4', 'SIL', 'Taxon', 'DIVISION', 'SUPERGROUP_CLASS',
'Supergroup', 'Division', 'Subdivision'. You can override using the `.groups`
argument.
Show in New Window

Show in New Window

Show in New Window
 [1] 265570.05        NA  13533.78 111118.32 147652.32 216730.38  41762.97  80694.98
 [9] 119139.90 162941.99 118969.16  79107.16  49021.73
Show in New Window
 [1] "11_1001" "11_1476" "11_2"    "11_31"   "11_351"  "11_601"  "11_74"   "11_800" 
 [9] "12_1200" "12_1400" "12_1599" "12_1799" "12_1999" "12_2"    "12_2245" "12_25"  
[17] "12_62"   "14_115"  "14_1298" "14_1647" "14_199"  "14_1996" "14_2344" "14_2827"
[25] "14_3"    "14_399"  "14_599"  "14_70"   "14_948"  "15_104"  "15_3"    "15_300" 
[33] "15_498"  "15_75"   "15_899"  "1_19"    "1_2"     "2_150"   "2_24"    "2_269"  
[41] "2_3"     "2_353"   "3_1098"  "3_1234"  "3_150"   "3_29"    "3_3"     "3_401"  
[49] "3_550"   "3_65"    "3_698"   "3_898"   "4_1250"  "4_1500"  "4_1724"  "4_2"    
[57] "4_25"    "4_70"    "4_750"   "4_999"   "5_1250"  "5_1498"  "5_1796"  "5_2"    
[65] "5_21"    "5_2198"  "5_2387"  "5_60"    "5_999"   "7_10"    "7_1148"  "7_1399" 
[73] "7_1648"  "7_3"     "7_399"   "7_59"    "7_649"   "7_899"   "7_98"    "8_109"  
[81] "8_15"    "8_3"     "8_46"    "8_499"   "8_648"   "8_76"    "8_800"   "8_925"  
[89] "9_2"     "9_65"   
Show in New Window

Call: rda(formula = clr_asv_matrix ~ STN_CATEGORY_3 + PROX_TOSHORE + DEPTH + NH4
+ NO2 + OXYGEN + PO4 + SALINITY + SIL + TEMP, data = metadata_normalized,
na.action = na.exclude)

                Inertia Proportion Rank
Total         2396.5883     1.0000     
Constrained    549.5480     0.2293   11
Unconstrained 1847.0402     0.7707   78

Inertia is variance

Eigenvalues for constrained axes:
  RDA1   RDA2   RDA3   RDA4   RDA5   RDA6   RDA7   RDA8   RDA9  RDA10  RDA11 
202.19  80.73  55.37  37.65  34.48  32.77  26.83  22.46  20.65  19.25  17.18 

Eigenvalues for unconstrained axes:
   PC1    PC2    PC3    PC4    PC5    PC6    PC7    PC8 
107.98  86.85  65.03  58.32  52.04  49.71  43.79  42.05 
(Showing 8 of 78 unconstrained eigenvalues)

Show in New Window
Description:df [6 × 10]
 
 
STN_CATEGORY_3
<chr>
PROX_TOSHORE
<int>
DEPTH
<dbl>
NH4
<dbl>
NO2
<dbl>
11_1001	Coastal	2	0.4450340	-0.005570151	-0.48008013	
11_1476	Coastal	2	1.0987758	-1.059178312	0.05088407	
11_2	Coastal	2	-0.9298016	-0.821266792	-0.48008013	
11_31	Coastal	2	-0.8904008	-0.277469031	-0.48008013	
11_351	Coastal	2	-0.4499153	-0.549367911	-0.48008013	
11_601	Coastal	2	-0.1051141	-0.787279432	-0.48008013	
6 rows | 1-6 of 10 columns
Show in New Window
`summarise()` has grouped output by 'TRANSECT', 'STN_ORDER', 'DEPTH_ORDER',
'FeatureID', 'stn', 'nisk', 'Station', 'Niskin', 'DEPTH', 'TEMP', 'SALINITY',
'OXYGEN', 'NH4', 'NO2', 'NO3', 'PO4', 'SIL', 'Taxon', 'DIVISION', 'SUPERGROUP_CLASS',
'Supergroup', 'Division', 'Subdivision'. You can override using the `.groups`
argument.
Show in New Window
[1] 0.1206165
Permutation test for rda under reduced model
Terms added sequentially (first to last)
Permutation: free
Number of permutations: 999

Model: rda(formula = clr_asv_matrix ~ STN_CATEGORY_3 + PROX_TOSHORE + DEPTH + NH4 + NO2 + OXYGEN + PO4 + SALINITY + SIL + TEMP, data = metadata_normalized, na.action = na.exclude)
               Df Variance      F Pr(>F)    
STN_CATEGORY_3  2    66.62 1.4067  0.012 *  
PROX_TOSHORE    1    23.51 0.9927  0.445    
DEPTH           1   117.02 4.9416  0.001 ***
NH4             1    30.17 1.2742  0.083 .  
NO2             1    63.66 2.6882  0.001 ***
OXYGEN          1    47.79 2.0183  0.002 ** 
PO4             1    80.07 3.3813  0.001 ***
SALINITY        1    49.62 2.0954  0.003 ** 
SIL             1    35.85 1.5140  0.016 *  
TEMP            1    35.24 1.4883  0.013 *  
Residual       78  1847.04                  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1


## Coastal and surface samples

RDA for JGI call, shallower, and coastal! 

```{r}
head(env_info_mod)
station_jgi

surf_categorical <- env_info_mod %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, PROX_TOSHORE) %>% distinct()

metadata_for_rda_0_surface <- env_info_mod %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, TEMP, SALINITY, OXYGEN, NH4, NO2, PO4, SIL, DIST_OUTFLOW) %>% distinct() %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(SAMPLE_NAME, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

# rownames(metadata_for_rda_0_surface)
standarize_metadata_surf <- decostand(metadata_for_rda_0_surface, MARGIN = 2, method = "standardize")
# range(standarize_metadata)

metadata_normalized_surf <- surf_categorical %>% 
  right_join(as.data.frame(standarize_metadata_surf) %>% rownames_to_column(var = "SAMPLE_NAME")) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME")

rownames(metadata_normalized_surf)

subset_sample_surf <- as.character(unique(surf_categorical$SAMPLE_NAME))
```


```{r}
# head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0_surf <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  filter(SAMPLE_NAME %in% subset_sample_surf) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>% 
  as.matrix
  
clr_asv_matrix_surf <- data.frame(compositions::clr(asv_wide_avg_rda_0_surf))
# station_jgi
rownames(clr_asv_matrix_surf)
rownames(metadata_normalized_surf)
```

```{r}
rda_surf_gom <- rda(clr_asv_matrix_surf ~ ., data = metadata_normalized_surf, na.action = na.exclude)
rda_surf_gom
```
60% of the environmental variables explained changes in community composition for the surface samples.

```{r}
adj_rda_surf <- RsquareAdj(rda_surf_gom)$adj.r.squared
adj_rda_surf
anova.cca(rda_surf_gom, step = 100, by = "term")
```


# Diversity - JGI

Set up core dataframe.

```{r}
head(asv_long_clean_wtax)
```



```{r}
info_jgi <- asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>%
  select(Station, DEPTH, Supergroup_simplified, SUPERGROUP_CLASS, TEMP, SALINITY, OXYGEN, DIST_OUTFLOW, stn, PROX_TOSHORE, DEPTH_ORDER) %>% 
  unite(SAMPLE_NAME, Station, DEPTH, sep = ";") %>% 
  distinct()

# unique(info_jgi$Supergroup_simplified)
# sort(unique(info_jgi$SUPERGROUP_CLASS))

rm <- c("TSAR-NA", "Eukaryota", "Obazoa-Unannotated", "Obazoa-Metazoa")
rm_2 <- c("Fungi-Fungi", "Ichthyosporea-Ichthyophonida", 
          "Chrompodellids-Colpodellidea", "Perkinsea-Perkinsida")

jgi_taxa_df <- asv_long_clean_wtax %>% 
  filter(Station %in% station_jgi) %>% 
  filter(DEPTH < 500) %>% 
  filter(!(Supergroup_simplified %in% rm)) %>% 
  filter(!grepl("-Unannotated", SUPERGROUP_CLASS)) %>% 
  filter(!(SUPERGROUP_CLASS %in% rm_2)) %>% 
  group_by(Station, DEPTH, FeatureID, Supergroup_simplified, SUPERGROUP_CLASS, Order, TEMP, SALINITY, OXYGEN, DIST_OUTFLOW, stn) %>% 
  # Average across replicates
  summarise(MEAN_SEQ_COUNT = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Average to taxa group of interest
  unite(SAMPLE_NAME, Station, DEPTH, sep = ";") %>%
  group_by(SAMPLE_NAME, SUPERGROUP_CLASS) %>% 
  summarise(MEAN_TAX_SEQ = mean(MEAN_SEQ_COUNT)) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_TAX_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SUPERGROUP_CLASS") %>% 
  as.matrix()

jgi_taxa_df_clr <- compositions::clr(t(jgi_taxa_df))

depth
```

```{r, fig.height=6, fig.width=9}
as.data.frame(jgi_taxa_df_clr) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "SUPERGROUP_CLASS", values_to = "CLR") %>%
  left_join(info_jgi) %>% 
  filter(!is.na(stn)) %>% 
  ggplot(aes(x = SUPERGROUP_CLASS, y = DEPTH_ORDER, fill = CLR)) +
  geom_tile(stat = "identity", color = "grey40", lwd = 0.1) +
    scale_fill_fermenter(type = "div", palette = 5) +
  facet_grid(cols = vars(Supergroup_simplified), rows = vars(stn), scales = "free", space = "free") +
  # coord_equal() +
  theme_classic() +
  theme(legend.title = element_blank(),
    axis.text.x = element_text(angle = 90, color = "black", hjust = 1, vjust = 0.5, size = 8),
    axis.text.y = element_text(color = "black", size = 8),
    strip.text.x = element_text(size = 6, angle = 0, color = "black"),
    strip.text.y = element_text(size = 8, angle = 0, color = "black"),
    strip.background = element_blank()) +
  labs(x = "", y = "Depth (m)")
```


# Explore diversity indices with env

```{r}
head(asv_long_avg_wtax)
```

```{r}
asv_long_avg_wtax %>%
  mutate(NO3_NO2 = (NO3 + NO2)) %>%
  group_by(SAMPLE_ID, stn, nisk, DEPTH, NO3_NO2, PO4) %>%
    summarise(SUM = sum(MEAN_REPS_seq),
              ASV_TOTAL = n()) %>%
  ggplot(aes(x = PO4, y = NO3_NO2, fill = DEPTH)) +
  geom_point(shape = 21, size = 4, color = "black") +
    theme_classic()
```


# Session Information

```{r}
sessionInfo()
```





