---
title: "Northern Gulf"
editor: source
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 3
---

# Set up R environment

Set up R & import starting data

```{r, warning=FALSE, message=FALSE}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

```{r}
load(file = "input-data/gulf-2023-sequence-analysis_10192025.RData", verbose = TRUE)
stn_info <- read.csv("input-data/manual_stn_classification.csv")
```

```{r}
#offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5","S9", "S8", "S7", "S11", "S12", "S14", "S15")
head(env_metadata_toplot)
head(asv_long_wenv_wtax)
# unique(asv_long_wenv_wtax$Station)
stns_in_study <- as.character(unique(asv_long_wenv_wtax$Station))
```

```{r}
glimpse(env_metadata_toplot)
```

# Marine Heat Wave?

# Discharge at the MS River

# Northern Gulf region map

> See work from Meagan.


# Map

## Extract lat long from data

#### Fxn for ther surface params, lat, long

Set up a data frame with all parameters as columns.

```{r}
head(env_metadata_toplot)
# range(env_metadata_toplot$Depth)
# unique(env_metadata_toplot$Station)
```

```{r}
stn_labels <- env_metadata_toplot %>% 
  ungroup() %>% 
  filter(Station %in% stns_in_study) %>% 
  select(Station, Latitude, Longitude, Depth) %>% distinct()

surface_labels <- stn_labels %>% 
  filter(Depth < 5) %>% 
  select(Station, Latitude, Longitude) %>% distinct()
```

## Get map params

```{r}
# install.packages("Rcpp")
# install.packages(c("googleway", "ggrepel",
# "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))

library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
```

## Map of whole Gulf region

```{r}
ggulf <- ggplot(data = world) +
  geom_sf(aes(fill = region_wb)) +
  annotate(geom = "text", x = -90, y = 26, label = "", 
     fontface = "italic", color = "grey22", size = 6) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  theme(legend.position = "none", axis.title.x = element_blank(), 
     axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
     panel.border = element_rect(fill = NA))

ggulf
```

```{r, fig.width=7, fig.height=7}
# MAP
LARGE <- ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show:
  # coord_sf(xlim = c(-90, -86.5), ylim = c(27.5, 30), expand = FALSE) + # Region of study
  coord_sf(xlim = c(-110, -70), ylim = c(5, 45), expand = FALSE) + # Larger region
  # Add data points for stations
  # geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10", alpha = 0.9) + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 8, colour="black"),
    axis.title = element_text(size = 8, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA))
LARGE
```

```{r, fig.width=7, fig.height=7}
# MAP
INSET <- ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show
  coord_sf(xlim = c(-91, -85), ylim = c(27, 31), expand = FALSE) + # Larger region
  # Add data points for stations
  geom_point(data = surface_labels, aes(x = Longitude, y = Latitude), size = 5, shape = 21, color = "grey20", fill = "white") +
  geom_text(data = surface_labels, aes(label = Station, x = Longitude, y = Latitude), size = 3) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10") + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 12, colour="black"),
    axis.title = element_text(size = 12, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
    plot.background = element_blank(),
        panel.border = element_rect(fill = NA))
INSET
```

Map to save

```{r, fig.width=8, fig.height=8}
# Left, bottom, right, top
INSET + inset_element(LARGE, 0.4, 0.4, 1.2, 1, align_to = "plot", clip = FALSE)
# ggsave("figures/map_overview.svg", width = 8, height = 8, device = "svg", limitsize = FALSE)
```

# ODV plots in R

Extract points you want and dimensions. Interpolate, subtract bathymetry if needed and move along from there.

```{r}
# Manually extracted hexidecimal ODV colour palette
ODV_colours <- c("#b0020c", "#d31f2a", "#ffb300","#ffc000",  "#75ab19","#27ab19", "#0db5e6", "#4626bd","#7139fe", "#aa24e0", "#d16cfa", "#d5a0eb")
```

Station points and labels.

```{r}
stn_labels <- env_metadata_toplot %>% 
  filter(Station %in% stns_in_study) %>% 
  select(Station, Latitude, Longitude, Depth) %>% distinct() %>% 
  mutate(TRANSECT = case_when(
    (Station < 6) ~ 1,
    (Station > 5 & Station < 10) ~ 2,
    (Station >= 10) ~ 3
  ))

# stn_labels

# max_stn_num <- max(stn_labels$Station)
stn_order_shore_offshore <- c(1, 2, 3, 4, 5, 9, 8, 7, 6, 10, 11, 12, 13, 15, 16, 17)

surf_labels <- stn_labels %>% 
  ungroup() %>% 
  filter(Depth < 5) %>% 
  select(Station, Latitude, Longitude) %>% distinct() %>% 
  mutate(TRANSECT = case_when(
    (Station < 6) ~ 1,
    (Station > 5 & Station < 10) ~ 2,
    (Station >= 10) ~ 3
  ))

```

Get surface profiles

```{r}
surface_odv <- env_metadata_toplot %>%
  # Isolate the three transects we're interested in and cover the study area.
  filter(Station < 18) %>% 
  mutate(TRANSECT = case_when(
    (Station < 6) ~ 1,
    (Station > 5 & Station < 10) ~ 2,
    (Station >= 10) ~ 3
  )) %>% 
  filter(Depth < 5 & Depth > 2) %>% 
  group_by(Station, Latitude, Longitude, ENV_VARIABLE, TRANSECT) %>% 
    summarise(SURFACE_VALUE = mean(value)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = ENV_VARIABLE, values_from = SURFACE_VALUE)
```

Get depth profiles

```{r}
depth_odv <- env_metadata_toplot %>%
  # Isolate the three transects we're interested in and cover the study area.
  filter(Station < 18) %>% 
  mutate(TRANSECT = case_when(
    (Station < 6) ~ 1,
    (Station > 5 & Station < 10) ~ 2,
    (Station >= 10) ~ 3
  )) %>% 
  group_by(Station, Latitude, Longitude, Depth, ENV_VARIABLE, TRANSECT) %>% 
    summarise(VALUE = mean(value)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = ENV_VARIABLE, values_from = VALUE)
```

## Depth profiles

### Temperature & Long

```{r}
head(depth_odv)
```

Non-interpolated scatterplot

```{r}
depth_odv %>%
  ggplot(aes(x = Longitude, y = Depth)) +
  geom_point(aes(color = Temperature), size = 2) +
  # geom_point(aes(color = PROK_COUNT), size = 2) +
  scale_colour_gradientn(colours = rev(ODV_colours),
                         limits = c(2, 32)) +
  labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_y_reverse() + scale_x_reverse() +
  theme_classic()
```

Interpolate step

```{r}
library(MBA)
# ?mba.surf
head(depth_odv)
# Interpolate
temp_mba <- mba.surf(depth_odv[c("Longitude", "Depth", "Temperature")], no.X = 300, no.Y = 300, extend = T)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(temp_mba$xyz.est$z) <- list(temp_mba$xyz.est$x, temp_mba$xyz.est$y)
# head(temp_mba)

temp_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c('LONGITUDE', 'DEPTH'), value.name = 'TEMPERATURE') %>% 
  filter(DEPTH > 0) %>%
  mutate(TEMPERATURE = round(TEMPERATURE, 1))

```

```{r, fig.height=3, fig.width=5}
ggplot(data = temp_interpol, aes(x = LONGITUDE, y = DEPTH)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  scale_fill_gradientn(colours = rev(ODV_colours),
                       limits = c(2, 32)) +
  labs(y = "Depth (m)", x = "Longitude (°W)") +
  scale_y_reverse() + scale_x_reverse() +
  theme_classic() +
  coord_cartesian(expand = F)
```

### Surface params, lat, long

```{r}
head(surface_odv)
```

Non-interpolated scatterplot

```{r}
surface_odv %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
  geom_point(aes(color = Temperature), size = 3) +
  scale_colour_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic()
```

Interpolate step

```{r}
library(MBA)
# ?mba.surf
head(surface_odv)

# Interpolate
temp_mba <- mba.surf(surface_odv[c("Longitude", "Latitude", "Temperature")], 
                     no.X = 300, no.Y = 300, extend = F)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(temp_mba$xyz.est$z) <- list(temp_mba$xyz.est$x, temp_mba$xyz.est$y)
# head(temp_mba)
# 
temp_surface_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c("Longitude", "Latitude"), value.name = 'TEMPERATURE') %>% 
  mutate(TEMPERATURE = round(TEMPERATURE, 1))
head(temp_surface_interpol)
```

```{r, fig.height=3, fig.width=6, warning=FALSE, message=FALSE}
surface_temperature <- temp_surface_interpol %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  # Add data points
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  # ggrepel::geom_text_repel(data = surf_labels, aes(x = Longitude, y = Latitude, label = Station), box.padding = 0.1) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = Station),
  position = position_dodge(width = 0.3),
  hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours),
                       breaks = c(24, 26, 28, 30, 32)) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  coord_cartesian(expand = F) +
  theme(axis.text = element_text(size = 10, colour="black"))
surface_temperature
```

##### Temperature with map

```{r, fig.width=6, fig.height=3}
# MAP
ggplot(data = (world %>% filter(subunit == "United States"))) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
  # Set lat long region to show:
  coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
  # Overlay with temperature 
  geom_raster(data = temp_surface_interpol, aes(fill = TEMPERATURE, x = Longitude, y = Latitude)) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
  # Add data points for stations
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  ggrepel::geom_label_repel(data = surf_labels, aes(x = Longitude, y = Latitude, label = Station)) +
  # geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = Station), position = position_dodge(width = 0.1),
  #                                                hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)

```

##### Temperature Surface

```{r, fig.width=6, fig.height=3, warning=FALSE}
# MAP
ggplot(data = (world %>% filter(subunit == "United States"))) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
  # Set lat long region to show:
  coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
  # Overlay with temperature 
  geom_raster(data = temp_surface_interpol, aes(fill = TEMPERATURE, x = Longitude, y = Latitude)) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
  # Add data points for stations
  geom_point(data = surf_labels, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surf_labels, aes(x = Longitude, y = Latitude, label = Station), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)
```

## Fxn for surface params, lat, long

Set up a data frame with all parameters as columns.

```{r}
set.seed(123)
# Put PARAM in quotes
#limits = c(2, 32)
interpolate_lat_long_env <- function(df, PARAM, min, max, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)], 
                     no.X = 300, no.Y = 300, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Latitude"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Latitude, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1))
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature 
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
    # Add data points for stations
    geom_point(data = surf_labels, aes(x = Longitude, y = Latitude)) +
    ggrepel::geom_text_repel(data = surf_labels, 
                             aes(x = Longitude, y = Latitude, label = Station),
                             direction = "y", nudge_x = 0.1, nudge_y = 0) +
    scale_fill_gradientn(colours = rev(ODV_colours), 
                         na.value = NA,
                         limits = c(min, max)) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1,
        legend.background = element_blank(),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 1, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.83,.75))
}
# interpolate_lat_long_env(surface_odv, "SURFACE_TEMP", "Temperature (°C)")
colnames(surface_odv)
```

Parameters to use in function: "NH4" "NO2" "NO3" "Oxygen"\
\[9\] "PO4" "SIL" "Salinity" "Temperature"

```{r, warning=FALSE}
# TEMP: 2,32 (18, 32)
# Salinity: 30, 38 (25, 37)
# o2: 2, 5
surf_temp <- (interpolate_lat_long_env(surface_odv, "Temperature", 25, 32,
                                       "Temperature (°C)")); surf_temp

surf_sal <- interpolate_lat_long_env(surface_odv, "Salinity", 25, 37, "Salinity"); surf_sal

surf_oxy <- interpolate_lat_long_env(surface_odv, "Oxygen", 2, 5, (expression(bold(O[2]*" "*"ml/L")))); surf_oxy
```

now: "NH4" "NO2" "NO3" "PO4" "SIL"

```{r, warning=FALSE}
# expression("S"["in"] ~ "[W" ~ m^-2~"]")
# NH4: 0, 1.5
# NO3: 0, 30
# NO2: 0, 0.5
# PO4: 0, 2
# Sio: 2, 25 (0,15)
surf_nh4 <- interpolate_lat_long_env(surface_odv, "NH4",0, 1.5, expression(bold("NH"["4"]*" µmol/L")))
surf_no2 <- interpolate_lat_long_env(surface_odv, "NO2", 0, 0.5, expression(bold("NO"["2"]*" µmol/L")))
surf_no3 <- interpolate_lat_long_env(surface_odv, "NO3", 0, 30, expression(bold("NO"["3"]*" µmol/L")))
surf_po4 <- interpolate_lat_long_env(surface_odv, "PO4", 0, 2, expression(bold("PO"["4"]*" µmol/L")))
surf_sil <- interpolate_lat_long_env(surface_odv, "SIL", 0, 15, expression(bold("SiO" ["4"] ^"2-"*" µmol/L"))); surf_sil
```

### Compile all surface plots

```{r, fig.height=10, fig.width=19, warning=FALSE}
surf_temp + surf_sal + surf_oxy + surf_sil + surf_nh4 + surf_no3 + surf_no2 + surf_po4 + plot_layout(ncol = 4) + plot_annotation(tag_levels = "A")
# ?plot_annotation
```

## Fxn for depth parameters

```{r}
# df_all <- depth_odv
# transect <- "transect1"
# PARAM <- "TEMP"
# PARAM
# as.name(PARAM)
# TITLE <- "Temperature"
# Put PARAM in quotes

interpolate_depth_long_env <- function(df_all, transect, PARAM, min, max, TITLE){
  param <- as.name(PARAM)
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect) %>% 
    select(Longitude, Depth, param) %>% 
    drop_na()
  #
  df_labels <- stn_labels %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Depth", PARAM)], 
                     no.X = 300, no.Y = 300, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = "black") +
    # Add data points
    geom_point(data = df_labels, aes(x = Longitude, y = Depth)) +
    geom_text(data = df_labels, 
              aes(x = Longitude, y = -80, label = Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_gradientn(colours = rev(ODV_colours), 
                         na.value = NA,
                         limits = c(min, max)) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse() + scale_y_reverse() +
    theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_rect(fill = NA),
          aspect.ratio = 1,
          legend.background = element_blank(),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 0, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.2,0.26))
}

# interpolate_depth_long_env(depth_odv, 1 ,"Temperature", "Temperature (°C)")
```

```{r}
colnames(depth_odv)
# unique(depth_odv$TRANSECT)
```

### Transect 1
```{r}
# TEMP: depth: 2,32 surface: (18, 32)
# Salinity: 30, 38 (25, 37)
# o2: 2, 5
# Sio: 2, 25 (0,15)
# NH4: 0, 1.5
# NO3: 0, 30
# NO2: 0, 0.5
# PO4: 0, 2
```

```{r, fig.height=10, fig.width=20, warning=FALSE}
(interpolate_depth_long_env(depth_odv, 1, "Temperature", 2, 32,"Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, 1, "Salinity", 25, 37, "Salinity")) +
  (interpolate_depth_long_env(depth_odv, 1, "Oxygen", 2, 5, (expression(bold(O[2]*" "*"ml/L"))))) +
  (interpolate_depth_long_env(depth_odv, 1, "SIL", 1, 28, expression(bold("SiO" ["4"] ^"2-"*" µmol/L")))) +
  #
  (interpolate_depth_long_env(depth_odv, 1, "NH4", 0, 1.5, expression(bold("NH"["4"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 1, "NO2", 0, 0.5, expression(bold("NO"["2"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 1, "NO3", 0, 31, expression(bold("NO"["3"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 1, "PO4", 0, 2.5, expression(bold("PO"["4"]*" µmol/L")))) +
    plot_annotation(tag_levels = "A", title = "Transect 1") +
  plot_layout(ncol = 4)
```

### Transect 2

```{r, fig.height=10, fig.width=20, warning=FALSE}
# (interpolate_depth_long_env(depth_odv, 2, "Temperature", "Temperature (°C)")) +
#   (interpolate_depth_long_env(depth_odv, 2, "Salinity", "Salinity")) +
#   (interpolate_depth_long_env(depth_odv, 2, "Oxygen", (expression(bold(O[2]*" "*"ml/L"))))) +
#   (interpolate_depth_long_env(depth_odv, 2, "SIL", expression(bold("SiO" ["4"] ^"2-"*" µmol/L")))) +
#   #
#   (interpolate_depth_long_env(depth_odv, 2, "NH4", expression(bold("NH"["4"]*" µmol/L")))) +
#   (interpolate_depth_long_env(depth_odv, 2, "NO2", expression(bold("NO"["2"]*" µmol/L")))) +
#   (interpolate_depth_long_env(depth_odv, 2, "NO3", expression(bold("NO"["3"]*" µmol/L")))) +
#   (interpolate_depth_long_env(depth_odv, 2, "PO4", expression(bold("PO"["4"]*" µmol/L")))) +
#     plot_annotation(tag_levels = "A", title = "Transect 1") +
#   plot_layout(ncol = 4)
```

### Transect 3

```{r, fig.height=10, fig.width=20, warning=FALSE}
# (interpolate_depth_long_env(depth_odv, 3, "Temperature", "Temperature (°C)")) +
#   (interpolate_depth_long_env(depth_odv, 3, "Salinity", "Salinity")) +
#   (interpolate_depth_long_env(depth_odv, 3, "Oxygen", (expression(bold(O[2]*" "*"ml/L"))))) +
#   (interpolate_depth_long_env(depth_odv, 3, "SIL", expression(bold("SiO" ["4"] ^"2-"*" µmol/L")))) +
#   #
#   (interpolate_depth_long_env(depth_odv, 3, "NH4", expression(bold("NH"["4"]*" µmol/L")))) +
#   (interpolate_depth_long_env(depth_odv, 3, "NO2", expression(bold("NO"["2"]*" µmol/L")))) +
#   (interpolate_depth_long_env(depth_odv, 3, "NO3", expression(bold("NO"["3"]*" µmol/L")))) +
#   (interpolate_depth_long_env(depth_odv, 3, "PO4", expression(bold("PO"["4"]*" µmol/L")))) +
#     plot_annotation(tag_levels = "A", title = "Transect 1") +
#   plot_layout(ncol = 4)
```

# Compiled physical parameters

```{r}
# TEMP: 2,32 (18, 32)
# Salinity: 30, 38 (25, 37)
# o2: 2, 5
# Sio: 2, 25 (0,15)
# NH4: 0, 1.5
# NO3: 0, 30
# NO2: 0, 0.5
# PO4: 0, 2

```


```{r, fig.width=20, fig.height=20, warning=FALSE, message=FALSE}
# Temperature
surf_temp + 
  (interpolate_depth_long_env(depth_odv, 1, "Temperature", 2, 32, "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, 2, "Temperature", 2, 32, "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, 3, "Temperature", 2, 32, "Temperature (°C)")) +
surf_sal + 
  (interpolate_depth_long_env(depth_odv, 1, "Salinity", 25, 37, "Salinity")) +
  (interpolate_depth_long_env(depth_odv, 2, "Salinity", 25, 37, "Salinity")) +
  (interpolate_depth_long_env(depth_odv, 3, "Salinity", 25, 37, "Salinity")) +
surf_oxy +
  (interpolate_depth_long_env(depth_odv, 1, "Oxygen", 2, 5, (expression(bold(O[2]*" "*"ml/L"))))) +
  (interpolate_depth_long_env(depth_odv, 2, "Oxygen", 2, 5, (expression(bold(O[2]*" "*"ml/L"))))) +
  (interpolate_depth_long_env(depth_odv, 3, "Oxygen", 2, 5, (expression(bold(O[2]*" "*"ml/L"))))) +
surf_sil + 
  (interpolate_depth_long_env(depth_odv, 1, "SIL", 1, 28, expression(bold("SiO" ["4"] ^"2-"*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 2, "SIL", 1, 28, expression(bold("SiO" ["4"] ^"2-"*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 3, "SIL", 1, 28, expression(bold("SiO" ["4"] ^"2-"*" µmol/L")))) +
  plot_annotation(tag_levels = "a", title = "") +
  plot_layout(ncol = 4)
# ggsave("figures/env-profiles-all.svg", width = 20, height = 20, device = "svg", limitsize = FALSE)
```

```{r, fig.width=20, fig.height=20, warning=FALSE, message=FALSE}
  surf_nh4 + 
  (interpolate_depth_long_env(depth_odv, 1, "NH4", 0, 1.5, expression(bold("NH"["4"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 2, "NH4", 0, 1.5,expression(bold("NH"["4"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 3, "NH4", 0, 1.5,expression(bold("NH"["4"]*" µmol/L")))) +
  #
  surf_no3 + 
   (interpolate_depth_long_env(depth_odv, 1, "NO3", 0, 31, expression(bold("NO"["3"]*" µmol/L")))) +
   (interpolate_depth_long_env(depth_odv, 2, "NO3", 0, 31, expression(bold("NO"["3"]*" µmol/L")))) +
   (interpolate_depth_long_env(depth_odv, 3, "NO3", 0, 31, expression(bold("NO"["3"]*" µmol/L")))) +
  #
  surf_no2 + 
  (interpolate_depth_long_env(depth_odv, 1, "NO2", 0, 0.5, expression(bold("NO"["2"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 2, "NO2", 0, 0.5, expression(bold("NO"["2"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 3, "NO2", 0, 0.5, expression(bold("NO"["2"]*" µmol/L")))) +
  #
  surf_po4 + 
  (interpolate_depth_long_env(depth_odv, 1, "PO4", 0, 2.8, expression(bold("PO"["4"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 2, "PO4", 0, 2.8, expression(bold("PO"["4"]*" µmol/L")))) +
  (interpolate_depth_long_env(depth_odv, 3, "PO4", 0, 2.8, expression(bold("PO"["4"]*" µmol/L")))) +
  #
  plot_annotation(tag_levels = "a") +
  plot_layout(ncol = 4)
# ggsave("figures/nutrient-profiles-all.svg", width = 20, height = 20, device = "svg", limitsize = FALSE)
```

# Determine depth features

```{r}
tmp <- read.csv("GradGoM23_MasterSpreadsheet_9-26-23.xlsx - Master.csv")
colnames(tmp)
# head(env_metadata_toplot)
```

```{r}
raw_env_data <- tmp %>% 
  select(Station, Niskin, Depth = `Bottle.Depth..m.`, Latitude, Longitude, Date, Time_UTC = `Time..UTC.`, value = `Fluorescence.CTD`) %>% 
  mutate(Station = as.integer(Station),
         Depth = as.numeric(Depth),
         value = as.numeric(value)) %>% 
  add_column(ENV_VARIABLE = "Fluorescence",
             depth_feature = "")
# colnames(raw_env_data)
# colnames(env_metadata_toplot)


stn_list <- as.character(unique(stn_info$Station))

env_metadata_toplot_flor <- rbind(env_metadata_toplot, raw_env_data) %>% 
  filter(Station %in% stn_list)
# glimpse(raw_env_data)
# glimpse(env_metadata_toplot)
```

```{r}
head(env_metadata_toplot_flor)
unique(env_metadata_toplot_flor$Station)
length(unique(env_metadata_toplot_flor$Station))
```

## O2 min

Assign oxygen mininum as the depth with the lowest oxygen below 30 meters.
```{r}
o2_min <- env_metadata_toplot_flor %>% 
  filter(ENV_VARIABLE == "Oxygen") %>% 
  filter(Depth > 30) %>% 
  group_by(Station) %>% 
  mutate(MIN = min(value)) %>% 
  filter(MIN == value) %>% 
  select(-MIN) %>% 
  add_column(FEATURE = "Oxygen minimum")
head(o2_min)
```

## DCM
```{r}
unique(env_metadata_toplot_flor$ENV_VARIABLE)
```

Assign oxygen minimum as the depth with the lowest oxygen below 30 meters.
```{r}
dcm <- env_metadata_toplot_flor %>% 
  filter(ENV_VARIABLE == "Fluorescence") %>% 
  group_by(Station) %>% 
  mutate(MAX = max(value)) %>% 
  filter(MAX == value) %>%
  select(-MAX) %>% 
  add_column(FEATURE = "DCM")
# head(o2_min)
```

## Salt max

```{r}
sal_max <- env_metadata_toplot_flor %>% 
  filter(ENV_VARIABLE == "Salinity") %>% 
  group_by(Station) %>% 
  mutate(MAX = max(value)) %>% 
  filter(MAX == value) %>%
  select(-MAX) %>% 
  add_column(FEATURE = "Salinity maximum")
```

## Depth profiles across transects


```{r}
feature_profiles <- rbind(sal_max, o2_min, dcm) %>% 
  select(FEATURE, Station, Niskin, Depth) %>% drop_na()

# Any duplicates?
length(feature_profiles$Depth)
length(unique(feature_profiles$Depth))
# feature_profiles
# write.csv(feature_profiles, file = "input-data/feature_profiles_depth.csv")
```


```{r}
# dim(env_metadata_toplot_flor)
# head(env_metadata_toplot_flor)
env_metadata_mod <- env_metadata_toplot_flor %>% 
  ungroup() %>% 
  select(-depth_feature) %>% 
  left_join(feature_profiles) %>% 
  left_join(stn_info)
# dim(env_metadata_mod)  
head(env_metadata_mod)
```
```{r, fig.height=3, fig.width=8}
env_metadata_mod %>% 
  ungroup() %>% 
  filter(!is.na(FEATURE)) %>% 
  select(FEATURE, Station, Depth, Transect) %>% 
  distinct() %>% 
  ggplot(aes(x = Station, y = Depth, fill = FEATURE, shape = FEATURE, color = FEATURE)) +
  geom_jitter(size = 2, width = 0.1) +
  scale_color_manual(values = c("#18545e", "#603655", "#c49235")) +
  geom_line() +
  scale_y_reverse() +
  facet_grid(cols = vars(Transect), scales = "free", space = "free") +
  theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_rect(fill = NA),
          legend.background = element_blank(),
          panel.grid.major = element_line(color = "grey90", linetype = "dotdash"),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_blank())
# ggsave("figures/depth_profiles.svg", width = 8, height = 3, device = "svg", limitsize = FALSE)
```


# ASVs on map

```{r}
load("input-data/gulf-2023-sequence-analysis_10252025.RData", verbose = TRUE)
head(asv_long_avg_wtax)
```

## Surface ASV profiles

Surface lat/long
```{r}
surface_asv_odv <- asv_long_avg_wtax %>%
  # Isolate surface & average across replicates
  filter(Depth < 5) %>%
  ungroup() %>%
  group_by(Station, Latitude, Longitude, TRANSECT) %>%
    summarise(ASV_COUNT = n())
```

Interpolate function for ASVs:

```{r}
## Testing
# df <- surface_asv_odv
# PARAM <- "ASV_COUNT"
# TITLE <- "Total ASVs"

# Put PARAM in quotes
interpolate_lat_long_asv <- function(df, PARAM, min, max, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)],
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z,
                                  varnames = c("Longitude", "Latitude"),
                                  value.name = PARAM) %>%
    select(Longitude, Latitude, VALUE = !!PARAM) %>%
    mutate(VALUE = round(VALUE, 1))
head(envs_interpol)
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = NA) +
    # Add data points for stations
    ggrepel::geom_text_repel(data = surf_labels, 
                             aes(x = Longitude, y = Latitude, label = Station),
                             direction = "y", nudge_x = 0.1, nudge_y = 0) +
    geom_point(data = surf_labels, aes(x = Longitude, y = Latitude)) +
    scale_fill_fermenter(type = "seq", palette = "Reds", 
                         direction = 1, na.value = NA,
                         limits = c(min, max)) +
    # scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1,
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.background = element_blank(),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 1, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.78,.75))
}
```


### Total ASVs

```{r, fig.height=4, fig.width=4, warning=FALSE}
surf_all_asv <- interpolate_lat_long_asv(surface_asv_odv, "ASV_COUNT", 1000, 3000, "Total ASVs")
surf_all_asv
```

### Taxa by ASVs

```{r}
bytaxa_surface_asvs <- asv_long_avg_wtax %>%
  # Isolate surface & average across replicates
  filter(Depth < 5) %>%
  ungroup() %>%
  group_by(SUPERGROUP_CLASS, Station, Latitude, Longitude, TRANSECT) %>%
    summarise(ASV_COUNT = n())
```


```{r, fig.height=4, fig.width=4, warning=FALSE}
# Dinoflagellata-Syndiniales
interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), "ASV_COUNT", "Syndiniales ASVs")
```


```{r, fig.height=4, fig.width=4, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
# Dinoflagellata-Syndiniales
interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Gyrista-Bacillariophyceae")), "ASV_COUNT", "Diatom ASVs")


interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Gyrista-Chrysophyceae")), "ASV_COUNT", "Chrysophyte ASVs")


interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Gyrista-Pelagophyceae")), "ASV_COUNT", "Pelagophyceae ASVs")

```
             
"Radiolaria-Polycystinea"           
"Radiolaria-RAD"                    
"Radiolaria-Radiolaria"   
```{r, fig.height=4, fig.width=4, warning=FALSE}
interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Radiolaria-Acantharea")), "ASV_COUNT", "Acantharea ASVs")

interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Radiolaria-Polycystinea")), "ASV_COUNT", "Polycystinea ASVs")

interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Radiolaria-RAD")), "ASV_COUNT", "RAD ASVs")

# interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  # filter(SUPERGROUP_CLASS == "Radiolaria-Radiolaria")), "ASV_COUNT", "Radiolaria ASVs")
```
## Depth ASV profiles

```{r}
depth_asv_odv <- asv_long_avg_wtax %>%
  filter(!is.na(Latitude)) %>% 
  ungroup() %>%
  group_by(Station, Latitude, Longitude, Depth, TRANSECT) %>%
    summarise(ASV_COUNT = n())
head(depth_asv_odv)
# unique(depth_asv_odv$TRANSECT)
```

```{r}
# Put PARAM in quotes
interpolate_depth_long_asv <- function(df_all, transect, PARAM, min, max, TITLE){
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Depth", PARAM)], 
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = NA) +
    # Add data points
    geom_point(data = df, aes(x = Longitude, y = Depth)) +
    geom_text(data = df, 
              aes(x = Longitude, y = -60, label = df$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_fermenter(type = "seq", palette = "Reds", 
                         direction = 1, na.value = NA,
                         limits = c(min, max)) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse() + scale_y_reverse() +
    theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_blank(),
          aspect.ratio = 1,
          legend.background = element_blank(),
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 0, vjust = 1),
        axis.line = element_line(),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.2,0.35))
}

interpolate_depth_long_asv(depth_asv_odv, "transect1","ASV_COUNT", 1000, 3000,"Total ASVs")
```

### Total ASVs

```{r, fig.height=5, fig.width=15, warning=FALSE}
interpolate_lat_long_asv(surface_asv_odv, "ASV_COUNT", 1000, 3000, "Total ASVs") +
interpolate_depth_long_asv(depth_asv_odv, "transect1","ASV_COUNT", 1000, 3000,"Total ASVs") +
  interpolate_depth_long_asv(depth_asv_odv, "transect2","ASV_COUNT", 1000, 3000, "Total ASVs") +
  interpolate_depth_long_asv(depth_asv_odv, "transect3","ASV_COUNT", 1000, 3000, "Total ASVs") +
  plot_layout(nrow = 1) + plot_annotation(tag_levels = c("a"))
# ggsave("figures/asv-map-all.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
### Taxa by ASVs

```{r}
bytaxa_depth_asvs <- asv_long_avg_wtax %>%
  ungroup() %>%
  group_by(SUPERGROUP_CLASS, Station, Latitude, Longitude, Depth, TRANSECT) %>%
    summarise(ASV_COUNT = n())

```


```{r, fig.height=4, fig.width=4, warning=FALSE}
# Dinoflagellata-Syndiniales
interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), 
  "transect1", "ASV_COUNT",
  100, 1500, "Syndiniales ASVs")
```
## ASV profiles by taxa
```{r}
# df_surf <- (bytaxa_surface_asvs %>% 
#   filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales"))
# 
# df <- (bytaxa_depth_asvs %>% 
#   filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales"))
```


```{r, fig.height=5, fig.width=17, warning=FALSE}

tax_whole_asv_profile <- function(SUBSET, MIN, MAX, TITLE){
interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "ASV_COUNT", MIN, MAX, TITLE) +
interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "transect1","ASV_COUNT", MIN, MAX, TITLE) +
  interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "transect2","ASV_COUNT", MIN, MAX, TITLE) +
  interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "transect3","ASV_COUNT", MIN, MAX, TITLE) +
  plot_layout(nrow = 1, guides = "collect") + plot_annotation(tag_levels = c("a"))
}
```

## Dinoflagellate ASV profiles
   
"Dinoflagellata-Syndiniales"          
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Dinoflagellata-Syndiniales", 100, 2000, "Dinoflagellata-Syndiniales")
```
            
"Dinoflagellata-Noctilucophyceae"      
```{r, fig.height=5, fig.width=17, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
tax_whole_asv_profile("Dinoflagellata-Noctilucophyceae", 0, 10, "Dinoflagellata-Noctilucophyceae")
```

"Dinoflagellata-Dinophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
tax_whole_asv_profile("Dinoflagellata-Dinophyceae", 50, 500, "Dinoflagellata-Dinophyceae")
```
"Dinoflagellata-Dinophyta_X"  
```{r, fig.height=5, fig.width=17, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
tax_whole_asv_profile("Dinoflagellata-Dinophyta_X", 0, 10, "Dinoflagellata-Dinophyta_X"  )
```

## Radiolaria ASV profiles

```{r}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
colnames(asv_long_avg_wtax)
asv_long_avg_wtax %>% 
  filter(grepl("Radiolaria", SUPERGROUP_CLASS)) %>% 
  ungroup() %>% 
  group_by(SUPERGROUP_CLASS) %>% 
  summarise(ASV = n())
  
```

"Radiolaria-Polycystinea"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Radiolaria-Polycystinea", 10, 200, "Radiolaria-Polycystinea")
```
"Radiolaria-Acantharea"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Radiolaria-Acantharea", 10, 200, "Radiolaria-Acantharea")
```
"Radiolaria-RAD-A"
"Radiolaria-RAD-B"
"Radiolaria-RAD-C"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Radiolaria-RAD-A", 10, 70, "Radiolaria-RAD-A")
tax_whole_asv_profile("Radiolaria-RAD-B", 10, 70, "Radiolaria-RAD-B")
tax_whole_asv_profile("Radiolaria-RAD-C", 10, 70, "Radiolaria-RAD-C")
```

## Gyrista ASV profiles

```{r}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
colnames(asv_long_avg_wtax)
asv_long_avg_wtax %>% 
  filter(grepl("Gyrista", SUPERGROUP_CLASS)) %>% 
  ungroup() %>% 
  group_by(SUPERGROUP_CLASS) %>% 
  summarise(ASV = n())
  
```

"Gyrista-Chrysophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Chrysophyceae", 0, 20, "Gyrista-Chrysophyceae")
```

"Gyrista-Bacillariophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Bacillariophyceae", 0, 20, "Gyrista-Bacillariophyceae")
```

"Gyrista-Coscinodiscophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Coscinodiscophyceae", 0, 20, "Gyrista-Coscinodiscophyceae")
```

"Gyrista-Dictyochophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Dictyochophyceae", 0, 20, "Gyrista-Dictyochophyceae")
```

# Session Information

```{r}
sessionInfo()
```

