---
title: "2023 Gulf data analysis"
editor: source
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 3
---

# Contents & introduction

## Set up R environment

Set up R & import starting data

```{r, warning=FALSE, message=FALSE}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

Below is direct from `GoM-2023-amplicons.qmd`
```{r}
load("input-data/gulf-2023-sequence-analysis_11192025.RData", verbose = TRUE)
```
## Ordering & factoring

```{r}
offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5","S9", "S8", "S7", "S11", "S12", "S14", "S15")

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(env_metadata_mod$Depth))))

depth_order_round <- as.character(sort(round(unique(env_metadata_mod$Depth), digits = 0)))

depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")

depth_bin_CHARC <- c("..5m","5.25m","25.60m",
                          "60.270m","270.400m","400.800m",
                          "800.1200m","1200.2000m","2000.3000m")
```


## Study stats

```{r}
(unique(asv_long_cleaned_wenv_mod$Station)) # 12 stations
length(unique(asv_long_cleaned_wenv_mod$Station)) # 12 stations
length(unique(asv_long_cleaned_wenv_mod$STN_NISKIN)) # 90 unique samples
# length(unique(asv_long_avg_wtax$STN_NISKIN))
```

```{r}
table1 <- asv_long_cleaned_wenv_mod %>%
  select(STN_NISKIN, Station, Depth, Date, Time_UTC, Latitude, Longitude, Transect, DEPTH_BIN, DIST_OUTFLOW, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature, Prok_count, DAY_NIGHT, HR_OF_DAY, DEPTH_MAX) %>%
  distinct()
# head(table1)
# write.csv(table1, file = "output-tables/table-samplelist.csv")
```


# Sample comparison & taxonomic classification

## PCA analysis

Create wide format matrix with sequence counts averaged across replicates (by ASV)

Station & depth info

```{r}
glimpse(env_metadata_mod)
```


```{r}
stn_depth_info <- env_metadata_mod |> 
  mutate(STN_NISKIN = as.character(paste("S", Station, "_N", Niskin, sep = ""))) %>%
  filter(STN_NISKIN != "S12_N10") %>% 
  mutate(DEPTH_BIN = case_when(
    Depth <= 5 ~ "< 5m",
    Depth > 5 & Depth <=25 ~ "5-25m",
    Depth > 25 & Depth <=60 ~ "25-60m",
    Depth > 60 & Depth <=270 ~ "60-270m",
    Depth > 270 & Depth <=400 ~ "270-400m",
    Depth > 400 & Depth <=800 ~ "400-800m",
    Depth > 800 & Depth <=1200 ~ "800-1200m",
    Depth > 1200 & Depth <=2000 ~ "1200-2000m",
    Depth > 2000 & Depth <=3000 ~ "2000-3000m",
  )) %>% 
  select(-ENV_VARIABLE, -value, -Pressure, -Date, -Time_UTC) %>% 
  distinct()
# glimpse(stn_depth_info)
# head(asv_long_avg_wtax$STN_NISKIN)
```

Make wide format and as a matrix.
```{r}
asv_wide_avg_matrix <- asv_long_avg_wtax %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_matrix))))
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

#### Plot PCA

```{r, fig.height=5, fig.width=6}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

PCA_PLOT <- data.frame(pca_clr$x, STN_NISKIN = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(stn_depth_info) %>% 
  # Factoring
  mutate(depth_order = factor(DEPTH_BIN, levels = depth_bin)) %>% 
  mutate(FEATURE = str_replace_na(FEATURE, replacement = "other")) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = FEATURE, 
             fill = depth_order)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_jitter(stroke = 0.6, size = 2.5, width = 0.8) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(24, 23, 21, 22)) +
  scale_fill_manual(values = blues) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
PCA_PLOT
# ggsave("figures/pca-plot.svg", width = 6, height = 5, device = "svg", limitsize = FALSE)
```

## Euclidean analysis

Start with matrix
```{r}
# head(asv_wide_avg_matrix)[1:4,1:3]
```

```{r}
# normed_gulf <- decostand(asv_wide_avg_matrix, method = "normalize", MARGIN = 2)
# range(normed_gulf) # Should be between 0 and 1.
```
Community distance metric.
```{r}
# dist_normed_gulf <- stats::dist(t(normed_gulf), method = "euclidean")
# rownames(as.data.frame(as.matrix(dist_normed_gulf)))
```

```{r, fig.height=15, fig.width=20}
# dim(table1)
# head(table1)
# orange_heat <- c("#fff2d4", "#f3d38a", "#ffc94a", "#c76549", "#915e0f", "#62410e")
# as.data.frame(as.matrix(dist_normed_gulf)) %>%
#   rownames_to_column(var = "STN_NISKIN") %>% 
#   pivot_longer(cols = -c(STN_NISKIN), names_to = "STN_NISKIN_Y", values_to = "VALUE") %>% 
#   left_join(table1) %>% 
#   separate(STN_NISKIN, into = c("STN", "NISKIN"), sep = "_", remove = FALSE) %>% 
#   mutate(stn_order = factor(STN, levels = offshore_on_shore_order)) %>% 
#   ggplot(aes(x = STN_NISKIN, y = STN_NISKIN_Y, fill = VALUE)) + 
#   geom_tile(stat = "identity", color = "white", lwd = 0.75) +
#   # scale_fill_gradientn(colors = red_darkred) +
#   scale_fill_stepsn(colors = orange_heat, breaks = c(0.1, 0.3, 0.5, 0.7, 0.9)) +
#   scale_x_discrete(position = "top") +
#   coord_equal() +
#   theme_classic()
```



# Combined dendrogram

Take the matrix of all data, divide by column total so dendrogram is based on relative sequence abundance by samples. 

```{r}
library(ggdendro)
# Get relative sequence abundance
rel_abun <- decostand(asv_wide_avg_matrix, MARGIN = 2, method = "total")
# sum((as.data.frame(rel_abun))$S8_N11)

# Calculate the maximum distance between each sample pair (dist( method = "maximum")). Then take dissimilarity structure and assemble using the average hierarchical clustering as a dendrogram.
cluster_gom <- hclust(dist(t(rel_abun), method = "maximum"), method = "average")

# Convert object to a dendrogram object in R
dendro <- as.dendrogram(cluster_gom)
gom_dendro <- dendro_data(dendro, type = "rectangle")
```

```{r, fig.height=18, fig.width=4}
gom_dendro_plot <- ggplot(segment(gom_dendro)) +
  geom_segment(aes(x = x, y = y, xend = xend, 
    yend = yend)) + 
  coord_flip() + 
  scale_x_continuous(expand = c(0, 0.4)) +
  scale_y_reverse(expand = c(0.4, 0), breaks = c(0, 0.2, 0.4, 0.6, 0.8)) +
  geom_text(size = 3, aes(x = x, y = y, label = label, angle = 0, hjust = 0), data = label(gom_dendro)) +
  theme_dendro() + labs(y = "Dissimilarity") +
    theme(axis.text.x = element_text(color = "black", size = 8),
          axis.line.x = element_line(color = "#252525"), 
        axis.ticks.x = element_line(), axis.title.x = element_text(color = "black", 
            size = 8))
# svg('figs/SUPPLEMENTARY-dendrogram-wreps.svg', w = 10, h = 8)
gom_dendro_plot
```

Isolate label names that correspond to dendrogram.

```{r}
tmp <- data.frame(gom_dendro$labels)
class(tmp)
labels_for_dendro <- as.character(tmp$label)
setdiff(labels_for_dendro, stn_depth_info$STN_NISKIN)
```

### Dendrogram annotation

Import Oil rig data.
```{r}
buffer_counts <- read.csv("input-data/OilRig_BufferCount - Sheet1.csv")
# head(buffer_counts)
```

> What other parameters for dendrogram?

Set up metadata for dendrogram annotation.
```{r}
# head(asv_long_avg_wtax)

dendro_annotation <- asv_long_avg_wtax %>%
  select(STN_NISKIN, Station, Niskin, Depth, DEPTH_ORDER, Pressure, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL, Prok_count, STN_CATEGORY_3, STN_CATEGORY_4, TRANSECT, DIST_OUTFLOW, DEPTH_MAX, DEPTH_BIN, DAY_NIGHT, HR_OF_DAY) %>% 
  left_join(buffer_counts) %>% 
  filter(STN_NISKIN %in% labels_for_dendro) %>% 
  # Factor labeling order
  mutate(label_order = factor(as.character(STN_NISKIN), levels = labels_for_dendro)) %>% 
  distinct()
```

```{r, fig.height=12, fig.width=3}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

DEPTH <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = DEPTH_BIN)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = blues) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
DEPTH
```

```{r, fig.height=12, fig.width=3}
DIST <- dendro_annotation %>%
  mutate(dist_outflow = DIST_OUTFLOW/1000) %>% 
  ggplot(aes(y = label_order, x = 1, fill = dist_outflow)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_continuous(option = "magma") +
  scale_fill_viridis_c(option = "magma", direction = -1) +
    # scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "x10^3 km", y = "")
DIST

```
Dendrogram annotation with DCM, O2, and Salt max
```{r}
feature_profiles <- read.csv("input-data/feature_profiles_depth.csv")
```

```{r, fig.height=12, fig.width=2}
PROFILES <- dendro_annotation %>%
  ungroup() %>% 
  left_join(feature_profiles) %>% 
  select(STN_NISKIN, label_order, Station, Niskin, Depth, FEATURE) %>% distinct() %>%
  mutate(VAL = case_when(
    !is.na(FEATURE) ~ 1,
    TRUE ~ 0
  )) %>%
  
  ggplot(aes(y = label_order, x = FEATURE, fill = 1)) +
    geom_point(color = "black", fill = "grey10", shape = 22, size = 4) +
  coord_fixed(ratio = 1) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed"),
    axis.line.x = element_line(color = "black", linewidth = 0.2),
    axis.line.y = element_line(color = "black", linewidth = 0.2),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_text(color = "black", size = 8, angle = 90, hjust = 1, vjust = 0.5)) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
PROFILES

```


```{r, fig.height=12, fig.width=2}
TRANSECT <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = TRANSECT)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  scale_fill_manual(values = c("#e0a8d8", "#9c388d", "#540548")) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
TRANSECT
```

```{r, fig.height=12, fig.width=2}
CATEGORY <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = STN_CATEGORY_4)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  scale_fill_manual(values = c("#e3c69d","#e3ab5d", "#c4770a", "#523711")) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
CATEGORY
```
```{r, fig.height=12, fig.width=2}
OIL <- dendro_annotation %>%
  mutate(OIL_COUNT = as.character(OilRig_Count)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = OIL_COUNT)) +
    geom_point(shape = 22, color = "black", size = 4) +
    scale_fill_manual(values = c("grey90","#e38aa6","#ba4e70","#a13356","#8a1a3d","#54051e"))+
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
OIL
```
```{r, fig.height=12, fig.width=2}
DAY <- dendro_annotation %>%
  ggplot(aes(y = label_order, x = 1, fill = DAY_NIGHT)) +
    geom_point(shape = 22, color = "black", size = 4) +
    scale_fill_manual(values = c("#afa7d1","#413c59"))+
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
DAY
```

```{r, fig.height=12, fig.width=4}
# range(dendro_annotation$DEPTH_MAX)
MAX <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = DEPTH_MAX)) +
    geom_point(shape = 22, color = "black", size = 4) +
    scale_fill_viridis_c(option = "mako", direction = -1, breaks = c(0, 50, 300, 500, 1000, 2000, 3000)) +
scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
MAX
```
```{r, fig.height=12, fig.width=4}
PROK <- dendro_annotation %>% 
  ggplot(aes(y = label_order, x = 1, fill = Prok_count)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_fermenter(na.value = "grey30", palette = "OrRd", direction = 1) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", size = 6),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
PROK
```

# Taxonomy

Richness by depth, station, and feature.

```{r, fig.height=5, fig.width=7}
asv_long_avg_wtax %>% 
  group_by(Station, DEPTH_BIN) %>% 
    summarise(SEQ_SUM = sum(MEAN_REPS_seq),
              ASV_COUNT = n()) %>% 
  mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin))) |> 
  ggplot(aes(x = as.factor(Station), y = depth_bin_order, 
             size = ASV_COUNT)) +
    geom_jitter(shape = 21, color = "black", stroke = 1.5) +
  scale_x_discrete(position = "top") +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(3,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted")
  ) +
  labs(x = "", y = "Depth (m)")
```

```{r}
tax_color  <- c("#db1f7d", "#2c5c1d", "#806f8c",
                "#f5bfb0", "#93c47d","#7c52bf", "#007a7a", 
                "#34495e", "#9b59b6", "#fea02f","#ee429b",
                "#ffe8a4","#ff6f61","#fcf517","#8a0622",
                "#ff7530", "#2986cc", "#e71414","#de6600", 
                "#1abc9c", "#a0b59f", "#d7b6d7", "#23b521", "#5b545c", "black", "white", "grey10", "grey60")


tax_simplified_list <- as.character((unique(asv_long_avg_wtax$Supergroup_simplified)))
names(tax_color) <- tax_simplified_list
```

Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level.

### For dendrogram

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
TAX <- asv_long_avg_wtax %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  group_by(STN_NISKIN, Station, Niskin, Depth, DEPTH_BIN,
           DEPTH_ORDER, Supergroup_simplified) |> 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro),
         tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) %>% 
  # mutate(label_order = str_remove(label_order, "_")) |> 
  ggplot(aes(y = (label_order), x = SUM_SEQ, fill = tax_order)) +
    geom_bar(stat = "identity", position = "fill", 
             color = "grey20", width = 0.7, linewidth = 0.3) +
  scale_fill_manual(values = tax_color) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.line.x = element_line(),
    
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 10),
    axis.text.x = element_text(color = "black", size = 6)) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
TAX

```

### Plot sequence relative abundance

```{r, fig.width=5, fig.height=14}
asv_long_avg_wtax %>%
  group_by(Station, DEPTH_BIN, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin)),
          tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = depth_bin_order, x = SUM_SEQ, fill = tax_order)) +
  geom_bar(stat = "identity", position = "fill", color = "white", width = 0.9) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(Station), scales = "free_y", space = "free", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

Plot ASV richness by taxonomic groups
```{r}
head(asv_long_avg_wtax)
```

```{r, fig.width=22, fig.height=5}
# tax_color
asv_long_avg_wtax %>%
  group_by(DEPTH_BIN, Supergroup_simplified, Class) %>%
  summarise(ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "Eukaryota_X") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin)),
          tax_order = factor(Supergroup_simplified, levels = rev(tax_simplified_list))) |> 
  ggplot(aes(y = depth_bin_order, x = Class, fill = tax_order, size = ASV_COUNT)) +
  geom_point(shape = 21, color = "grey10", alpha = 0.7, stroke = 0.5) +
  scale_fill_manual(values = tax_color) +
  scale_size_continuous(range = c(1,10), breaks = c(1, 10, 50, 100, 500, 1000, 2000)) +
  facet_grid(cols = vars(tax_order), scales = "free_x", space = "free_x", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
    strip.text.y = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 8, color = "black", face = "bold"),
    strip.text.x = element_text(angle = 90, hjust = 0, vjust = 0.5, size = 8, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold", angle = 90, 
                               hjust = 1, vjust = 0, size = 8)) +
  guides(fill = "none",
         size = guide_legend(nrow = 1, shape = 21, color = "black", fill = "black")) +
  labs(x = "", y = "Depth (m)")

# ggsave("figures/tax-dotplot.svg", width = 22, height = 5, device = "svg", limitsize = FALSE)
```


# Diversity indices

Use wide dataframe from above.

```{r}
head(asv_wide_avg_matrix)[,1:3]
```

```{r}
shannon_all <- vegan::diversity(asv_wide_avg_matrix, index = "shannon", MARGIN = 2)
```

```{r, fig.height=20, fig.width=2}
SHANNON <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(dendro_annotation) %>% 
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(x = shannon_all, y = label_order)) +
  geom_point(shape = 21, fill = "grey30", color = "black", size = 2, stroke = 1) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 6),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Shannon Diversity", y = "")
SHANNON
```

```{r}
invsimp_all <- vegan::diversity(asv_wide_avg_matrix, index = "invsimpson", MARGIN = 2)
# head(data.frame(invsimp_all))
```

```{r, fig.height=20, fig.width=2}
INVSIMP <- data.frame(invsimp_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(stn_depth_info) |> 
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(x = invsimp_all, y = label_order)) +
  geom_point(shape = 21, fill = "grey80", color = "black", size = 2, stroke = 1) +
    theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 6),
    panel.grid.major = element_line(color = "grey70", linetype = "dashed")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Inverse Simpson", y = "")
INVSIMP
```

## Plot Dendrogram compilation

```{r, fig.height=13, fig.width=28}
gom_dendro_plot + DEPTH + PROFILES + 
  DAY + MAX + DIST + 
  CATEGORY + TRANSECT + OIL + 
  PROK + SHANNON + INVSIMP + TAX +
  plot_layout(nrow = 1, 
              widths = c(2.5, 1, 1,
                         1, 1, 1,
                         1, 1, 1,
                         1, 1, 1, 1.6))
# ggsave("figures/dendrogram-compiled.svg", width = 28, height = 13, device = "svg", limitsize = FALSE)
```

# Taxa and depth

```{r}
# head(asv_long_avg_wtax)
asv_long_avg_wtax %>%
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>%
  group_by(Class, Depth) %>%
  summarise(ASV_COUNT = n()) %>%
  ggplot(aes(x = ASV_COUNT, y = Depth, fill = Class)) +
  geom_point(stat = "identity", shape = 21, color = "black") +
  scale_y_reverse() +
  theme_classic()
```


# CLR targets & plots

Confirming what core protistan groups are the most prominent members of the community across all of the samples. 
```{r}
# head(asv_long_avg_wtax)
top_taxa <- asv_long_avg_wtax %>% 
  filter(!(grepl("Obazoa", Supergroup_simplified))) %>% 
  group_by(Supergroup_simplified, STN_NISKIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>% 
  ungroup() %>% 
  # Subset the top 2 groups with the highest ASV count
  group_by(STN_NISKIN) %>% 
  slice_max(order_by = ASV_COUNT, n = 3)
  # slice_max(order_by = SUM_SEQ, n = 3)
# top_taxa
list_top <- unique(top_taxa$Supergroup_simplified)
list_top
```

Write a function to go through each of these taxa
```{r}
head(asv_long_avg_wtax)
```
```{r}
# View(asv_long_avg_wtax)
list_top
```

## Dinoflagellates
```{r}
class_dino_mat <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Dinoflagellata-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix

head(class_dino_mat)

class_dino_clr <- vegan::decostand(class_dino_mat, MARGIN = 1, method = "rclr")
##
colnames(class_dino_clr) <- colnames(class_dino_mat)
rownames(class_dino_clr) <- rownames(class_dino_mat)
# data.frame(class_dino_clr)
```

```{r, fig.height=10, fig.width=3}
plot_dino_clr <- data.frame(class_dino_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
plot_dino_clr
```

## Radiolaria
```{r}
list_top
# tmp <- asv_long_avg_wtax %>% 
  # filter(STN_NISKIN != "S12_N10") %>% 
  # filter(Supergroup_simplified == "TSAR-Radiolaria"); unique(tmp$Class)
  
class_rad_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Radiolaria-Unannotated",
    (is.na(Order) & Class == "X") ~ "Radiolaria-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_rad_mat)
class_rad_clr <- vegan::decostand(class_rad_mat, MARGIN = 1, method = "rclr")
##
colnames(class_rad_clr) <- colnames(class_rad_mat)
rownames(class_rad_clr) <- rownames(class_rad_mat)

# data.frame(class_rad_clr)
```

```{r, fig.height=10, fig.width=3}
plot_rad_clr <- data.frame(class_rad_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```


## Ciliate
```{r}
list_top
tmp <- asv_long_avg_wtax %>%
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Ciliophora"); unique(tmp$Class)
  
class_cil_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Ciliophora-Unannotated",
    (is.na(Order) & Class == "X") ~ "Ciliophora-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_cil_mat)
class_cil_clr <- vegan::decostand(class_cil_mat, MARGIN = 1, method = "rclr")
##
colnames(class_cil_clr) <- colnames(class_cil_mat)
rownames(class_cil_clr) <- rownames(class_cil_mat)

# data.frame(class_rad_clr)
```

```{r, fig.height=13, fig.width=5}
plot_cil_clr <- data.frame(class_cil_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```


## Gyrista
```{r}
list_top
tmp <- asv_long_avg_wtax %>%
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Gyrista"); unique(tmp$Class)
  
class_gyrista_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Gyrista-Unannotated",
    (is.na(Order) & Class == "X") ~ "Gyrista-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_gyrista_mat)
class_gyrista_clr <- vegan::decostand(class_gyrista_mat, MARGIN = 1, method = "rclr")
##
colnames(class_gyrista_clr) <- colnames(class_gyrista_mat)
rownames(class_gyrista_clr) <- rownames(class_gyrista_mat)
```

```{r, fig.height=13, fig.width=5}
plot_gyrista_clr <- data.frame(class_gyrista_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```





## Bigyra
```{r}
list_top
tmp <- asv_long_avg_wtax %>%
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Bigyra"); unique(tmp$Class)
  
class_bigyra_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Bigyra-Unannotated",
    (is.na(Order) & Class == "X") ~ "Bigyra-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_bigyra_mat)
class_bigyra_clr <- vegan::decostand(class_bigyra_mat, MARGIN = 1, method = "rclr")
##
colnames(class_bigyra_clr) <- colnames(class_bigyra_mat)
rownames(class_bigyra_clr) <- rownames(class_bigyra_mat)
```

```{r, fig.height=13, fig.width=3}
plot_bigya_clr <- data.frame(class_bigyra_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```




## Haptista
```{r}
# list_top
class_hapto_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Haptista-Unannotated",
    (is.na(Order) & Class == "X") ~ "Haptista-Unannotated",
         TRUE ~ Order)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_hapto_mat)
class_hapto_clr <- vegan::decostand(class_hapto_mat, MARGIN = 1, method = "rclr")
##
colnames(class_hapto_clr) <- colnames(class_hapto_mat)
rownames(class_hapto_clr) <- rownames(class_hapto_mat)

# data.frame(class_hapto_clr)
```

```{r, fig.height=10, fig.width=3}
plot_hapto_clr <- data.frame(class_hapto_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```

Try boxplot version. 
```{r, fig.height=5, fig.width=4.5}
data.frame(class_hapto_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, fill = stn, group = DEPTH_BIN_ORDER)) +
  geom_vline(xintercept = 0, color = "grey40") +
  ggdist::stat_halfeye(side = "top", adjust = 0.5, width = 0.2, 
                       justification = 0, .width = 0, fill = "black", alpha = 0.6,
                       overflow = "compress") +
  # geom_boxplot(width = 0.2, fill = "grey20", color = "grey10",
  #              alpha = 0.6, linewidth = 0.3) +
  ggdist::stat_dots(side = "bottom", layout = "bin",
                    justification = 1.0,
                    overflow = "compress", dotsize = 0.4,
                    fill = "grey40", color = "grey10") +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "", title = "Haptophyte ASVs (clr)")
```


## Archaeplastida
```{r}
class_arch_mat <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "Archaeplastida") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Dinoflagellata-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix

head(class_arch_mat)

class_arch_clr <- vegan::decostand(class_arch_mat, MARGIN = 1, method = "rclr")
##
colnames(class_arch_clr) <- colnames(class_arch_mat)
rownames(class_arch_clr) <- rownames(class_arch_mat)
# data.frame(class_dino_clr)
```

```{r, fig.height=10, fig.width=5}
plot_arch_clr <- data.frame(class_arch_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4, breaks = c(-3,-1,0,1,3)) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
plot_arch_clr
```




## Compile taxa CLR plots
```{r, fig.width=22, fig.height=11}
plot_dino_clr + plot_hapto_clr + plot_rad_clr + plot_cil_clr + plot_gyrista_clr + plot_bigya_clr + plot_layout(nrow = 1, widths = c(1, 2.6, 1.2, 2.6, 3.5, 0.8))
# ggsave("figures/clr-top-taxa.svg", width = 22, height = 11, device = "svg", limitsize = FALSE)
```



# Taxa by depth - curated

```{r}
unique(top_taxa$Supergroup_simplified)
head(asv_long_avg_wtax)
```
## Dinoflagellata
```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Dinoflagellata") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```



## Radiolaria

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Radiolaria") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Acantharea", TAXA) ~ "Acantharea",
    # grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```

## Ciliate

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Ciliophora") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("CONThreeP", TAXA) ~ paste("CONThreeP",Family, sep = ";"),
    grepl("CONTH_", TAXA) ~ "CONTH unannotated",
    grepl("Phyllopharyngea;Cyrtophoria", TAXA) ~ "Phyllopharyngea;Cyrtophoria",
    grepl("Spirotrichea;Choreotrichida", TAXA) ~ "Spirotrichea;Choreotrichida",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```


## Gyrista

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Gyrista") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
  # unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = Class) %>% 
  # mutate(TAXA = case_when(
  #   grepl("Acantharea", TAXA) ~ "Acantharea",
  #   # grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
  #   TRUE ~ TAXA
  # )) %>% 
  # filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```


## Bigyra
```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "TSAR-Bigyra") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
  unite(TAXA, Order, Family, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Labyrinthulomycetes", TAXA) ~ "Labyrinthulomycetes",
    grepl("Pseudodendromonadales", TAXA) ~ "Pseudodendromonadales",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```




## Haptista

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "Haptista") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
  unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  mutate(TAXA = case_when(
    grepl("Pterocystida", TAXA) ~ "Pterocystida",
    grepl("Haptophyta_Clade_", TAXA) ~ "Haptophyta Clades-3/4",
    TRUE ~ TAXA
  )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```




## Archaeplastida

```{r, fig.width=13, fig.height=4}
mat_tmp <- asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "Archaeplastida") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
  unite(TAXA, FeatureID, Family, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  colnames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_violin(width = 0.4, fill = "grey80", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
      geom_jitter(height = 0.04,  alpha = 0.2, shape = 21, fill = "white", color = "black", size = 0.75) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    scale_x_continuous(limits = c(-4, 4)) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right") +
    labs(x = "CLR of all ASVs", y = "", title = "Archaeplastida") +
    ##
    asv_long_avg_wtax %>% 
  filter(Supergroup_simplified == "Archaeplastida") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(DEPTH_BIN) %>% 
      summarise(ASV_COUNT = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = 1, size = ASV_COUNT)) +
  geom_point(shape = 21, fill = "black", color = "black") +
  scale_size_continuous(range = c(2, 10)) +
  theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          axis.ticks = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Total ASVs", y = "") + 
    ##
    asv_long_avg_wtax %>% 
    mutate(TAXA = Class) %>% 
  # unite(TAXA, Class, Order, sep = ";", remove = FALSE) %>% 
  # mutate(TAXA = case_when(
  #   grepl("Pterocystida", TAXA) ~ "Pterocystida",
  #   grepl("Haptophyta_Clade_", TAXA) ~ "Haptophyta Clades-3/4",
  #   TRUE ~ TAXA
  # )) %>% 
  filter(!is.na(Order)) %>% 
  filter(Supergroup_simplified == "Archaeplastida") %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    group_by(TAXA, DEPTH_BIN) %>% 
      summarise(SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
  ggplot(aes(y = DEPTH_BIN_ORDER, x = SEQ_SUM, fill = TAXA)) +
  geom_bar(stat = "identity", color = "black", position = "fill", width = 0.6) +
  theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "right",
          legend.title = element_blank()) +
    labs(x = "Sequence Relative Abundance", y = "") +
    plot_layout(nrow = 1, widths = c(1, 0.2, 1))

```



## Box plot top taxa

```{r}
# glimpse(asv_long_avg_wtax)
# SUPERGROUP_SIMPLIFIED <- "TSAR-Gyrista"
clr_dist_fxn <- function(SUPERGROUP_SIMPLIFIED){
  mat_tmp <- asv_long_avg_wtax %>% 
    filter(Supergroup_simplified == SUPERGROUP_SIMPLIFIED) %>% 
    filter(MEAN_REPS_seq >= 100) %>% 
    ungroup() %>% 
    # unite(FAMILY_GENUS_SPECIES, Family, Genus, Species, sep = ";") %>% 
    unite(TAXA, SUPERGROUP_CLASS, sep = ";") %>% 
    unite(SAMPLE, DEPTH_BIN, Station, sep = "_") %>%
    group_by(TAXA, SAMPLE) %>% 
      summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>% 
    pivot_wider(names_from = TAXA, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
    column_to_rownames(var = "SAMPLE") %>% 
    as.matrix
  # colnames(mat_tmp)
  # rownames(mat_tmp)
  clr_tmp <- vegan::decostand(mat_tmp, MARGIN = 1, method = "rclr")
  ##
  colnames(clr_tmp) <- colnames(mat_tmp)
  rownames(clr_tmp) <- rownames(mat_tmp)
  ##
  data.frame(clr_tmp) %>% 
    rownames_to_column(var = "SAMPLE") %>% 
    pivot_longer(cols = -c(SAMPLE), names_to = "TAXA", values_to = "CLR") %>% 
    # separate(FAMILY_GENUS_SPECIES, into = c("Family", "Genus", "Species"), sep = ";", remove = FALSE) %>% 
    separate(SAMPLE, into = c("DEPTH_BIN", "Station"), sep = "_") %>% 
    # mutate(stn = paste("S", Station, sep = "")) %>% 
    # mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
    mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = CLR, group = DEPTH_BIN_ORDER)) +
    geom_vline(xintercept = 0, color = "grey90") +
    geom_jitter(height = 0.04,  alpha = 0.7, shape = 21, color = "black", size = 0.5) +
    # geom_violin(fill = NA) +
    # ggdist::stat_halfeye(side = "top", adjust = 0.6, width = 0.1,
    #                      justification = -0.1,
    #                      fill = "black", alpha = 0.6) +
    geom_boxplot(width = 0.2, fill = "grey20", color = "grey10",
    alpha = 0.8, linewidth = 0.3) +
    theme_classic() +
    theme(axis.text.x = element_text(color = "black", angle = 0, size = 8,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black", size = 8),
          plot.title = element_text(color = "black", size = 8, face = "bold"),
          panel.grid.major = element_line(color = "grey90", linetype = "dashed"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank(),
          legend.position = "none") +
    labs(x = "CLR", y = "", title = SUPERGROUP_SIMPLIFIED)
}
```

```{r}
top_taxa

unique(top_taxa$Supergroup_simplified)
```

```{r, fig.height=15, fig.width=20}
clr_dist_fxn("TSAR-Dinoflagellata") + clr_dist_fxn("TSAR-Ciliophora") + clr_dist_fxn("TSAR-Bigyra") + clr_dist_fxn("TSAR-Radiolaria") + clr_dist_fxn("Haptista") + plot_layout(ncol = 3)
```




# Linear regression

```{r}
head(asv_long_avg_wtax)
# PARAMS: DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL
```

```{r}
library(broom)
regression_gom_tmp <- asv_long_avg_wtax %>%
  filter(STN_NISKIN != "S12_N10") %>%
  select(Supergroup_simplified, MEAN_REPS_seq, Station, Niskin, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
  group_by(Station, Niskin, Supergroup_simplified, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
summarise(seq_sum = sum(MEAN_REPS_seq),
          asv_rich = n()) %>% 
  # filter(ASV_rich > 1 &)
  pivot_longer(cols = c(DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL), names_to = "PARAMS", values_to = "VALUES") %>% 
  ungroup() 
dim(regression_gom_tmp)
head(regression_gom_tmp)
# unique(regression_gom_tmp$Class)
```

```{r}
part1 <- regression_gom_tmp %>% 
# For taxonomic groups, group linear regression by ASV Richness
 group_by(Supergroup_simplified, PARAMS) %>% 
  nest(-Supergroup_simplified, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
    tidied = map(lm_fit, tidy)) %>% unnest(tidied) %>%
  select(Supergroup_simplified, PARAMS, term, estimate, p.value) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
    select(everything(), SLOPE = VALUES, p.value) %>% data.frame
# ?pivot_wider
```

```{r}
part2 <- regression_gom_tmp %>% 
# For taxonomic groups, group linear regression by ASV Richness
 group_by(Supergroup_simplified, PARAMS) %>% 
  nest(-Supergroup_simplified, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
    glanced = map(lm_fit, glance)) %>% unnest(glanced) %>%
  select(Supergroup_simplified, PARAMS, r.squared, adj.r.squared) %>%
  right_join(part1) %>% 
  right_join(regression_gom_tmp)
```

```{r}
# head(part2)
# unique(part2$Supergroup_simplified)
```

```{r, fig.width=10, fig.height=5}
part2 %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "Ciliate")
```


```{r, fig.width=10, fig.height=5}
part2 %>% 
  filter(Supergroup_simplified == "Haptista") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "Haptista")
```

```{r}
#TSAR-Dinoflagellata
part2 %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "TSAR-Dinoflagellata")
```
## Linear regression function

Where R2 is the proportion of the variation in ASV richness predictable from the environmental parameter supplied. Closer to 1 is a better fit for the relationship. 

`function(df, TITLE)`

```{r}
library(broom)
lm_gom <- function(df, TITLE) {
  df_tmp <- df %>%
  select(TAXA, MEAN_REPS_seq, Station, Niskin, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
  group_by(TAXA, Station, Niskin, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
  summarise(seq_sum = sum(MEAN_REPS_seq),
            asv_rich = n()) %>% 
  pivot_longer(cols = c(DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL), names_to = "PARAMS", values_to = "VALUES") %>% 
  ungroup() 
#
df_tmp_tidy <- df_tmp %>% 
  # For taxonomic groups, group linear regression by ASV Richness
  group_by(TAXA, PARAMS) %>% 
  nest(-TAXA, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
         tidied = map(lm_fit, tidy)) %>% unnest(tidied) %>%
  select(TAXA, PARAMS, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
  select(everything(), SLOPE = VALUES) %>% data.frame

df_lm <- df_tmp %>% 
  # For taxonomic groups, group linear regression by ASV Richness
  group_by(TAXA, PARAMS) %>% 
  nest(-TAXA, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
         glanced = map(lm_fit, glance)) %>% unnest(glanced) %>%
  select(TAXA, PARAMS, r.squared, adj.r.squared) %>%
  right_join(df_tmp_tidy) %>% 
  right_join(df_tmp)

df_lm %>% 
  ggplot(aes(x = VALUES, y = asv_rich)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
                                3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
  theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
        strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black",                                                     size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = TITLE)
}
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  add_column(TAXA = "All ASVs")), "All ASVs")

lm_gom((asv_long_avg_wtax %>% 
          filter(Depth < 10) %>% 
  add_column(TAXA = "Surface ASVs")), "Surface ASVs")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  filter(Class == "Syndiniales") %>% 
  add_column(TAXA = "Syndiniales")), "Syndiniales")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 50) %>% 
  filter(Class == "Dinophyceae") %>% 
  add_column(TAXA = "Dinophyceae")), "Dinophyceae")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Subdivision == "Gyrista") %>% 
  add_column(TAXA = "Gyrista")), "Gyrista")
#Bacillariophyceae
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Bacillariophyceae") %>% 
  add_column(TAXA = "Bacillariophyceae")), "Bacillariophyceae")


lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Syndiniales") %>% 
  add_column(TAXA = "Syndiniales")), "Syndiniales")

lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Dinophyceae") %>% 
  add_column(TAXA = "Dinophyceae")), "Dinophyceae")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  filter(Subdivision == "Ciliophora") %>% 
  add_column(TAXA = "Ciliates")), "Ciliates")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH > 400) %>%
  filter(Subdivision == "Cercozoa") %>% 
  add_column(TAXA = "Cercozoa")), "Cercozoa")

lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH > 400) %>%
  filter(Subdivision == "Radiolaria") %>% 
  add_column(TAXA = "Radiolaria")), "Radiolaria")
```


# RDA analysis

RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it.

```{r}
set.seed(1234)
head(table1)
buffer_counts <- read.csv("input-data/OilRig_BufferCount - Sheet1.csv")
# head(buffer_counts)
features <- read.csv(file = "input-data/feature_profiles_depth.csv")
# head(features)
```

```{r}
# table(features %>% select(FEATURE, Station))
features %>% select(-X, -Niskin) %>% 
  pivot_wider(names_from = FEATURE, values_from = Depth) %>% 
  left_join(buffer_counts)

# head(table1)
table1_wfeature <- table1 %>% 
  left_join((features %>% select(-X, -Niskin) %>% 
  pivot_wider(names_from = FEATURE, values_from = Depth))) %>% 
  left_join(buffer_counts) %>% 
  mutate(SALT_MAX = case_when(
    Depth >= `Salinity maximum` ~ "Below salinity maximum", #At and below the salinity maximum
    Depth < `Salinity maximum` ~ "Above salinity maximum",
  )) %>% 
  mutate(DCM = case_when(
    Depth > DCM ~ "Below DCM",
    Depth <= DCM ~ "Above DCM", #At and above the DCM
  )) %>% 
  mutate(O2_MIN = case_when(
    Depth > `Oxygen minimum` ~ "Below oxygen minimum",
    Depth <= `Oxygen minimum` ~ "Above oxygen minimum", #At and above the O2 minimum
  )) %>% 
  select(-`Oxygen minimum`, -`Salinity maximum`)

head(table1_wfeature)
```

## All

```{r}
# Make categorical data table from metadata (Table 1)
categorical <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  select(STN_NISKIN, Transect, OilRig_Count) %>% distinct()
# str(categorical)
# Generate a matrix of numerical data from metadata
metadata_for_rda_0 <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  select(STN_NISKIN, DIST_OUTFLOW, Depth, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature) %>% distinct() %>% 
  pivot_longer(cols = -c(STN_NISKIN), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(STN_NISKIN, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>%
  as.matrix
```

Normalize matrix of numeric metadata.
```{r}
standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")

# Combine normalized metadata with categorical data
metadata_normalized <- categorical %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN")
# head(metadata_normalized)
rownames(metadata_normalized) # Output from here should be the same rownames for ASV table below.
```
Generate a matrix from ASV counts
```{r}
# head(asv_long_avg_wtax)
# hist(asv_long_avg_wtax$MEAN_REPS_seq)
asv_wide_avg_reps_matrix <- asv_long_avg_wtax %>% 
  # Filter low abundance ASVs
  filter(MEAN_REPS_seq >= 100) %>% 
  ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
# colnames(asv_wide_avg_reps_matrix)
```

```{r}
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix

clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))

# Confirm that all 
asv_list <- rownames(asv_wide_avg_rda_0)
meta_list <- rownames(metadata_normalized)
clr_list <- rownames(clr_asv_matrix)
# setdiff(meta_list, clr_list)
```
Run RDA analysis
```{r}
# | eval: false
rda_all_gom <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
rda_all_gom
```
```{r}
# | eval: false
adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
# RsquareAdj(rda_all_gom)
adj_rda_out
anova.cca(rda_all_gom, step = 100, by = "term")
```

### RDA function
First create a function for running RDA analysis.

```{r}
# table1
# asv_long_avg_wtax

# How to use:
## Modify and subsame table1 or asv_long_avg_wtax completely before use.
## Check to make sure the same STN_NISKIN IDs are present in both.
run_rda <- function(mod_metadata, mod_asv_long){
  # Select categorical data
  # mod_metadata <- surf_table
  # mod_asv_long <- surf_asv
  # 
  # CATEGORIES <- c("Transect") #Character list
  categorical <- mod_metadata %>% 
    select(STN_NISKIN, Transect, OilRig_Count) %>% distinct()
  # Isolate numerical data as matrix
  metadata_for_rda_0 <- mod_metadata %>% 
    select(STN_NISKIN, DIST_OUTFLOW, Depth, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature) %>% distinct() %>% 
    pivot_longer(cols = -c(STN_NISKIN), names_to = "VARIABLES", values_to = "VALUE") %>% 
    group_by(STN_NISKIN, VARIABLES) %>% 
      summarise(MEAN_VAL = mean(VALUE)) %>% 
    pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
    arrange(STN_NISKIN) %>% 
    column_to_rownames(var = "STN_NISKIN") %>% as.matrix
  # Normalize
  standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")
  # Combine normalized numerical data with categorical
  metadata_normalized <- categorical %>% 
    right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
    arrange(STN_NISKIN) %>% 
    column_to_rownames(var = "STN_NISKIN")
  #
  ## Prepare ASV data
  asv_wide_avg_reps_matrix <- mod_asv_long %>% 
    ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
  # Transform
  asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix
  #
  # CLR normalization
  clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))
  #
  # setdiff(rownames(metadata_normalized), (rownames(clr_asv_matrix)))
  # RDA step
  rda_output <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
  rda_output
  return(rda_output)
}
```


### Above | Below DCM 

Generate a matrix from ASV counts
```{r}
surf_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  # Filter low abundance ASVs
  filter(DCM == "Above DCM") %>% ungroup()

surf_subset <- as.character(unique(surf_table$STN_NISKIN))
# surf_subset

surf_asv <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 50) %>% 
  filter(STN_NISKIN %in% surf_subset) %>% ungroup()

# How to use:
## Modify and subsame table1 or asv_long_avg_wtax completely before use.
## Check to make sure the same STN_NISKIN IDs are present in both.
above_dcm_rda <- run_rda(surf_table, surf_asv); above_dcm_rda
# above_dcm_rda
RsquareAdj(above_dcm_rda)$adj.r.squared
anova.cca(above_dcm_rda, step = 100, by = "term")
```
```{r}
below_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  # Filter low abundance ASVs
  filter(DCM != "Above DCM") %>% ungroup()

below_subset <- as.character(unique(below_table$STN_NISKIN))
# surf_subset

below_asv <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 50) %>% 
  filter(STN_NISKIN %in% below_subset) %>% ungroup()

below_dcm_rda <- run_rda(below_table, below_asv); below_dcm_rda

RsquareAdj(below_dcm_rda)$adj.r.squared
anova.cca(below_dcm_rda, step = 100, by = "term")
```


### Above | Below O2 minimum

```{r}
# head(table1_wfeature)
below_02_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  # Filter low abundance ASVs
  filter(O2_MIN == "Below oxygen minimum") %>% ungroup()

below_02_subset <- as.character(unique(below_02_table$STN_NISKIN))
# surf_subset

below_02_asv <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 50) %>% 
  filter(STN_NISKIN %in% below_02_subset) %>% ungroup()

below_02_rda <- run_rda(below_02_table, below_02_asv); below_02_rda

RsquareAdj(below_02_rda)$adj.r.squared
anova.cca(below_02_rda, step = 100, by = "term")
```

```{r}
# head(table1_wfeature)
above_02_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  # Filter low abundance ASVs
  filter(O2_MIN != "Below oxygen minimum") %>% ungroup()

above_02_subset <- as.character(unique(above_02_table$STN_NISKIN))
# surf_subset

above_02_asv <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 50) %>% 
  filter(STN_NISKIN %in% above_02_subset) %>% ungroup()

above_02_rda <- run_rda(above_02_table, above_02_asv); above_02_rda

RsquareAdj(above_02_rda)$adj.r.squared
anova.cca(above_02_rda, step = 100, by = "term")
```

## RDA with taxa


### Dinoflagellates

```{r}
dino_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  ungroup()

dino_subset <- as.character(unique(dino_table$STN_NISKIN))

dino_asv <- asv_long_avg_wtax %>%
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  filter(MEAN_REPS_seq >= 50) %>% 
  filter(STN_NISKIN %in% dino_subset) %>% ungroup()

dino_rda <- run_rda(dino_table, dino_asv); dino_rda

RsquareAdj(dino_rda)$adj.r.squared
anova.cca(dino_rda, step = 100, by = "term")
```

### Syndiniales

```{r}
syn_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  ungroup()

syn_subset <- as.character(unique(syn_table$STN_NISKIN))

syn_asv <- asv_long_avg_wtax %>%
  filter(Class == "Syndiniales") %>% 
  filter(MEAN_REPS_seq >= 50) %>%
  filter(STN_NISKIN %in% syn_subset) %>% ungroup()

syn_rda <- run_rda(syn_table, syn_asv); syn_rda

RsquareAdj(syn_rda)$adj.r.squared
anova.cca(syn_rda, step = 100, by = "term")
```


# Dinoflagellate functional traits

```{r}
dino_fxn <- read.csv("input-data/dino_fxn_trait.csv")
# head(dino_fxn)
dino_trait_asv <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  left_join(dino_fxn) %>% 
  select(-starts_with("X"), -depth_feature)
# glimpse(dino_trait_asv)
```

```{r, fig.height=4, fig.width=10}
# unique(dino_trait_asv$Feeding.Mode)
# unique(dino_trait_asv$Morphotype)

dino_trait_asv %>% 
  ungroup() %>% 
  # filter(!is.na(Feeding.Mode)) %>% 
  group_by(STN_NISKIN, Station, Depth, DEPTH_BIN, Feeding.Mode, Morphotype) %>% 
  summarise(ASV_TOTAL = n(),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  ggplot(aes(x = ASV_TOTAL, y = Depth, fill = Feeding.Mode, size = SEQ_SUM)) +
    geom_jitter(stat = "identity", shape = 21, width = 0.9) +
  facet_grid(cols = vars(Feeding.Mode),  scale = "free") +
  scale_y_reverse() +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(2,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 12),
    axis.text.y = element_text(color = "black", size = 12),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted"),
    strip.background = element_blank(),
    strip.text.x = element_text(color = "black", size = 12)
  ) +
  labs(x = "Total ASVs", y = "Depth (m)")
```

```{r}
head(dino_trait_asv)
```


```{r, fig.height=3, fig.width=8}
dino_trait_asv %>% 
  ungroup() %>% 
  filter(!is.na(Feeding.Mode)) %>%
  filter(!is.na(Order)) %>%
  unite(DINO_TAX, Class, Order, sep = "; ") %>% 
  group_by(DINO_TAX, Feeding.Mode, Morphotype) %>% 
  summarise(ASV_TOTAL = n(),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  ggplot(aes(x = Morphotype, y = DINO_TAX, size = ASV_TOTAL, fill = Feeding.Mode)) +
    geom_point(shape = 21) +
    facet_grid(cols = vars(Feeding.Mode), space = "free", scale = "free") +
  scale_x_discrete(position = "top") +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(2,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", size = 10),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted"),
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.x = element_text(color = "black", size = 10)
  ) + labs(x = "", y = "")
```
### Gymnodiniales & Peridiniales

4% of all Gymnodiniales ASVs and 5.5% of all sequences.

```{r}
head(dino_trait_asv)
total_dino_asvs <- length(unique(dino_trait_asv$FeatureID))
total_dino_seqs <- sum(dino_trait_asv$MEAN_REPS_seq)
gymnod <- dino_trait_asv %>% 
  filter(Order == "Gymnodiniales")
(length(unique(gymnod$FeatureID)) / total_dino_asvs) * 100
((sum(gymnod$MEAN_REPS_seq)) / total_dino_seqs) * 100


```

```{r}
perid <- dino_trait_asv %>% 
  filter(Order == "Peridiniales")
(length(unique(perid$FeatureID)) / total_dino_asvs) * 100
((sum(perid$MEAN_REPS_seq)) / total_dino_seqs) * 100
"Peridiniales" #Order
```

```{r, fig.width=14, fig.height=4}
dino_trait_asv %>% 
  filter(Order == "Peridiniales" | Order == "Gymnodiniales") %>% 
  filter(!is.na(Feeding.Mode)) %>% 
  group_by(DEPTH_BIN, Order, Genus, Feeding.Mode, Morphotype) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            ASV = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>%
  ggplot(aes(x = SUM, y = DEPTH_BIN_ORDER, fill = Genus)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.7) +
  facet_grid(cols = vars(Order, Feeding.Mode, Morphotype), space = "free", scales = "free") +
  theme_classic() +
  theme(strip.background = element_blank(), 
        strip.text = element_text(color = "black", face = "bold", hjust = 0),
        axis.text = element_text(color = "black", face = "bold")) +
  labs(x = "Proportion of sequences", y = "Depth (m)")
```


### Parasites
```{r}
dino_trait_asv %>% 
  filter(Class == "Syndiniales") %>% 
  # filter(!is.na(Feeding.Mode)) %>% 
  group_by(DEPTH_BIN, Class, Order) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            ASV = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>%
  ggplot(aes(x = SUM, y = DEPTH_BIN_ORDER, fill = Order)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.7) +
  facet_grid(cols = vars(Class), space = "free", scales = "free") +
  theme_classic() +
  theme(strip.background = element_blank(), 
        strip.text = element_text(color = "black", face = "bold", hjust = 0),
        axis.text = element_text(color = "black", face = "bold")) +
  labs(x = "Proportion of sequences", y = "Depth (m)")
```
```{r, fig.height=5, fig.width=18}
dino_trait_asv %>% 
  filter(Class == "Syndiniales") %>% 
  # filter(!is.na(Feeding.Mode)) %>% 
  filter(!is.na(Family)) %>% 
  group_by(DEPTH_BIN, Order, Family) %>% 
  summarise(SUM = sum(MEAN_REPS_seq),
            ASV = n()) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>%
  ggplot(aes(x = Family, y = DEPTH_BIN_ORDER, fill = log(SUM))) +
  # geom_bar(stat = "identity", position = "fill", color = "black", width = 0.7) +
  geom_tile(color = "black") +
  scale_fill_fermenter(type = "seq", palette = 4, direction = 1) +
  facet_grid(cols = vars(Order), space = "free", scales = "free") +
  theme_classic() +
  theme(strip.background = element_blank(), 
        strip.text = element_text(color = "black", face = "bold", hjust = 0),
        axis.text.y = element_text(color = "black", face = "bold"),
        axis.text.x = element_text(color = "black", face = "bold", angle = 45, hjust = 1, vjust = 1)) +
  labs(x = "Syndiniales Families", y = "Depth (m)")
```

# Indicator Species

## indicspecies
https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html

```{r}
library(indicspecies)
```

multipatt is the name of the command from the indicspecies package. The mulitpatt command results in lists of species that are associated with a particular group of samples. If your group has more than 2 categories, multipatt will also identify species that are statistically more abundant in combinations of categories.

The parameters for multipatt are as follows:

the community abundance matrix:"abund"
the vector that contains your sample grouping information: "time"
the function that multipatt is using to identify indicator species: func = "r.g"
the number of permutations used in the statistical test: control = how(nperm=9999)

https://cran.r-universe.dev/indicspecies/doc/manual.html

```{r}
unique(asv_long_avg_wtax$Supergroup_simplified)
```
```{r}
hist(asv_long_avg_wtax$Temperature)
hist(asv_long_avg_wtax$NO3)
```


```{r}
# | eval: false
# | echo: true
indic_input <- asv_long_avg_wtax %>%
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(!grepl("_X", Supergroup_simplified)) %>%
  # Add additional bins
  mutate(TEMP_BIN = case_when(
    Temperature < 8 ~ "< 8",
    Temperature >=8 & Temperature < 15 ~ "8-15",
    Temperature >= 15 & Temperature < 20 ~ "15-20",
    Temperature >= 20 & Temperature < 25 ~ "20-25",
    Temperature > 25 ~ "> 25")) %>% 
  mutate(NO3_BIN = case_when(
    NO3 < 10 ~ "< 10",
    NO3 >= 10 & NO3 < 20 ~ "10-20",
    NO3 >= 20 ~ "> 20")) %>% 
  left_join(table1_wfeature %>% select(STN_NISKIN, DCM, OilRig_Count, SALT_MAX, O2_MIN)) %>% 
  filter(MEAN_REPS_seq >= 25) %>%
  mutate(OIL_BIN = case_when(
      OilRig_Count > 0 ~ "Oil Rig",
      TRUE ~ "No oil rig"
    )) %>% 
  select(STN_NISKIN, DEPTH_BIN, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, TEMP_BIN, Temperature, DCM, NO3_BIN, OIL_BIN, OilRig_Count, SALT_MAX, O2_MIN, MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, names_prefix = "ASV_", values_from = MEAN_REPS_seq, values_fill = 0)
```

Ahead of determining indicator species, ASVs were filtered to only include those that had more than or equal to 25 sequences.

## IndVal

two species are not strongly correlated and their frequency in the data set is larger than the frequency of sites belonging to the site group, the joint occurrence of the two species has higher positive predictive value for the site group than the two species taken independently

‘d’ of species as the product of the relative frequency and relative average abundance in clusters.

```{r, eval=FALSE}
# | eval: false
# | echo: true
# STN_NISKIN, DEPTH_BIN, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature, DCM, OilRig_Count, SALT_MAX, O2_MIN, MEAN_REPS_seq, FeatureID

# clust <- indic_input$DEPTH_BIN
# clust <- indic_input$DCM
# clust <- indic_input$O2_MIN
# clust <- indic_input$SALT_MAX
# clust <- indic_input$NO3_BIN
# clust <- indic_input$TEMP_BIN
clust <- indic_input$OIL_BIN
# unique(indic_input$TEMP_BIN)
abun <- indic_input %>% 
  select(STN_NISKIN, starts_with("ASV_")) %>% 
  column_to_rownames(var = "STN_NISKIN")
```

```{r, eval=FALSE}
# | eval: false
# | echo: true
INV <- multipatt(abun, clust, func = "r.g", control = how(nperm = 99))
summary(INV)

# ?multipatt
depth_bin_result <- INV
# dcm_result <- INV
# o2_result <- INV
# salt_result <- INV
# no3_result <- INV
# temp_bin_result <- INV
oil_result <- INV
# unique(indic_input$NO3_BIN)
```


```{r, eval=FALSE}
# | eval: false
# | echo: true
process_inv_output <- function(INV){
  A <- colnames(as.data.frame(INV$sign))[1]
  B <- colnames(as.data.frame(INV$sign))[2]
  # C <- colnames(as.data.frame(INV$sign))[3]
  # D <- colnames(as.data.frame(INV$sign))[4]
  # E <- colnames(as.data.frame(INV$sign))[5]
  # F_ <- colnames(as.data.frame(INV$sign))[6]
  # G <- colnames(as.data.frame(INV$sign))[7]
  # H <- colnames(as.data.frame(INV$sign))[8]
  # I <- colnames(as.data.frame(INV$sign))[9]
  tmp_indv <- as.data.frame(INV$sign) %>% 
  mutate(INV_SP_RESULT = case_when(
    (.[[1]] == 1 & p.value <= 0.05) ~ A,
    (.[[2]] == 1 & p.value <= 0.05) ~ B
    # (.[[3]] == 1 & p.value <= 0.05) ~ C,
    # (.[[4]] == 1 & p.value <= 0.05) ~ D,
    # (.[[5]] == 1 & p.value <= 0.05) ~ E,
    # (.[[6]] == 1 & p.value <= 0.05) ~ F_,
    # (.[[7]] == 1 & p.value <= 0.05) ~ G,
    # (.[[8]] == 1 & p.value <= 0.05) ~ H,
    # (.[[9]] == 1 & p.value <= 0.05) ~ I
  )) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  select(FeatureID, p.value, INV_SP_RESULT) %>% 
  pivot_wider(names_from = INV_SP_RESULT, values_from = p.value) %>% 
    select(-`NA`)
}
# colnames(as.data.frame(depth_bin_result$sign))[1:9]
# depth_inv <- process_inv_output(depth_bin_result) # Need all nine
# dcm_inv <- process_inv_output(dcm_result)
# o2_inv <- process_inv_output(o2_result)
# salt_inv <- process_inv_output(salt_result)
# oil_inv <- process_inv_output(oil_result) 
# no3_inv <- process_inv_output(no3_result) # need 3
# temp_inv <- process_inv_output(temp_bin_result) # need 5

asv_list_feature <- asv_long_avg_wtax %>% 
  mutate(FeatureID = paste("ASV_", FeatureID, sep = "")) %>% 
  select(FeatureID) %>% distinct() %>%
  left_join(depth_inv) %>% 
  left_join(dcm_inv) %>% 
  left_join(o2_inv) %>% 
  left_join(salt_inv) %>% 
  left_join(oil_inv) %>% 
  left_join(no3_inv) %>% 
  left_join(temp_inv) %>% 
  pivot_longer(cols = starts_with("s."), names_to = "Indicative_of", values_to = "p_value") %>% 
  filter(!is.na(p_value)) %>% 
  mutate(FeatureID = str_remove_all(FeatureID, "ASV_"))
```

```{r}
# save(asv_list_feature, file = "input-data/asv_list_feature.RData")
load(file = "input-data/asv_list_feature.RData")
# head(asv_long_avg_wtax)
# head(asv_list_feature)
```


```{r, fig.height=5, fig.width=45}
# Sum by taxonomic groups
asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 100) %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  left_join(asv_list_feature) %>% 
  filter(!is.na(p_value)) %>% 
  group_by(FeatureID, Supergroup_simplified, Indicative_of) |>
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV = n()) %>%
  add_column(PRESENCE = 1) %>% 
  # # Factoring and ordering
  # mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>%
  ggplot(aes(x = FeatureID, y = Indicative_of, fill = PRESENCE)) +
    geom_tile() +
  facet_grid(cols = vars(Supergroup_simplified), space = "free", scale = "free") +
  # scale_x_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.line.x = element_line(),

    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 10),
    axis.text.x = element_blank()) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")

```
```{r, fig.height=5, fig.width=20}
# Sum by taxonomic groups
asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 100) %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified != "Eukaryota") %>%
  filter(Supergroup_simplified == "Haptista") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  left_join(asv_list_feature) %>% 
  filter(!is.na(p_value)) %>% 
  group_by(FeatureID, Supergroup_simplified, Indicative_of) |>
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV = n()) %>%
  add_column(PRESENCE = 1) %>% 
  # # Factoring and ordering
  # mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>%
  ggplot(aes(x = FeatureID, y = Indicative_of, fill = PRESENCE)) +
    geom_tile(color = "black") +
  facet_grid(cols = vars(Supergroup_simplified), space = "free", scale = "free") +
  # scale_x_discrete(position = "right") +
  theme_classic() + theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.line.x = element_line(),

    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 10),
    axis.text.x = element_blank()) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")

```
Table to demonstrate how many ASVs and what taxa could be indicator species. 
```{r}
total_seq <- sum(asv_long_avg_wtax$MEAN_REPS_seq)
total_asv <- length(unique(asv_long_avg_wtax$FeatureID))

table2 <- asv_long_avg_wtax %>%
  filter(STN_NISKIN != "S12_N10") %>% 
  left_join(asv_list_feature) %>% 
  filter(!is.na(Indicative_of)) %>% 
  group_by(Supergroup_simplified, Indicative_of) %>% 
  summarise(SEQ_SUM = sum(MEAN_REPS_seq),
            ASV_COUNT = n(),
            SEQ_PERC = round((SEQ_SUM/total_seq)*100, digits = 1),
            ASV_PERC = round((ASV_COUNT/total_asv)*100, digits = 1)) %>% 
  select(Supergroup_simplified, Indicative_of, ASV_COUNT) %>% 
  pivot_wider(names_from = "Supergroup_simplified", values_from = "ASV_COUNT")
table2
# write.csv(table2, file = "output-tables/table-indicator-spp.csv")
```


# Session information

```{r}
sessionInfo()
```
