---
title: "Northern Gulf"
editor: source
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 3
---

# Set up R environment

Set up R & import starting data

```{r, warning=FALSE, message=FALSE}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

```{r}
load("input-data/gulf-2023-sequence-analysis_02062026.RData", verbose = TRUE)
```

```{r}
offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5","S9", "S8", "S7", "S11", "S12", "S14", "S15")
stn_nisk_in_study <- as.character(unique(asv_long_avg_wtax_clean$STN_NISKIN))
```


## Change Lat/Long to surface stations
```{r}
head(metadata_allsamples) # All stations are present.
# unique(env_metadata_mod$Station)

tmp_lat_long <- metadata_allsamples %>% 
  filter(Depth < 5) %>% 
  ungroup() %>%
  group_by(Station) %>% 
  slice_sample(n = 1) %>% 
  select(Station, Latitude, Longitude)

metadata_allsamples_mod <- metadata_allsamples %>% 
  select(-Latitude, -Longitude) %>% 
  left_join(tmp_lat_long)
```

## Factoring

```{r}
offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5",
                             "S9", "S8", "S7", 
                             "S11", "S12", "S14", "S15")
stn_order_shore_offshore <- c(1, 2, 3, 4, 5, 
                              9, 8, 7, 6, 
                              10, 11, 12, 13, 14, 15, 16, 17)

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

watercol_features <- c("other", "DCM", "Salinity maximum", "Oxygen minimum")
watercol_features_label <- c("Depth with no feature", "DCM", "Salinity maximum", "Oxygen minimum")
watercol_shape <- c(21, 23, 24, 22)
names(watercol_shape) <- watercol_features_label
```

# Map

#### Fxn for ther surface params, lat, long

Set up a data frame with all parameters as columns.

```{r}
head(metadata_allsamples_mod)
```

```{r}
stn_labels <- metadata_seqsamples %>% 
  ungroup() %>% 
  select(Station, Latitude, Longitude, Depth) %>% distinct()
stn_labels

surface_labels <- stn_labels %>% 
  filter(Depth < 5) %>% 
  select(Station, Latitude, Longitude) %>% distinct()
```

## Get map params

```{r}
# install.packages("Rcpp")
# install.packages(c("googleway", "ggrepel",
# "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))

library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
```

## Map of whole Gulf region

```{r}
ggulf <- ggplot(data = world) +
  geom_sf(aes(fill = region_wb)) +
  annotate(geom = "text", x = -90, y = 26, label = "", 
     fontface = "italic", color = "grey22", size = 6) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  theme(legend.position = "none", axis.title.x = element_blank(), 
     axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
     panel.border = element_rect(fill = NA))

ggulf
```

```{r, fig.width=7, fig.height=7}
# MAP
LARGE <- ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show:
  # coord_sf(xlim = c(-90, -86.5), ylim = c(27.5, 30), expand = FALSE) + # Region of study
  coord_sf(xlim = c(-110, -70), ylim = c(5, 45), expand = FALSE) + # Larger region
  # Add data points for stations
  # geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10", alpha = 0.9) + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 8, colour="black"),
    axis.title = element_text(size = 8, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA))
LARGE
```

```{r, fig.width=7, fig.height=7}
# MAP
# surface_labels
INSET <- ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show
  coord_sf(xlim = c(-91, -85), ylim = c(27, 31), expand = FALSE) + # Larger region
  # Add data points for stations
  geom_point(data = surface_labels, aes(x = Longitude, y = Latitude), size = 5, shape = 21, color = "grey20", fill = "white") +
  geom_text(data = surface_labels, aes(label = Station, x = Longitude, y = Latitude), size = 3) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10") + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 12, colour="black"),
    axis.title = element_text(size = 12, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
    plot.background = element_blank(),
        panel.border = element_rect(fill = NA))
INSET
```

Map to save

```{r, fig.width=8, fig.height=8}
# Left, bottom, right, top
INSET + inset_element(LARGE, 0.4, 0.4, 1.2, 1, align_to = "plot", clip = FALSE)
# ggsave("figures/map_overview.svg", width = 8, height = 8, device = "svg", limitsize = FALSE)
```

Station points and labels.

```{r}
stn_labels <- metadata_seqsamples %>% 
  select(Station, Latitude, Longitude, Depth) %>% distinct() %>% 
  mutate(TRANSECT = case_when(
    (Station < 6) ~ 1,
    (Station > 5 & Station < 10) ~ 2,
    (Station >= 10) ~ 3
  ))

stn_labels

# max_stn_num <- max(stn_labels$Station)
stn_order_shore_offshore <- c(1, 2, 3, 4, 5, 9, 8, 7, 6, 10, 11, 12, 13, 14, 15, 16, 17)

surf_labels <- stn_labels %>% 
  ungroup() %>% 
  filter(Depth < 5) %>% 
  select(Station, Latitude, Longitude, TRANSECT) %>% 
  distinct()
surf_labels
```

# ASVs on map

```{r}
set.seed(123)
library(MBA)
# ?mba.surf
```


```{r}
load("input-data/gulf-2023-sequence-analysis_02062026.RData", verbose = TRUE)
head(asv_long_avg_wtax_clean)
```

## Surface ASV profiles


Surface lat/long
```{r}
surface_asv_odv <- asv_long_avg_wtax_clean %>% ungroup() %>% 
  left_join(metadata_seqsamples) %>%
  # Isolate surface & average across replicates
  filter(Depth < 5) %>%
  select(FeatureID, Station, Latitude, Longitude, TRANSECT, Depth) %>% distinct() %>% 
  ungroup() %>%
  group_by(Station, Latitude, Longitude, TRANSECT) %>%
    summarise(ASV_COUNT = n())
```

Interpolate function for ASVs:

```{r}
# Put PARAM in quotes
interpolate_lat_long_asv <- function(df, PARAM, min, max, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)],
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z,
                                  varnames = c("Longitude", "Latitude"),
                                  value.name = PARAM) %>%
    select(Longitude, Latitude, VALUE = !!PARAM) %>%
    mutate(VALUE = round(VALUE, 1))
head(envs_interpol)
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = NA) +
    # Add data points for stations
    ggrepel::geom_text_repel(data = surf_labels, 
                             aes(x = Longitude, y = Latitude, label = Station),
                             direction = "y", nudge_x = 0.1, nudge_y = 0) +
    geom_point(data = surf_labels, aes(x = Longitude, y = Latitude)) +
    scale_fill_fermenter(type = "seq", palette = "Reds", 
                         direction = 1, na.value = NA,
                         limits = c(min, max)) +
    # scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse(n.breaks = 4) +
  theme_classic() +
  theme(axis.text = element_text(size = 12, face = "bold", colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1,
        legend.text = element_text(size = 10, colour="black", face = "bold"),
        legend.background = element_blank(),
        legend.title = element_text(size = 10, colour="black", face = "bold", hjust = 1, vjust = 1),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.80,.75))
}
```


### Total ASVs

```{r, fig.height=4, fig.width=4, warning=FALSE}
surf_all_asv <- interpolate_lat_long_asv(surface_asv_odv, "ASV_COUNT", 300, 2000, "Total ASVs")
surf_all_asv
```

### Taxa by ASVs

```{r}
bytaxa_surface_asvs <- asv_long_avg_wtax_clean %>% 
  left_join(metadata_seqsamples) %>%
  # Isolate surface & average across replicates
  filter(Depth < 5) %>%
    select(SUPERGROUP_CLASS, FeatureID, Station, Latitude, Longitude, TRANSECT, Depth) %>% distinct() %>% 
  ungroup() %>%
  group_by(SUPERGROUP_CLASS, Station, Latitude, Longitude, TRANSECT) %>%
    summarise(ASV_COUNT = n())
```


```{r, fig.height=4, fig.width=4, warning=FALSE}
# Dinoflagellata-Syndiniales
syndiniales_asvs <- interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), "ASV_COUNT", 100, 1000, "Syndiniales ASVs")
```


```{r, fig.height=4, fig.width=4, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
# Dinoflagellata-Syndiniales
Bacillario <- interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Gyrista-Bacillariophyceae")), "ASV_COUNT", 5, 20,"Diatom ASVs")


Chrysophy<-interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Gyrista-Chrysophyceae")), "ASV_COUNT", 4, 16,"Chrysophyte ASVs")


interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Gyrista-Pelagophyceae")), "ASV_COUNT", 2, 10,"Pelagophyceae ASVs")

```
             
"Radiolaria-Polycystinea"           
"Radiolaria-RAD"                    
"Radiolaria-Radiolaria"   
```{r, fig.height=4, fig.width=4, warning=FALSE}
interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Radiolaria-Acantharea")), "ASV_COUNT", 0, 100, "Acantharea ASVs")

interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == "Radiolaria-Polycystinea")), "ASV_COUNT", 0, 300, "Polycystinea ASVs")

# interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
#   filter(SUPERGROUP_CLASS == "Radiolaria-RAD")), "ASV_COUNT", 0, 10, "RAD ASVs")

# interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  # filter(SUPERGROUP_CLASS == "Radiolaria-Radiolaria")), "ASV_COUNT", "Radiolaria ASVs")
```
## Depth ASV profiles

```{r}
# head(asv_long_avg_wtax_clean)
```


```{r}
depth_asv_odv <- asv_long_avg_wtax_clean %>% 
  left_join((metadata_seqsamples %>% select(STN_NISKIN, Station, Latitude, Longitude, Depth, TRANSECT) %>% distinct())) %>%
  filter(!is.na(Latitude)) %>% 
  ungroup() %>%
  group_by(Station, Latitude, Longitude, Depth, TRANSECT) %>%
    summarise(ASV_COUNT = n())
head(depth_asv_odv)
# unique(depth_asv_odv$TRANSECT)
```

```{r}
# Put PARAM in quotes
interpolate_depth_long_asv <- function(df_all, transect, PARAM, min, max, TITLE){
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Depth", PARAM)], 
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = NA) +
    # Add data points
    geom_point(data = df, aes(x = Longitude, y = Depth)) +
    geom_text(data = df, 
              aes(x = Longitude, y = -60, label = df$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_fermenter(type = "seq", palette = "Reds", 
                         direction = 1, na.value = NA,
                         limits = c(min, max)) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse(n.breaks = 4, expand = c(0,0)) + scale_y_reverse(position = "right") +
    theme_classic() +
    theme(axis.text = element_text(size = 10, face = "bold", colour="black"),
          panel.border = element_blank(),
          aspect.ratio = 1,
          legend.background = element_blank(),
        legend.text = element_text(size = 14, colour="black", face = "bold"),
        legend.title = element_text(size = 14, colour="black", face = "bold", hjust = 0, vjust = 1),
        axis.line = element_line(),
        legend.ticks = element_line(color = "black"),
        legend.position=c(0.16,0.31))
}

interpolate_depth_long_asv(depth_asv_odv, "transect1","ASV_COUNT", 400, 2000,"Total ASVs")
```

### Total ASVs panel

```{r, fig.height=5, fig.width=15, warning=FALSE}
(interpolate_lat_long_asv(surface_asv_odv, "ASV_COUNT", 400, 2000, "Total ASVs") + theme(legend.position = "none")) +
interpolate_depth_long_asv(depth_asv_odv, "transect1","ASV_COUNT", 400, 2000,"Total ASVs") +
  (interpolate_depth_long_asv(depth_asv_odv, "transect2","ASV_COUNT", 400, 2000, "Total ASVs") + theme(legend.position = "none")) +
  (interpolate_depth_long_asv(depth_asv_odv, "transect3","ASV_COUNT", 400, 2000, "Total ASVs") + theme(legend.position = "none")) +
  plot_layout(nrow = 1) + plot_annotation(tag_levels = c("a"))
# ggsave("figures/asv-map-all.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```

```{r}
stns_in_study <- as.character(unique(metadata_seqsamples$Station))
shallow_stn <- (surf_labels %>% filter(Station %in% stns_in_study))

interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect1","ASV_COUNT", 400, 2000,"Total ASVs") + 
  geom_text(data = (shallow_stn %>% filter(TRANSECT == 1)), 
              aes(x = Longitude, y = 0, label = (shallow_stn %>% filter(TRANSECT == 1))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(0,200))
```


```{r, fig.height=5, fig.width=15, warning=FALSE}
(interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect1","ASV_COUNT", 400, 2000,"Total ASVs") + geom_text(data = (shallow_stn %>% filter(TRANSECT == 1)), 
              aes(x = Longitude, y = -3, label = (shallow_stn %>% filter(TRANSECT == 1))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(-3,200)))+
  (interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect2","ASV_COUNT", 400, 2000, "Total ASVs") + geom_text(data = (shallow_stn %>% filter(TRANSECT == 2)), 
              aes(x = Longitude, y = -3, label = (shallow_stn %>% filter(TRANSECT == 2))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(-3,200))) +
  (interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect3","ASV_COUNT", 400, 2000, "Total ASVs") + geom_text(data = (shallow_stn %>% filter(TRANSECT == 3)), 
              aes(x = Longitude, y = -3, label = (shallow_stn %>% filter(TRANSECT == 3))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(-3,200))) +
  plot_layout(nrow = 1, guides = "collect") + plot_annotation(tag_levels = c("a"))
# ggsave("figures/asv-map-all.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

```

```{r, fig.height=5, fig.width=15, warning=FALSE}
(interpolate_lat_long_asv(surface_asv_odv, "ASV_COUNT", 400, 2000, "Total ASVs") + theme(legend.position = "none")) |
((interpolate_depth_long_asv(depth_asv_odv, "transect1","ASV_COUNT", 400, 2000,"Total ASVs") +
  (interpolate_depth_long_asv(depth_asv_odv, "transect2","ASV_COUNT", 400, 2000, "Total ASVs") + theme(legend.position = "none")) +
  (interpolate_depth_long_asv(depth_asv_odv, "transect3","ASV_COUNT", 400, 2000, "Total ASVs") + theme(legend.position = "none"))) /
((interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect1","ASV_COUNT", 400, 2000,"Total ASVs") + geom_text(data = (shallow_stn %>% filter(TRANSECT == 1)), 
              aes(x = Longitude, y = -3, label = (shallow_stn %>% filter(TRANSECT == 1))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(-3,200))) +
  (interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect2","ASV_COUNT", 400, 2000, "Total ASVs") + geom_text(data = (shallow_stn %>% filter(TRANSECT == 2)), 
              aes(x = Longitude, y = -3, label = (shallow_stn %>% filter(TRANSECT == 2))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(-3,200))) +
  (interpolate_depth_long_asv((depth_asv_odv %>% filter(Depth <= 200)), "transect3","ASV_COUNT", 400, 2000, "Total ASVs") + geom_text(data = (shallow_stn %>% filter(TRANSECT == 3)), 
              aes(x = Longitude, y = -3, label = (shallow_stn %>% filter(TRANSECT == 3))$Station), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
  scale_y_reverse(position = "right", limits = c(-3,200))))) +
  plot_layout(guides = "collect", axis_titles = "collect") + plot_annotation(tag_levels = c("a"))
# ggsave("figures/asv-map-all.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

```

### Taxa by ASVs

```{r}
bytaxa_depth_asvs <- asv_long_avg_wtax_clean %>% 
  left_join((metadata_seqsamples %>% select(STN_NISKIN, Station, Latitude, Longitude, Depth, TRANSECT) %>% distinct())) %>%
  ungroup() %>%
  group_by(SUPERGROUP_CLASS, Station, Latitude, Longitude, Depth, TRANSECT) %>%
    summarise(ASV_COUNT = n())

```


```{r, fig.height=4, fig.width=4, warning=FALSE}
# Dinoflagellata-Syndiniales
interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), 
  "transect1", "ASV_COUNT",
  100, 1500, "Syndiniales ASVs")
```

## ASV profiles by taxa
```{r}
# df_surf <- (bytaxa_surface_asvs %>% 
#   filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales"))
# 
# df <- (bytaxa_depth_asvs %>% 
#   filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales"))
```


```{r, fig.height=5, fig.width=17, warning=FALSE}

tax_whole_asv_profile <- function(SUBSET, MIN, MAX, TITLE){
interpolate_lat_long_asv((bytaxa_surface_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "ASV_COUNT", MIN, MAX, TITLE) +
interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "transect1","ASV_COUNT", MIN, MAX, TITLE) +
  interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "transect2","ASV_COUNT", MIN, MAX, TITLE) +
  interpolate_depth_long_asv((bytaxa_depth_asvs %>% 
  filter(SUPERGROUP_CLASS == SUBSET)), "transect3","ASV_COUNT", MIN, MAX, TITLE) +
  plot_layout(nrow = 1, guides = "collect")
}

# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
```



## Dinoflagellate ASV profiles

            
"Dinoflagellata-Noctilucophyceae"      
```{r, fig.height=5, fig.width=17, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
# tax_whole_asv_profile("Dinoflagellata-Noctilucophyceae", 0, 10, "Dinoflagellata-Noctilucophyceae")
```

"Dinoflagellata-Dinophyceae"
```{r, fig.height=5, fig.width=15, warning=FALSE}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
tax_whole_asv_profile("Dinoflagellata-Dinophyceae", 50, 800, "Dinoflagellata-Dinophyceae")
# ggsave("figures/dinophyceae.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
"Dinoflagellata-Syndiniales"          
```{r, fig.height=5, fig.width=15, warning=FALSE}
tax_whole_asv_profile("Dinoflagellata-Syndiniales", 50, 800, "Dinoflagellata-Syndiniales")
# ggsave("figures/syndiniales.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
## Radiolaria ASV profiles

```{r}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
colnames(asv_long_avg_wtax_clean)
asv_long_avg_wtax_clean %>% 
  left_join((metadata_seqsamples %>% select(STN_NISKIN, Station, Latitude, Longitude, Depth, TRANSECT) %>% distinct())) %>%
  filter(grepl("Radiolaria", SUPERGROUP_CLASS)) %>% 
  ungroup() %>% 
  group_by(SUPERGROUP_CLASS) %>% 
  summarise(ASV = n())
  
```

"Radiolaria-Polycystinea"
```{r, fig.height=5, fig.width=15, warning=FALSE}
tax_whole_asv_profile("Radiolaria-Polycystinea", 10, 200, "Radiolaria-Polycystinea")
# ggsave("figures/rad-polycystinea.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
"Radiolaria-Acantharea"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Radiolaria-Acantharea", 10, 100, "Radiolaria-Acantharea")
# ggsave("figures/rad-acantharia.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
"Radiolaria-RAD-A"
"Radiolaria-RAD-B"
"Radiolaria-RAD-C"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Radiolaria-RAD-A", 10, 70, "Radiolaria-RAD-A")
# ggsave("figures/rad-A.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

tax_whole_asv_profile("Radiolaria-RAD-B", 10, 70, "Radiolaria-RAD-B")
# ggsave("figures/rad-B.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

tax_whole_asv_profile("Radiolaria-RAD-C", 10, 70, "Radiolaria-RAD-C")
# ggsave("figures/rad-C.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```

## Gyrista ASV profiles

```{r}
# unique(bytaxa_surface_asvs$SUPERGROUP_CLASS)
colnames(asv_long_avg_wtax_clean)
asv_long_avg_wtax_clean %>% 
  left_join(metadata_seqsamples) %>% 
  filter(grepl("Gyrista", SUPERGROUP_CLASS)) %>% 
  ungroup() %>% 
  group_by(SUPERGROUP_CLASS) %>% 
  summarise(ASV = n())
```


"Gyrista-Bacillariophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Bacillariophyceae", 0, 20, "Gyrista-Bacillariophyceae")
ggsave("figures/gyrista-bacil.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```

"Gyrista-Chrysophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Chrysophyceae", 0, 20, "Gyrista-Chrysophyceae")
# ggsave("figures/gyrista-chryso.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```

"Gyrista-Dictyochophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Dictyochophyceae", 0, 20, "Gyrista-Dictyochophyceae")
# ggsave("figures/gyrista-dictyocho.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```

"Gyrista-Coscinodiscophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Coscinodiscophyceae", 0, 20, "Gyrista-Coscinodiscophyceae")
# ggsave("figures/gyrista-coscinod.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
"Gyrista-Mediophyceae"
```{r, fig.height=5, fig.width=17, warning=FALSE}
tax_whole_asv_profile("Gyrista-Mediophyceae", 0, 20, "Gyrista-Mediophyceae")
# ggsave("figures/gyrista-medioph.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```
"Ciliophora-CONThreeP"
```{r, fig.height=5, fig.width=15, warning=FALSE}
tax_whole_asv_profile("Ciliophora-CONThreeP", 5, 50, "Ciliophora-CONThreeP")
# ggsave("figures/ciliate-con3p.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

tax_whole_asv_profile("Ciliophora-Oligohymenophorea", 5, 50, "Ciliophora-Oligohymenophorea")
# ggsave("figures/ciliate-oligo.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

tax_whole_asv_profile("Ciliophora-Phyllopharyngea", 5, 50, "Ciliophora-Phyllopharyngea")
# ggsave("figures/ciliate-phyllo.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)

tax_whole_asv_profile("Ciliophora-Spirotrichea", 5, 50, "Ciliophora-Spirotrichea")
# ggsave("figures/ciliate-spirot.svg", width = 15, height = 5, device = "svg", limitsize = FALSE)
```


# K-means clustering on map

```{r}
load("input-data/kmeans-membership-seqs-5.RData", verbose = TRUE)
head(kmeans_membership_5)
```

```{r}
CLUSTER_COLOR <- c("#c85e82", "#efc85a", "#df9147","#6e9b4a", "#606aa9")
CLUSTER_NUM <- c(1, 2, 3, 4, 5)
names(CLUSTER_NUM) <- CLUSTER_COLOR
```

## Surface kmeans
```{r}
(metadata_seqsamples %>% 
   filter(Depth < 6) %>% 
   select(Station, Latitude, Longitude) %>% distinct())
```


```{r}
surface_kmeans <- kmeans_membership_5 %>% 
  separate(STN_NISKIN, into = c("STN", "NISKIN"), sep = "_", remove = FALSE) %>% 
  mutate(Station = as.integer(str_remove(STN, "S"))) %>% 
  select(Station, CLUSTER) %>% distinct() %>% 
  right_join(metadata_seqsamples) %>% distinct() %>% 
  filter(Depth < 6) %>% 
   select(Station, CLUSTER, Latitude, Longitude) %>% distinct() %>% 
  mutate(CLUSTER_ORDER = factor(CLUSTER, levels = CLUSTER_NUM))
surface_kmeans
```

```{r, fig.width=7, fig.height=7}
ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show
  coord_sf(xlim = c(-91, -85), ylim = c(27, 31), expand = FALSE) + # Larger region
  # Add data points for stations
  geom_point(data = surface_kmeans, aes(x = Longitude, y = Latitude, fill = CLUSTER_ORDER), size = 5, shape = 21, color = "grey20") +
  scale_fill_manual(values = CLUSTER_COLOR) +
  geom_text(data = surface_labels, aes(label = Station, x = Longitude, y = Latitude), size = 3) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10") + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 12, colour="black"),
    axis.title = element_text(size = 12, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
    plot.background = element_blank(),
        panel.border = element_rect(fill = NA))
```
## Depth k-means

```{r}
depth_kmeans <- kmeans_membership_5 %>% 
  separate(STN_NISKIN, into = c("STN", "NISKIN"), sep = "_", remove = FALSE) %>% 
  mutate(Station = as.integer(str_remove(STN, "S"))) %>%
  mutate(Niskin = as.integer(str_remove(NISKIN, "N"))) %>% 
  select(Station, Niskin, CLUSTER) %>% distinct() %>%
  left_join((metadata_seqsamples %>% select(Station, Niskin, Depth, Latitude, Longitude, TRANSECT) %>% distinct())) %>% 
  filter(!is.na(Latitude)) %>% 
  mutate(CLUSTER_ORDER = factor(CLUSTER, levels = CLUSTER_NUM))

depth_kmeans
```

```{r}
vert_label <- depth_kmeans %>% filter(Depth < 6) %>% select(CLUSTER_ORDER, Station, Longitude, TRANSECT) %>% distinct()
vert_label
max(depth_kmeans$Depth)
```


```{r, fig.height=4, fig.width=9}
depth_kmeans %>% 
  ggplot() +
  geom_segment(data = vert_label, linetype = "dashed", aes(x = Longitude, y = 3000, yend = 0, xend = Longitude)) +
    # Add data points
    geom_jitter(aes(x = Longitude, y = Depth, fill = CLUSTER_ORDER), size = 4, shape = 21, color = "grey20", width = 0.1) +
  scale_fill_manual(values = CLUSTER_COLOR) +
    geom_text(data = vert_label,
              aes(x = Longitude, y = -60, label = Station),
              hjust = 0.5, vjust = -0.5, color= "black") +
    labs(y = "Depth (m)", x = "Longitude (°W)") +
    scale_x_reverse(n.breaks = 4, expand = c(0.1,0.1)) + 
  scale_y_reverse(position = "right", expand = c(0.1,0)) +
    theme_classic() +
  facet_grid(cols = vars(TRANSECT), scales = "free", space = "free") +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_blank(),
          legend.background = element_blank(),
        legend.text = element_text(size = 14, colour="black", face = "bold"),
        legend.title = element_blank(),
        axis.line = element_line(),
        legend.ticks = element_line(color = "black"),
        legend.position="left")
```


# Session Information

```{r}
sessionInfo()
```

