---
title: "2023 Gulf data analysis"
editor: source
format:
  html:
    theme: flatly
    toc: true
    toc-location: left
    number-sections: true
    number-depth: 3
---

# Contents & introduction

## Set up R environment

Set up R & import starting data

```{r, warning=FALSE, message=FALSE}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

Import ASV tables as R objects.

```{r}
load(file = "input-data/asv_wtax_qc_GoM23_102025.RData", verbose = TRUE)
```

Import & compile metadata

```{r}
master_metadata <- read.csv("input-data/grad23_master.csv", header = TRUE, 
                            check.names = FALSE, na.strings = "--")
# head(master_metadata)
cellcounts <- read.csv("cellcounts-sept22.csv")
```

## Modify metadata file

```{r}
colnames(master_metadata)
# str(master_metadata)
# str(cellcounts)
env_metadata_toplot <- master_metadata %>%
  mutate(depth_feature = case_when(
    grepl("O2 min", Notes) ~ "O2 min",
    grepl("DCM", Notes) ~ "DCM"
  )) %>% 
  mutate(Station = as.integer(Station)) %>% 
  select(Station, Niskin, Depth = `CTD Recorded Depth (m)`, depth_feature, Latitude = `NMEA Latitude`, Longitude = `NMEA Longitude`,
         Date, Time_UTC = `Time (UTC)`, Temperature = `CTD Temperature`, Salinity = `Corrected CTD Salinity`, 
         Oxygen = `Corrected CTD Oxygen`, NO3, PO4, SIL, NO2, NH4) %>% 
  left_join(cellcounts %>% select(Station, Niskin, Prok_count = `prok.count`)) %>% 
  separate_longer_delim(c(NO3, PO4, SIL, NO2, NH4), delim = " / ") %>%
  pivot_longer(cols = c(Temperature, Salinity, Oxygen, NO3, PO4, SIL, NO2, NH4, Prok_count), names_to = "ENV_VARIABLE", values_to = "VALUE", values_drop_na = TRUE, values_transform = as.numeric) %>% 
  group_by(Station, Niskin, Depth, depth_feature, Latitude, Longitude, Date, Time_UTC, ENV_VARIABLE) %>% 
    summarise(value = mean(VALUE))
```

### Create metabarcoding table with metadata

```{r}
# head(asv_long_cleaned)
# head(env_metadata_toplot)
# Widen environmental metadata to add to ASV df.
tmp_forasv <- env_metadata_toplot %>% 
  filter(!is.na(ENV_VARIABLE)) |> 
  pivot_wider(names_from = ENV_VARIABLE, values_from = value)
  
asv_long_cleaned_wenv <- asv_long_cleaned %>% 
  separate(SAMPLES, into = c("gom", "STATION", "NISKIN", "extra"), sep = "_") %>% 
  mutate(Station = as.integer(str_remove(STATION, "S")),
         Niskin = as.integer(str_remove(NISKIN, "N"))) %>% 
  unite(STN_NISKIN, STATION, NISKIN, sep = "_") %>% 
  select(-gom, -extra) %>% 
  left_join(tmp_forasv) 

head(asv_long_cleaned_wenv)
```

## Calculate distance from shore

Distance from shore

```{r}
library(geosphere)
# Long, Lat order
outflow_lat <- 28.95
outflow_long <- 89.39
outflow <- c(outflow_long, outflow_lat)
outflow
```

Set up ordering of stations and locations.

```{r}
stn_info <- read.csv("input-data/manual_stn_classification.csv")

offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5","S9", "S8", "S7", "S11", "S12", "S14", "S15")

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(env_metadata_toplot$Depth))))

depth_order_round <- as.character(sort(round(unique(env_metadata_toplot$Depth), digits = 0)))

depth_bin <- c("< 5m","5-25m","25-60m",
                          "60-270m","270-400m","400-800m",
                          "800-1200m","1200-2000m","2000-3000m")

depth_bin_CHARC <- c("..5m","5.25m","25.60m",
                          "60.270m","270.400m","400.800m",
                          "800.1200m","1200.2000m","2000.3000m")

asv_long_cleaned_wenv_mod <- asv_long_cleaned_wenv %>% 
  left_join(stn_info) %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order),
         TRANSECT = factor(stn, levels = offshore_on_shore_order, labels = transect_labels),
         DEPTH_ORDER = factor(Depth, levels = rev(depth_order), labels = rev(depth_order_round))) %>% 
  mutate(DIST_OUTFLOW = distHaversine(cbind(Longitude, Latitude), cbind(outflow_long, outflow_lat))) %>% 
  mutate(DEPTH_BIN = case_when(
    Depth <= 5 ~ "< 5m",
    Depth > 5 & Depth <=25 ~ "5-25m",
    Depth > 25 & Depth <=60 ~ "25-60m",
    Depth > 60 & Depth <=270 ~ "60-270m",
    Depth > 270 & Depth <=400 ~ "270-400m",
    Depth > 400 & Depth <=800 ~ "400-800m",
    Depth > 800 & Depth <=1200 ~ "800-1200m",
    Depth > 1200 & Depth <=2000 ~ "1200-2000m",
    Depth > 2000 & Depth <=3000 ~ "2000-3000m",
  ))
# dim(asv_long_cleaned_wenv)
# dim(asv_long_cleaned_wenv_mod)
```

## Study stats

```{r}
range(asv_long_cleaned_wenv_mod$Station)
length(unique(asv_long_cleaned_wenv_mod$Station)) # 12 stations
(unique(asv_long_cleaned_wenv_mod$Station)) # 12 stations
length(unique(asv_long_cleaned_wenv_mod$STN_NISKIN)) # 91 unique samples
```

```{r}
table1 <- asv_long_cleaned_wenv_mod %>% 
  select(STN_NISKIN, Station, Depth, depth_feature, Date, Time_UTC, Latitude, Longitude, Transect, DEPTH_BIN, DIST_OUTFLOW, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature, Prok_count) %>%
  distinct()
# head(table1)
write.csv(table1, file = "output-tables/table-samplelist.csv")
```

# Curate taxonomic assignment

```{r}
# head(asv_long_cleaned_wenv_mod)
# unique(asv_long_cleaned_wenv_mod$Domain) # take EUKS
# unique(asv_long_cleaned_wenv_mod$Supergroup)
# unique(asv_long_cleaned_wenv_mod$Division)
# View(asv_long_cleaned_wenv_mod %>% 
#        select(Domain, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct())
```

```{r}
asv_long_wenv_wtax <- asv_long_cleaned_wenv_mod %>% 
  filter(Domain == "Eukaryota") %>% 
  filter(Supergroup != "nucl") %>% 
  mutate(DIVISION = case_when(
    Subdivision == "X" ~ "Unannotated",
    is.na(Subdivision) ~ "Unannotated",
    TRUE ~ Subdivision
  )) %>% 
  # Goal is to combine Subdivision-Class
  mutate(SUPERGROUP_CLASS = case_when(
    ## High level NAs at the supergroup and division level
    (is.na(Division) & !is.na(Supergroup)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Division) & is.na(Supergroup)) ~ "Eukaryote-Unannotated",
    (is.na(Subdivision)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Class) & !is.na(Subdivision)) ~ paste(Subdivision, "Unannotated", sep = "-"),
    #
    (Subdivision == "X") ~ paste(Division, "Unannotated", sep = "-"),
    (Class == "X") ~ paste(Subdivision, "Unannotated", sep = "-"),
    TRUE ~ paste(Subdivision, Class, sep = "-"))) %>% 
  mutate(Supergroup_simplified = case_when(
    Supergroup == "TSAR" ~ paste("TSAR", Subdivision, sep = "-"),
    Supergroup == "Obazoa" ~ paste("Obazoa", DIVISION, sep = "-"),
    TRUE ~ Supergroup
  ))

# 
# sort(unique(tmp$DIVISION))
# sort(unique(tmp$Subdivision))
# unique(tmp$SUPERGROUP_CLASS)
# head(asv_long_clean_wtax)
tax_key <- asv_long_wenv_wtax %>% 
  select(FeatureID, Taxon, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct()
```

## Average across replicates

```{r}
head(asv_long_wenv_wtax)
```

```{r}
asv_long_avg_wtax <- asv_long_wenv_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(FeatureID, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup_simplified, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species,
           STN_NISKIN, Station, Niskin, Depth, depth_feature, Latitude, Longitude, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL, Prok_count,
           STN_CATEGORY_3, STN_CATEGORY_4, STN_ORDER, TRANSECT, DEPTH_ORDER, DIST_OUTFLOW, DEPTH_BIN) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
 filter(STN_NISKIN != "S12_N10") |>
  ungroup()

# head(asv_long_avg_wtax)
# unique(asv_long_avg_wtax$STN_CATEGORY_4)
```

# Data frames for analysis

Central data frames:

-   `asv_long_wenv_wtax`
-   `tax_key`
-   `env_metadata_toplot`
-   `asv_long_avg_wtax`

```{r}
# save(asv_long_wenv_wtax, tax_key, table1, env_metadata_toplot, asv_long_avg_wtax, file = "input-data/gulf-2023-sequence-analysis_10252025.RData")
```

# Sample comparison & taxonomic classification

## PCA analysis

Create wide format matrix with sequence counts averaged across replicates (by ASV)

```{r}
head(asv_long_avg_wtax)
```

Station & depth info

```{r}
stn_depth_info <- asv_long_avg_wtax |> 
  select(STN_NISKIN, Station, Niskin, Depth, depth_feature, Latitude, Longitude, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL, Prok_count,
           STN_CATEGORY_3, STN_CATEGORY_4, STN_ORDER, TRANSECT, DEPTH_ORDER, DIST_OUTFLOW, DEPTH_BIN) |> 
  filter(STN_NISKIN != "S12_N10") |> 
  distinct()
```

Make wide format and as a matrix

```{r}
asv_wide_avg_matrix <- asv_long_avg_wtax %>% 
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_matrix))))
# ?prcomp
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

#### Plot PCA

```{r, fig.height=5, fig.width=5.5}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

PCA_PLOT <- data.frame(pca_clr$x, STN_NISKIN = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, STN_NISKIN) %>% 
  left_join(stn_depth_info) |> 
  mutate(depth = as.numeric(DEPTH_ORDER)) |> 
  mutate(depth_feature = str_replace_na(depth_feature, replacement = "other")) |> 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, 
             shape = depth_feature, 
             fill = DEPTH_BIN)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_point(stroke = 0.6, alpha = 0.8, size = 4) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(24, 23, 21)) +
  scale_fill_manual(values = blues) +
  # scale_size_continuous(range = c(2, 6)) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=12),
        legend.title = element_blank(),
        panel.grid.major = element_line(color = "grey90", linetype = "dotted"),
        axis.title = element_text(color="black", size= 12),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(fill = guide_legend(override.aes = list(
    color = "black", 
    shape = 22, 
    size = 5)),
    shape = guide_legend(override.aes = list(
    color = "black", 
    size = 5)))
PCA_PLOT
# ggsave("figures/pca-plot.svg", width = 5.5, height = 5, device = "svg", limitsize = FALSE)
```

## Combined dendrogram

```{r}
library(ggdendro)
rel_abun <- decostand(asv_wide_avg_matrix, MARGIN = 2, method = "total")
# Cluster dendrogram (average hierarchical clustering)
cluster_gom <- hclust(dist(t(rel_abun)), method = "average")
dendro <- as.dendrogram(cluster_gom)
gom_dendro <- dendro_data(dendro, type = "rectangle")
```

```{r, fig.height=18, fig.width=4}
gom_dendro_plot <- ggplot(segment(gom_dendro)) +
  geom_segment(aes(x = x, y = y, xend = xend, 
    yend = yend)) + 
  coord_flip() + 
  scale_y_reverse(expand = c(0.2, 0.5), breaks = c(0, 0.2, 0.4, 0.6, 0.8)) +
  geom_text(size = 3, aes(x = x, y = y, label = label, angle = 0, hjust = 0), data = label(gom_dendro)) +
  theme_dendro() + labs(y = "Dissimilarity") +
    theme(axis.text.x = element_text(color = "black", size = 10),
          axis.line.x = element_line(color = "#252525"), 
        axis.ticks.x = element_line(), axis.title.x = element_text(color = "black", 
            size = 10))
# svg('figs/SUPPLEMENTARY-dendrogram-wreps.svg', w = 10, h = 8)
gom_dendro_plot
```

Isolate label names that correspond to dendrogram.

```{r}
tmp <- data.frame(gom_dendro$labels)
class(tmp)
labels_for_dendro <- as.character(tmp$label)
# labels_for_dendro
```

### Dendrogram annotation

```{r, fig.height=12, fig.width=4}
blues <- c("#dde4f0","#b7cae8","#96b3e3",
           "#6b94d6","#3473d9", "#0b4fbd",
           "#0b429c", "#07214a", "#081730")

names(blues) <- depth_bin

DEPTH <- stn_depth_info %>%
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = DEPTH_BIN)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = blues) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
DEPTH
```

```{r, fig.height=12, fig.width=4}
DIST <- stn_depth_info %>%
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  mutate(dist_outflow = DIST_OUTFLOW/1000) %>% 
  ggplot(aes(y = label_order, x = 1, fill = dist_outflow)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_continuous(option = "magma") +
  scale_fill_viridis_c(option = "magma", direction = -1) +
    # scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "x10^3 km", y = "")
DIST

```
dendrogram annotation with DCM, O2, and Salt max
```{r}
feature_profiles <- read.csv("input-data/feature_profiles_depth.csv")
```

```{r, fig.height=12, fig.width=2}
PROFILES <- stn_depth_info %>%
  ungroup() %>% 
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  left_join(feature_profiles) %>% 
  select(STN_NISKIN, label_order, Station, Niskin, Depth, FEATURE) %>% distinct() %>%
  mutate(VAL = case_when(
    !is.na(FEATURE) ~ 1,
    TRUE ~ 0
  )) %>%
  ggplot(aes(y = label_order, x = FEATURE, fill = 1)) +
    geom_tile(color = "black") +
  coord_fixed(ratio = 1) +
  theme_void() + theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted"),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_text(color = "black", angle = 90, hjust = 1, vjust = 0, size = 5)) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
PROFILES

```


```{r, fig.height=12, fig.width=4}
TRANSECT <- stn_depth_info %>%
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = TRANSECT)) +
    geom_point(shape = 22, color = "black", size = 4) +
  scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
TRANSECT
```

```{r, fig.height=12, fig.width=4}
PROK <- stn_depth_info %>%
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(y = label_order, x = 1, fill = Prok_count)) +
    geom_point(shape = 22, color = "black", size = 4) +
  # scale_fill_manual(values = c("#ff8787", "#eaf576", "#76aff5")) +
  scale_fill_fermenter(na.value = "grey30", palette = 3, direction = 1) +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.ticks = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.text.x = element_blank()) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
PROK
```

# Taxonomy

Richness by depth, station, and feature.

```{r, fig.height=8, fig.width=8}
asv_long_avg_wtax %>% 
  group_by(Station, depth_feature, DEPTH_BIN) %>% 
    summarise(SEQ_SUM = sum(MEAN_REPS_seq),
              ASV_COUNT = n()) %>% 
  mutate(depth_feature = str_replace_na(depth_feature, replacement = "other")) |> 
  mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin))) |> 
  ggplot(aes(x = as.factor(Station), y = depth_bin_order, 
             size = ASV_COUNT, fill = depth_feature)) +
    geom_jitter(shape = 21, color = "black", stroke = 1.5) +
  scale_x_discrete(position = "top") +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(3,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted")
  ) +
  labs(x = "", y = "Depth (m)")
```

```{r}
tax_color  <- c("#db1f7d", "#2c5c1d", "#806f8c",
                "#f5bfb0", "#93c47d","#7c52bf", "#007a7a", 
                "#34495e", "#9b59b6", "#fea02f","#ee429b",
                "#ffe8a4","#ff6f61","#fcf517","#8a0622",
                "#ff7530", "#2986cc", "#e71414","#de6600", 
                "#1abc9c", "#a0b59f", "#d7b6d7", "#23b521", "#5b545c", "black", "white", "grey10", "grey60")


tax_simplified_list <- as.character((unique(asv_long_avg_wtax$Supergroup_simplified)))
names(tax_color) <- tax_simplified_list
```

Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level.

### For dendrogram

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
TAX <- asv_long_avg_wtax %>%
   filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  group_by(STN_NISKIN, Station, Niskin, Depth, DEPTH_BIN,
           DEPTH_ORDER, Supergroup_simplified) |> 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  # Factoring and ordering
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  # mutate(label_order = str_remove(label_order, "_")) |> 
  ggplot(aes(y = (label_order), x = SUM_SEQ, fill = Supergroup_simplified)) +
    geom_bar(stat = "identity", position = "fill", 
             color = "grey20", width = 0.8, linewidth = 0.5) +
  scale_fill_manual(values = tax_color) +
  scale_y_discrete(position = "right") +
  theme_void() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    axis.line.x = element_line(),
    
    axis.title = element_text(color = "black"),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 10),
    axis.text.x = element_text(color = "black", size = 8)) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "", y = "")
TAX

```

### Plot sequence relative abundance

```{r, fig.width=5, fig.height=14}
asv_long_avg_wtax %>%
  group_by(Station, DEPTH_BIN, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
   mutate(depth_bin_order = factor(DEPTH_BIN, levels = rev(depth_bin))) |> 
  ggplot(aes(y = depth_bin_order, x = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "white", width = 0.9) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(Station), scales = "free_y", space = "free", axis.labels = "all_x") +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

# Diversity indices

Use wide dataframe from above.

```{r}
head(asv_wide_avg_matrix)[,1:3]
```

```{r}
shannon_all <- vegan::diversity(asv_wide_avg_matrix, index = "shannon", MARGIN = 2)
```

```{r, fig.height=20, fig.width=2}
SHANNON <- data.frame(shannon_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(stn_depth_info) |> 
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(x = shannon_all, y = label_order)) +
  geom_point(shape = 21, fill = "grey30", color = "black", size = 2, stroke = 1) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 10),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Shannon Diversity", y = "")
```

```{r}
invsimp_all <- vegan::diversity(asv_wide_avg_matrix, index = "invsimpson", MARGIN = 2)
# head(data.frame(invsimp_all))
```

```{r, fig.height=20, fig.width=2}
INVSIMP <- data.frame(invsimp_all) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  left_join(stn_depth_info) |> 
  mutate(label_order = factor(STN_NISKIN, levels = labels_for_dendro)) %>% 
  ggplot(aes(x = invsimp_all, y = label_order)) +
  geom_point(shape = 21, fill = "grey80", color = "black", size = 2, stroke = 1) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black"),
    strip.background = element_blank(),
    # axis.ticks = element_blank(),
    axis.title = element_text(color = "black", size = 8),
    axis.text.y = element_text(color = "black", vjust = 0.5, hjust = 0, size = 10),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
    labs(x = "Inverse Simpson", y = "")
INVSIMP
```

## Plot Dendrogram compilation

```{r, fig.height=13, fig.width=22}
gom_dendro_plot + DEPTH + PROFILES + DIST + TRANSECT + PROK + SHANNON + INVSIMP + TAX +
  plot_layout(nrow = 1, 
              widths = c(1.3, 0.5, 0.5, 0.5, 0.5, 0.5, 
                         0.4, 0.4, 0.7))
# ggsave("figures/dendrogram-compiled.svg", width = 22, height = 13, device = "svg", limitsize = FALSE)
```

# Taxa and depth

```{r}
# head(asv_long_avg_wtax)
asv_long_avg_wtax %>%
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>%
  group_by(Class, Depth) %>%
  summarise(ASV_COUNT = n()) %>%
  ggplot(aes(x = ASV_COUNT, y = Depth, fill = Class)) +
  geom_point(stat = "identity", shape = 21, color = "black") +
  scale_y_reverse() +
  theme_classic()
```



# CLR targets & plots

TSAR-Dinoflagellata
TSAR-Radiolaria
TSAR-Gyrista


Confirming what core protistan groups are the most prominent members of the community across all of the samples. 
```{r}
# head(asv_long_avg_wtax)
top_taxa <- asv_long_avg_wtax %>% 
  filter(!(grepl("Obazoa", Supergroup_simplified))) %>% 
  group_by(Supergroup_simplified, STN_NISKIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>% 
  ungroup() %>% 
  # Subset the top 2 groups with the highest ASV count
  group_by(STN_NISKIN) %>% 
  slice_max(order_by = ASV_COUNT, n = 3)
  # slice_max(order_by = SUM_SEQ, n = 3)
# top_taxa
list_top <- unique(top_taxa$Supergroup_simplified)
```

```{r}
head(asv_long_avg_wtax)
```
```{r}
# View(asv_long_avg_wtax)
list_top
```

## Dinoflagellates
```{r}
class_dino_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Dinoflagellata-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix

head(class_dino_mat)

class_dino_clr <- vegan::decostand(class_dino_mat, MARGIN = 1, method = "rclr")
##
colnames(class_dino_clr) <- colnames(class_dino_mat)
rownames(class_dino_clr) <- rownames(class_dino_mat)
# data.frame(class_dino_clr)
```

```{r, fig.height=10, fig.width=3}
plot_dino_clr <- data.frame(class_dino_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```



## Haptista
```{r}
# list_top

class_hapto_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "Haptista") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Haptista-Unannotated",
    (is.na(Order) & Class == "X") ~ "Haptista-Unannotated",
         TRUE ~ Order)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_hapto_mat)
class_hapto_clr <- vegan::decostand(class_hapto_mat, MARGIN = 1, method = "rclr")
##
colnames(class_hapto_clr) <- colnames(class_hapto_mat)
rownames(class_hapto_clr) <- rownames(class_hapto_mat)

# data.frame(class_hapto_clr)
```

```{r, fig.height=10, fig.width=3}
plot_hapto_clr <- data.frame(class_hapto_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```



## Radiolaria
```{r}
list_top
# tmp <- asv_long_avg_wtax %>% 
  # filter(STN_NISKIN != "S12_N10") %>% 
  # filter(Supergroup_simplified == "TSAR-Radiolaria"); unique(tmp$Class)
  
class_rad_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Radiolaria-Unannotated",
    (is.na(Order) & Class == "X") ~ "Radiolaria-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_rad_mat)
class_rad_clr <- vegan::decostand(class_rad_mat, MARGIN = 1, method = "rclr")
##
colnames(class_rad_clr) <- colnames(class_rad_mat)
rownames(class_rad_clr) <- rownames(class_rad_mat)

# data.frame(class_rad_clr)
```

```{r, fig.height=10, fig.width=3}
plot_rad_clr <- data.frame(class_rad_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```



## Ciliate
```{r}
list_top
tmp <- asv_long_avg_wtax %>%
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Ciliophora"); unique(tmp$Class)
  
class_cil_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Ciliophora-Unannotated",
    (is.na(Order) & Class == "X") ~ "Ciliophora-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_cil_mat)
class_cil_clr <- vegan::decostand(class_cil_mat, MARGIN = 1, method = "rclr")
##
colnames(class_cil_clr) <- colnames(class_cil_mat)
rownames(class_cil_clr) <- rownames(class_cil_mat)

# data.frame(class_rad_clr)
```

```{r, fig.height=13, fig.width=5}
plot_cil_clr <- data.frame(class_cil_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```




## Gyrista
```{r}
list_top
tmp <- asv_long_avg_wtax %>%
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Gyrista"); unique(tmp$Class)
  
class_gyrista_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Gyrista-Unannotated",
    (is.na(Order) & Class == "X") ~ "Gyrista-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_gyrista_mat)
class_gyrista_clr <- vegan::decostand(class_gyrista_mat, MARGIN = 1, method = "rclr")
##
colnames(class_gyrista_clr) <- colnames(class_gyrista_mat)
rownames(class_gyrista_clr) <- rownames(class_gyrista_mat)
```

```{r, fig.height=13, fig.width=5}
plot_gyrista_clr <- data.frame(class_gyrista_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```





## Bigyra
```{r}
list_top
tmp <- asv_long_avg_wtax %>%
filter(STN_NISKIN != "S12_N10") %>%
filter(Supergroup_simplified == "TSAR-Bigyra"); unique(tmp$Class)
  
class_bigyra_mat <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Bigyra") %>% 
  ungroup() %>% 
  mutate(CLASS = case_when(
    is.na(Class) ~ "Bigyra-Unannotated",
    (is.na(Order) & Class == "X") ~ "Bigyra-Unannotated",
         TRUE ~ Class)) %>% 
  group_by(CLASS, Station, DEPTH_BIN) %>% 
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  unite(SAMPLE, Station, DEPTH_BIN, sep = ";") %>% 
  pivot_wider(names_from = CLASS, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "SAMPLE") %>%
  as.matrix
#
head(class_bigyra_mat)
class_bigyra_clr <- vegan::decostand(class_bigyra_mat, MARGIN = 1, method = "rclr")
##
colnames(class_bigyra_clr) <- colnames(class_bigyra_mat)
rownames(class_bigyra_clr) <- rownames(class_bigyra_mat)
```

```{r, fig.height=13, fig.width=3}
plot_bigya_clr <- data.frame(class_bigyra_clr) %>% 
  rownames_to_column(var = "SAMPLE") %>% 
  pivot_longer(cols = -c(SAMPLE), names_to = "Class", values_to = "CLR") %>% 
  separate(SAMPLE, into = c("Station", "DEPTH_BIN"), sep = ";") %>% 
  mutate(stn = paste("S", Station, sep = "")) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order)) %>% 
  mutate(DEPTH_BIN_ORDER = factor(DEPTH_BIN, levels = rev(depth_bin))) %>% 
    ggplot(aes(y = DEPTH_BIN_ORDER, x = Class, fill = CLR)) +
    geom_tile(color = "grey30", linewidth = 0.4) +
    facet_grid(rows = vars(STN_ORDER), space = "free_y", scale = "free_y") +
    scale_x_discrete(position = "top") +
    scale_fill_fermenter(type = "div", palette = 4) +
    theme_classic() +
    theme(axis.ticks = element_blank(),
          axis.text.x = element_text(color = "black", angle = 45, size = 7,
                                     hjust = 0, vjust = 0.5),
          axis.text.y = element_text(color = "black"),
          strip.text.y = element_text(color = "black", angle = 0),
          strip.background = element_blank()) +
    labs(x = "", y = "")
```

## Compile taxa CLR plots
```{r, fig.width=20, fig.height=11}
plot_dino_clr + plot_hapto_clr + plot_rad_clr + plot_cil_clr + plot_gyrista_clr + plot_bigya_clr + plot_layout(nrow = 1, widths = c(0.5, 0.55, 0.4, 1, 1.5, 0.4))
```



# Linear regression

```{r}
head(asv_long_avg_wtax)
# PARAMS: DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL
```

```{r}
library(broom)
regression_gom_tmp <- asv_long_avg_wtax %>%
  filter(STN_NISKIN != "S12_N10") %>%
  select(Supergroup_simplified, MEAN_REPS_seq, Station, Niskin, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
  group_by(Station, Niskin, Supergroup_simplified, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
summarise(seq_sum = sum(MEAN_REPS_seq),
          asv_rich = n()) %>% 
  # filter(ASV_rich > 1 &)
  pivot_longer(cols = c(DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL), names_to = "PARAMS", values_to = "VALUES") %>% 
  ungroup() 
dim(regression_gom_tmp)
head(regression_gom_tmp)
# unique(regression_gom_tmp$Class)
```

```{r}
part1 <- regression_gom_tmp %>% 
# For taxonomic groups, group linear regression by ASV Richness
 group_by(Supergroup_simplified, PARAMS) %>% 
  nest(-Supergroup_simplified, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
    tidied = map(lm_fit, tidy)) %>% unnest(tidied) %>%
  select(Supergroup_simplified, PARAMS, term, estimate, p.value) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
    select(everything(), SLOPE = VALUES, p.value) %>% data.frame
# ?pivot_wider
```

```{r}
part2 <- regression_gom_tmp %>% 
# For taxonomic groups, group linear regression by ASV Richness
 group_by(Supergroup_simplified, PARAMS) %>% 
  nest(-Supergroup_simplified, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
    glanced = map(lm_fit, glance)) %>% unnest(glanced) %>%
  select(Supergroup_simplified, PARAMS, r.squared, adj.r.squared) %>%
  right_join(part1) %>% 
  right_join(regression_gom_tmp)
```

```{r}
# head(part2)
# unique(part2$Supergroup_simplified)
```

```{r, fig.width=10, fig.height=5}
part2 %>% 
  filter(Supergroup_simplified == "TSAR-Ciliophora") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "Ciliate")
```


```{r, fig.width=10, fig.height=5}
part2 %>% 
  filter(Supergroup_simplified == "Haptista") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "Haptista")
```

```{r}
#TSAR-Dinoflagellata
part2 %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>%
  ggplot(aes(x = VALUES, y = asv_rich, fill = Supergroup_simplified)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  # scale_fill_manual(values = samplecolor) + 
   # scale_shape_manual(values = shapes) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
    3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
    theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
    strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black", 
        size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = "TSAR-Dinoflagellata")
```
## Linear regression function

Where R2 is the proportion of the variation in ASV richness predictable from the environmental parameter supplied. Closer to 1 is a better fit for the relationship. 

`function(df, TITLE)`

```{r}
library(broom)
lm_gom <- function(df, TITLE) {
  df_tmp <- df %>%
  select(TAXA, MEAN_REPS_seq, Station, Niskin, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
  group_by(TAXA, Station, Niskin, DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL) %>% 
  summarise(seq_sum = sum(MEAN_REPS_seq),
            asv_rich = n()) %>% 
  pivot_longer(cols = c(DIST_OUTFLOW, Depth, Temperature, Salinity, Oxygen, NH4, NO2, NO3, PO4, SIL), names_to = "PARAMS", values_to = "VALUES") %>% 
  ungroup() 
#
df_tmp_tidy <- df_tmp %>% 
  # For taxonomic groups, group linear regression by ASV Richness
  group_by(TAXA, PARAMS) %>% 
  nest(-TAXA, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
         tidied = map(lm_fit, tidy)) %>% unnest(tidied) %>%
  select(TAXA, PARAMS, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>% 
  select(everything(), SLOPE = VALUES) %>% data.frame

df_lm <- df_tmp %>% 
  # For taxonomic groups, group linear regression by ASV Richness
  group_by(TAXA, PARAMS) %>% 
  nest(-TAXA, -PARAMS) %>% 
  mutate(lm_fit = map(data, ~lm(asv_rich ~ VALUES, data = .)), 
         glanced = map(lm_fit, glance)) %>% unnest(glanced) %>%
  select(TAXA, PARAMS, r.squared, adj.r.squared) %>%
  right_join(df_tmp_tidy) %>% 
  right_join(df_tmp)

df_lm %>% 
  ggplot(aes(x = VALUES, y = asv_rich)) +
  geom_abline(aes(slope = SLOPE, 
                  intercept = X.Intercept.),
              color = "black", linetype = "dashed", lwd = 0.5) +
  geom_point(color = "black", 
             size = 2) + geom_smooth(method = lm) + 
  facet_wrap(. ~ PARAMS + round(r.squared, 
                                3), scales = "free", ncol = 5, strip.position = "bottom", labeller = label_parsed) + 
  theme_bw() + 
  theme(strip.background = element_blank(), 
        strip.placement = "outside", 
        legend.position = "none",               
        strip.text = element_text(color = "black", size = 10), axis.title = element_text(color = "black",                                                     size = 10), legend.title = element_blank()) + labs(y = "ASV richness", x = "", title = TITLE)
}
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  add_column(TAXA = "All ASVs")), "All ASVs")

lm_gom((asv_long_avg_wtax %>% 
          filter(Depth < 10) %>% 
  add_column(TAXA = "Surface ASVs")), "Surface ASVs")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  filter(Class == "Syndiniales") %>% 
  add_column(TAXA = "Syndiniales")), "Syndiniales")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 50) %>% 
  filter(Class == "Dinophyceae") %>% 
  add_column(TAXA = "Dinophyceae")), "Dinophyceae")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Subdivision == "Gyrista") %>% 
  add_column(TAXA = "Gyrista")), "Gyrista")
#Bacillariophyceae
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Bacillariophyceae") %>% 
  add_column(TAXA = "Bacillariophyceae")), "Bacillariophyceae")


lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Syndiniales") %>% 
  add_column(TAXA = "Syndiniales")), "Syndiniales")

lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH <= 10) %>%
  filter(Class == "Dinophyceae") %>% 
  add_column(TAXA = "Dinophyceae")), "Dinophyceae")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
  filter(Subdivision == "Ciliophora") %>% 
  add_column(TAXA = "Ciliates")), "Ciliates")
```

```{r, fig.width=13, fig.height=7}
lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH > 400) %>%
  filter(Subdivision == "Cercozoa") %>% 
  add_column(TAXA = "Cercozoa")), "Cercozoa")

lm_gom((asv_long_avg_wtax %>% 
          # filter(DEPTH > 400) %>%
  filter(Subdivision == "Radiolaria") %>% 
  add_column(TAXA = "Radiolaria")), "Radiolaria")
```


# RDA analysis

RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it.

> This is where we can add future oil spill work

```{r}
set.seed(1234)
head(table1)
```

```{r}
features <- read.csv(file = "input-data/feature_profiles_depth.csv")
head(features)
# table(features %>% select(FEATURE, Station))
features %>% select(-X, -Niskin) %>% 
  pivot_wider(names_from = FEATURE, values_from = Depth)

table1_wfeature <- table1 %>% select(-depth_feature) %>% 
  left_join((features %>% select(-X, -Niskin) %>% 
  pivot_wider(names_from = FEATURE, values_from = Depth))) %>% 
  mutate(SALT_MAX = case_when(
    Depth >= `Salinity maximum` ~ "Below salinity maximum", #At and below the salinity maximum
    Depth < `Salinity maximum` ~ "Above salinity maximum",
  )) %>% 
  mutate(DCM = case_when(
    Depth > DCM ~ "Below DCM",
    Depth <= DCM ~ "Above DCM", #At and above the DCM
  )) %>% 
  mutate(O2_MIN = case_when(
    Depth > `Oxygen minimum` ~ "Below oxygen minimum",
    Depth <= `Oxygen minimum` ~ "Above oxygen minimum", #At and above the O2 minimum
  )) %>% 
  select(-`Oxygen minimum`, -`Salinity maximum`)

```

## All

```{r}
# Make categorical data table from metadata (Table 1)
categorical <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  select(STN_NISKIN, Transect, DCM, O2_MIN, SALT_MAX) %>% distinct()

# Generate a matrix of numerical data from metadata
metadata_for_rda_0 <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  select(STN_NISKIN, DIST_OUTFLOW, Depth, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature) %>% distinct() %>% 
  pivot_longer(cols = -c(STN_NISKIN), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(STN_NISKIN, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>%
  as.matrix
```

Normalize matrix of numeric metadata.
```{r}
standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")

# Combine normalized metadata with categorical data
metadata_normalized <- categorical %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN")
# head(metadata_normalized)
rownames(metadata_normalized) # Output from here should be the same rownames for ASV table below.
```
Generate a matrix from ASV counts
```{r}
# head(asv_long_avg_wtax)
# hist(asv_long_avg_wtax$MEAN_REPS_seq)
asv_wide_avg_reps_matrix <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  # Filter low abundance ASVs
  filter(MEAN_REPS_seq >= 50) %>% 
  ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
# colnames(asv_wide_avg_reps_matrix)
```

```{r}
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix

clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))

# Confirm that all 
asv_list <- rownames(asv_wide_avg_rda_0)
meta_list <- rownames(metadata_normalized)
clr_list <- rownames(clr_asv_matrix)
# asv_list %in% meta_list
# meta_list %in% clr_list
# setdiff(meta_list, clr_list)
```
Run RDA analysis
```{r}
# | eval: false
rda_all_gom <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
rda_all_gom
```
```{r}
# | eval: false
adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
# RsquareAdj(rda_all_gom)
adj_rda_out
anova.cca(rda_all_gom, step = 100, by = "term")
```

### RDA function
First create a function for running RDA analysis.

```{r}
# table1
# asv_long_avg_wtax

# How to use:
## Modify and subsame table1 or asv_long_avg_wtax completely before use.
## Check to make sure the same STN_NISKIN IDs are present in both.
run_rda <- function(mod_metadata, mod_asv_long, CATEGORIES){
  # Select categorical data
  # mod_metadata <- surf_table
  # mod_asv_long <- surf_asv
  # CATEGORIES <- c("Transect") #Character list
  categorical <- mod_metadata %>% 
    select(STN_NISKIN, all_of(CATEGORIES)) %>% distinct()
  # Isolate numerical data as matrix
  metadata_for_rda_0 <- mod_metadata %>% 
    select(STN_NISKIN, DIST_OUTFLOW, Depth, NH4, NO2, NO3, Oxygen, PO4, SIL, Salinity, Temperature) %>% distinct() %>% 
    pivot_longer(cols = -c(STN_NISKIN), names_to = "VARIABLES", values_to = "VALUE") %>% 
    group_by(STN_NISKIN, VARIABLES) %>% 
      summarise(MEAN_VAL = mean(VALUE)) %>% 
    pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
    arrange(STN_NISKIN) %>% 
    column_to_rownames(var = "STN_NISKIN") %>% as.matrix
  # Normalize
  standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")
  # Combine normalized numerical data with categorical
  metadata_normalized <- categorical %>% 
    right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "STN_NISKIN")) %>%
    arrange(STN_NISKIN) %>% 
    column_to_rownames(var = "STN_NISKIN")
  #
  ## Prepare ASV data
  asv_wide_avg_reps_matrix <- mod_asv_long %>% 
    ungroup() %>%  
  select(FeatureID, STN_NISKIN, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = STN_NISKIN, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
  # Transform
  asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "STN_NISKIN") %>% 
  arrange(STN_NISKIN) %>% 
  column_to_rownames(var = "STN_NISKIN") %>% 
  as.matrix
  #
  # CLR normalization
  clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))
  #
  # setdiff(rownames(metadata_normalized), (rownames(clr_asv_matrix)))
  # RDA step
  rda_output <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
  rda_output
  return(rda_output)
}
```


### Above DCM 

Generate a matrix from ASV counts
```{r}
surf_table <- table1_wfeature %>% 
  filter(STN_NISKIN != "S12_N10") |> 
  # Filter low abundance ASVs
  filter(DCM == "Above DCM") %>% ungroup()

surf_subset <- as.character(unique(surf_table$STN_NISKIN))
# surf_subset

surf_asv <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 50) %>% 
  filter(STN_NISKIN %in% surf_subset) %>% ungroup()

# How to use:
## Modify and subsame table1 or asv_long_avg_wtax completely before use.
## Check to make sure the same STN_NISKIN IDs are present in both.
above_DCM <- c("Transect")
above_dcm_rda <- run_rda(surf_table, surf_asv, above_DCM)
```

```{r}
RsquareAdj(above_dcm_rda)$adj.r.squared
anova.cca(above_dcm_rda, step = 100, by = "term")
```

## Dinoflagellates, Diatoms, other abundant?

# Dinoflagellate functional traits

```{r}
dino_fxn <- read.csv("input-data/dino_fxn_trait.csv")
# head(dino_fxn)
dino_trait_asv <- asv_long_avg_wtax %>% 
  filter(STN_NISKIN != "S12_N10") %>% 
  filter(Supergroup_simplified == "TSAR-Dinoflagellata") %>% 
  left_join(dino_fxn) %>% 
  select(-starts_with("X"), -depth_feature)
glimpse(dino_trait_asv)
```
```{r, fig.height=5, fig.width=8}
# unique(dino_trait_asv$Feeding.Mode)
# unique(dino_trait_asv$Morphotype)

dino_trait_asv %>% 
  ungroup() %>% 
  filter(!is.na(Feeding.Mode)) %>% 
  group_by(STN_NISKIN, Station, Depth, DEPTH_BIN, Feeding.Mode, Morphotype) %>% 
  summarise(ASV_TOTAL = n(),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  ggplot(aes(x = ASV_TOTAL, y = Depth, fill = Feeding.Mode, size = SEQ_SUM)) +
    geom_jitter(stat = "identity", shape = 21, width = 0.9) +
  facet_grid(cols = vars(Feeding.Mode), space = "free_x") +

  scale_y_reverse() +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(2,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 12),
    axis.text.y = element_text(color = "black", size = 12),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted"),
    strip.background = element_blank(),
    strip.text.x = element_text(color = "black", size = 12)
  ) +
  labs(x = "Total ASVs", y = "Depth (m)")
```

```{r}
head(dino_trait_asv)
```


```{r, fig.height=3, fig.width=8}
dino_trait_asv %>% 
  ungroup() %>% 
  filter(!is.na(Feeding.Mode)) %>%
  filter(!is.na(Order)) %>%
  unite(DINO_TAX, Class, Order, sep = "; ") %>% 
  group_by(DINO_TAX, Feeding.Mode, Morphotype) %>% 
  summarise(ASV_TOTAL = n(),
            SEQ_SUM = sum(MEAN_REPS_seq)) %>% 
  ggplot(aes(x = Morphotype, y = DINO_TAX, size = ASV_TOTAL, fill = Feeding.Mode)) +
    geom_point(shape = 21) +
    facet_grid(cols = vars(Feeding.Mode), space = "free", scale = "free") +
  scale_x_discrete(position = "top") +
  scale_fill_manual(values = c("#6edb8f", "#ab70e6", "#ebb923")) +
  scale_size(range=c(2,7), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 10),
    axis.text.y = element_text(color = "black", size = 10),
    panel.grid.major = element_line(color = "grey90", linetype = "dotted"),
    strip.background = element_blank(),
    strip.placement = "outside",
    strip.text.x = element_text(color = "black", size = 10)
  ) + labs(x = "", y = "")
```


# Indicator Species

# CRT

```{r}
# install.packages("tree")
library(rpart)
library(tree)
```

```{r}
data("iris")
head(iris)
iris_df <- iris
colnames(iris_df) <- c("sepall", "sepalw", "petall", "petalw", "class")
```

CRT, on the other side, focuses on creating predictive models for classifying and prioritizing data as “key indicator variables”. CRT employs a decision tree approach, partitioning the dataset based on informative environmental variables to create a hierarchical structure of decision rules (Alkhasawneh et al., 2014). We used CRT to identify specific taxonomic groups as “key indicator variables” that show ability to indicate different DLs.

```{r}
xtabs( ~ class, data = iris_df)
```

```{r}
# execute tree classification
iris.tree <- tree::tree(class ~ ., data = iris_df)
# ?tree()
head(iris_df)
```

Above step fits a classification or regression tree with data.

```{r}
summary(iris.tree)
# Only 3 of 4 variables were used for the tree.
# misclassification error rate? 
```

```{r}
plot(iris.tree); text(iris.tree, all = T)
```

For this classification tree, petall has a value \< 2.45, therefore suggesting it has a more important classification. Sepall \<5.15 indicates that is has a less important criterion

Next we "prune" the tree to reduce complexity. For this, we use a function to identify the number of significant notes. The cross validation will find the number of misclassifications.

```{r}
tree::cv.tree(iris.tree)
```

We prune the small branches. Therefore, we want to remove the "Branches" from the tree with the fewest data points. Therefore, we look at the output values of k and see when the values start to really increase. In this example, it is 3. A tree size of 3 is therefore suitable. THe *K* is considered the "cost-complexity parameter" (which is the "Best" number of nodes/branches to include. )

Pruning the tree:

```{r}
iris.tree.prune <- prune.misclass(iris.tree, best = 3)
summary(iris.tree.prune)
```

```{r}
plot(iris.tree.prune);text(iris.tree.prune, all = T)
```

The tree is much less complex.

### CRT all samples

```{r}
# head(asv_long_avg_wtax)
```

```{r}
# | eval: false
# input_CRT <- asv_long_avg_wtax %>% 
#   select(FeatureID, DEPTH, TEMP, SALINITY, OXYGEN, NH4, MEAN_REPS_seq) %>% distinct()
# 
# toy_input_CRT <- input_CRT %>% 
#   sample_n(1000)
# 
# test_tree <- tree(FeatureID ~ , toy_input_CRT)
# summary(test_tree)

# test_tree$frame
# plot(test_tree); text(test_tree, all = T)
```

```{r}
# install.packages("devtools")
# require("devtools")
# devtools::install_github(repo = 'zellerlab/siamcat')
```


# Session information

```{r}
sessionInfo()
```
