---
title: "Analysis of 18S rRNA metabarcoding"
editor: visual
---

# Contents & introduction

# Set up R environment

Set up R & import starting data

```{r}
library(tidyverse);library(phyloseq); library(decontam)
library(compositions); library(patchwork); 
library(ggupset); library(gt)
library(plotly); library(viridis); library(vegan)
```

# Import prior data

From sequencing

```{r}
# | eval: false
# save(samples_too_low, env_info, asv_wide_cleaned, asv_long_cleaned, file =  "input-data/asv_wtax_qc_GoM23_062025.RData")
```

Import sequence R objects

```{r}
load(file = "input-data/asv_wtax_qc_GoM23_062025.RData", verbose = TRUE)
# env_info <- read.csv("input-data/metadata-gom-23.csv")
tmp <- read.csv("cellcounts-sept22.csv")
prok <- tmp %>% select(Station, Niskin, PROK_COUNT = `prok.count`)
head(env_info)
```

```{r}
# head(prok)
env_info <- env_info %>% 
  left_join(prok)
```

## Station information, ordering, location

Distance from shore

```{r}
library(geosphere)
# Long, Lat order
outflow_lat <- 28.95
outflow_long <- 89.39
outflow <- c(outflow_long, outflow_lat)
outflow
```

Set up ordering of stations and locations.

```{r}
stn_info <- read.csv("input-data/manual_stn_classification.csv")

offshore_on_shore_order <- c("S1", "S2", "S3", "S4", "S5","S9", "S8", "S7", "S11", "S12", "S14", "S15")

transect_labels <- c("transect1", "transect1", "transect1", "transect1", "transect1", "transect2", "transect2", "transect2", "transect3", "transect3", "transect3", "transect3")

depth_order <- as.character(sort((unique(env_info$DEPTH))))
depth_order_round <- as.character(sort(round(unique(env_info$DEPTH), digits = 0)))

env_info_mod <- env_info %>% 
  left_join(stn_info) %>% 
  mutate(STN_ORDER = factor(stn, levels = offshore_on_shore_order),
         TRANSECT = factor(stn, levels = offshore_on_shore_order, labels = transect_labels),
         DEPTH_ORDER = factor(DEPTH, levels = rev(depth_order), labels = rev(depth_order_round))) %>% 
  mutate(DIST_OUTFLOW = distHaversine(cbind(Longitude, Latitude), cbind(outflow_long, outflow_lat)))
head(env_info_mod)
# # unique(env_info_mod$DIST_OUTFLOW)
table(env_info_mod$DEPTH)
# unique(env_info_mod$STN_ORDER)
```

# Curate taxonomic assignment

```{r}
# head(asv_long_cleaned)
# unique(asv_long_cleaned$Domain) # take EUKS
# unique(asv_long_cleaned$Supergroup)
# unique(asv_long_cleaned$Division)
# View(asv_long_cleaned %>% 
#        select(Domain, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct())
```

```{r}
asv_long_clean_wtax <- asv_long_cleaned %>% 
  filter(Domain == "Eukaryota") %>% 
  filter(Supergroup != "nucl") %>% 
  mutate(DIVISION = case_when(
    Subdivision == "X" ~ "Unannotated",
    is.na(Subdivision) ~ "Unannotated",
    TRUE ~ Subdivision
  )) %>% 
  # Goal is to combine Subdivision-Class
  mutate(SUPERGROUP_CLASS = case_when(
    ## High level NAs at the supergroup and division level
    (is.na(Division) & !is.na(Supergroup)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Division) & is.na(Supergroup)) ~ "Eukaryote-Unannotated",
    (is.na(Subdivision)) ~ paste(Supergroup, "Unannotated", sep = "-"),
    (is.na(Class) & !is.na(Subdivision)) ~ paste(Subdivision, "Unannotated", sep = "-"),
    #
    (Subdivision == "X") ~ paste(Division, "Unannotated", sep = "-"),
    (Class == "X") ~ paste(Subdivision, "Unannotated", sep = "-"),
    TRUE ~ paste(Subdivision, Class, sep = "-"))) %>% 
  mutate(Supergroup_simplified = case_when(
    Supergroup == "TSAR" ~ paste("TSAR", Subdivision, sep = "-"),
    Supergroup == "Obazoa" ~ paste("Obazoa", DIVISION, sep = "-"),
    TRUE ~ Supergroup
  )) %>% 
  left_join(env_info_mod) 

# 
# sort(unique(tmp$DIVISION))
# sort(unique(tmp$Subdivision))
# unique(tmp$SUPERGROUP_CLASS)
# head(asv_long_clean_wtax)
tax_key <- asv_long_clean_wtax %>% 
  select(FeatureID, Supergroup, Division, Subdivision, Class, Order, Family, Genus, Species) %>% distinct()
```

## Normalization and averaging across replicates

```{r}
colnames(asv_long_clean_wtax)
```

```{r}
range(asv_long_clean_wtax$DEPTH)
```

```{r}
asv_long_avg_wtax <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, PROX_TOSHORE, STN_ORDER, STN_CATEGORY_3, STN_CATEGORY_4, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup_simplified, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_ID, stn, nisk, sep = " ", remove = FALSE)

head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$STN_CATEGORY_4)
```

> Create a figure with data points for each station and depth. Size of data point = \# of ASVs

```{r, fig.height=10, fig.width=8}
asv_long_avg_wtax %>% 
  group_by(DEPTH_ORDER, STN_ORDER, Station, DEPTH) %>% 
    summarise(SEQ_SUM = sum(MEAN_REPS_seq),
              ASV_COUNT = n()) %>% 
  ggplot(aes(x = STN_ORDER, y = DEPTH, size = ASV_COUNT)) +
    geom_jitter(shape = 21, color = "black", stroke = 2) +
  scale_y_reverse() +
  scale_x_discrete(position = "top") +
     scale_size(range=c(1,12), guide="legend") +
  theme_classic() +
  theme(
    axis.text.x = element_text(color = "black", size = 14),
    axis.text.y = element_text(color = "black", size = 14),
    panel.grid.major.x = element_line()
  ) +
  labs(x = "", y = "Depth (m)")
```

```{r}
tax_color  <- c("#db1f7d", "#2c5c1d", "#806f8c",
                "#f5bfb0", "#93c47d","#7c52bf", "#007a7a", 
                "#34495e", "#9b59b6", "#fea02f","#ee429b",
                "#ffe8a4","#ff6f61","#fcf517","#8a0622",
                "#ff7530", "#2986cc", "#e71414","#de6600", 
                "#1abc9c", "#a0b59f", "#d7b6d7", "#23b521", "#5b545c", "black", "white", "grey10", "grey60")


tax_simplified_list <- as.character((unique(asv_long_avg_wtax$Supergroup_simplified)))
names(tax_color) <- tax_simplified_list
```

Low taxonomic resolution should be at DIVISION level. Then to SUPERGROUP_CLASS level.

```{r, fig.height=20, fig.width=7}
# Sum by taxonomic groups
asv_long_avg_wtax %>%
  group_by(PROX_TOSHORE, TRANSECT, STN_ORDER, DEPTH_ORDER, SAMPLE_ID, DEPTH, TEMP, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  ggplot(aes(y = (DEPTH_ORDER), x = SUM_SEQ, fill = Supergroup_simplified)) +
    geom_bar(stat = "identity", position = "fill", color = "black") +
  facet_grid(rows = vars(PROX_TOSHORE, STN_ORDER), space = "free", scales = "free") +
  scale_fill_manual(values = tax_color) +
  theme_classic() +theme(
    strip.background = element_blank(),
    axis.text = element_text(color = "black")
  )
```

```{r, fig.width=4, fig.height=7}
asv_long_avg_wtax %>%
  group_by(STN_ORDER, DEPTH_ORDER, TEMP, Supergroup, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  ggplot(aes(y = (DEPTH_ORDER), x = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.8) +
  scale_fill_manual(values = tax_color) +
  scale_x_continuous(expand = c(0,0)) +
  facet_grid(rows = vars(STN_ORDER), scales = "free_y", space = "free", axis.labels = "all_x") +
  # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
  theme_classic() + theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

## Only surface parameters

```{r, fig.width=8, fig.height=3}
asv_long_avg_wtax %>%
  filter(DEPTH < 5) %>% 
  group_by(stn, STN_ORDER, Supergroup_simplified, TRANSECT) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq),
            ASV_COUNT = n()) %>%
  filter(Supergroup_simplified != "Eukaryota") %>% 
  filter(Supergroup_simplified != "TSAR-NA") %>% 
  filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
  ggplot(aes(x = STN_ORDER, y = SUM_SEQ, fill = Supergroup_simplified)) +
  geom_bar(stat = "identity", position = "fill", color = "black", width = 0.6) +
  scale_fill_manual(values = tax_color) +
  scale_y_continuous(expand = c(0,0)) +
  facet_grid(cols = vars(TRANSECT), scales = "free", space = "free") +
  # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
  theme_classic() + theme(
    legend.position = "left",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
  # guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
  labs(x = "Station", y = "Relative Sequence Abundance")
```

```{r, fig.width=6, fig.height=18}
# asv_long_avg_wtax %>%
#   # filter(STN_ORDER %in% c(1:5)) %>% 
#   group_by(STN_ORDER, DEPTH, DEPTH_ORDER, TEMP, Supergroup, Supergroup_simplified) %>%
#   summarise(SUM_SEQ = sum(MEAN_REPS_seq),
#             ASV_COUNT = n()) %>%
#   filter(Supergroup_simplified != "Eukaryota") %>% 
#   filter(Supergroup_simplified != "TSAR-NA") %>% 
#   filter(Supergroup_simplified != "Obazoa-Metazoa") %>% 
#   ggplot(aes(x = as.numeric(DEPTH), y = SUM_SEQ, fill = Supergroup_simplified)) +
#   geom_area(stat = "identity", position = "fill", color = "black") +
#   coord_flip() +
#   # geom_bar(stat = "identity", position = "fill", color = "black", width = 0.8) +
#   scale_fill_manual(values = tax_color) +
#   scale_x_reverse() +
#   # scale_x_continuous(expand = c(0,0)) +
#   facet_grid(rows = vars(STN_ORDER), scales = "free_y", space = "free", axis.labels = "all_x") +
#   # facet_wrap(vars(STN_ORDER), ncol = 5, scales = "free_y", shrink = TRUE) +
#   theme_classic() + theme(
#     legend.position = "right",
#     legend.title = element_blank(),
#     legend.text = element_text(color = "black", face = "bold"),
#     strip.background = element_blank(),
#     strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
#     axis.title = element_text(color = "black", face = "bold"),
#     axis.text.y = element_text(color = "black", face = "bold"),
#     axis.text.x = element_text(color = "black", face = "bold")) +
#   guides(fill = guide_legend(reverse = TRUE, ncol = 1)) +
#   labs(x = "Sequence Relative Abundace", y = "Depth (m)")
```

# PCA analysis

Create wide format matrix with sequence counts averaged across replicates (by ASV)

```{r}
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
pca_clr <- prcomp(data.frame(compositions::clr(t(asv_wide_avg_reps_matrix))))
# ?prcomp
```

```{r}
variance_clr <- (pca_clr$sdev^2)/sum(pca_clr$sdev^2)

barplot(variance_clr, main = "Log-Ratio PCA Screeplot", xlab = "PC Axis", ylab = "% Variance", 
        cex.names = 1.5, cex.axis = 1.5, cex.lab = 1.5, cex.main = 1.5)

```

```{r, fig.height=5, fig.width=6}
blues <- c("#d0d1e6","#a6bddb","#74a9cf","#3690c0", "#0570b0", "#045a8d", "#023858")

data.frame(pca_clr$x, SAMPLE = rownames(pca_clr$x)) %>% 
  select(PC1, PC2, PC3, SAMPLE) %>% 
  separate(SAMPLE, c("STN_ORDER", "DEPTH_ORDER"), sep = "_", remove = FALSE) %>% left_join(env_info_mod) %>% 
  mutate(Depth = as.numeric(DEPTH_ORDER)) %>% 
  ## Generate PCA plot
  ggplot(aes(x = PC1, y = PC2, shape = TRANSECT, fill = Depth)) +
  geom_hline(yintercept = 0) + geom_vline(xintercept = 0, color = "#525252") +
  geom_point(size = 2.5, stroke = 0.4) +
  ylab(paste0('PC2 ',round(variance_clr[2]*100,2),'%')) +
  xlab(paste0('PC1 ',round(variance_clr[1]*100,2),'%')) +
  scale_shape_manual(values = c(21, 23, 24, 25)) +
  scale_fill_stepsn(colors = blues, breaks = c(50, 100, 300, 500, 1000, 1500, 2000)) +
  theme_bw() +
  theme(axis.text = element_text(color="black", size=10),
        legend.title = element_blank(),
        axis.title = element_text(color="black", size=8),
        legend.text = element_text(color = "black", size = 8, face = "bold"),
        plot.margin = margin(2, 1, 2, 1, "cm")) +
  guides(shape = guide_legend(override.aes = list(fill = "black")))
```

> PCA results to consider. Communities are more similar to one another at depth, surface samples are influenced by other parameters. Should do RDA with all data and then shallower than 500 meters.

# RDA analysis

RDA reveals relationships among the sites and the environmental parameters. We can explore the relationship among all ASV signatures by site, and then model the effect that each environmental parameter has on it.

### All samples

```{r}
# | eval: false
head(env_info_mod)

categorical <- env_info_mod %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, TRANSECT) %>% distinct()

metadata_for_rda_0 <- env_info_mod %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, PO4, SIL) %>% distinct() %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(SAMPLE_NAME, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

standarize_metadata <- decostand(metadata_for_rda_0, MARGIN = 2, method = "standardize")
range(standarize_metadata)
# head(categorical)
metadata_normalized <- categorical %>% 
  right_join(as.data.frame(standarize_metadata) %>% rownames_to_column(var = "SAMPLE_NAME")) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME")
head(metadata_normalized)
```

Isolate all samples from ASV data

```{r}
# | eval: false
asv_wide_avg_reps_matrix <- asv_long_clean_wtax %>% 
  # First average ASV sequence count across replicates
  group_by(TRANSECT, STN_ORDER, DEPTH_ORDER, FeatureID, stn, nisk, Station, Niskin, DEPTH, TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT, Taxon, DIVISION, SUPERGROUP_CLASS, Supergroup, Division, Subdivision, Class) %>% 
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>% 
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>% 
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix()
```

```{r}
# | eval: false
# head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>% 
  as.matrix
  
clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))

rownames(clr_asv_matrix)
rownames(metadata_normalized)
```

```{r}
# | eval: false
rda_all_gom <- rda(clr_asv_matrix ~ ., data = metadata_normalized, na.action = na.exclude)
rda_all_gom
```

22% of the environmental variables explained changes in community composition for all the samples.

```{r}
# | eval: false
adj_rda_out <- RsquareAdj(rda_all_gom)$adj.r.squared
# RsquareAdj(rda_all_gom)
adj_rda_out
anova.cca(rda_all_gom, step = 100, by = "term")
```

> Model: rda(formula = clr_asv_matrix \~ TRANSECT + DEPTH + NH4 + NO2 + OXYGEN + PO4 + SALINITY + SIL + TEMP, data = metadata_normalized, na.action = na.exclude) 
Df Variance F Pr(\>F)\
> TRANSECT 2 68.59 1.4530 0.003 \*\* 
DEPTH 1 115.91 4.9113 0.001 ***
NH4 1 27.20 1.1523 0.149\
> NO2 1 61.65 2.6121 0.002** 
OXYGEN 1 52.45 2.2225 0.001* **
PO4 1 82.89 3.5120 0.001** *
SALINITY 1 50.12 2.1237 0.001* **
SIL 1 38.13 1.6155 0.011 \*\
> TEMP 1 35.13 1.4886 0.019 \*\
> Residual 79 1864.52\
> --- Signif. codes: 0 ‘**\*’ 0.001 ‘\*\*’ 0.01 ‘\*’ 0.05 ‘.’ 0.1 ‘ ’ 1

So many parameters are significant here.

## Coastal and surface samples

RDA for shallow and transect 1

```{r}
head(env_info_mod)
# station_jgi

surf_categorical <- env_info_mod %>% 
  filter(TRANSECT == "transect1") %>% 
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, PROX_TOSHORE) %>% distinct()

metadata_for_rda_0_surface <- env_info_mod %>% 
  filter(TRANSECT == "transect1") %>%  
  filter(Sample_Ctrl == "Sample") %>% 
  filter(!(SAMPLES %in% samples_too_low)) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>% 
  select(SAMPLE_NAME, TEMP, SALINITY, OXYGEN, NH4, NO2, PO4, SIL, PROK_COUNT,DIST_OUTFLOW) %>% distinct() %>% 
  pivot_longer(cols = -c(SAMPLE_NAME), names_to = "VARIABLES", values_to = "VALUE") %>% 
  group_by(SAMPLE_NAME, VARIABLES) %>% 
    summarise(MEAN_VAL = mean(VALUE)) %>% 
  pivot_wider(names_from = VARIABLES, values_from = MEAN_VAL) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

# rownames(metadata_for_rda_0_surface)
standarize_metadata_surf <- decostand(metadata_for_rda_0_surface, MARGIN = 2, method = "standardize")
# range(standarize_metadata)

metadata_normalized_surf <- surf_categorical %>% 
  right_join(as.data.frame(standarize_metadata_surf) %>% rownames_to_column(var = "SAMPLE_NAME")) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME")

rownames(metadata_normalized_surf)

subset_sample_surf <- as.character(unique(surf_categorical$SAMPLE_NAME))
```

```{r}
# head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0_surf <- data.frame(t(asv_wide_avg_reps_matrix)) %>% 
  rownames_to_column(var = "SAMPLE_NAME") %>% 
  filter(SAMPLE_NAME %in% subset_sample_surf) %>% 
  arrange(SAMPLE_NAME) %>% 
  column_to_rownames(var = "SAMPLE_NAME") %>% 
  as.matrix
  
clr_asv_matrix_surf <- data.frame(compositions::clr(asv_wide_avg_rda_0_surf))

rownames(clr_asv_matrix_surf)
rownames(metadata_normalized_surf)
```

```{r}
# rda_surf_gom <- rda(clr_asv_matrix_surf ~ ., data = metadata_normalized_surf, na.action = na.exclude)
# rda_surf_gom
```

60% of the environmental variables explained changes in community composition for the surface samples.

```{r}
# adj_rda_surf <- RsquareAdj(rda_surf_gom)$adj.r.squared
# adj_rda_surf
# anova.cca(rda_surf_gom, step = 100, by = "term")
```

# Map

```{r}
# install.packages(c("googleway", "ggrepel", 
# "ggspatial", "libwgeom", "sf", "rnaturalearth", "rnaturalearthdata"))

library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

world <- ne_countries(scale='medium',returnclass = 'sf')
class(world)
```

## Map of whole Gulf region

```{r}
ggulf <- ggplot(data = world) +
  geom_sf(aes(fill = region_wb)) +
  annotate(geom = "text", x = -90, y = 26, label = "", 
     fontface = "italic", color = "grey22", size = 6) +
  coord_sf(xlim = c(-102.15, -74.12), ylim = c(7.65, 33.97), expand = FALSE) +
  scale_fill_viridis_d(option = "plasma") +
  theme(legend.position = "none", axis.title.x = element_blank(), 
     axis.title.y = element_blank(), panel.background = element_rect(fill = "azure"), 
     panel.border = element_rect(fill = NA))
ggulf
```

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = world) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
  # Set lat long region to show:
  # coord_sf(xlim = c(-90, -86.5), ylim = c(27.5, 30), expand = FALSE) + # Region of study
  coord_sf(xlim = c(-110, -70), ylim = c(7, 47), expand = FALSE) + # Larger region
  # Add data points for stations
  # geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = "pink", color = "grey10", alpha = 0.5) + #inset
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(legend.position = "none",
    axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA))
```

```{r, fig.width=7, fig.height=7}
# MAP
# ggplot(data = world) +
#   geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.8) +
#   # Set lat long region to show
#   # coord_sf(xlim = c(-90, -86.5), ylim = c(27.5, 30), expand = FALSE) + # Region of study
#   coord_sf(xlim = c(-100, -84), ylim = c(21, 33), expand = FALSE) + # Larger region
#   # Add data points for stations
#   geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
#   geom_rect(xmin = -90, xmax = -86.5, ymin = 27.5, ymax = 30, fill = NA, color = "grey10") + #inset
#   labs(y = "Latitude (°N)", x = "Longitude (°W)") +
#   scale_x_reverse() +
#   theme_classic() +
#   theme(legend.position = "none",
#     axis.text = element_text(size = 10, colour="black"),
#         panel.background = element_rect(fill = "#B4C7D9"),
#         panel.border = element_rect(fill = NA))
```

# ODV plots in R

Extract points you want and dimensions. Interpolate, subtract bathymetry if needed and move along from there.

```{r}
# Manually extracted hexidecimal ODV colour palette
ODV_colours <- c("#feb483", "#d31f2a", "#ffc000", "#27ab19", "#0db5e6", "#7139fe", "#d16cfa")
```

## Surface

### Temperature & Long

```{r}
head(asv_long_clean_wtax)
```

```{r}
temp_odv <- asv_long_clean_wtax %>% 
  select(Transect, TRANSECT, Station, DEPTH, stn, Latitude, Longitude, 
         TEMP, SALINITY, OXYGEN, NH4, NO2, NO3, PO4, SIL, PROK_COUNT) %>% 
  distinct()
```

Non-interpolated scatterplot

```{r}
# temp_odv %>%
#   ggplot(aes(x = Longitude, y = DEPTH)) +
#   # geom_point(aes(color = TEMP), size = 2) +
#   # geom_point(aes(color = PROK_COUNT), size = 2) +
#   scale_colour_gradientn(colours = rev(ODV_colours)) +
#   labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
#   scale_y_reverse() + scale_x_reverse() +
#   theme_classic()
```

Interpolate step

```{r}
library(MBA)
# ?mba.surf
head(temp_odv)
# Interpolate
temp_mba <- mba.surf(temp_odv[c("Longitude", "DEPTH", "TEMP")], no.X = 300, no.Y = 300, extend = T)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(temp_mba$xyz.est$z) <- list(temp_mba$xyz.est$x, temp_mba$xyz.est$y)
# head(temp_mba)

temp_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c('LONGITUDE', 'DEPTH'), value.name = 'TEMPERATURE') %>% 
  filter(DEPTH > 0) %>%
  mutate(TEMPERATURE = round(TEMPERATURE, 1))

```

```{r}
ggplot(data = temp_interpol, aes(x = LONGITUDE, y = DEPTH)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  scale_fill_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_y_reverse() + scale_x_reverse() +
  theme_classic() +
  coord_cartesian(expand = F)
```

### Surface params, lat, long

```{r}
head(asv_long_clean_wtax)
```

```{r}
surface_odv <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, SURFACE_TEMP, Transect, TRANSECT) %>%
  summarise(ASV_COUNT = n())
```

Non-interpolated scatterplot

```{r}
surface_odv %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
  geom_point(aes(color = SURFACE_TEMP), size = 2) +
  scale_colour_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic()
```

Interpolate step

```{r}
library(MBA)
# ?mba.surf
head(surface_odv)

# Interpolate
temp_mba <- mba.surf(surface_odv[c("Longitude", "Latitude", "SURFACE_TEMP")], 
                     no.X = 300, no.Y = 300, extend = F)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(temp_mba$xyz.est$z) <- list(temp_mba$xyz.est$x, temp_mba$xyz.est$y)
# head(temp_mba)
# 
temp_surface_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c("Longitude", "Latitude"), value.name = 'TEMPERATURE') %>% 
  mutate(TEMPERATURE = round(TEMPERATURE, 1))
head(temp_surface_interpol)
```

```{r}
surface_temperature <- temp_surface_interpol %>% 
  ggplot(aes(x = Longitude, y = Latitude)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  # Add data points
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  coord_cartesian(expand = F) +
  theme(axis.text = element_text(size = 10, colour="black"))
  
```

##### Temperature with map

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = (world %>% filter(subunit == "United States"))) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
  # Set lat long region to show:
  coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
  # Overlay with temperature 
  geom_raster(data = temp_surface_interpol, aes(fill = TEMPERATURE, x = Longitude, y = Latitude)) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
  # Add data points for stations
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)

```

##### Temperature Surface

```{r, fig.width=7, fig.height=7}
# MAP
ggplot(data = (world %>% filter(subunit == "United States"))) +
  geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
  # Set lat long region to show:
  coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
  # Overlay with temperature 
  geom_raster(data = temp_surface_interpol, aes(fill = TEMPERATURE, x = Longitude, y = Latitude)) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = temp_surface_interpol, aes(z = TEMPERATURE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
  # Add data points for stations
  geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
  geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 1, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)

```

#### Fxn for ther surface params, lat, long

Set up a data frame with all parameters as columns.

```{r}
surface_odv <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SAL = mean(SALINITY),
            SURFACE_OXY = mean(OXYGEN),
            SURFACE_NH4 = mean(NH4),
            SURFACE_NO2 = mean(NO2),
            SURFACE_PO4 = mean(PO4),
            SURFACE_PROK = mean(PROK_COUNT),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4, SURFACE_PROK, Transect, TRANSECT) %>%
  summarise(ASV_COUNT = n())
```

```{r}
# Put PARAM in quotes
interpolate_lat_long_env <- function(df, PARAM, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)], 
                     no.X = 300, no.Y = 300, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Latitude"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Latitude, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1))
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature 
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = "black") +
    # Add data points for stations
    geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
    geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 1, color= "black") +
    scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)
}
```

Parameters to use in function: SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4

```{r}
surf_temp <- interpolate_lat_long_env(surface_odv, "SURFACE_TEMP", "Temperature (°C)")
surf_sal <- interpolate_lat_long_env(surface_odv, "SURFACE_SAL", "Salinity")
surf_oxy <- interpolate_lat_long_env(surface_odv, "SURFACE_OXY", "Oxygen")
```

now: SURFACE_NH4, SURFACE_NO2, SURFACE_PO4

```{r}
# expression("S"["in"] ~ "[W" ~ m^-2~"]")
surf_nh4 <- interpolate_lat_long_env(surface_odv, "SURFACE_NH4", expression("NH"["4"]))
surf_no2 <- interpolate_lat_long_env(surface_odv, "SURFACE_NO2", expression("NO"["2"]))
surf_po4 <- interpolate_lat_long_env(surface_odv, "SURFACE_PO4", expression("PO"["4"]))
```

```{r}
# surf_prok <- interpolate_lat_long_env(surface_odv, "SURFACE_PROK", expression("PO"["4"]))
# surf_prok
```

```{r, fig.height=12, fig.width=17}
surf_temp + surf_sal + surf_oxy + surf_nh4 + surf_no2 + surf_po4 + plot_layout(guides = "collect")
```

```{r}
# envs_surface_interpol <- reshape2::melt(temp_mba$xyz.est$z, varnames = c("Longitude", "Latitude"), value.name = 'VALUE') %>% 
#   mutate(VALUE = round(VALUE, 1))
# head(envs_surface_interpol)
# salinity_surf <- envs_surface_interpol %>% 
#   ggplot(aes(x = Longitude, y = Latitude)) +
#   geom_raster(aes(fill = VALUE)) +
#   geom_contour(aes(z = VALUE), binwidth = 2, colour = "black", alpha = 0.2) +
#   geom_contour(aes(z = VALUE), breaks = 20, colour = "black") +
#   # Add data points
#   geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
#   geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
#                                                  hjust = 0.5, vjust = 1, color= "black") +
#   scale_fill_gradientn(colours = rev(ODV_colours)) +
#   labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = "Salinity") +
#   scale_x_reverse() +
#   theme_classic() +
#   coord_cartesian(expand = F) +
#   theme(axis.text = element_text(size = 10, colour="black"))
# salinity_surf

# (class(envs_surface_interpol$VALUE))
```

## Repeat with ASV

Surface lat/long

```{r}
surface_odv <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SAL = mean(SALINITY),
            SURFACE_OXY = mean(OXYGEN),
            SURFACE_NH4 = mean(NH4),
            SURFACE_NO2 = mean(NO2),
            SURFACE_PO4 = mean(PO4),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4, Transect, TRANSECT) %>%
  summarise(ASV_COUNT = n())
```

Interpolate function for ASVs:

```{r}
# Put PARAM in quotes
interpolate_lat_long_asv <- function(df, PARAM, TITLE){
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "Latitude", PARAM)], 
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Latitude"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Latitude, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1))
  # MAP
  ggplot(data = (world %>% filter(subunit == "United States"))) +
    geom_sf(color = "black", linewidth = 0.6, fill = "grey10", alpha = 0.6) +
    # Set lat long region to show:
    coord_sf(xlim = c(-90, -86.75), ylim = c(27.5, 29.5), expand = FALSE) +
    # Overlay with temperature 
    geom_raster(data = envs_interpol, aes(fill = VALUE, x = Longitude, y = Latitude)) +
    geom_contour(data = envs_interpol, aes(z = VALUE, x = Longitude, y = Latitude), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(data = envs_interpol, aes(z = VALUE,x = Longitude, y = Latitude), breaks = 20, colour = NA) +
    # Add data points for stations
    geom_point(data = surface_odv, aes(x = Longitude, y = Latitude)) +
    geom_text(data = surface_odv, aes(x = Longitude, y = Latitude, label = surface_odv$stn), position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 1, color= "black") +
    scale_fill_fermenter(type = "seq", palette = "Reds", direction = 1, na.value = NA) +
    # scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Latitude (°N)", x = "Longitude (°W)", fill = TITLE) +
  scale_x_reverse() +
  theme_classic() +
  theme(axis.text = element_text(size = 10, colour="black"),
        panel.background = element_rect(fill = "#B4C7D9"),
        panel.border = element_rect(fill = NA),
        aspect.ratio = 1)
}
```

### Total ASVs

```{r}
surf_all_asv<- interpolate_lat_long_asv(surface_odv, "ASV_COUNT", "Total ASVs")
surf_all_asv
```

```{r}
surface_odv_tax_simple <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, Transect, TRANSECT, Supergroup_simplified) %>%
  summarise(ASV_COUNT = n())

unique(surface_odv_tax_simple$Supergroup_simplified)
```

#### Taxa specific ASV profiles

```{r, fig.height=10, fig.width=18}
interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Gyrista")), 
                         "ASV_COUNT", "Gyrista") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Dinoflagellata")), 
                         "ASV_COUNT", "Dinoflagellate") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Ciliophora")), 
                         "ASV_COUNT", "Ciliates") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Radiolaria")), 
                         "ASV_COUNT", "Radiolaria") +
  interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "Haptista")), 
                         "ASV_COUNT", "Haptista") + plot_layout()
```

```{r}
dinos <- interpolate_lat_long_asv((surface_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Ciliophora")), 
                         "ASV_COUNT", "Total ASVs")
```

```{r}
surface_odv_tax_class <- asv_long_clean_wtax %>% 
  # Isolate surface & average across replicates
  filter(DEPTH < 5) %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, stn, Station, Latitude, Longitude, Transect, TRANSECT) %>% 
  summarise(SURFACE_TEMP = mean(TEMP),
            SURFACE_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, Transect, TRANSECT, SUPERGROUP_CLASS) %>%
  summarise(ASV_COUNT = n())

sort(unique(surface_odv_tax_class$SUPERGROUP_CLASS))
```

```{r}
# "Dinoflagellata-Syndiniales"
# "Dinoflagellata-Dinophyceae"
# "Dinoflagellata-Dinophyta" 
# "Dinoflagellata-Noctilucophyceae"
interpolate_lat_long_asv((surface_odv_tax_class %>% 
                            filter(SUPERGROUP_CLASS == "Dinoflagellata-Dinophyceae")), 
                         "ASV_COUNT", "Total Dinophyceae ASVs") +
  interpolate_lat_long_asv((surface_odv_tax_class %>% 
                            filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), 
                         "ASV_COUNT", "Total Syndiniales ASVs") +
    interpolate_lat_long_asv((surface_odv_tax_class %>% 
                            filter(SUPERGROUP_CLASS == "Dinoflagellata-Syndiniales")), 
                         "ASV_COUNT", "Total Syndiniales ASVs")
```

## Repeat with cell counts

```{r}

```

## Depth

Depth params, lat, long

```{r}
head(asv_long_clean_wtax)
```

```{r}
depth_odv <- asv_long_clean_wtax %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT, DEPTH) %>% 
  summarise(TEMP = mean(TEMP),
            SAL = mean(SALINITY),
            OXY = mean(OXYGEN),
            NH4 = mean(NH4),
            NO2 = mean(NO2),
            PO4 = mean(PO4),
            MEAN_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, TEMP, SAL, 
           OXY, NH4, NO2, PO4, Transect, TRANSECT, DEPTH) %>%
  summarise(ASV_COUNT = n())
head(depth_odv)
```

Non-interpolated scatterplot

```{r}
depth_odv %>% 
  filter(TRANSECT == "transect1") %>% 
  ggplot(aes(x = Longitude, y = DEPTH)) +
  geom_point(aes(color = TEMP), size = 2) +
  scale_colour_gradientn(colours = rev(ODV_colours)) +
  labs(y = "Depth (m)", x = "Longitude (°W)", colour = "Temperature (°C)") +
  scale_x_reverse() + scale_y_reverse() +
  theme_classic()
```

Interpolate step

```{r}
# library(MBA)
# ?mba.surf
head(depth_odv)

one <- depth_odv %>% 
  filter(TRANSECT == "transect1")
# head(one)
# Interpolate
trn1_mba <- mba.surf(one[c("Longitude", "DEPTH", "TEMP")], 
                     no.X = 300, no.Y = 300, extend = F)

# range(one$Longitude)
max_long <- max(one$Longitude)
min_long <- min(one$Longitude)
max_depth <- max(one$DEPTH)
# max_depth <- max(one$DEPTH)
```

X, y, z are Longitude", "DEPTH", "TEMP"

```{r}
dimnames(trn1_mba$xyz.est$z) <- list(trn1_mba$xyz.est$x, trn1_mba$xyz.est$y)
# head(temp_mba)
# 
temp_trn1_interpol <- reshape2::melt(trn1_mba$xyz.est$z, varnames = c("Longitude", "Depth"), value.name = 'TEMPERATURE') %>% 
  mutate(TEMPERATURE = round(TEMPERATURE, 1)) %>% 
  filter(Longitude > 88 & Longitude < 90)

# 88.136 89.460
# head(temp_trn1_interpol)
# range(temp_trn1_interpol$Depth)
```

```{r}
temp_trn1_interpol %>% 
  ggplot(aes(x = Longitude, y = Depth)) +
  geom_raster(aes(fill = TEMPERATURE)) +
  geom_contour(aes(z = TEMPERATURE), binwidth = 2, colour = "black", alpha = 0.2) +
  geom_contour(aes(z = TEMPERATURE), breaks = 20, colour = "black") +
  # Add data points
  geom_point(data = one, aes(x = Longitude, y = DEPTH)) +
  geom_text(data = one, aes(x = Longitude, y = -30, label = one$stn), position = position_dodge(width = 0.1),
                                                 hjust = 0.5, vjust = 0, color= "black") +
  scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
  labs(y = "Depth (m)", x = "Longitude (°W)", fill = "Temperature (°C)") +
  scale_x_reverse() + scale_y_reverse() +
  theme_classic() +
  # coord_cartesian(expand = F) +
  theme(axis.text = element_text(size = 10, colour="black"))
  
```

### Fxn for depth parameters

```{r}

```

```{r}
# df_all <- depth_odv
# transect <- "transect1"
# PARAM <- "TEMP"
# TITLE <- "Temperature"
# Put PARAM in quotes

interpolate_depth_long_env <- function(df_all, transect, PARAM, TITLE){
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "DEPTH", PARAM)], 
                     no.X = 300, no.Y = 300, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = "black", alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = "black") +
    # Add data points
    geom_point(data = df, aes(x = Longitude, y = DEPTH)) +
    geom_text(data = df, 
              aes(x = Longitude, y = -30, label = df$stn), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_gradientn(colours = rev(ODV_colours), na.value=NA) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse() + scale_y_reverse() +
    theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_rect(fill = NA),
          aspect.ratio = 1)
  }
```

Parameters to use in function: SURFACE_TEMP, SURFACE_SAL, SURFACE_OXY, SURFACE_NH4, SURFACE_NO2, SURFACE_PO4

df_all, transect, PARAM, TITLE

```{r}
colnames(depth_odv)
unique(depth_odv$TRANSECT)
```

Transect 1

```{r, fig.height=12, fig.width=17}
(interpolate_depth_long_env(depth_odv, "transect1","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, "transect1","SAL", "Salinity")) +
  (interpolate_depth_long_env(depth_odv, "transect1","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(depth_odv, "transect1","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(depth_odv, "transect1","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(depth_odv, "transect1","PO4", expression("PO"["4"]))) +
    plot_layout(guides = "collect")
```

Transect 2

```{r, fig.height=12, fig.width=17}
(interpolate_depth_long_env(depth_odv, "transect2","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(depth_odv, "transect2","SAL", "Salinity")) +
  (interpolate_depth_long_env(depth_odv, "transect2","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(depth_odv, "transect2","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(depth_odv, "transect2","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(depth_odv, "transect2","PO4", expression("PO"["4"]))) +
    plot_layout(guides = "collect")
```

### Repeat for upper water column

```{r}
# head(depth_odv)
# range(depth_odv$DEPTH)
shallow <- depth_odv %>% 
  filter(DEPTH <= 300)
```

Transect 1

```{r, fig.height=12, fig.width=17}
(interpolate_depth_long_env(shallow, "transect1","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(shallow, "transect1","SAL", "Salinity")) +
  (interpolate_depth_long_env(shallow, "transect1","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(shallow, "transect1","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(shallow, "transect1","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(shallow, "transect1","PO4", expression("PO"["4"]))) +
    plot_layout(guides = "collect")

```

Transect 2

```{r, fig.height=12, fig.width=17}
(interpolate_depth_long_env(shallow, "transect2","TEMP", "Temperature (°C)")) +
  (interpolate_depth_long_env(shallow, "transect2","SAL", "Salinity")) +
  (interpolate_depth_long_env(shallow, "transect2","OXY", "Oxygen")) +
  #
  (interpolate_depth_long_env(shallow, "transect2","NH4", expression("NH"["4"]))) +
  (interpolate_depth_long_env(shallow, "transect2","NO2", expression("NO"["2"]))) +
  (interpolate_depth_long_env(shallow, "transect2","PO4", expression("PO"["4"]))) +
    plot_layout(guides = "collect")

```

## Repeat with ASVs

```{r}
# Put PARAM in quotes
interpolate_depth_long_asv <- function(df_all, transect, PARAM, TITLE){
  # Isolate transect of interest
  df <- df_all %>% 
    filter(TRANSECT == transect)
  #
  # Get specific transect Longitude
  max_long <- round(max(df$Longitude), 0) + 1
  min_long <- round(min(df$Longitude), 0) - 1
  #
  # Interpolate
  output_mba <- mba.surf(df[c("Longitude", "DEPTH", PARAM)], 
                     no.X = 150, no.Y = 150, extend = F)
  # Extract interpolated values
  dimnames(output_mba$xyz.est$z) <- list(output_mba$xyz.est$x, output_mba$xyz.est$y)
  #
  envs_interpol <- reshape2::melt(output_mba$xyz.est$z, 
                                  varnames = c("Longitude", "Depth"), 
                                  value.name = PARAM) %>% 
    select(Longitude, Depth, VALUE = !!PARAM) %>% 
    mutate(VALUE = round(VALUE, 1)) %>% 
    filter(Longitude > min_long & Longitude < max_long)
  #
  # Plot
  envs_interpol %>% 
    ggplot(aes(x = Longitude, y = Depth)) +
    geom_raster(aes(fill = VALUE)) +
    geom_contour(aes(z = VALUE), binwidth = 2, colour = NA, alpha = 0.2) +
    geom_contour(aes(z = VALUE), breaks = 20, colour = NA) +
    # Add data points
    geom_point(data = df, aes(x = Longitude, y = DEPTH)) +
    geom_text(data = df, 
              aes(x = Longitude, y = -30, label = df$stn), 
              position = position_dodge(width = 0.1),
              hjust = 0.5, vjust = 0, color= "black") +
    scale_fill_fermenter(type = "seq", palette = "Reds", direction = 1, na.value = NA) +
    labs(y = "Depth (m)", x = "Longitude (°W)", fill = TITLE) +
    scale_x_reverse() + scale_y_reverse() +
    theme_classic() +
    theme(axis.text = element_text(size = 10, colour="black"),
          panel.border = element_rect(fill = NA),
          aspect.ratio = 1)
}
```

Transect 1 with all ASVs

```{r}
interpolate_depth_long_asv(depth_odv, "transect1","ASV_COUNT", "Total ASVs")
```

Transects

```{r, fig.width=16, fig.height=7}
interpolate_depth_long_asv(depth_odv, "transect1","ASV_COUNT", "Total ASVs") +
interpolate_depth_long_asv(depth_odv, "transect2","ASV_COUNT", "Total ASVs") + 
  interpolate_depth_long_asv(depth_odv, "transect3","ASV_COUNT", "Total ASVs") +
  plot_layout(nrow = 1)

```

```{r}
depth_odv_tax_simple <- asv_long_clean_wtax %>% 
  group_by(FeatureID, Taxon, SUPERGROUP_CLASS, Supergroup_simplified, stn, Station, Latitude, Longitude, Transect, TRANSECT, DEPTH) %>% 
  summarise(MEAN_SEQ = mean(SEQUENCE_COUNT)) %>% 
  ungroup() %>% 
  # Get totals
  group_by(stn, Station, Latitude, Longitude, Transect, TRANSECT, DEPTH, Supergroup_simplified) %>%
  summarise(ASV_COUNT = n())

unique(depth_odv_tax_simple$Supergroup_simplified)
```

interpolate_depth_long_asv

by taxa, use `depth_odv_tax_simple`

```{r, fig.height=10, fig.width=18}
interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Gyrista")), "transect1", 
                         "ASV_COUNT", "Gyrista") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Dinoflagellata")), "transect1",
                         "ASV_COUNT", "Dinoflagellate") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Ciliophora")), "transect1",
                         "ASV_COUNT", "Ciliates") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "TSAR-Radiolaria")), "transect1",
                         "ASV_COUNT", "Radiolaria") +
  interpolate_depth_long_asv((depth_odv_tax_simple %>% 
                            filter(Supergroup_simplified == "Haptista")), "transect1",
                         "ASV_COUNT", "Haptista") + plot_layout()
```

# Identify ASVs with highest number of sequences throughout data

```{r}
dominant_taxa <- asv_long_avg_wtax %>% 
  group_by(Taxon, Supergroup_simplified, DEPTH_ORDER) %>% 
  summarise(SUM_SEQS = sum(MEAN_REPS_seq)) %>% 
  ungroup() %>% 
  group_by(Supergroup_simplified) %>% 
  slice_max(order_by = SUM_SEQS, n = 15)

```

# Shannon Diversity

Surface stations only

```{r}
asv_wide_surf_matrix <- asv_long_clean_wtax %>%
  filter(DEPTH < 5) %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  # First average ASV sequence count across replicates
  group_by(FeatureID, SAMPLE_NAME) %>%
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>%
  as.matrix()

asv_wide_all_matrix <- asv_long_clean_wtax %>%
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  # First average ASV sequence count across replicates
  group_by(FeatureID, SAMPLE_NAME) %>%
  summarise(MEAN_REPS_seq = mean(SEQUENCE_COUNT)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, MEAN_REPS_seq) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = MEAN_REPS_seq, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>%
  as.matrix()
```

Calculate Shannon diversity

```{r, fig.width=5, fig.height=3}
# ?diversity()
surface_shannon <- diversity(asv_wide_surf_matrix, index = "shannon", MARGIN = 2)
data.frame(surface_shannon) %>% 
  rownames_to_column(var = "stn_niskin") %>% 
  separate(stn_niskin, into = c("STN", "NISKIN"), remove = FALSE) %>% 
  mutate(STN_ORDER = factor(STN, levels = offshore_on_shore_order),
         TRANSECT = factor(STN, levels = offshore_on_shore_order, labels = transect_labels)) %>% 
  ggplot(aes(x = STN_ORDER, y = surface_shannon)) +
  geom_point(shape = 21, fill = "grey30", color = "black", size = 3) +
  facet_grid(cols = vars(TRANSECT), scales = "free", space = "free") +
  theme_classic() + 
  theme(
    legend.position = "none",
    legend.title = element_blank(),
    legend.text = element_text(color = "black", face = "bold"),
    strip.background = element_blank(),
    panel.background = element_rect(color = "grey80"),
    strip.text.y = element_text(angle = 0, color = "black", face = "bold"),
    axis.title = element_text(color = "black", face = "bold"),
    axis.text.y = element_text(color = "black", face = "bold"),
    axis.text.x = element_text(color = "black", face = "bold")) +
    labs(x = "Station", y = "Shannon Diversity")
  
  
```

# Targeted CLR for taxonomic groups

```{r}
head(asv_long_avg_wtax)
unique(asv_long_avg_wtax$Supergroup_simplified)
```


Make matrix:
TSAR-Gyrista
TSAR-Radiolaria
```{r}
asv_wide_gyrista_matrix <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 150) %>% 
  filter(Supergroup_simplified == "TSAR-Gyrista") %>% 
  filter(TRANSECT == "transect1") %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  group_by(FeatureID, SAMPLE_NAME, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, SUM_SEQ) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix


asv_wide_radio_matrix <- asv_long_avg_wtax %>%
  filter(MEAN_REPS_seq >= 500 ) %>% 
  filter(Supergroup_simplified == "TSAR-Radiolaria") %>% 
  filter(TRANSECT == "transect1") %>% 
  unite(SAMPLE_NAME, STN_ORDER, DEPTH_ORDER, sep = "_") %>%
  group_by(FeatureID, SAMPLE_NAME, Supergroup_simplified) %>%
  summarise(SUM_SEQ = sum(MEAN_REPS_seq)) %>%
  ungroup() %>%
  select(FeatureID, SAMPLE_NAME, SUM_SEQ) %>%
  pivot_wider(names_from = SAMPLE_NAME, values_from = SUM_SEQ, values_fn = mean, values_fill = 0) %>%
  column_to_rownames(var = "FeatureID") %>% 
  as.matrix
# head(asv_wide_gyrista_matrix)
```
```{r}
clr_gyrista <- vegan::decostand(asv_wide_gyrista_matrix, MARGIN = 2, method = "rclr")
colnames(clr_gyrista) <- colnames(asv_wide_gyrista_matrix)
rownames(clr_gyrista) <- rownames(asv_wide_gyrista_matrix)

clr_radio <- vegan::decostand(asv_wide_radio_matrix, MARGIN = 2, method = "rclr")
colnames(clr_radio) <- colnames(asv_wide_radio_matrix)
rownames(clr_radio) <- rownames(asv_wide_radio_matrix)

colnames(asv_long_avg_wtax)
```

```{r}
head(env_info_mod)
```

CLR plot
```{r}
clr_radio_df <- data.frame(clr_radio) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  pivot_longer(cols = -c(FeatureID), names_to = "STN_NI", values_to = "CLR") %>% 
  separate(STN_NI, into = c("STN", "DEPTH_ORDER"), remove = FALSE) %>% 
  left_join(tax_key) %>% distinct()
# unique(clr_radio_df$Family)
   
dim(clr_radio);dim(clr_radio_df)

clr_gyrista_df <- data.frame(clr_gyrista) %>% 
  rownames_to_column(var = "FeatureID") %>% 
  pivot_longer(cols = -c(FeatureID), names_to = "STN_NI", values_to = "CLR") %>% 
  separate(STN_NI, into = c("STN", "DEPTH_ORDER"), remove = FALSE) %>% 
  left_join(tax_key) %>% distinct() %>% 
  filter(!is.na(Class))
  
dim(clr_gyrista);dim(clr_gyrista_df)
head(clr_gyrista_df)
unique(clr_gyrista_df$Class)
```

```{r, fig.height=5, fig.width=12}
depth_order <- as.character(unique(sort(as.numeric(clr_gyrista_df$DEPTH_ORDER))))
# View(clr_gyrista_df)
unique(clr_gyrista_df$Class)
rm <- c("Hyphochytriomyceta","MOCH","Pinguiophyceae","Pelagophyceae")
clr_gyrista_df %>% 
  filter(!(Class %in% rm)) %>% 
  # left_join(env_info_mod) %>% 
  # filter(TRANSECT == "transect1") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "white") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```

```{r, fig.height=5, fig.width=13}
depth_order <- as.character(unique(sort(as.numeric(clr_radio_df$DEPTH_ORDER))))

unique(clr_radio_df$Class)
# View(clr_radio_df)
clr_radio_df %>% 
  # left_join(env_info_mod) %>% 
  # filter(TRANSECT == "transect1") %>% 
  filter(Class != "RAD") %>% 
  filter(Class != "Radiolaria") %>% 
  mutate(depth = factor(DEPTH_ORDER, levels = rev(depth_order))) %>%
  ggplot(aes(x = depth, y = FeatureID, fill = CLR)) +
  geom_tile(color = "white") +
  facet_grid(rows = vars(STN), cols = vars(Class), space = "free", scale = "free") +
  theme(axis.text.x = element_blank()) +
  scale_fill_fermenter(type = "div", palette = 4) +
  coord_flip() +
  theme(axis.ticks = element_blank(),
        strip.background = element_blank()) +
  labs(x = "Depth (m)", y = "")

```


# Indicator Species

https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html

```{r}
library(indicspecies)
```

Format input table

```{r}
head(asv_long_avg_wtax)
```

```{r}
# save(asv_long_avg_wtax, file = "asv_gom_input_indicator.RData")
head(asv_long_avg_wtax)
# hist(asv_long_avg_wtax$MEAN_REPS_seq)
```

Below is taking much too long on the HPRC and not working.

```{r}

```


```{r}
indic_input <- asv_long_avg_wtax %>%
  # filter(DEPTH < 5) %>%
  filter(MEAN_REPS_seq > 5) %>%
  filter(TRANSECT == "transect1") %>%
  # filter(stn == "S1" | stn == "S2") %>% #TESTING
  select(SAMPLE_ID, DEPTH_ORDER, TEMP, SALINITY, MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fill = 0)
head(indic_input)
```

Try paring down the data more.

```{r}
indic_input <- asv_long_avg_wtax %>%
  filter(DEPTH < 500) %>% # Surface only
  filter(MEAN_REPS_seq >= 10) %>% # Subset of tASVs
  filter(TRANSECT == "transect1") %>% # Only the 1 transect
  # filter(stn == "S1" | stn == "S2") %>% #TESTING
  select(SAMPLE_ID, DEPTH, TEMP, SALINITY, MEAN_REPS_seq, FeatureID) %>%
  distinct() %>%
  pivot_wider(names_from = FeatureID, values_from = MEAN_REPS_seq, values_fill = 0)
head(indic_input)
# ?multipatt
```

```{r}
# | eval: false
head(asv_wide_avg_reps_matrix)
asv_wide_avg_rda_0 <- data.frame(t(asv_wide_avg_reps_matrix)) %>%
  rownames_to_column(var = "SAMPLE_NAME") %>%
  arrange(SAMPLE_NAME) %>%
  column_to_rownames(var = "SAMPLE_NAME") %>%
  as.matrix

clr_asv_matrix <- data.frame(compositions::clr(asv_wide_avg_rda_0))
# 
rownames(clr_asv_matrix)
rownames(metadata_normalized)
head(clr_asv_matrix)
```

```{r}

# clr_asv_matrix
# data = metadata_normalized, na.action = na.exclude)
```

```{r}
asv_abun <- indic_input %>%
  select(-SAMPLE_ID, -DEPTH, -TEMP, -SALINITY)
# head(asv_abun)

head(clr_asv_matrix)
head(metadata_normalized)
tmp_100_asv <- clr_asv_matrix %>% select(1:50)
tmp_100_meta <- metadata_normalized

# inc_sampleid <- multipatt(tmp_100_asv, tmp_100_meta$DEPTH, func = "r.g", control = how(nperm = 10))
# # inc_sampleid <- multipatt(asv_abun, indic_input$DEPTH_ORDER, func = "IndVal.g", control = how(nperm = 10))
# # inc_sampleid <- multipatt(asv_abun, indic_input$TEMP, func = "r.g", control = how(nperm = 100))
# 
# summary(inc_sampleid)
```

```{r}
# library(stats)
# output_stats <- as.data.frame(inc_sampleid$sign, keep.rownames = TRUE)
# # Add adj p-values
# output_stats[ ,p.value.bh = p.adjust(p.value, method="BH")]
# # range(output_stats$p.value)
# ??p.adjust()
```

`r.g` - indicator species test called a "point biserial correlation coefficient". It measures the correlation between two binary vectors.

```{r}
#extract table of stats
# indisp.sign<-as.data.table(indisp$sign, keep.rownames=TRUE)
# #add adjusted p-value
# indisp.sign[ ,p.value.bh:=p.adjust(p.value, method="BH")]
# #now can select only the indicators with adjusted significant p-values
# indisp.sign[p.value.bh<=0.05, ]
```

# Linear Discriminent Analysis Effect Size (LDA-LEfSe)

-   https://www.bioconductor.org/packages/release/bioc/vignettes/lefser/inst/doc/lefser.html

-   chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://www.bioconductor.org/packages/release/bioc/manuals/lefser/man/lefser.pdf

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("lefser")
library(lefser)

```

# CRT

```{r}
# install.packages("tree")
library(rpart)
library(tree)
```

```{r}
data("iris")
head(iris)
iris_df <- iris
colnames(iris_df) <- c("sepall", "sepalw", "petall", "petalw", "class")
```


CRT, on the other side, focuses on creating predictive models for classifying and prioritizing data as “key indicator variables”. CRT employs a decision tree approach, partitioning the dataset based on informative environmental variables to create a hierarchical structure of decision rules (Alkhasawneh et al., 2014). We used CRT to identify specific taxonomic groups as “key indicator variables” that show ability to indicate different DLs.

```{r}
xtabs( ~ class, data = iris_df)
```

```{r}
# execute tree classification
iris.tree <- tree::tree(class ~ ., data = iris_df)
# ?tree()
head(iris_df)
```

Above step fits a classification or regression tree with data. 

```{r}
summary(iris.tree)
# Only 3 of 4 variables were used for the tree.
# misclassification error rate? 
```

```{r}
plot(iris.tree); text(iris.tree, all = T)
```
For this classification tree, petall has a value < 2.45, therefore suggesting it has a more important classification. 
Sepall <5.15 indicates that is has a less important criterion

Next we "prune" the tree to reduce complexity. For this, we use a function to identify the number of significant notes. The cross validation will find the number of misclassifications.

```{r}
tree::cv.tree(iris.tree)
```

We prune the small branches. Therefore, we want to remove the "Branches" from the tree with the fewest data points. Therefore, we look at the output values of k and see when the values start to really increase. In this example, it is 3. A tree size of 3 is therefore suitable. THe _K_ is considered the "cost-complexity parameter" (which is the "Best" number of nodes/branches to include. )

Pruning the tree:
```{r}
iris.tree.prune <- prune.misclass(iris.tree, best = 3)
summary(iris.tree.prune)
```
```{r}
plot(iris.tree.prune);text(iris.tree.prune, all = T)
```
The tree is much less complex. 

### CRT all samples

```{r}
# head(asv_long_avg_wtax)
```


```{r}
# | eval: false
# input_CRT <- asv_long_avg_wtax %>% 
#   select(FeatureID, DEPTH, TEMP, SALINITY, OXYGEN, NH4, MEAN_REPS_seq) %>% distinct()
# 
# toy_input_CRT <- input_CRT %>% 
#   sample_n(1000)
# 
# test_tree <- tree(FeatureID ~ , toy_input_CRT)
# summary(test_tree)

# test_tree$frame
# plot(test_tree); text(test_tree, all = T)
```

```{r}
# install.packages("devtools")
# require("devtools")
# devtools::install_github(repo = 'zellerlab/siamcat')
```


# Session Information

```{r}
sessionInfo()
```
